### 2015-10-20 IAU_proceedings/ms.tex
adding ms

% This paper is part of the k2-paper project.
% Copyright 2015 Dan Foreman-Mackey (NYU) and the co-authors listed below.
%
%  RULES OF THE GAME
%
%  * 80 characters
%  * line breaks at the ends of sentences
%  * eqnarrys ONLY
%  * ``light curve'' not ``light-curve'' or ``lightcurve''
%  * Do not put in any comments that might get tweeted by @OverheardOnAph
%    (or maybe do put in a few....)
%  * ``percent'' (not \%) is a unit, as is ppm, so 5~percent.
%  * that is all.
%

\documentclass[11pt,preprint]{aastex}

\pdfoutput=1

\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0,0.5}
\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
            filecolor=linkcolor,urlcolor=linkcolor}
\usepackage{url}
\usepackage{amssymb,amsmath}
\usepackage{subfigure}
\usepackage{booktabs}

\usepackage{natbib}
\bibliographystyle{apj}

\newcommand{\project}[1]{\textsl{#1}} % hogg say
\newcommand{\kepler}{\project{Kepler}}
\newcommand{\KT}{\project{K2}}
\newcommand{\tess}{\project{TESS}}
\newcommand{\jwst}{\project{JWST}}
\newcommand{\license}{MIT License}
\newcommand{\paper}{\textsl{Article}}
\newcommand{\foreign}[1]{\emph{#1}}
\newcommand{\etal}{\foreign{et\,al.}}
\newcommand{\etc}{\foreign{etc.}}
\definecolor{mygreen}{rgb}{0, 0.50196, 0}
\newcommand{\response}[1]{{\color{mygreen} {\bf #1}}}
\newcommand{\figlabel}[1]{\label{fig:#1}}

\begin{document}

\title{%
	Stellar rotation periods inference with Gaussian processes
}

\newcommand{\oxford}{1}
\newcommand{\UW}{2}
\author{%
    Ruth~Angus\altaffilmark{1,\oxford},
    Suzanne~Aigrain\altaffilmark{\oxford}
    Daniel~Foreman-Mackey\altaffilmark{\UW},
}
\altaffiltext{1}         {To whom correspondence should be addressed:
                          \url{ruth.angus@astro.ox.ac.uk}}
\altaffiltext{\oxford}      {Subdepartment of Astrophysics, Denys-Wilkinson
	building, Keble road, Oxford, OX1 3RH}

\begin{abstract}

The light curves of spotted, rotating stars are often non-sinusoidal and
Quasi-Periodic (QP) and a sinusoid is therefore not a representative generative
model for stellar light curves.
Sine-fitting periodograms only provide the period of the best fitting
sinusoid, not the period of the best fitting rotating star model.
Ideally, a physical model of the stellar surface would be conditioned on the
data, however the parameters of such physical models can be highly degenerate.
Instead, we use an appropriate {\it effective} model: a Gaussian Process (GP)
with a QP covariance kernel function.
By modelling the covariance matrix of the light curve with a QP kernel
function, we remain agnostic about model choice, whilst sampling directly from
the posterior probability distribution function of the periodic parameter and
marginalising over the other kernel hyperparameters.
We simulated light curves with a range of rotation periods and spot lifetimes
and attempted to recover the rotation periods using three methods: our GP
method, a sine-fitting periodogram method and an AutoCorrelation Function (ACF)
method.
The GP method produces rotation period measurements that are more precise than
the periodogram and both more accurate and precise than the ACF method.

\end{abstract}

\section{Introduction}

\section{Method}

\section{Results}

\section{Conclusion}

\begin{figure}[p]
\begin{center}
\includegraphics{compare.png}
\end{center}
\caption{%
Measured vs true rotation periods for 300 simulations of light curves from
spotted, rotating stars.
Three different methods were tested: the ACF method, a Lomb-Scargle
periodogram (sine-fitting) method and my new GP method.
The GP method measures the most precise and accurate rotation periods
and is expected to perform even better on real data.
\figlabel{rotation}}
\end{figure}

{\it Facilities:} \facility{Kepler}

% \bibliography{GP}

\end{document}

### 2015-11-05 IAU_proceedings/ms.tex
adding proceedings docs

@@
-% This paper is part of the k2-paper project.
-% Copyright 2015 Dan Foreman-Mackey (NYU) and the co-authors listed below.
-%
-%  RULES OF THE GAME
-%
-%  * 80 characters
-%  * line breaks at the ends of sentences
-%  * eqnarrys ONLY
-%  * ``light curve'' not ``light-curve'' or ``lightcurve''
-%  * Do not put in any comments that might get tweeted by @OverheardOnAph
-%    (or maybe do put in a few....)
-%  * ``percent'' (not \%) is a unit, as is ppm, so 5~percent.
-%  * that is all.
-%
+% iaus2esa.tex -- sample pages for Proceedings IAU Symposium document class
+% (based on v1.0 cca2esam.tex)
+% v1.04 released 17 May 2004 by TechBooks
+%% small changes and additions made by KAvdH/IAU 4 June 2004
+% Copyright (2004) International Astronomical Union
 
-\documentclass[11pt,preprint]{aastex}
+\NeedsTeXFormat{LaTeX2e}
 
-\pdfoutput=1
+\documentclass{iau_FM}
+\usepackage{graphicx}
+\usepackage{natbib}
 
-\usepackage{color,hyperref}
-\definecolor{linkcolor}{rgb}{0,0,0.5}
-\hypersetup{colorlinks=true,linkcolor=linkcolor,citecolor=linkcolor,
-            filecolor=linkcolor,urlcolor=linkcolor}
-\usepackage{url}
-\usepackage{amssymb,amsmath}
-\usepackage{subfigure}
-\usepackage{booktabs}
+\title[Gaussian processes for stellar rotation]
+{Stellar rotation period inference with Gaussian processes}
 
-\usepackage{natbib}
-\bibliographystyle{apj}
+\author[Ruth Angus, Suzanne Aigrain \& Daniel Foreman-Mackey]
+{Ruth Angus$^1$,
+Susanne Aigrain$^1$
+\and Daniel Foreman-Mackey$^2$}
 
-\newcommand{\project}[1]{\textsl{#1}} % hogg say
-\newcommand{\kepler}{\project{Kepler}}
-\newcommand{\KT}{\project{K2}}
-\newcommand{\tess}{\project{TESS}}
-\newcommand{\jwst}{\project{JWST}}
-\newcommand{\license}{MIT License}
-\newcommand{\paper}{\textsl{Article}}
-\newcommand{\foreign}[1]{\emph{#1}}
-\newcommand{\etal}{\foreign{et\,al.}}
-\newcommand{\etc}{\foreign{etc.}}
-\definecolor{mygreen}{rgb}{0, 0.50196, 0}
-\newcommand{\response}[1]{{\color{mygreen} {\bf #1}}}
-\newcommand{\figlabel}[1]{\label{fig:#1}}
+\affiliation{$^1$Subdepartment of Astrophysics, University of Oxford, Oxford,
+	UK \\
+email: {\tt ruth.angus@astro.ox.ac.uk}, {\tt suzanne.aigrain@astro.ox.ac.uk} \\[\affilskip]
+$^2$Sagan Fellow, Department of Astronomy, University of Washington, Seattle
+\\ email: {\tt danfm@uw.edu}}
 
+\pubyear{2015}
+\setcounter{page}{1}
+\jname{Astronomy in Focus, Volume 1}
+\editors{Piero Benvenuti, ed.}
 \begin{document}
 
-\title{%
-	Stellar rotation periods inference with Gaussian processes
-}
+\maketitle
 
-\newcommand{\oxford}{1}
-\newcommand{\UW}{2}
-\author{%
-    Ruth~Angus\altaffilmark{1,\oxford},
-    Suzanne~Aigrain\altaffilmark{\oxford}
-    Daniel~Foreman-Mackey\altaffilmark{\UW},
-}
-\altaffiltext{1}         {To whom correspondence should be addressed:
-                          \url{ruth.angus@astro.ox.ac.uk}}
-\altaffiltext{\oxford}      {Subdepartment of Astrophysics, Denys-Wilkinson
-	building, Keble road, Oxford, OX1 3RH}
-
-\begin{abstract}
-
+% \begin{abstract}
 The light curves of spotted, rotating stars are often non-sinusoidal and
-Quasi-Periodic (QP) and a sinusoid is therefore not a representative generative
-model for stellar light curves.
-Sine-fitting periodograms only provide the period of the best fitting
-sinusoid, not the period of the best fitting rotating star model.
+Quasi-Periodic (QP) and a strictly periodic sinusoid is therefore not
+a representative generative model.
 Ideally, a physical model of the stellar surface would be conditioned on the
-data, however the parameters of such physical models can be highly degenerate.
+data, however the parameters of such models can be highly degenerate.
 Instead, we use an appropriate {\it effective} model: a Gaussian Process (GP)
 with a QP covariance kernel function.
-By modelling the covariance matrix of the light curve with a QP kernel
-function, we remain agnostic about model choice, whilst sampling directly from
-the posterior probability distribution function of the periodic parameter and
-marginalising over the other kernel hyperparameters.
-We simulated light curves with a range of rotation periods and spot lifetimes
-and attempted to recover the rotation periods using three methods: our GP
-method, a sine-fitting periodogram method and an AutoCorrelation Function (ACF)
-method.
-The GP method produces rotation period measurements that are more precise than
-the periodogram and both more accurate and precise than the ACF method.
+% \begin{equation}
+% 	k_{i, j} = A \exp \left[ - \frac{\sin^2(\pi(x_i-x_j)/P)} \
+% 	{2g^2}-\frac{(x_i-x_j)^2}{2l^2}\right],
+% \end{equation}
+% with
+By modelling the covariance matrix of the light curve with a QP GP, we remain
+agnostic about model choice, whilst sampling directly from the posterior
+probability distribution function of the periodic parameter and marginalising
+over the other kernel hyperparameters.
+We simulated 300 light curves with a range of rotation periods and spot
+lifetimes and attempted to recover the rotation periods using three methods:
+our GP method, a sine-fitting periodogram method and an AutoCorrelation
+Function (ACF) method.
+% Function (ACF) method \citep{amy}.
+The posterior probability distribution of the rotation period parameter was
+sampled using the affine invariant ensemble MCMC sampler, {\tt emcee}
+% \footnote{
+% Foreman-Mackey {\it et al.}, 2013, PASP, 125, 306}
+and the
+GP operations were performed using the {\tt george} python package
+% \footnote{https://github.com/dfm/george}.
+This method produces rotation periods that are more precise than the
+periodogram and both more accurate and precise than the ACF method.
+Furthermore, the improvement is expected to be even more dramatic when applied
+to real, noisy {\it Kepler} light curves, since the GP method is well suited to
+modelling rotation signals and correlated noise simultaneously.
 
-\end{abstract}
+\keywords{Stars: rotation, Methods: data analysis, Methods: statistical,
+techniques:photometric}
+% \end{abstract}
 
-\section{Introduction}
+% \firstsection % if your document starts with a section,
+%               % remove some space above using this command.
+% \section{Introduction}
 
-\section{Method}
-
-\section{Results}
-
-\section{Conclusion}
-
-\begin{figure}[p]
+\begin{figure}[b]
+% \vspace*{-2.0 cm}
 \begin{center}
-\includegraphics{compare.png}
+	\includegraphics[width=3in,trim={0 1.2cm 0 1.2cm},clip]{compare2.eps}
+% \vspace*{-1.0 cm}
+ \caption{Measured vs true rotation periods for 300 simulations of light curves
+	 from spotted, rotating stars.
+	 Three different methods were tested: the ACF method, a Lomb-Scargle
+	 periodogram (sine-fitting) method and our new GP method.
+	 The GP method measures the most precise and accurate rotation periods
+	 and is expected to perform even better on real data.}
+   \label{fig1}
 \end{center}
-\caption{%
-Measured vs true rotation periods for 300 simulations of light curves from
-spotted, rotating stars.
-Three different methods were tested: the ACF method, a Lomb-Scargle
-periodogram (sine-fitting) method and my new GP method.
-The GP method measures the most precise and accurate rotation periods
-and is expected to perform even better on real data.
-\figlabel{rotation}}
 \end{figure}
 
-{\it Facilities:} \facility{Kepler}
+% \begin{thebibliography}
 
-% \bibliography{GP}
+% \bibitem[Mcquillan {\it et al.} (2014)]{amy}
+% McQuillan, A., Mazeh, T. \& Aigrain, S., ApJ, 211, 24
+
+% \bibitem[Foreman-Mackey {\it et al.} (2013)]{emcee}
+% Foreman-Mackey, D., Hogg, D.~W., Lang, D., Goodman, J., 2013, PASP, 125, 306
+
+% \bibitem[Foreman-Mackey (2015)]{george}
+% Foreman-Mackey, D.
+
+% \end{thebibliography}
 

### 2016-01-18 code/kepler_injections.py
making plots of light curves to choose quiet ones

@@ -1,0 +1,73 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import mklc
+import pyfits
+import fitsio
+import scipy.interpolate as spi
+import george
+from george.kernels import ExpSine2Kernel, ExpSquaredKernel, WhiteKernel
+
+def simulate(id, x, pmin=.5, pmax=100., amin=1e-3, amax=1e-1, nsim=100,
+             gen_type="s", kepler=False, plot=False):
+    """
+    Simulate noise free light curves with the same time-sampling as the real
+    target that you are going to inject into.
+    pmin and pmax in days, amin and amax in ppm.
+    """
+    periods = np.exp(np.random.uniform(np.log(pmin), np.log(pmax), nsim))
+    amps = np.exp(np.random.uniform(np.log(amin), np.log(amax), nsim))
+    np.savetxt("simulations/true_periods.txt", np.transpose((np.arange(nsim),
+               periods, amps)))
+
+    for i, p in enumerate(periods):
+        print i, "of ", len(periods), "\n"
+        print "amps = ", amps[i]
+        if gen_type == "s":
+            res0, res1 = mklc.mklc(time, p=p)
+            nspot, ff, amp_err = res0
+            time, area_tot, dF_tot, dF_tot0 = res1
+            simflux = dF_tot0 / np.median(dF_tot0) - 1
+
+        elif gen_type == "GP":  # if performing sims with GPs
+            thetas = np.zeros((nsim, 5))
+            thetas[:, 0] = np.exp(np.random.uniform(-6, -4))
+            thetas[:, 1] = np.exp(np.random.uniform(4, 7))
+            thetas[:, 2] = np.exp(np.random.uniform(-1.5, 1.5))
+            thetas[:, 3] = np.exp(np.random.uniform(-11, -8))
+            thetas[:, 4] = periods
+            k = thetas[i, 0] * ExpSquaredKernel(thetas[i, 1]) \
+                    * ExpSine2Kernel(thetas[i, 2], thetas[i, 4]) \
+                    + WhiteKernel(std)
+            gp = george.gp()
+            gp.compute(time, yerr)
+            simflux = gp.sample()
+
+        np.savetxt("simulations/%s_%s.txt" % (str(int(i)).zfill(4), gen_type),
+                   np.transpose((time, simflux)))
+
+        if plot:
+            plt.clf()
[…]

### 2016-01-18 documents/ms.tex

\documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
\usepackage{cite, natbib}
% \documentclass[onecolumn, 12pt, usenatbib, preprint]{emulateapj}
\usepackage{float}
\usepackage{epsfig}
\usepackage{cases}
\usepackage[section]{placeins}
\usepackage{graphicx, subfigure}
\usepackage{color}

\newcommand{\article}{article}
\newcommand{\oxford}{1}
\newcommand{\uw}{2}
\newcommand{\sagan}{3}
\newcommand{\ktwo}{{\it K2}}
\newcommand{\kepler}{{\it Kepler}}

\begin{document}

\title{Stellar rotation period inference with Gaussian processes}

\author{%
   Ruth Angus\altaffilmark{\oxford}
   Suzanne Aigrain\altaffilmark{\oxford}
   \& Daniel Foreman-Mackey\altaffilmark{\uw}\altaffilmark{\sagan}
}

\altaffiltext{\oxford}{Subdepartment of Astrophysics, University of Oxford,
OX1 3RH, UK}
\altaffiltext{\uw}{Department of Astronomy, University of Washington, Seattle,
WA, USA}
\altaffiltext{\sagan}{Sagan fellow}

\begin{abstract}

blah
% Aims
% Methods
% Results

\end{abstract}

\keywords{blah}

\section{Introduction}
\label{Introduction}

\section{Method}
\label{sec:Method}

\subsection{Results and discussion}
\label{sec:discussion}

\section{Conclusions}
\label{sec:conclusions}

The data presented in this paper were obtained from the Mikulski Archive for
Space Telescopes (MAST).
STScI is operated by the Association of Universities for Research in Astronomy,
Inc., under NASA contract NAS5-26555.
Support for MAST for non-HST data is provided by the NASA Office of Space
Science via grant NNX09AF08G and by other grants and contracts.
This paper includes data collected by the Kepler mission.
Funding for the Kepler mission is provided by the NASA Science Mission
directorate.

\bibliographystyle{plainnat}
\bibliography{k2rotation}

\end{document}

### 2016-01-29 code/GProtation.py
making python3 compatible

@@ -5,7 +5,8 @@ import george
 from george.kernels import ExpSine2Kernel, ExpSquaredKernel, WhiteKernel
 import glob
 import emcee
-import triangle
+# import triangle
+import corner
 import h5py
 import subprocess
 from plotstuff import params, colours
@@ -14,12 +15,6 @@ cols = colours()
 import scipy.optimize as spo
 import time
 
-plotpar = {"axis.labelsize": 20,
-           "font.size": 20,
-           "legend.fontsize": 20,
-           "xtick.labelsize": 20,
-           "text.usetex": True}
-plt.rcParams.update(plotpar)
 
 def lnprior(theta, plims):
     """
@@ -31,6 +26,7 @@ def lnprior(theta, plims):
     and -20 < theta[3] < 20 and plims[0] < theta[4] < plims[1]:
         return 0.
     return -np.inf
+
 
 def Glnprior(theta, plims):
     """
@@ -100,7 +96,8 @@ def make_plot(sampler, xb, yb, yerrb, x, y, yerr, ID, DIR, traces=False,
     if tri:
         print("Making triangle plot")
         flat[:, -1] = np.exp(flat[:, -1])
-        fig = triangle.corner(flat, labels=fig_labels)
+        # fig = triangle.corner(flat, labels=fig_labels)
+        fig = corner.corner(flat, labels=fig_labels)
         fig.savefig("%s/%s_triangle" % (DIR, ID))
         print("%s/%s_triangle.png" % (DIR, ID))
 

### 2016-01-29 code/Kepler_ACF.py

@@ -11,7 +11,7 @@ import atpy
 import pylab
 import copy
 import glob
-from sets import Set
+# from sets import Set
 import collections
 no_rpy = True
 import scipy.io
@@ -114,7 +114,7 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
     max_psearch_len = len(lc_tab.flux) / 2.0
 
     # Calculate ACF
-    print 'Calculating ACF...'
+    print('Calculating ACF...')
 
     acf_tab, acf_per_pos, acf_per_height, acf_per_err, locheight, asym,  = \
         acf_calc(time = lc_tab.time, flux = lc_tab.flux, interval = gap_days, \
@@ -151,7 +151,7 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         n_s = 'k'
 
         period[x] = acf_peak_per[x]
-#         print 'PEAK RATIO = ', peak_ratio
+#         print('PEAK RATIO = ', peak_ratio)
 
         # plot period lines on full plot
         pylab.figure(1)
@@ -168,7 +168,7 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         pylab.xlim(period[x] * 10)
 
         # variability stats
-        print 'calculating var for P_med...'
+        print('calculating var for P_med...')
         amp_all[x], amp_per[x], per_cent, var_arr_real = \
             calc_var(kid = x, time_in = lc_tab.time, \
                      flux = lc_tab.flux, period = acf_peak_per[x])
@@ -186,8 +186,8 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         pylab.xlim(period[x] * 10)
 
         if saveplot:
-            print "saving figure", "%s/%s_full.png" % (savedir,
-                                                str(int(id_list[0])).zfill(4))
+            print("saving figure", "%s/%s_full.png" % (savedir,
+                                                str(int(id_list[0])).zfill(4)))
             pylab.savefig('%s/%s_full.png' %(savedir, str(int(id_list[0])).zfill(4)))
 
         maxpts = 40.0
@@ -195,8 +195,8 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
             maxpts = float(scipy.floor(lc_tab.time.max() / acf_peak_per[x]))
[…]

### 2016-01-29 code/filter.py

@@ -13,7 +13,7 @@ def filt1d(array, nmed, nlin, fill = False, circ = False):
     """Nonlinear (median+boxcar) filter with edge reflection."""
     nmed = 2 * int(nmed/2) + 1
     N = len(array)
-    if N < (3*nmed): 
+    if N < (3*nmed):
         return array
     # Check for NaNs
     lnan = numpy.isnan(array)
@@ -30,7 +30,7 @@ def filt1d(array, nmed, nlin, fill = False, circ = False):
         il = numpy.min(lnotnan)
         ih = numpy.max(lnotnan)
         xnew = numpy.arange(ih-il+1) + il
-        f = interpolate.interp1d(lnotnan, array[lnotnan])  
+        f = interpolate.interp1d(lnotnan, array[lnotnan])
         work[il:ih+1] = f(xnew)
         # ... or extending slope of nearest data points if at the edges
         if (il!=0):
@@ -44,7 +44,7 @@ def filt1d(array, nmed, nlin, fill = False, circ = False):
     # Edge reflection
     nr = min(20, nlin)
     sz = max([nmed, nlin])
-    if sz >= (N-1): 
+    if sz >= (N-1):
         sz = N-2
     if circ != False:
         wl = work[N-sz:]
@@ -63,7 +63,7 @@ def filt1d(array, nmed, nlin, fill = False, circ = False):
     work2[sz:sz+N] = work
     work2[sz+N:2*sz+N] = wr
     # Filter
-    if nmed > 1: work = signaltools.medfilt(work2, nmed) 
+    if nmed > 1: work = signaltools.medfilt(work2, nmed)
     else: work = work2
     box = scipy.signal.boxcar(2*nlin+1)
     work = signaltools.convolve(work, box) / float(2*nlin+1)
@@ -86,7 +86,7 @@ def NIF(array, nmed, nlin, nsig = 3.0, prefilt = False, \
     Irwin 2004)."""
     # Pre-filter if requested
     if prefilt != False:
-        wd = filt1d(array[:], 7, 3) 
+        wd = filt1d(array[:], 7, 3)
     else:
         wd = array[:]
     wd2 = wd
@@ -99,7 +99,7 @@ def NIF(array, nmed, nlin, nsig = 3.0, prefilt = False, \
             out = numpy.isnan(wd) + (abs(wd2-wd) > (nsig*sigma))
             irej = sum(out)
             wd = array[:]
-            print iter, irej, sigma
+            print(iter, irej, sigma)
[…]

### 2016-01-29 code/kepler_simulations.py

@@ -65,8 +65,8 @@ def run(pmin, pmax, amin, amax, nsim, plot=False):
         r_as.append(amps)
 
         for i, p in enumerate(periods):  # loop over periods
-            print i, "of", len(periods), "periods", number, "of", len(ids), \
-                    "stars"
+            print(i, "of", len(periods), "periods", number, "of", len(ids), \
+                    "stars")
             new_y = simulate(id, x, p, plot=True)  # sim
             new_std = np.std(new_y)
             new_y /= new_std

### 2016-01-29 code/mklc.py

@@ -23,7 +23,7 @@ def mklc(t, nspot=200, incl=(scipy.pi)*5./12., amp=1., tau=30.5, p=10.0):
               rotation period
     (unit of time is equatorial rotation period)'''
 
-    print 'Period = ', p
+    print('Period = ', p)
     dur = (max(t) - min(t))
 
     # (crude estimate of) total number of spots needed during entire
@@ -56,7 +56,7 @@ def mklc(t, nspot=200, incl=(scipy.pi)*5./12., amp=1., tau=30.5, p=10.0):
     pk = scipy.rand(nspot_tot) * (dur + 2 * extra) - extra
 
     # COMPUTE THE LIGHT CURVE
-    print "Computing light curve..."
+    print("Computing light curve...")
     time = numpy.array(t - min(t))
 
     area_tot = scipy.zeros_like(time)
@@ -99,9 +99,9 @@ def mklc(t, nspot=200, incl=(scipy.pi)*5./12., amp=1., tau=30.5, p=10.0):
     res1[2,:] = dF_tot
     res1[3,:] = dF_tot0
 
-    print 'Used %d spots in total over %d rotation periods.' % (nspot_tot, dur)
-    print 'Mean filling factor of individual spots was %.4f.' % ff
-    print 'Desired amplitude was %.4f, actual amplitude was %.4f.' \
-            % (amp, amp_eff)
-    print 'Desired number of spots at any one time was %d.' % nspot
+    print('Used %d spots in total over %d rotation periods.' % (nspot_tot, dur))
+    print('Mean filling factor of individual spots was %.4f.' % ff)
+    print('Desired amplitude was %.4f, actual amplitude was %.4f.' \
+            % (amp, amp_eff))
+    print('Desired number of spots at any one time was %d.' % nspot)
     return res0, res1

### 2016-01-29 code/mpfit.py

@@ -32,12 +32,12 @@ import numpy
       p1 = numpy.array([6,0.2])
       fa = {'x':x, 'y':y + noise, 'err':err}
       m = mpfit(mpfit.example_func, p1, functkw = fa)
-      print 'parameters = ', m.params
-      print 'parameters = ', m.pcerror 
+      print('parameters = ', m.params)
+      print 'parameters = ', m.pcerror
 # Note how the errors are underestimated because the two parameters
 # are strongly correlated!
       pylab.plot(x, mpfit.F(x, m.params))
-      
+
    The user function should be built according to a specific set of
    rules, which are described below and illustrated in the example_func
    function.
@@ -59,7 +59,7 @@ import numpy
    fjac, which is set if analytical derivatives are desired (see
    below). Additional keywords can be used to pass the data, and
    optionally measurement errors. These are passed to the user-defined
-   function via the functkw keyword to mpfit. 
+   function via the functkw keyword to mpfit.
 
    The user-defined function should return a 2-element list
    containing:
@@ -73,7 +73,7 @@ import numpy
    example are suggestive but not required. There are no restrictions
    on the number of dimensions in x, y or err, but the deviates
    must*be returned in a one-dimensional numpy array.
- 
+
    ANALYTICAL DERIVATIVES:
 
    In the search for the best-fit solution, mpfit.mpfit by default
@@ -83,7 +83,7 @@ import numpy
    faster than computing the derivatives analytically. However, if you
    wish to compute them analytically, then set the autoderivative
    keyword of mpfitfun to 0.
-  
+
    If analytical derivatives are requested, they should be returned in
    an M x N array, where M is the number of data points and N is the
    number of parameters. If this array was called dp, then dp[i,j]
@@ -93,11 +93,11 @@ import numpy
    length as p, with a value of 1 for a parameter which is free, and a
    value of zero for a parameter which is fixed (and hence no
    derivative needs to be calculated).
- 
+
    REFERENCES
 
    MINPACK-1, Jorge More, available from netlib (www.netlib.org).
-   "Optimization Software Guide," Jorge More and Stephen Wright, 
+   "Optimization Software Guide," Jorge More and Stephen Wright,
       SIAM, *Frontiers in Applied Mathematics*, Number 14.
    More', Jorge J., "The Levenberg-Marquardt Algorithm:
       Implementation and Theory," in *Numerical Analysis*, ed. Watson,
@@ -110,7 +110,7 @@ import numpy
[…]

### 2016-01-29 code/simple_acf.py

@@ -89,7 +89,7 @@ def make_plot(acf_smooth, lags, id, fn):
         if acf_smooth[peaks[0]] > acf_smooth[peaks[1]]:
             period = lags[peaks[0]]
         else: period = lags[peaks[1]]
-        print period
+        print(period)
 
         plt.clf()
         for i in peaks:

### 2016-01-29 code/Kepler_ACF.py
merging

@@ -186,9 +186,8 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         pylab.xlim(period[x] * 10)
 
         if saveplot:
-            print "saving figure", "%s/%s_full.png" % (savedir,
-                                                str(int(id_list[0])).zfill(4))
-            pylab.savefig('%s/%s_full.png' %(savedir, str(int(id_list[0])).zfill(4)))
+            print "saving figure", "%s/%s_full.png" % (savedir, id_list[0])
+            pylab.savefig('%s/%s_full.png' %(savedir, id_list[0]))
 
         maxpts = 40.0
         if scipy.floor(lc_tab.time.max() / acf_peak_per[x]) < maxpts:
@@ -210,8 +209,8 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         else: error = dlag_per_err[x]
 
         print 'PERIOD = ', period[x], '+/-', error,
-        print 'saving as', '%s/%s_result.txt'%(savedir, str(int(id_list[0])).zfill(4))
-        np.savetxt('%s/%s_result.txt'%(savedir, str(int(id_list[0])).zfill(4)),
+        print 'saving as', '%s/%s_result.txt'%(savedir, id_list[0])
+        np.savetxt('%s/%s_result.txt'%(savedir, id_list[0]),
                    np.transpose((period[x], error)))
     else:
         blank = np.array([0,0])

### 2016-01-29 code/kepler_simulations.py

@@ -94,6 +94,7 @@ def run(pmin, pmax, amin, amax, nsim, plot=False):
     r_as = np.array([i for j in r_as for i in j])
     data = np.vstack((np.arange(len(r_ps)), r_ps, r_as))
     np.savetxt("simulations/kepler_injections/true_periods_amps.txt", data.T)
+    np.savetxt("simulations/noise-free/true_periods_amps.txt", data.T)
 
 if __name__ == "__main__":
     pmin, pmax =.5, 100.

### 2016-01-29 code/recover_kepler_injections.py

@@ -11,6 +11,8 @@ import sys
 from simple_acf import simple_acf, make_plot
 from Kepler_ACF import corr_run
 from multiprocessing import Pool
+import glob
+from kepler_data import load_kepler_data
 
 plotpar = {'axes.labelsize': 22,
            'font.size': 22,
@@ -30,7 +32,7 @@ def my_acf(id, x, y, yerr, fn, plot=False, amy=False):
         period, period_err = corr_run(x, y, yerr, id, fn, saveplot=plot)
     else:
         period, acf, lags = simple_acf(x, y)
-        np.savetxt("{0}/{1}_result.txt".format(fn, str(id).zfill(4)), period)
+        np.savetxt("{0}/{1}_result.txt".format(fn, id, period))
         if plot:
             make_plot(acf, lags, id, fn)
     return period
@@ -41,7 +43,6 @@ def periodograms(id, x, y, yerr, fn, plot=False, savepgram=True):
     results.
     (the files are saved in a directory that is a global variable).
     """
-    id = str(int(id)).zfill(4)
     # initialise with acf
     try:
         p_init, _ = np.genfromtxt("{0}/{1}_result.txt".format(fn, id))
@@ -125,8 +126,7 @@ def acf_pgram_GP(id):
     """
     id = str(int(id)).zfill(9)
     p = "/home/angusr/.kplr/data/lightcurves"
-    fits_path = "{0}/{1}/*llc.fits".format(p, id)
-    fnames = np.sort(glob.glob(fits_path))
+    fnames = np.sort(glob.glob("{0}/{1}/*llc.fits".format(p, id)))
     x, y, yerr = load_kepler_data(fnames)  # load data
     path = "real_lcs"
     periodograms(id, x, y, yerr, path, plot=True)  # pgram
@@ -150,16 +150,18 @@ def acf_pgram_GP_sim(id):
 
 if __name__ == "__main__":
 
-    # noise-free simulations
-    N = 2
-    ids = range(N)
-    pool = Pool()
+#     # noise-free simulations
+#     N = 2
+#     ids = range(N)
+#     ids = [str(int(i)).zfill(4) for i in ids]
+#     pool = Pool()
 #     pool.map(acf_pgram_GP_sim, ids)
-    acf_pgram_GP_sim(0)
+#     acf_pgram_GP_sim(0)
 
 #     # noisy simulations
 #     N = 2
 #     ids = range(N)
[…]

### 2016-01-29 code/Kepler_ACF.py
removing zfill

@@ -186,9 +186,8 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         pylab.xlim(period[x] * 10)
 
         if saveplot:
-            print("saving figure", "%s/%s_full.png" % (savedir,
-                                                str(int(id_list[0])).zfill(4)))
-            pylab.savefig('%s/%s_full.png' %(savedir, str(int(id_list[0])).zfill(4)))
+            print("saving figure", "%s/%s_full.png" % (savedir, id_list[0]))
+            pylab.savefig('%s/%s_full.png' %(savedir, id_list[0]))
 
         maxpts = 40.0
         if scipy.floor(lc_tab.time.max() / acf_peak_per[x]) < maxpts:
@@ -210,8 +209,8 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         else: error = dlag_per_err[x]
 
         print('PERIOD = ', period[x], '+/-', error)
-        print('saving as', '%s/%s_result.txt'%(savedir, str(int(id_list[0])).zfill(4)))
-        np.savetxt('%s/%s_result.txt'%(savedir, str(int(id_list[0])).zfill(4)),
+        print('saving as', '%s/%s_result.txt'%(savedir, id_list[0]))
+        np.savetxt('%s/%s_result.txt'%(savedir, id_list[0])),
                    np.transpose((period[x], error)))
     else:
         blank = np.array([0,0])

### 2016-01-29 code/Kepler_ACF.py
debugging

@@ -210,7 +210,7 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
 
         print('PERIOD = ', period[x], '+/-', error)
         print('saving as', '%s/%s_result.txt'%(savedir, id_list[0]))
-        np.savetxt('%s/%s_result.txt'%(savedir, id_list[0])),
+        np.savetxt('%s/%s_result.txt'%(savedir, id_list[0]),
                    np.transpose((period[x], error)))
     else:
         blank = np.array([0,0])

### 2016-01-29 code/download_kepler_data.py
adding download_kepler_Data

@@ -3,8 +3,9 @@ import numpy as np
 import kplr
 client = kplr.API()
 
-data = np.genfromtxt("data/garcia.txt", skip_header=1).T
-kid = data[0]
+# data = np.genfromtxt("data/garcia.txt", skip_header=1).T
+# kid = data[0]
+kid = np.genfromtxt("data/quiet_kepler_ids.txt", skip_header=1).T
 for i, id in enumerate(kid):
     print(i, "of", len(kid))
     star = client.star(str(int(id)))

### 2016-09-02 code/plot_lcs.py
small changes

@@ -4,7 +4,7 @@ import matplotlib.pyplot as plt
 from plotstuff import params, colours
 cols = colours()
 reb = params()
-from simple_acf import dan_acf
+from simple_acf import dan_acf, simple_acf
 
 plotpar = {'axes.labelsize': 20,
            'text.fontsize': 20,
@@ -29,8 +29,9 @@ lilac = '#CC99FF'
 ids, ps, amps = np.genfromtxt("simulations/noise-free/true_periods_amps.txt").T
 id = "0006"
 x, y = np.genfromtxt("simulations/noise-free/{}.txt".format(id)).T
+period, _, _ = simple_acf(x, y)
 plt.clf()
-plt.plot(x, y, "k-", label = "$\mathrm{Period} = %.1f~\mathrm{days}$" % ps[int(id)])
+plt.plot(x, y, "k-", label = "$\mathrm{Period} = %.1f~\mathrm{days}$" % period)
 plt.xlabel("$\mathrm{Time~(days)}$")
 plt.ylabel("$\mathrm{Normalised~flux}$")
 plt.xlim(0, 200)

### 2016-09-02 documents/ms.tex

@@
 \newcommand{\ktwo}{{\it K2}}
-\newcommand{\kepler}{{\it Kepler}}
+\newcommand{\eg}{{\it e.g.}}
+\newcommand{\ie}{{\it i.e.}}
+\newcommand{\nlightcurves}{300}
+\newcommand{\Kepler}{{\it Kepler}}
+\newcommand{\kepler}{\Kepler}
 
@@
 \begin{abstract}
-
-blah
-% Aims
-% Methods
-% Results
+The light curves of spotted, rotating stars are often non-sinusoidal and
+quasi-periodic---spots are not static on the stellar surface and have finite
+lifetimes causing stellar flux variations to slowly shift in phase.
+A strictly periodic sinusoid is therefore not a representative generative
+model for the light curve of star with a rotational signal.
+Ideally, a physical model of the stellar surface would be conditioned on the
+data, however the parameters of such models can be highly degenerate
+\citep[\eg][]{Russell1906, Jeffers2009}.
+Instead, we use an appropriate {\it effective} model: a Gaussian Process (GP)
+with a quasi-periodic covariance kernel function.
+By modelling the covariance matrix of the light curve with a GP, a highly
+flexible semi-parametric function, we avoid the necessity to choose a `best
+fitting' functional form, whilst sampling directly from the posterior
+probability distribution function of the periodic parameter and marginalising
+over the other kernel hyperparameters.
+We used \nlightcurves\ simulated light curves with a range of rotation periods
+to test the GP model.
+We attempted to recover the rotation periods using three methods: our GP
+method, a sine-fitting periodogram method and an autocorrelation function
+method.
+The posterior probability distribution of the rotation period parameter was
+sampled using the affine invariant ensemble MCMC sampler {\tt emcee}
+\citep{Foreman-Mackey2013}, and the GP operations were performed using the
+{\tt george} python package \citep{George}.
+Rotation periods inferred via this method are more precise and accurate than
+the periodogram and ACF methods.
+Furthermore, the improvement is expected to be even more dramatic when applied
+to real, noisy {\it Kepler} light curves, since the GP method is well suited
+to modelling rotation signals and correlated noise simultaneously.
 
@@
 
-\keywords{blah}
+\section{Introduction}
 
-\section{Introduction}
-\label{Introduction}
+The light curves of spotted, rotating stars are often non-sinusoidal and
+Quasi-Periodic (QP).
+These stars vary in brightness due to active regions on their surfaces which
+rotate in and out of view.
+The non-sinusoidal quality is caused by the complicated surface spot patterns
+and the quasi-periodicity is produced both by the finite lifetimes of these
+active regions and the presence of differential rotation on the stellar
+surface.
+A strictly periodic sinusoid is therefore not a good model for stellar light
+curves.
+In an ideal world, a physical model of the stellar surface would be
+conditioned on the data.
+A physically realistic, generative model would perfectly capture the complexity
+of shapes within stellar light curves as well as the quasi-periodic nature,
+allowing for extremely precise probabilistic period recovery.
+However, such physical models require many free parameters in order to
+accurately represent a stellar surface and some of these parameters are
+extremely degenerate \citep[\eg][]{Russell1906, Jeffers2009}.
+In addition to global stellar parameters such as inclination and rotation
+period, each spot or active region should have (at minimum) a longitude,
+latitude, size, temperature and lifetime.
+Considering that many stars have on the order of hundreds of spots, the number
+of free parameters quickly becomes unwieldy, especially if the posterior PDFs
+of these parameters are explored with MCMC.
+Simplified spot models, such as the one described in \citet{Lanza2014} where
+only two spots are modelled, have produced successful results, however these
+simplified models sacrifice some precision due to lack of model flexibility.
+Instead of using a physical model for stellar light curves, we choose to use
+an {\it effective} model: one which captures the behaviour but is not
+physically motivated, although the parameters of this model may be {\it
+interpreted} as physical ones.
+An ideal effective model for the light curve of a spotted, rotating star is
+one with a small number of non-degenerate parameters that is flexible enough
+to perfectly capture non-sinusoidal, QP behaviour.
+These requirements are perfectly fulfilled by a Gaussian process (GP) model.
 
-\section{Method}
-\label{sec:Method}
+The standard methods used for measuring rotation periods include detecting
+peaks in a Lomb-Scargle \citep{Lomb1976, Scargle1982} (LS) periodogram
+\citep[e.g.][]{Reinhold2013}, Auto-Correlation Functions (ACFs)
+\citep{Mcquillan2013} and wavelet transforms \citep{Garcia2014}.
+The precision of the LS periodogram and wavelet methods are limited by the
+suitability of the model choice.
+A sinusoid is used in the case of the LS periodogram and the wavelet method
+relies on a choice of mother wavelet that is assumed to describe the data over
+a range of transpositions \citep[see, \eg][]{Carter2010}.
+In contrast, the ACF method is much better suited to signals that are
+non-sinusoidal.
+In fact, it doesn't matter what shape the signal is: as long as it is
+approximately periodic the ACF will display peaks located at the rotation
+period.
+A drawback of the ACF method however, is that it requires data to be
+evenly-spaced\footnote{\citet{Edelson1988} describe a method for computing
+ACFs for unevenly-spaced data.} which is not the case with \Kepler\ light
+curves (although in many cases it can be approximated as uniformly sampled).
+An ACF is also an operation performed on the data, not a generative model of
+the data and is therefore not inherently probabilistic.
+For this reason it is very difficult to estimate the uncertainty on an ACF
+rotation period measurement.
+Many rotation periods in the literature have been inferred by measuring the
+position of the first peak in an ACF, however this approach can be dangerous.
+The exponential decay in correlation can shift the peak position short-wards
+of its true value, leading to an underestimate of the period.
+We return to this point in section \textsection \ref{sec:discussion}.
+% In order to avoid this, a period should be inferred by modelling the entire
+% ACF, not just measuring the position of the first peak.
 
-\subsection{Results and discussion}
+The motivation for developing a GP-based method for rotation period inference
+is, firstly to measure more accurate and precise rotation periods using a
+better-suited generative model than a sinusoid for the reasons explained
+above.
+Secondly, in order to infer {\it probabilistic} periods, i.e. to estimate the
+posterior PDF of the rotation period and thereby measure a realistic and
+representative uncertainty.
+And thirdly, to allow for an additional correlated noise model to be included
+during regression, the parameters of which could be marginalised over.
+% And fourthly to provide some way of determining whether a periodic model is
+% supported by the data over a purely stochastic one.
+
+GPs are commonly used in the machine learning community and increasingly
+in other scientific fields, for example biology and geophysics (where GP
+regression is called `kriging').
+More recently, GPs have been used in the astronomy literature \citep[see
+\eg][]{Gibson2012, Haywood2014, Haywood2015, Evans2015, Rajpaul2015,
+Rajpaul2016, Aigrain2016}.
+% FIXME: CITATIONS Gibson, Aigrain, Evans, Rajpaul, Waldman(?), Cosmology,
+% geology, biology, Rasmussen & Williams, etc.
+They are useful in regression problems involving any stochastic process,
+specifically when the probability distribution for the process is a
+multi-variate Gaussian.
+If the probability of obtaining a data-set is a Gaussian in $N$ dimensions,
+where $N$ is the number of data points, that data-set can be described as, and
+with, a Gaussian process.
+An in-depth introduction to Gaussian processes is provided by
+\citet{Rasmussen2005}.
+
+GP models parameterise the covariance between data points and a kernel
+function provides the form of covariance matrix parameterisation.
+For example, take the time-series in figure \ref{fig:GP_example}.
+% This is a \kepler\ light curve of Earth-like planet host Kepler-452: a G-type
+% star that rotates once every $\sim$ 15 days.
+This is the \kepler\ light curve of KID 1430163, a relatively active star that
+rotates once every $\sim$ 4 days.
+The stochastic variability in this time-series is typical of \kepler\ FGK
+stars.
+Clearly, data points in this light curve are correlated.
+Points that are close together in time are tightly correlated and those that
+are far apart are loosely correlated.
+It is the way in which the covariance varies with the separation between data
+points that is modelled when using GP regression.
+It is not the data but the {\it covariance structure} that is modelled.
+This fact is what gives GPs their flexibility---they can model any time
+series with a similar covariance structure.
+In addition, a very simple function can usually capture the covariance
+structure of a light curve, whereas a much more complicated one may be
+required to model the time-series itself.
+% The light curve of \Kepler-452 has been modelled with a GP in figure
+% \ref{fig:GP_example}, shown in orange.
+The light curve of this star has been modelled with two different GP kernel
+functions in figure \ref{fig:GP_example}, shown in blue and pink.
+Both provide adequate fits to the data, however only the periodic kernel
+function, `QP' is a useful model because it has a periodic parameter.
+I return to this point shortly.
+
+A range of kernel functions could be used to describe stellar variability.
+For example, the simplest and most commonly used kernel function, the `Squared
+Exponential' (SE) produces an adequate fit to the KID 1430163 light curve.
+The SE kernel function is defined as,
+\begin{equation}
+k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2} \right).
+\end{equation}
+\label{eq:SE}
+Here $A$ is the amplitude of covariance, $l$ is the length scale of
+exponential decay and $x_i-x_j$ is the separation between data points.
+The SE kernel function has the advantage of being very simple---it has just
+two parameters, a covariance amplitude and length scale: $A$ and $l$.
+If $l$ is large two data points far apart in $x$ will be tightly correlated,
+and if small they will be loosely correlated.
+Another property of the SE kernel function is that it produces functions that
+are infinitely differentiable.
+It is therefore possible to model a data set and its derivatives
+simultaneously.
+% \citep[see \eg][]{Rajpaul2015}
+The SE kernel function may be an adequate model of the covariance in stellar
+light curves, but it is not a {\it useful} one for the problem of rotation
+period inference because it does not have a parameter that describes a period.
+In order to infer rotation periods it is necessary to use a periodic kernel
+function.
+For this reason, we use the `Quasi-Periodic' kernel function.
+The QP kernel function is defined as
+\begin{equation}
+k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2} -
+\frac{\sin^2(\frac{2\pi}{P})}{\Gamma^2} \right).
+\end{equation}
+\label{eq:QP}
+It is the product of the SE kernel function, which describes the overall
+covariance decay, and an exponentiated, squared, sinusoidal kernel function
+that describes the periodic covariance structure.
+$P$ can be interpreted as the rotation period of the star and $\Gamma$
+controls the number of zero crossings per period.
+This kernel function allows two data points that are separated in time by one
+rotation period to be tightly correlated, while points separated by half a
+period can be weakly correlated.
+Both the QP and SE kernel functions were used to produce the blue and pink
+models shown in figure ~\ref{fig:GP_example}.
+
+In order to infer a stellar rotation period from a light curve, we fit a GP
+model with a QP kernel function to the data.
+The likelihood of the model, conditioned on the data could then be maximised
+in order to find the maximum likelihood value for $P$.
+In this study however, the full posterior PDFs of the parameters are explored
+using MCMC.
+This latter approach comes at a cost: a GP model is expensive to compute once
+as it requires a matrix inversion and determinant evaluation, let alone the
+many thousands of times that is necessary to fully explore the posteriors of
+the parameters.
+However, fully exploring the posterior PDF of $P$, allows for an accurate
+estimate of the uncertainty on the rotation period.
+
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=6in, clip=true]{figures/Kepler452b.pdf}
+% \caption[A light curve with a GP model.]
+% {Light curve of Kepler-452 b, a habitable-zone, Earth-sized planet
+% hosting G star \citep{Jenkins2015}. The orange line shows a fit to the data using
+% a Gaussian process model with a QP covariance kernel function.}
+% \label{fig:GP_example}
+% \end{center}
+% \end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/001430163.pdf}
+\caption[A light curve with a GP model.]
+{Light curve of KID 1430163, an active star with a rotation period of $\sim$ 4
+days.
+The blue solid line shows a fit to the data using
+a Gaussian process model with a QP covariance kernel function and the pink
+broken line shows a fit with a SE kernel function.
+Both kernel functions provide an adequate fit to the data, however we use the
+QP function because it has a useful parameter that corresponds to the star's
+rotation period.}
+\label{fig:GP_example}
+\end{center}
+\end{figure}
+
+\section{Rotation period injection and recovery}
+
+In order to test the GP method we attempted to the measure rotation periods of
+a set of simulated light curves.
+We simulated \nlightcurves\ light curves using a spot model similar to that
+used
+% We used light curves simulated
+for the \citet{Aigrain2015} `hare and hounds' rotation period recovery
+experiment.
+These light curves were simulated by placing dark, circular spots with slowly
+evolving size, on the surface of bright, rotating spheres, ignoring
+limb-darkening effects.
+\citet{Aigrain2015} simulated one thousand light curves in order to test the
+ability of participating teams to recover both the stellar rotation periods
+and the rotational shear: the amplitude of surface differential rotation.
+% Since we are not interested in recovering differential rotation we only used
+% those light curves simulated {\it without} differential rotation, of which
+% there are \nlightcurves.
+Unlike in their study we are not interested in recovering differential
+rotation in this work, so did not include it in our simulations.
+We opted to use only solid-body rotators because differential rotation may
+produce some additional scatter in the rotation period measurements recovered.
+This code can also be adjusted to produce more realistic light curves by
+altering spot lifetimes.
+Stars with spot lifetimes that are long relative to their rotation periods
+will have highly periodic light curves.
+Those with short spot lifetimes will be more quasi-periodic.
+We fixed the mean spot lifetime at an arbitrary value of 30.5 days for all
+light curves in these initial tests.
+In future we plan to include both differential rotation and variable spot
+lifetimes in our light curve model.
+Light curves were simulated with a real \Kepler\ long-cadence time array, with
+one data point every thirty minutes over a four year duration.
+The rotation periods of the simulations were randomly drawn from a log-uniform
+distribution between 0.5-60 days.
+Figure \ref{fig:noise-free_lc} shows an example of a simulated, noise-free
+light curve with a period of 4.2 days.
+
+% The ranges and distributions of the physical stellar parameters used in the
+% simulated light curves are tabulated below, in table
+% \ref{tab:simulation_parameters}.
+
+% \begin{table*}
+% \begin{center}
+% \caption{Ranges and distributions of parameters used to simulate light curves
+% in \citet{aigrain2015}}
+% \begin{tabular}{lcc}
+% \hline\hline
+%     Parameter & Range & Distribution \\
+%     \hline
+%     Rotation period, $P_{rot}$ & 10 - 50 days (90\%) & log uniform \\
+%     & 1 - 10 days (10\%) & log uniform \\
+%     Activity cycle length & 1 - 10 years & log uniform \\
+%     Inclination & 0 - 90$^\circ$ & Uniform in $\sin^2i$ \\
+%     Decay timescale & (1 - 10) $\times P_{rot}$ & log uniform \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table*}
+% \label{tab:simulation_parameters}
+
+% and figure \ref{fig:period_hist} shows the
+% distribution of rotation periods in the \citet{aigrain15} sample.
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=6in, clip=true]{figures/acf_compare_noise-free_hist.pdf}
+% \caption{A histogram of the rotation periods used to generate the simulated
+% light curves in \citet{aigrain15}.}
+% \label{fig:period_hist}
+% \end{center}
+% \end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/noise-free_lc.pdf}
+% \includegraphics[width=6in, clip=true]{figures/thesis_plot.pdf}
+\caption[A simulated light curve.]
+{An example simulated, noise-free light curve. This `star' has a
+rotation period of 4.2 days.}
+\label{fig:noise-free_lc}
+\end{center}
+\end{figure}
+
+Initial tests were conducted on noise-free light curves, in order to provide
+proof-of-concept.
+We attempted to recover the rotation periods of these \nlightcurves\ light
+curves
+% , both with and without noise, (\ie before {\it and} after being injected
+% into \kepler light curves)
+using three rotation period recovery methods: the ACF method, the LS
+periodogram method and the GP method.
+
+\subsection{ACF}
+
+We calculated an ACF for each light curve using the method of
+\citet{Mcquillan2013}.
+In this method, an ACF is calculated for each light curve and smoothed by
+convolving with a Gaussian with $\sigma=9$ days.
+A rotation period is estimated as the lag-time of the first peak in the ACF,
+unless the second peak is larger in which case {\it that} lag-time is
+interpreted as the true period.
+The second peak in an ACF can be larger than the first if there are two active
+regions at or near opposite longitudes on the surface of the star---this would
+produce a light curve with two dips per rotation period.
+% By extension, if three active longitudes existed at 60$^\circ$ separations on
+% the stellar surface, the first peak in the ACF may be at one third of the true
+% rotation period and so on.
+An example ACF of the light curve in figure~\ref{fig:noise-free_lc} is shown
+in figure \ref{fig:ACF_example}.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/noise-free_acf.pdf}
+\caption[An ACF of a simulated light curve.]
+{The autocorrelation function of the simulated, noise-free light
+curve depicted in figure \ref{fig:noise-free_lc}.}
+\label{fig:ACF_example}
+\end{center}
+\end{figure*}
+
+The ACF method has proven to be extremely useful for measuring rotation
+periods.
+The catalogue of rotation periods of \Kepler\ stars provided in
+\citet{Mcquillan2013} has been widely used by the community and has provided
+ground-breaking results for stellar and exoplanetary science.
+The results of the ACF method as tested in \citet{Aigrain2015} were positive
+(see, for example their figure 8) as it produced a large number of accurate
+rotation period measurements.
+Another advantage of the ACF method is that it is conveniently fast to
+implement.
+
+We applied the ACF method to our sample of \nlightcurves\ noise-free,
+simulated light curves.
+Periods measured using the ACF method are plotted against the original
+rotation period values used to generate the light curves in figure
+\ref{fig:compare_noise_free}.
+% 73\% of the injected rotation periods were recovered with a value lying
+% within 10\% of the truth and 83\% within 20\%.
+A noteworthy feature of this figure is that many of the points fall below the
+$x=y$ line, \ie\ the recovered rotation periods are a little shorter than the
+true periods.
+We believe this is a feature of the peak position measurement method that is
+performed on the ACF.
+ACFs of stellar light curves can be roughly described as a cosine function
+super-imposed on top of a decaying exponential.
+In such a function the peak positions can be shifted towards the left--the
+short period end--because the decaying exponential raises the left side of
+each peak more than the right.
+It is possible to model this effect of course, however the standard practice
+is to simply measure the peak position without taking it into account.
+We are still investigating the origin and implications of this effect.
+% To demonstrate that this effect reproduces the underestimated periods seen in
+% figure \ref{fig:compare_noise_free}, we simulated 1000 ACFs by generating
+% cosine plus exponential functions with a range of periods, decay timescales
+% and relative amplitudes.
+% The measured position of the first peak in each simulated ACF is compared to
+% the true position in figure \ref{fig:exp_sine_test}.
+% The same underestimation of the peak position is apparent in this figure.
+% Dashed lines show harmonics of $2P_{\mathrm{rot}}$ and $1/2P_{\mathrm{rot}}$.
+% In several cases, twice the true rotation period is measured.
+% The RMS of residuals is 1.59 days.
+
+% \begin{figure*}
+% \begin{center}
+% \includegraphics[width=6in, clip=true]{figures/exp_sine_test.pdf}
+% \caption[A flaw in the ACF method.]
+% {Recovered peak position as a function of the `true' period used to
+% generate simulated ACFs. In many cases the position is measured short-wards of
+% the truth, as also seen in figure \ref{fig:compare_noise_free}. The dashed
+% lines are at $x=2x$, $x=x$ and $x=\frac{1}{2}x$.}
+% \end{center}
+% \end{figure*}
+% \label{fig:exp_sine_test}
+
+\begin{figure*}
+\begin{center}
+% \includegraphics[width=6in, clip=true]{figures/acf_compare_noise-free.pdf}
+\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
+\caption[ACF results.]
+{Rotation periods measured using the ACF method as a function of the
+`true' period used to generate \nlightcurves\ simulated, noise-free light
+curves.}
+\label{fig:compare_noise_free}
+\end{center}
+\end{figure*}
+
+\subsection{The sine-fitting periodogram method}
+
+For each simulated light curve, a Lomb-Scargle (LS) periodogram
+\citep{Lomb1976, Scargle1982} was computed over a grid of 10,000 periods,
+evenly spaced in frequency, between 1 and 100 days.
+The period of the highest peak in the periodogram was adopted as the rotation
+period.
+The resulting recovered rotation periods are plotted as a function of the true
+periods in figure \ref{fig:pgram_compare_noise_free}.
+These recovered rotation periods are in general more accurate than the ACF
+results: they do not systematically over or under predict rotation period.
+They are however less precise than the ACF results---the 8.03 day RMS of
+residuals is five times larger.
+The longer-period measurements provide the dominant contribution to this RMS.
+% We believe that the unsuccessful recoveries occurring preferentially for
+% shorter rotation periods are produced by additional, longer timescale
+% variations in some of the light curves.
+% Specifically, by fluctuations produced by changes in the total spot coverage
+% on the stellar surface.
+% These fluctuations can occasionally be larger in magnitude than those produced
+% by the rotational signal itself and are present in the majority of outlying
+% light curves that occupy the top left of figure
+% \ref{fig:pgram_compare_noise-free}.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
+\caption[LS periodogram results.]
+{Rotation periods measured using the LS periodogram method as a
+function of the `true' rotation period used to generate \nlightcurves\
+simulated light curves.}
+\label{fig:pgram_compare_noise_free}
+\end{center}
+\end{figure*}
+
+\subsection{The GP method}
+
+In order to recover the rotation periods of the simulated light curves using
+Gaussian processes, we sampled the posterior PDFs of the kernel
+hyperparameters described in equation \ref{eq:QP}.
+The likelihood function for a GP is similar to the simple Gaussian likelihood
+function that is used for optimisation problems where the uncertainties are
+Gaussian and uncorrelated.
+The latter can be written
+\begin{equation}
+\ln \mathcal{L} = -\frac{1}{2}\sum_{n=1}^N\frac{(y_n-\mu)^2}{\sigma_n^2}
+    - \frac{N}{2}\ln(2\pi\sigma_n^2),
+\end{equation}
+\label{eq:chi2}
+where $y_n$ are the data, $\mu$ is the mean model and $\sigma_n$ are the
+Gaussian uncertainites on the data.
+The equivalent equation in matrix notation is
+\begin{equation}
+\ln \mathcal{L} = -\frac{1}{2}\bf{r}^T\bf{C}^{-1}\bf{r}-\ln|\bf{C}|
+    + \mathrm{constant},
+\end{equation}
+\label{eq:lhf1}
+where $\bf{r}$ is the vector of residuals and $\bf{C}$ is the covariance
+matrix,
+\begin{eqnarray}
+	\mathbf{C} &=& \left (\begin{array}{cccc}
+	\sigma^2_1 & \sigma_{2, 1} & \cdots & \sigma_{N, 1} \\
+	\sigma_{1, 2} & \sigma^2_2 & \cdots & \sigma_{N, 2} \\
+    && \vdots & \\
+	\sigma_{1, N} & \sigma_{2, N} & \cdots & \sigma^2_N
+\end{array}\right )
+\end{eqnarray}
+In the case where the uncertainties are uncorrelated, the noise is `white',
+(which is a frequent assumption made by astronomers and is sometimes
+justified) and the off-diagonal elements of the covariance matrix are zero.
+However, in the case where there is evidence for correlated
+`noise'\footnote{In our case the `noise' is actually the model!}, as in the
+case of \Kepler\ light curves, those off-diagonal elements are non-zero.
+With GP regression, a covariance matrix generated by the kernel function,
+${\bf K}$ replaces ${\bf C}$ in the above equation.
+Incidentally, this approach is the reverse of the regression techniques
+usually employed by astronomers.
+In most problems in astronomy one tries to infer the parameters that describe
+the mean model and, if correlated noise is present, to marginalise over that
+noise.
+Here, the parameters describing the correlated noise are what we are
+interested in and our mean model is simply a straight line at $y=0$.
+
+One could either maximise this likelihood function in order to find the
+best-fit hyper-parameters of the covariance kernel function or, as in our
+approach, sample from the posterior PDFs of the hyper-parameters.
+The advantage of the maximum-likelihood method is that the best-fit parameters
+will be found much faster, but the uncertainities on the rotation period
+will not be constrained.
+In order to measure accurate uncertainties, the posterior PDFs of the
+parameters must be explored.
+Since obtaining accurate uncertainties on rotation periods is one of our main
+motivations for this method, we use MCMC despite the added computational
+expense.
+
+We use the ACF period as an initial guess for the rotation period (this
+decision is discussed further in section \textsection \ref{sec:discussion}).
+We use a uniform prior over rotation period with bounds described below and
+assert that the covariance decay timescale parameter, $l$ must be greater than
+the rotation period.
+This represents our assumption that the evolution timescale of active regions
+is greater than stellar rotation periods.
+This assumption may be more valid for late spectral types---hot stars have
+smaller active regions which are likely to be shorter-lived.
+For the remaining hyperparameters we use the following initial values and
+log-uniform prior distributions:
+\begin{eqnarray}
+ 	&	A_{initial} = e^{-5}, \sim \exp(U[-20:20]) \\ \nonumber
+ 	&	l_{initial} = e^{7}, \sim \exp(U[-20:20]), l<P \\ \nonumber
+ 	&	\Gamma_{initial} = e^{0.6}, \sim \exp(U[-20:20]) \\ \nonumber
+ 	&	\sigma_{initial} = e^{-16}, \sim \exp(U[-20:20]) \\ \nonumber
+ 	&	P_{initial} = P_{ACF}, \sim U([1 - 0.4]P_{ACF}<P_{ACF}<[1 +
+0.4]P_{ACF}).
+ \end{eqnarray}
+ \label{eq:initialisation}
+$\sigma$ is an additional white noise term added to the diagonal elements of
+the covariance matrix. It is the fraction by which the observational
+uncertainties have been underestimated.
+If the errorbars reported on the data are too small, this parameter will be
+non-zero.
+In practice this parameter should always be non-zero when performing GP
+regression.
+This is because the covariance matrix must be positive definite, however
+matrix inversion performed using most solvers is approximate, not exact,
+therefore slight deviations from positive definitism can arise.
+Including a small amount of extra variance in the model allows enough
+flexibility that the matrix inversion algorithms do not run into numerical
+issues.
+
+We subsampled the light curves in order to reduce computation time.
+The subsampling is altered for each light curve and depends, again, on the ACF
+period estimate.
+We found that retaining only 20 data points per rotation period, based on the
+ACF period estimate, significantly reduced computation time whilst maintaining
+performance.
+The {\tt george} \citep{George} python package was used to implement the GP
+model which uses the fast matrix solver, HODLR \citep{Ambikasaran2014}.
+Matrix operations performed with HODLR scale as $N \log ^2 (N)$, where $N$ is
+the number of data points.
+We used {\tt emcee} \citep{Foreman-Mackey2013}, an affine invariant ensemble
+sampler to explore the posterior PDFs of the model parameters.
+The rotation period was taken as the median value of the posterior PDF.
+The resulting measured rotation periods are compared to the true rotation
+periods in figure \ref{fig:GP_compare_noise_free}.
+The root-mean-square (RMS) of residuals for the GP results is 0.26 days.
+The results of all three methods are plotted on the same axes in figure
+\ref{fig:compare_noise_free}.
+
+% The subsampling reduces computation time simply by contracting the data
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare_gp.pdf}
+\caption[GP results.]
+{Measured vs true rotation period for \nlightcurves\ simulated light
+curves using the GP method.}
+\label{fig:GP_compare_noise_free}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare2.pdf}
+\caption[All results.]
+{Measured vs true rotation period for \nlightcurves\ simulated light
+curves from the three different methods described in the text.}
+\label{fig:compare_noise_free}
+\end{center}
+\end{figure*}
+
+The marginal posterior distributions of the QP kernel hyperparameters for an
+example noise-free light curve  with a 14.4 day rotation period, are shown in
+figure \ref{fig:gp_posteriors}.
+The light curve itself with the best fit GP model is shown in figure
+\ref{fig:demo_lc_GP}.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/demo_lc_GP.pdf}
+\caption{A simulated light curve with a rotation period of 14.4 days.
+The pink curve shows the GP model with the best-fit parameters.
+The marginal posteriors of the GP hyper-parameters used to produce this fit
+are shown in figure \ref{fig:gp_posteriors}.}
+\label{fig:demo_lc_GP}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/demo_triangle.pdf}
+\caption{Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
+additional white noise term added to the diagonal elements of the covariance
+matrix. It is the fraction by which the observational uncertainties have been
+underestimated. If the errorbars reported on the data are too small, this
+parameter will be non-zero. Since this light curve was simulated without
+noise, this parameter is very close to zero.
+The true rotation period of this star is 14.4 days.}
+\label{fig:gp_posteriors}
+\end{center}
+\end{figure*}
+
+\subsection{Real \kepler\ data}
+
+The noise-free light curves were injected into real \kepler\ light curves in
+order to test the performance of the GP method in a realistic case.
+Ideally, simulated light curves would have been injected into the \kepler\
+pixel-level data, then the same detrending method that is applied to the
+\kepler\ light curves would be applied to these light curves before attempting
+to measure rotation periods for these stars.
+However, since this work is designed to be more of a demonstration of the
+overall efficacy of the GP periodogram, rather than the development of a
+\Kepler\ specific method, this is beyond the scope of our demonstrations.
+
+Unfortunately, automating the GP method so that it can run on large numbers of
+real \kepler\ light curves is difficult.
+The added noise reduces the acceptance fraction within the MCMC chains,
+significantly increasing the required time for convergence.
+Despite this, the method works well on individual targets where chains can be
+run for a long time and subsampling can be reduced.
+We used a simulated, noisy light curve from \citet{Aigrain2015} that was
+generated by injecting a noise-free simulation into a real \kepler\ light
+curve in order to preserve realistic \kepler\ noise properties.
+
+\kepler\ light curves are big data and extracting the maximum amount of
+information from them requires either large numbers of CPU hours, or using
+work-arounds.
+One feature of \kepler\ light curves that works to our advantage is that they
+are naturally broken up into smaller time units by quarterly (three month)
+breaks.
+Splitting the data set into quarters, rather than modelling the entire time
+series contiguously speeds up the computation time as inverting several
+small matrices is faster than inverting one large one.
+The \kepler\ quarter divisions are natural places to split the time series
+because the spacecraft rotates by ninety degrees every quarter (three months),
+placing each star on a new CCD module.
+Pixel response functions and background flux differs from pixel-to-pixel and
+module-to-module, so noise properties of \kepler\ light curves change every
+quarter.
+Additionally, changes in the spacecraft's orientation and position during
+quarterly re-pointings temporarily affect the temperature of the CCD,
+producing systematic features in the light curves at the start of some
+quarters.
+We model each quarter separately but the parameters of the GP kernel function
+are global, {\it i.e.} we do not use a separate period parameter for each
+quarter---there is just one period parameter for an entire light curve.
+It would be possible to model the time series with a mixture of some global
+parameters and some quarter-specific parameters, for example one might expect
+that the amplitude of covariance, $A$ or white noise level, $\sigma$ to vary
+on a quarterly basis.
+However, since there are seventeen quarters this would lead to thirty-seven
+parameters, and in the interest of minimising computation time (adding
+parameters leads to longer MCMC burn in and convergence time), we choose to
+use global parameters only.
+Unfortunately, the application of this method to noisy test cases is still
+under development and cannot provide results here.
+We hope to develop this method further and apply it to real light curves in
+the near future.
+
+% The marginal posterior distributions of the QP kernel hyperparameters for an
+% example noisy light curve  with a ... day rotation period, are shown in
+% figure \ref{fig:gp_posteriors}.
+% The light curve itself with the best fit GP model is shown in figure
+% \ref{fig:demo_lc_GP}.
+
+% \begin{figure*}
+% \begin{center}
+% \includegraphics[width=6in, clip=true]{figures/demo_lc_GP.pdf}
+% \caption{A simulated light curve with a rotation period of 14.4 days.
+% The pink curve shows the GP model with the best-fit parameters.
+% The marginal posteriors of the GP hyper-parameters used to produce this fit
+% are shown in figure \ref{fig:gp_posteriors}.}
+% \label{fig:demo_lc_GP}
+% \end{center}
+% \end{figure*}
+
+% \begin{figure*}
+% \begin{center}
+% \includegraphics[width=6in, clip=true]{figures/demo_triangle.pdf}
+% \caption{Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
+% additional white noise term added to the diagonal elements of the covariance
+% matrix. It is the fraction by which the observational uncertainties have been
+% underestimated. If the errorbars reported on the data are too small, this
+% parameter will be non-zero. Since this light curve was simulated without
+% noise, this parameter is very close to zero.
+% The true rotation period of this star is 14.4 days.}
+% \label{fig:gp_posteriors}
+% \end{center}
+% \end{figure*}
+
+% \begin{figure*}
+% \begin{center}
+% \includegraphics[width=6in, clip=true]{figures/prediction.pdf}
+% \caption{A noisy simulated light curve with a rotation period of 17.04 days.
+% The blue curve shows the GP model with the best-fit parameters.}
+% \end{center}
+% \end{figure*}
+% \label{fig:compare_noise_free}
+
+\section{Discussion}
 \label{sec:discussion}
 
+The main drawback of the GP method is computation time.
+Because GPs are so expensive to compute, it is necessary to come up with
+shortcuts in order to perform inference on \kepler\ light curves, each of
+which compromises accuracy.
+The shortcuts that we employ here include:
+\begin{itemize}
+\item{Initialisation.
+The choice of initial parameter values makes a difference to computation time:
+the closer they are to the maximum likelihood parameters the shorter the MCMC
+burn-in time.
+Using the ACF period to initialise is not ideal as results will not be fully
+independent of the ACF period unless MCMC chains are run for an infinitely
+long time.} \item{Subsampling.
+Not only is some information always lost by subsampling, but the choice of
+subsampling frequency will also effect the results.
+We chose to retain only 20 data points per period --- based on the ACF
+period --- which, again means that our results are not independent of the ACF
+results.}
+\item{Priors.
+We chose to use uniform, bounded priors to reduce parameter space and
+therefore convergence time.
+Unfortunately, if the true period lies outside the bounded region, it will
+never be found.}
+\end{itemize}
+
+These shortcuts are only necessary when working with a large number of stars
+(of order hundreds or thousands).
+If only interested in single or small numbers of targets it may be possible to
+reduce the extremity of these shortcuts or to avoid them altogether.
+
+\subsection{The ACF method}
+The main advantage of the ACF method over the GP method is speed: it is
+extremely fast and therefore useful to get a quick estimate of a period.
+We have demonstrated that the ACF method may produce rotation periods that are
+slightly systematically biased towards lower rotation periods.
+This effect comes from the process of extracting the period from the ACF
+itself: typically superimposed onto a decaying exponential, that first peak in
+the ACF gets shifted towards faster periods.
+One may be able to get around this by modelling the ACF as a sum of a cosine
+and exponential function (a similar function to GP kernel function) but
+in practice ACFs tend to have a more complicated structure and it is not
+possible to model them with a simple function.
+This observation leads to the idea of modelling the ACF with a Gaussian
+processes, which then logically progresses to a using a Gaussian process with
+a Gaussian process kernel function.
+Although this may be an attractive idea, in practice a GP would not
+necessarily produce a positive semi-definite covariance kernel\footnote{One
+that can be inverted; a necessary regression operation.}.
+An alternative approach may be to correct for rotation period bias after the
+peak position has been measured by inferring a correction factor.
+Yet another approach could be to measure the mean separation between adjacent
+peaks in the ACF: the first peak should be most strongly affected by the
+exponential decay, the second peak less so, and so on.
+
+\subsection{Initialisation}
+Using the ACF to initialise the MCMC chains is not ideal because, of course,
+one becomes reliant on the assumption that the ACF period is close to the true
+period.
+Of course, if you had infinite CPU time, this would not be a problem as you
+would eventually sample the entire posterior PDF of the period parameter,
+however in practice this is likely to be an issue.
+The only way to get around this problem is to run the MCMC chains for as long
+as possible, or, alternatively to use a sampler that is designed to move
+around the parameter space much more quickly than {\tt emcee}; nested sampling
+for example.
+We chose to use the ACF periods rather than the periodogram periods to
+initialise as, although there was some systematic bias present in these
+results, that bias was small and there were fewer large outliers.
+
+\subsection{Kernel function choice and interpretation of hyper-parameters}
+The QP kernel function represents a simplistic effective model of a stellar
+light curve.
+It adequately describes the data, captures that all-important periodic quality
+and is relatively simple, with only a few hyper-parameters.
+It also satisfies the requirement to produce positive semi-definite covariance
+matrices.
+Whilst the QP kernel function evidently captures the periodic qualities of
+light curves adequately, it is still a somewhat arbitrary choice.
+Another valid choice would be a squared cosine function multiplied by a
+squared exponential,
+\begin{equation}
+k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2}\right)
+\cos^2\left(\frac{2\pi}{P}\right)
+\end{equation}
+\label{eq:cos_kernel}
+This function produces a positive semi-definite matrix and has the $P$
+parameter of interest.
+It may in fact be even {\it more} suited to modelling stellar
+light curves as it describes a Gaussian in frequency space.
+It is easy to imagine a differentially rotating star with a period that is a
+Gaussian in frequency space: the mean frequency would be the frequency at the
+most active latitude, at or near which spots spend the majority of their time
+and the tails would be occupied by spots that drift near the equator or poles.
+The main difference between this cosine and the QP function is that the cosine
+function allows negative covariances and the QP function does not.
+Is is realistic to allow negative covariances?
+In practice, the ACFs of \Kepler\ light curves often go negative.
+However, many stars have two active regions on opposite hemispheres that
+produce two brightness dips per rotation.
+If the covariance is forced to be negative for two data points that are
+separated by half a rotation period, those light curves with two peaks per
+rotation period may not be well modelled.
+It would be very worthwhile to test this assumption and this alternative
+kernel function in future.
+If CPU time were not limited there may be some benefit to performing formal
+model comparison with different kernel functions.
+However, the evidence integral is an ambitious calculation for models with
+likelihood functions that take milliseconds to compute, let alone those
+involving GPs with light curves containing thousands of data points which can
+take minutes.
+
+Clearly, stellar rotation periods are well represented by the $P$ parameter in
+our QP kernel function, as evidenced by the its impressive ability to recover
+the true rotation periods from the simulated light curves in this work.
+However, it is not clear whether the QP kernel function is the {\it best}
+function to use.
+There may be an alternative function which is better suited to capturing
+stellar variability and is able to recover periods even more precisely than
+the QP kernel.
+There may also be an alternative function which is more physically motivated,
+that captures not just the rotation period but also (for example) the spot
+lifetime or differential surface rotation.
+This is beyond the scope of this work but we hope that these questions will be
+answered in the near future.
+
+\subsection{Future work}
+The next stages of improving this method are listed as follows:
+\begin{itemize}
+\item{To optimise the tradeoff between computational efficiency and accuracy.
+We have performed limited tests to explore the optimum subsampling strategy
+and prior bounds.
+These should be explored more thoroughly in future.
+For example, it would be useful to know how the precision and accuracy of the
+recovered rotation period varies as a function of subsampling frequency}
+\item{To perform model selection with different kernel functions. Again, these
+tests have not been performed due to the cost of calculating the fully
+marginalised likelihood with GPs. This calculation may be prohibitively
+expensive, however simpler model selection tests can be performed. For
+example, the relative precision of rotation periods recovered using
+alternative kernel functions could be tested.}
+\item{To design and implement a physically motivated kernel function. We have
+only explored the physical interpretation of one parameter in our kernel
+function, $P$.
+However, the other parameters may also be related to some physical processes.
+For example, the overall timescale for covariance fall-off, $l$ may be
+related to spot lifetimes.
+A star with long spot lifetimes will show little variation in the overall
+shape and amplitude of its light curve between rotations and $l$ will be
+large.
+In contrast, the light curve of a star with short spot lifetimes may display
+non-repeating patterns and amplitudes that vary rapidly between rotations.
+In this case $l$ will be small.
+The $\Gamma$ parameter is related to the number of zero crossings within one
+rotation period: when $\Gamma$ is small there are many zero crossings and
+vice versa.
+Since the number of zero crossings per rotation period is related to the
+number of active regions on the surface of the star, this parameter may also
+be of physical interest.
+In addition, instead of interpreting the parameters of the QP kernel function
+used here, it may be possible to design an entirely new kernel function, based
+on the physical processes that drive the light curve variability.
+This idea is being explored by another member of my research group.}
+\item{To develop a detection criterion to assess whether a rotation period was
+measured.
+Another general problem in rotation period inference is deciding whether a
+real rotation period was measured at all.
+Detection thresholds are usually set using the amplitudes of peaks in LS
+periodograms or ACFs in order to eliminate contaminants.
+Using a GP model it may be possible to do this via model selection: \ie\ are
+the data better described by a periodic model, or a non-periodic model?
+There is more than one way to implement such a model comparison, one way could
+be to use a sum of two kernel functions: one periodic and one non-periodic,
+each with its own amplitude parameter.
+If there is periodicity in the data, the amplitude of the periodic kernel will
+be larger than the amplitude of the non-periodic kernel and vice versa.}
+\item{To build in a noise model for \kepler\ data.
+Another huge advantage of the GP method is, because it is a {\it generative}
+model of the data, the rotation period signal can be modelled at the same time
+as systematic noise features.
+One can then marginalise over the parameters of the noise model.
+This approach would be extremely advantageous for \kepler\ data since
+long-term trends are often removed by the \kepler\ detrending pipeline.
+Marginalising over the noise model at the same time as inferring the
+parameters of interest will insure that the periodic signal is preserved.}
+\end{itemize}
+
 \section{Conclusions}
-\label{sec:conclusions}
+
+We simulated \nlightcurves\ noise-free, \kepler-like light curves using a spot
+model and attempted to recover the rotation periods used to generate them.
+Three methods were compared: a LS periodogram method, an ACF method and our GP
+method.
+The GP method produced the most precise and accurate rotation periods of the
+three techniques.
+This is because a GP is a semi-parametric model that is well suited to signals
+that are non-sinusoidal and quasi-periodic, because it does not require the
+assumption of regular time sampling and because a full posterior PDF of the
+rotation period parameter is produced.
+Not only this, unlike the other two methods, the GP method provides accurate
+rotation period uncertainties.
+
+The main drawback of this method is the computation time.
+We demonstrated that the GP method is capable of recovering signals from
+noise-free light curves, but were unable to automate the GP method for light
+curves with realistic noise properties.
+It may be that this method must remain a `boutique' method: ideal for small
+numbers of targets but not well suited to the entire \kepler\ catalogue.
+We continue to develop this method and remain hopeful that we can, eventually
+infer a probabilistic rotation period for thousands of \kepler\ stars.
 
@@
 \bibliographystyle{plainnat}
-\bibliography{k2rotation}
+\bibliography{GProtation}
 

### 2016-09-02 ruth/simulate_GP.py
adding simulate_GP

@@ -1,0 +1,34 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import george
+from george.kernels import ExpSine2Kernel, ExpSquaredKernel, WhiteKernel
+
+
+def generate_GP(theta, x, yerr):
+    """
+    Using the parameters provided in theta, generate a GP with x time sampling.
+    theta: array of log(A), log(l), log(g), log(sigma), log(P).
+    x: array of time values.
+    yerr: array of y uncertainties
+    """
+    theta = np.exp(theta)
+    k = theta[0] * ExpSquaredKernel(theta[1]) \
+            * ExpSine2Kernel(theta[2], theta[4]) + WhiteKernel(theta[3])
+    gp = george.GP(k, solver=george.HODLRSolver)
+    gp.compute(x, np.sqrt(theta[3]+yerr**2))
+    return gp.sample(x)
+
+
+if __name__ == "__main__":
+    x = np.arange(0, 20, .1)
+    yerr = np.ones_like(x) * 1e-5
+
+    theta = np.log([np.exp(-5), np.exp(7), np.exp(.6), np.exp(-16), 4])
+    ys = generate_GP(theta, x, yerr)
+
+    data = np.vstack((x, ys))
+    np.savetxt("simulations/0000_gp.txt", data.T)
+
+    plt.clf()
+    plt.errorbar(x, ys, yerr=yerr, fmt="k.")
+    plt.savefig("simulations/0000_gp")

### 2016-11-01 gprotation/comparison_plots.py
making more compatible with gp runs

@@ -105,22 +105,10 @@ def make_new_df(truths, R_DIR):
     return truths_s
 
 
-def mcmc_plots(truths, DIR):
+def mcmc_plots(N, true, med, med_errp, med_errm, maxlike, amp, DIR):
     """
     Plot the GP results.
     """
-
-    truths_e = make_new_df(truths, DIR)
-    m = (truths_e.DELTA_OMEGA.values == 0) \
-            * (truths_e.acf_period.values > 0)
-
-    N = truths_e.N.values[m]
-    true = truths_e.P_MIN.values[m]
-    med = np.exp(truths_e.sigma.values[m])  # period and sigma names swapped
-    med_errp = np.exp(truths_e.sigma_errp.values[m])
-    med_errm = np.exp(truths_e.sigma_errm.values[m])
-    maxlike = np.exp(truths_e.sigma_max[m])
-    amp = truths_e.AMP.values[m]
 
     plt.clf()
     plt.plot(truths_e.NSPOT, np.exp(truths_e.gamma_max), "k.")
@@ -165,20 +153,10 @@ def mcmc_plots(truths, DIR):
     return (np.median((maxlike - true)**2))**.5
 
 
-def acf_plot(truths, DIR):
+def acf_plot(truths, N, true, acfs, acf_errs, DIR):
     """
     Plot the acf results.
     """
-    truths_e = make_new_df(truths, DIR)
-    m = (truths_e.DELTA_OMEGA.values == 0) \
-            * (truths_e.acf_period.values > 0)
-
-    N = truths_e.N.values[m]
-    true = truths_e.P_MIN.values[m]
-    acfs = truths_e.acf_period.values[m]
-    acf_errs = truths_e.acf_period_err.values[m]
-    amp = truths_e.AMP.values[m]
-
     # acf plot
     plt.clf()
     xs = np.arange(0, 100, 1)
@@ -197,19 +175,10 @@ def acf_plot(truths, DIR):
     plt.savefig(os.path.join(DIR, "compare_acf"))
     return (np.median((acfs - true)**2))**.5
 
-def pgram_plot(truths, DIR):
+def pgram_plot(truths_e, N, true, pgram, DIR):
     """
     Plot the pgram results.
     """
-
-    truths_e = make_new_df(truths, DIR)
-    m = (truths_e.DELTA_OMEGA.values == 0) \
-            * (truths_e.acf_period.values > 0)
-
-    N = truths_e.N.values[m]
[…]

### 2016-11-01 gprotation/GProtation.py
bug fix

@@ -82,7 +82,7 @@ def neglnlike(theta, x, y, yerr):
     return -gp.lnlikelihood(y, quiet=True)
 
 # make various plots
-def make_plot(sampler, x, y, yerr, ID, RESULTS_DIR, truths, traces=False,
+def make_plot(sampler, x, y, yerr, ID, RESULTS_DIR, trths, traces=False,
               tri=False, prediction=True):
 
     nwalkers, nsteps, ndims = np.shape(sampler)
@@ -131,7 +131,7 @@ def make_plot(sampler, x, y, yerr, ID, RESULTS_DIR, truths, traces=False,
         print("Making triangle plot")
         fig = corner.corner(flat[:, :-1], labels=fig_labels,
                             quantiles=[.16, .5, .84], show_titles=True,
-                            truths=truths,)
+                            truths=trths,)
         fig.savefig(os.path.join(RESULTS_DIR, "{0}_triangle".format(ID)))
         print(os.path.join("{0}_triangle.png".format(ID)))
 

### 2016-11-01 gprotation/recover_gp.py

@@ -28,8 +28,8 @@ def recover(i):
     truths = pd.read_csv("gp_truths.csv")
 
     # select either periodic or non-periodic
-    RESULTS_DIR = "results_periodic_gp"
-#     RESULTS_DIR = "results_aperiodic_gp"
+#     RESULTS_DIR = "results_periodic_gp"
+    RESULTS_DIR = "results_aperiodic_gp"
 
     DATA_DIR = "periodic_gp_simulations"
     trths = [truths.lnA.values[i], truths.lnl_p.values[i],
@@ -41,7 +41,7 @@ def recover(i):
                    truths.lnsigma.values[i], None]
 
     sid = str(int(i)).zfill(4)
-    print(sid, "of", 160)
+    print(sid, "of", 333)
     x, y = np.genfromtxt(os.path.join(DATA_DIR, "{0}.txt".format(sid))).T
     yerr = np.ones_like(y) * 1e-5
     x_n, y_n = np.genfromtxt(os.path.join(DATA_DIR,
@@ -75,9 +75,9 @@ def recover(i):
     print("p_init = ", p_init, "p_init_n = ", p_init_n)
 
     # fast settings:
-#     burnin, nwalkers, nruns, full_run = 2, 12, 2, 50
-#     x, y, yerr = x[::100], y[::100], yerr[::100]
-#     x_n, y_n, yerr_n = x_n[::100], y_n[::100], yerr_n[::100]
+    burnin, nwalkers, nruns, full_run = 2, 12, 2, 50
+    x, y, yerr = x[::100], y[::100], yerr[::100]
+    x_n, y_n, yerr_n = x_n[::100], y_n[::100], yerr_n[::100]
 
     # set prior bounds
     plims = np.log([.5*p_init, 1.5*p_init])
@@ -86,14 +86,14 @@ def recover(i):
     # run on noisy and noise free
     mcmc_fit(x, y, yerr, p_init, plims, sid, RESULTS_DIR, trths,
 	     burnin=burnin, nwalkers=nwalkers, nruns=nruns, full_run=full_run)
-    mcmc_fit(x_n, y_n, yerr_n, p_init_n, plims_n, "{0}_n".format(sid), trths,
-             RESULTS_DIR, burnin=burnin, nwalkers=nwalkers, nruns=nruns,
+    mcmc_fit(x_n, y_n, yerr_n, p_init_n, plims_n, "{0}_n".format(sid),
+             RESULTS_DIR, trths, burnin=burnin, nwalkers=nwalkers, nruns=nruns,
[…]

### 2016-11-01 documents/ms.tex

@@
 \usepackage{cite, natbib}
-% \documentclass[onecolumn, 12pt, usenatbib, preprint]{emulateapj}
 \usepackage{float}
@@
 
-\newcommand{\article}{article}
-\newcommand{\oxford}{1}
-\newcommand{\uw}{2}
-\newcommand{\sagan}{3}
-\newcommand{\ktwo}{{\it K2}}
-\newcommand{\eg}{{\it e.g.}}
-\newcommand{\ie}{{\it i.e.}}
-\newcommand{\nlightcurves}{300}
+\newcommand{\columbia}{1}
+\newcommand{\oxford}{2}
+\newcommand{\uw}{3}
+\newcommand{\nlightcurves}{333}
 \newcommand{\Kepler}{{\it Kepler}}
 \newcommand{\kepler}{\Kepler}
+\newcommand{\corot}{{\it CoRoT}}
+\newcommand{\Ktwo}{{\it K2}}
+\newcommand{\ktwo}{\Ktwo}
+\newcommand{\TESS}{{\it TESS}}
+\newcommand{\LSST}{{\it LSST}}
+\newcommand{\SDSS}{{\it SDSS}}
+\newcommand{\PLATO}{{\it PLATO}}
+\newcommand{\Teff}{$T_{\mathrm{eff}}$}
+\newcommand{\teff}{$T_{\mathrm{eff}}$}
+\newcommand{\FeH}{[Fe/H]}
+\newcommand{\feh}{[Fe/H]}
+\newcommand{\ie}{{\it i.e.}}
+\newcommand{\eg}{{\it e.g.}}
+\newcommand{\logg}{log \emph{g}}
+\newcommand{\dnu}{$\Delta \nu$}
+\newcommand{\numax}{$\nu_{\mathrm{max}}$}
 
@@
 
-\title{Stellar rotation period inference with Gaussian processes}
+\title{Inferring stellar rotation periods with Gaussian processes}
 
 \author{%
-   Ruth Angus\altaffilmark{\oxford}
+   Ruth Angus\altaffilmark{\columbia}
    Suzanne Aigrain\altaffilmark{\oxford}
-   \& Daniel Foreman-Mackey\altaffilmark{\uw}\altaffilmark{\sagan}
+   \& Daniel Foreman-Mackey\altaffilmark{\uw}
 }
 
+\altaffiltext{\columbia}{Simons Fellow, Columbia University, NY, NY,
+ruthangus@gmail.com}
 \altaffiltext{\oxford}{Subdepartment of Astrophysics, University of Oxford,
@@
 \altaffiltext{\uw}{Department of Astronomy, University of Washington, Seattle,
-WA, USA}
-\altaffiltext{\sagan}{Sagan fellow}
+WA}
 
@@
 estimate of the uncertainty on the rotation period.
-
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/Kepler452b.pdf}
-% \caption[A light curve with a GP model.]
-% {Light curve of Kepler-452 b, a habitable-zone, Earth-sized planet
-% hosting G star \citep{Jenkins2015}. The orange line shows a fit to the data using
-% a Gaussian process model with a QP covariance kernel function.}
-% \label{fig:GP_example}
-% \end{center}
-% \end{figure}
 
@@
 
-The data presented in this paper were obtained from the Mikulski Archive for
-Space Telescopes (MAST).
-STScI is operated by the Association of Universities for Research in Astronomy,
-Inc., under NASA contract NAS5-26555.
-Support for MAST for non-HST data is provided by the NASA Office of Space
-Science via grant NNX09AF08G and by other grants and contracts.
-This paper includes data collected by the Kepler mission.
-Funding for the Kepler mission is provided by the NASA Science Mission
-directorate.
+% acknowledgements
 
@@
 \bibliography{GProtation}
-
 \end{document}

### 2016-11-11 code/GPr.py
adding generate_GP_lc.py

@@ -1,196 +1,0 @@
-from __future__ import print_function
-import numpy as np
-import george
-from george.kernels import ExpSine2Kernel, ExpSquaredKernel
-import emcee
-import triangle
-import h5py
-from Kepler_ACF import corr_run
-from tools import bin_data
-
-class GProtation(object):
-
-    def __init__(self, x, y, yerr):
-        self.x = x
-        self.y = y
-        self.yerr = yerr
-
-    def easy_fit(self):
-        """
-        This function spits out a posterior for Kepler data
-        """
-        kepler_cadence = .02043365
-        p_init, acf, lags = simple_acf(x, y, kepler_cadence)
-
-    def MCMC(self, theta_init, plims, burnin, run, nwalkers):
-        """
-        This function runs emcee in order to sample the posterior of your
-        hyperparameters or find the maxima, etc.
-        theta_init (array): logarithmic initial parameters
-        [A, gamma, l, sigma, period]
-        A = amplitude.
-        gamma = scaling factor in the exponential-sine kernel.
-        l = length scale of the squared-exponential kernel
-        sigma = white noise multiplier
-        period = period (obv). In the same units as your 'xs'.
-        plims (tuple): the upper and lower bounds for a flat (in log-space)
-        prior.
-        burnin (integer) is the number of steps during burn-in.
-        run (integer) is the number of steps for the full run.
-        nwalkers is the number of walkers.
-        This function returns the emcee sampler object.
-        See dfm.io/emcee for emcee documentation.
-        """
-
-        # run MCMC
-        p_init = theta_init[-1]
-        ndim, nwalkers = len(theta_init), 32
-        p0 = [theta_init+1e-4*np.random.rand(ndim) for i in range(nwalkers)]
-        args = [plims]
-        sampler = emcee.EnsembleSampler(nwalkers, ndim, self.lnprob, args=args)
-
-        print("burning in...")
-        p0, lp, state = sampler.run_mcmc(p0, burnin)
-        sampler.reset()
-        print("production run...")
-        p0, lp, state = sampler.run_mcmc(p0, run)
-        return sampler
-
-    def lnprior(self, theta, plims):
-        """
-        Here is a prior function if you want to use it!
-        theta is the array of parameters.
-        plims is a tuple containing the (log) lower and upper limits for
-        the rotation period. These are logarithmic!
[…]

### 2016-11-11 code/GProtation.py

@@ -1,177 +1,0 @@
-from __future__ import print_function
-import numpy as np
-import matplotlib.pyplot as plt
-import george
-from george.kernels import ExpSine2Kernel, ExpSquaredKernel, WhiteKernel
-import glob
-import emcee
-try:
-    import corner
-except:
-    import triangle
-import h5py
-import subprocess
-from plotstuff import params, colours
-reb = params()
-cols = colours()
-import scipy.optimize as spo
-import time
-
-
-def lnprior(theta, plims):
-    """
-    plims is a tuple, list or array containing the lower and upper limits for
-    the rotation period. These are logarithmic!
-    theta = A, l, Gamma, s, P
-    """
-    if -20 < theta[0] < 20 and theta[4] < theta[1] and -20 < theta[2] < 20 \
-    and -20 < theta[3] < 20 and plims[0] < theta[4] < plims[1]:
-        return 0.
-    return -np.inf
-
-
-def Glnprior(theta, plims):
-    """
-    plims is a tuple, list or array containing the lower and upper limits for
-    the rotation period. These are not logarithmic!
-    """
-    if -20 < theta[0] < 16 and -10 < theta[1] < 1 and -20 < theta[2] < 20 \
-    and -20 < theta[3] < 20 and np.log(plims[0]) < theta[4] < np.log(plims[1]):
-        return -.5 * ((theta[4] - np.log(plims[2]))/(.5*np.log(plims[2])))**2 \
-                - .5 * ((theta[1] - .1*np.log(plims[2]))/(.1*plims[2]))**2
-    return -np.inf
-
-# lnprob
-def lnprob(theta, x, y, yerr, plims):
-    return lnlike(theta, x, y, yerr) + lnprior(theta, plims)
-
-# lnlike
-def lnlike(theta, x, y, yerr):
-    theta = np.exp(theta)
-    k = theta[0] * ExpSquaredKernel(theta[1]) \
-            * ExpSine2Kernel(theta[2], theta[4])
-    gp = george.GP(k, solver=george.HODLRSolver)
-    try:
-        gp.compute(x, np.sqrt(theta[3]+yerr**2))
-    except (ValueError, np.linalg.LinAlgError):
-        return 10e25
-    return gp.lnlikelihood(y, quiet=True)
-
-# lnlike
-def neglnlike(theta, x, y, yerr):
[…]

### 2016-11-11 code/find_amplitude_distribution.py

@@ -1,24 +1,0 @@
-from __future__ import print_function
-import numpy as np
-import matplotlib.pyplot as plt
-from kepler_data import load_kepler_data
-import glob
-
-# ids = np.genfromtxt("data/NGC6819_members.txt", dtype=int).T
-ids = np.genfromtxt("data/astero_ids.txt", dtype=int).T
-var = np.zeros(len(ids))
-for j, i in enumerate(ids):
-    id = str(i)
-    print(id, j, "of", len(ids))
-    d = "/home/angusr/.kplr/data/lightcurves"
-    fnames = np.sort(glob.glob("{0}/{1}/*llc.fits".format(d, id.zfill(9))))
-    x, y, yerr = load_kepler_data(fnames)
-    var[j] = np.var(y)
-
-print(var)
-print(min(var), max(var))
-
-plt.clf()
-plt.hist(np.log10(var))
-# plt.savefig("var_NGC6819")
-plt.savefig("var_astero")

### 2016-11-11 code/generate_GP_lc.py

@@ -1,0 +1,39 @@
+# Code for generating periodic and non-periodic GP light curves.
+# Find the appropriate parameters from the real light curves and THEN do this
+# step.
+
+# coding: utf-8
+import numpy as np
+import matplotlib.pyplot as plt
+import george
+from george.kernels import ExpSine2Kernel, ExpSquaredKernel
+
+SIM_DIR = "../simulations"
+
+theta_init = np.log([np.exp(-5), np.exp(7), np.exp(.6), np.exp(-16),
+                        p_init])
+
+# Periodic GPs
+# Periods log uniform from .5 to 100 days
+periods = np.random.uniform(np.log(.5), np.log(100))
+
+# Amplitudes log uniform from -6 to -3
+
+
+# Gammas
+
+# ls
+
+# save truth file
+
+# generate and save GPs.
+
+# Non-Periodic GPs
+
+# Amplitudes
+
+# ls
+
+# generate and save GPs.
+
+# save truth file.

### 2016-11-11 code/plot_lcs.py

@@ -1,66 +1,0 @@
-# Script for generating thesis figures.
-import numpy as np
-import matplotlib.pyplot as plt
-from plotstuff import params, colours
-cols = colours()
-reb = params()
-from simple_acf import dan_acf, simple_acf
-
-plotpar = {'axes.labelsize': 20,
-           'text.fontsize': 20,
-           'legend.fontsize': 20,
-           'xtick.labelsize': 20,
-           'ytick.labelsize': 20,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-orange = '#FF9933'
-lightblue = '#66CCCC'
-blue = '#0066CC'
-pink = '#FF33CC'
-turquoise = '#3399FF'
-lightgreen = '#99CC99'
-green = '#009933'
-maroon = '#CC0066'
-purple = '#9933FF'
-red = '#CC0000'
-lilac = '#CC99FF'
-
-ids, ps, amps = np.genfromtxt("simulations/noise-free/true_periods_amps.txt").T
-id = "0006"
-x, y = np.genfromtxt("simulations/noise-free/{}.txt".format(id)).T
-period, _, _ = simple_acf(x, y)
-plt.clf()
-plt.plot(x, y, "k-", label = "$\mathrm{Period} = %.1f~\mathrm{days}$" % period)
-plt.xlabel("$\mathrm{Time~(days)}$")
-plt.ylabel("$\mathrm{Normalised~flux}$")
-plt.xlim(0, 200)
-plt.legend()
-plt.subplots_adjust(bottom=.12)
-plt.savefig("../figures/noise-free_lc.pdf")
-
-# ACF figure
-gap_days = .02043365
-acf = dan_acf(y)
-lags = np.arange(len(acf)) * gap_days
-plt.clf()
-plt.plot(lags, acf, "k")
-plt.xlim(0, 100)
-plt.xlabel("$\mathrm{Time~(days)}$")
-plt.ylabel("$\mathrm{Correlation}$")
-_, true_periods, _ = np.genfromtxt("simulations/true_periods.txt").T
-# plt.title("$\mathrm{P}_{\mathrm{rot}} = {0}\mathrm{days}$".format(true_periods[0]))
-plt.savefig("../figures/noise-free_acf.pdf")
-
-x2, y2, yerr = np.genfromtxt("simulations/kepler_injections/{}.txt".format(id)).T
-plt.clf()
[…]

### 2016-11-11 code/plotsims.py

@@ -1,9 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import sys
-id = sys.argv[1]
-
-x, y = np.genfromtxt("simulations/%s.txt" % id).T
-plt.clf()
-plt.plot(x[:300], y[:300], "k.")
-plt.savefig("test")

### 2016-11-11 code/test_on_GP_simulations.py

@@ -1,57 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import george
-from george.kernels import ExpSine2Kernel, ExpSquaredKernel, CosineKernel
-
-def generate_GP_sim(pars, sine_kernel=True):
-    '''
-    A function to generate a set of test data.
-    '''
-    print "pars = ", pars
-
-    x = np.arange(0, 10, .1)
-    yerr = np.ones_like(x) * .001
-
-    if sine_kernel:
-        k = pars[0] * ExpSquaredKernel(pars[1]) \
-                * ExpSine2Kernel(pars[2], pars[4])
-    else:
-        k = pars[0] * ExpSquaredKernel(pars[1]) \
-                * CosineKernel(pars[3])
-    gp = george.GP(k)
-    gp.compute(x, yerr)  # no added noise
-    ys = gp.sample(x, 1) + np.random.randn(len(x))*yerr
-
-    plt.clf()
-    plt.plot(x, ys, "k.")
-    mu, cov = gp.predict(ys, x)
-    plt.plot(x, mu)
-    plt.savefig("test_data")
-    return x, ys, yerr
-
-def fit(theta_init, x, ys, yerr, plims):
-    """
-    fitting the GP. theta_init should be in linear space, not log.
-    as should plims
-    """
-    theta_init = np.log(theta_init)
-    DIR = "../figs"
-    sampler = MCMC(theta_init, x, ys, yerr, plims, 200, 500, "test", DIR)
-    make_plot(sampler, x, ys, yerr, 1, DIR, traces=True)
-
-if __name__ == "__main__":
-    # generate a fake lc and see if you can recover it
-    plims = [2., 30.]
-    sine_kernel = False
-
-    if sine_kernel:
-        pars = [1., 1., 1., 1., 7.]
-        theta_init = [1., 1., 1., 1., 7.]
-        from GProtation import MCMC, make_plot
-    else:
-        pars = [1., 1., 1., 7.]
-        theta_init = [1., 1., 1., 7.]
-        from GProtation_cosine import MCMC, make_plot
-
-    x, y, yerr = generate(pars, sine_kernel=False)
-    fit(theta_init, x, y, yerr, plims)

### 2016-11-11 code/generate_GP_lc.py
adding all code from /GProtation/code/tests

@@ -1,39 +1,0 @@
-# Code for generating periodic and non-periodic GP light curves.
-# Find the appropriate parameters from the real light curves and THEN do this
-# step.
-
-# coding: utf-8
-import numpy as np
-import matplotlib.pyplot as plt
-import george
-from george.kernels import ExpSine2Kernel, ExpSquaredKernel
-
-SIM_DIR = "../simulations"
-
-theta_init = np.log([np.exp(-5), np.exp(7), np.exp(.6), np.exp(-16),
-                        p_init])
-
-# Periodic GPs
-# Periods log uniform from .5 to 100 days
-periods = np.random.uniform(np.log(.5), np.log(100))
-
-# Amplitudes log uniform from -6 to -3
-
-
-# Gammas
-
-# ls
-
-# save truth file
-
-# generate and save GPs.
-
-# Non-Periodic GPs
-
-# Amplitudes
-
-# ls
-
-# generate and save GPs.
-
-# save truth file.

### 2016-11-11 code/tests/GProtation.py

@@ -1,0 +1,175 @@
+from __future__ import print_function
+import numpy as np
+import matplotlib.pyplot as plt
+import george
+from george.kernels import ExpSine2Kernel, ExpSquaredKernel, WhiteKernel
+import glob
+import emcee
+# import triangle
+import corner
+import h5py
+import subprocess
+from plotstuff import params, colours
+reb = params()
+cols = colours()
+import scipy.optimize as spo
+import time
+
+
+def lnprior(theta, plims):
+    """
+    plims is a tuple, list or array containing the lower and upper limits for
+    the rotation period. These are logarithmic!
+    theta = A, l, Gamma, s, P
+    """
+    if -20 < theta[0] < 20 and theta[4] < theta[1] and -20 < theta[2] < 20 \
+    and -20 < theta[3] < 20 and plims[0] < theta[4] < plims[1]:
+        return 0.
+    return -np.inf
+
+
+def Glnprior(theta, plims):
+    """
+    plims is a tuple, list or array containing the lower and upper limits for
+    the rotation period. These are not logarithmic!
+    """
+    if -20 < theta[0] < 16 and -10 < theta[1] < 1 and -20 < theta[2] < 20 \
+    and -20 < theta[3] < 20 and np.log(plims[0]) < theta[4] < np.log(plims[1]):
+        return -.5 * ((theta[4] - np.log(plims[2]))/(.5*np.log(plims[2])))**2 \
+                - .5 * ((theta[1] - .1*np.log(plims[2]))/(.1*plims[2]))**2
+    return -np.inf
+
+# lnprob
+def lnprob(theta, x, y, yerr, plims):
+    return lnlike(theta, x, y, yerr) + lnprior(theta, plims)
+
+# lnlike
+def lnlike(theta, x, y, yerr):
+    theta = np.exp(theta)
+    k = theta[0] * ExpSquaredKernel(theta[1]) \
+            * ExpSine2Kernel(theta[2], theta[4])
+    gp = george.GP(k, solver=george.HODLRSolver)
+    try:
+        gp.compute(x, np.sqrt(theta[3]+yerr**2))
+    except (ValueError, np.linalg.LinAlgError):
+        return 10e25
+    return gp.lnlikelihood(y, quiet=True)
+
+# lnlike
+def neglnlike(theta, x, y, yerr):
+    theta = np.exp(theta)
[…]

### 2016-11-11 code/tests/plotstuff.py

@@ -1,0 +1,28 @@
+import matplotlib.pyplot as plt
+from colours import plot_colours
+ocols = plot_colours()
+
+def params():
+    plotpar = {'axes.labelsize': 12,
+               'text.fontsize': 10,
+               'legend.fontsize': 15,
+               'xtick.labelsize': 12,
+               'ytick.labelsize': 12,
+               'text.usetex': True}
+    plt.rcParams.update(plotpar)
+    return {'capsize':0, 'fmt':'k.', 'ecolor':'.8'}
+
+class colours(object):
+    def __init__(self):
+
+        self.orange = '#FF9933'
+        self.lightblue = '#66CCCC'
+        self.blue = '#0066CC'
+        self.pink = '#FF33CC'
+        self.turquoise = '#3399FF'
+        self.lightgreen = '#99CC99'
+        self.green = '#009933'
+        self.maroon = '#CC0066'
+        self.purple = '#9933FF'
+        self.red = '#CC0000'
+        self.lilac = '#CC99FF'

### 2016-11-11 code/tests/recover_kepler_injections.py

@@ -1,0 +1,172 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from measure_GP_rotation import fit
+from Kepler_ACF import corr_run
+import h5py
+from plotstuff import params, colours
+reb = params()
+cols = colours()
+from gatspy.periodic import LombScargle
+import sys
+from simple_acf import simple_acf, make_plot
+from Kepler_ACF import corr_run
+from multiprocessing import Pool
+import glob
+from kepler_data import load_kepler_data
+
+plotpar = {'axes.labelsize': 22,
+           'font.size': 22,
+           'legend.fontsize': 22,
+           'xtick.labelsize': 22,
+           'ytick.labelsize': 22,
+           'text.usetex': True}
+plt.rcParams.update(plotpar)
+
+def my_acf(id, x, y, yerr, fn, plot=False, amy=False):
+    """
+    takes id of the star, returns an array of period measurements and saves the
+    results.
+    (the files are saved in a directory that is a global variable).
+    """
+    if amy:
+        period, period_err = corr_run(x, y, yerr, id, fn, saveplot=plot)
+    else:
+        period, acf, lags = simple_acf(x, y)
+        np.savetxt("{0}/{1}_result.txt".format(fn, id, period))
+        if plot:
+            make_plot(acf, lags, id, fn)
+    return period
+
+def periodograms(id, x, y, yerr, fn, plot=False, savepgram=True):
+    """
+    takes id of the star, returns an array of period measurements and saves the
+    results.
+    (the files are saved in a directory that is a global variable).
+    """
+    # initialise with acf
+    try:
+        p_init, _ = np.genfromtxt("{0}/{1}_result.txt".format(fn, id))
+    except:
+        corr_run(x, y, yerr, id, fn, saveplot=False)
+        p_init, _ = np.genfromtxt("{0}/{1}_result.txt".format(fn, id))
+    print("acf period, err = ", p_init)
+
+    ps = np.linspace(p_init*.1, p_init*4, 1000)
+    model = LombScargle().fit(x, y, yerr)
+    pgram = model.periodogram(ps)
+
+    # find peaks
+    peaks = np.array([i for i in range(1, len(ps)-1) if pgram[i-1] <
[…]

### 2016-11-11 code/tests/simple_acf.py

@@ -1,0 +1,114 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import emcee
+import glob
+
+# dan's acf function
+def dan_acf(x, axis=0, fast=False):
+    """
+    Estimate the autocorrelation function of a time series using the FFT.
+    :param x:
+        The time series. If multidimensional, set the time axis using the
+        ``axis`` keyword argument and the function will be computed for every
+        other axis.
+    :param axis: (optional)
+        The time axis of ``x``. Assumed to be the first axis if not specified.
+    :param fast: (optional)
+        If ``True``, only use the largest ``2^n`` entries for efficiency.
+        (default: False)
+    """
+    x = np.atleast_1d(x)
+    m = [slice(None), ] * len(x.shape)
+
+    # For computational efficiency, crop the chain to the largest power of
+    # two if requested.
+    if fast:
+        n = int(2**np.floor(np.log2(x.shape[axis])))
+        m[axis] = slice(0, n)
+        x = x
+    else:
+        n = x.shape[axis]
+
+    # Compute the FFT and then (from that) the auto-correlation function.
+    f = np.fft.fft(x-np.mean(x, axis=axis), n=2*n, axis=axis)
+    m[axis] = slice(0, n)
+    acf = np.fft.ifft(f * np.conjugate(f), axis=axis)[m].real
+    m[axis] = 0
+    return acf / acf[m]
+
+def simple_acf(x, y):
+
+    # fit and subtract straight line
+    AT = np.vstack((x, np.ones_like(x)))
+    ATA = np.dot(AT, AT.T)
+    m, b = np.linalg.solve(ATA, np.dot(AT, y))
+    y -= m*x + b
+
+    # perform acf
+    acf = dan_acf(y)
+
+    # smooth with Gaussian kernel convolution
+    Gaussian = lambda x, sig: 1./(2*np.pi*sig**.5) * \
+                 np.exp(-0.5*(x**2)/(sig**2))
+    conv_func = Gaussian(np.arange(-28,28,1.), 9.)
+    acf_smooth = np.convolve(acf, conv_func, mode='same')
+
+    # create 'lags' array
+    gap_days = 0.02043365
+    lags = np.arange(len(acf))*gap_days
+
+    # find all the peaks
+    peaks = np.array([i for i in range(1, len(lags)-1)
+                     if acf_smooth[i-1] < acf_smooth[i] and
[…]

### 2016-11-11 code/tests/GProtation.py
adding all notebooks in GProtation/code/tests

@@ -1,175 +1,0 @@
-from __future__ import print_function
-import numpy as np
-import matplotlib.pyplot as plt
-import george
-from george.kernels import ExpSine2Kernel, ExpSquaredKernel, WhiteKernel
-import glob
-import emcee
-# import triangle
-import corner
-import h5py
-import subprocess
-from plotstuff import params, colours
-reb = params()
-cols = colours()
-import scipy.optimize as spo
-import time
-
-
-def lnprior(theta, plims):
-    """
-    plims is a tuple, list or array containing the lower and upper limits for
-    the rotation period. These are logarithmic!
-    theta = A, l, Gamma, s, P
-    """
-    if -20 < theta[0] < 20 and theta[4] < theta[1] and -20 < theta[2] < 20 \
-    and -20 < theta[3] < 20 and plims[0] < theta[4] < plims[1]:
-        return 0.
-    return -np.inf
-
-
-def Glnprior(theta, plims):
-    """
-    plims is a tuple, list or array containing the lower and upper limits for
-    the rotation period. These are not logarithmic!
-    """
-    if -20 < theta[0] < 16 and -10 < theta[1] < 1 and -20 < theta[2] < 20 \
-    and -20 < theta[3] < 20 and np.log(plims[0]) < theta[4] < np.log(plims[1]):
-        return -.5 * ((theta[4] - np.log(plims[2]))/(.5*np.log(plims[2])))**2 \
-                - .5 * ((theta[1] - .1*np.log(plims[2]))/(.1*plims[2]))**2
-    return -np.inf
-
-# lnprob
-def lnprob(theta, x, y, yerr, plims):
-    return lnlike(theta, x, y, yerr) + lnprior(theta, plims)
-
-# lnlike
-def lnlike(theta, x, y, yerr):
-    theta = np.exp(theta)
-    k = theta[0] * ExpSquaredKernel(theta[1]) \
-            * ExpSine2Kernel(theta[2], theta[4])
-    gp = george.GP(k, solver=george.HODLRSolver)
-    try:
-        gp.compute(x, np.sqrt(theta[3]+yerr**2))
-    except (ValueError, np.linalg.LinAlgError):
-        return 10e25
-    return gp.lnlikelihood(y, quiet=True)
-
-# lnlike
-def neglnlike(theta, x, y, yerr):
-    theta = np.exp(theta)
[…]

### 2016-11-11 code/tests/measure_GP_rotation.py

@@ -1,93 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-# import kplr
-# client = kplr.API()
-import pyfits
-import glob
-from Kepler_ACF import corr_run
-import h5py
-
-def get_data(id):
-    fnames = glob.glob("/Users/angusr/.kplr/data/lightcurves/%s/*" %
-                       str(int(id)).zfill(9))
-    hdulist = pyfits.open(fnames[0])
-    t = hdulist[1].data
-    time = t["TIME"]
-    flux = t["PDCSAP_FLUX"]
-    flux_err = t["PDCSAP_FLUX_ERR"]
-    q = t["SAP_QUALITY"]
-    m = np.isfinite(time) * np.isfinite(flux) * np.isfinite(flux_err) * (q==0)
-    x = time[m]
-    med = np.median(flux[m])
-    y = flux[m]/med - 1
-    yerr = flux_err[m]/med
-    for fname in fnames[1:]:
-        hdulist = pyfits.open(fname)
-        t = hdulist[1].data
-        time = t["TIME"]
-        flux = t["PDCSAP_FLUX"]
-        flux_err = t["PDCSAP_FLUX_ERR"]
-        q = t["SAP_QUALITY"]
-        m = np.isfinite(time) * np.isfinite(flux) * np.isfinite(flux_err) * \
-                (q==0)
-        x = np.concatenate((x, time[m]))
-        med = np.median(flux[m])
-        y = np.concatenate((y, flux[m]/med - 1))
-        yerr = np.concatenate((yerr, flux_err[m]/med))
-    return x, y, yerr
-
-def bin_data(x, y, yerr, npts):
-    mod, nbins = len(x) % npts, len(x) / npts
-    if mod != 0:
-        x, y, yerr = x[:-mod], y[:-mod], yerr[:-mod]
-    xb, yb, yerrb = [np.zeros(nbins) for i in range(3)]
-    for i in range(npts):
-        xb += x[::npts]
-        yb += y[::npts]
-        yerrb += yerr[::npts]**2
-        x, y, yerr = x[1:], y[1:], yerr[1:]
-    return xb/npts, yb/npts, yerrb**.5/npts
-
-def fit(x, y, yerr, id, p_init, plims, DIR, burnin=500, run=1500, npts=48,
-        cutoff=100, sine_kernel=False, acf=False, runMCMC=True, plot=False):
-    """
-    takes x, y, yerr and initial guesses and priors for period and does
-    the full GP MCMC.
[…]

### 2016-11-11 code/tests/mpfit.py

@@ -1,1298 +1,0 @@
-import numpy
-
-"""mpfit: Non-linear least squares minimisation using the
-   Levenberg-Marquardt technique with optional fixed, limited or tied
-   parameters.
-
-   This module is intended primarily for fitting non-linear models to
-   data.  The user should supply a Python function which, given the
-   data and parameters, computes an array of weighted deviations
-   between model and data. The user can then use the function
-   mpfit.mpfit to minimise the sum of the squares of the deviates,
-   i.e. the chi-squared.
-
-   The function mpfit.mpfit takes as arguments the user defined
-   function and a numpy array containing inital guesses for the
-   parameters. The data, and any measurement uncertainties, are passed
-   to the user-defined function using the functkw keyword of
-   mpfit.mpfit. It returns an instance of the mpfit.results class,
-   which contains the best-fit values of the parameters as well as the
-   covariance matrix and formal parameter uncertainties.
-
-   Here is a quick example using a user function defined in this module:
-      import mpfit
-      import numpy
-      import pylab
-      N = 100
-      x = numpy.r_[0:10:N*1j]
-      p0 = numpy.array([5.7, 0.22])
-      noise = mpfit.F(x, p0) + numpy.random.randn(N) * 10
-      err = numpy.ones(N) * 10
-      pylab.errorbar(x, y, yerr = err)
-      p1 = numpy.array([6,0.2])
-      fa = {'x':x, 'y':y + noise, 'err':err}
-      m = mpfit(mpfit.example_func, p1, functkw = fa)
-      print('parameters = ', m.params)
-      print 'parameters = ', m.pcerror
-# Note how the errors are underestimated because the two parameters
-# are strongly correlated!
-      pylab.plot(x, mpfit.F(x, m.params))
-
-   The user function should be built according to a specific set of
-   rules, which are described below and illustrated in the example_func
-   function.
-
-   Simple constraints can be placed on parameter values by using the
-   fixed, limited & limits, and/or tied keywords of mpfit.mpfit, as
-   described in the documentation of function mpfit itself.
-   *** NB: in the present version, only the fixed keyword is
-       functional and tested. The limited / limits keywords don't work
-       properly and the tied keyword is untested. ***
-
-   See the documentation for the individual functions and classes for
-   more details, including some keywords not described here.
-
-   USER_DEFINED FUNCTION:
-
[…]

### 2016-11-11 code/tests/plotstuff.py

@@ -1,28 +1,0 @@
-import matplotlib.pyplot as plt
-from colours import plot_colours
-ocols = plot_colours()
-
-def params():
-    plotpar = {'axes.labelsize': 12,
-               'text.fontsize': 10,
-               'legend.fontsize': 15,
-               'xtick.labelsize': 12,
-               'ytick.labelsize': 12,
-               'text.usetex': True}
-    plt.rcParams.update(plotpar)
-    return {'capsize':0, 'fmt':'k.', 'ecolor':'.8'}
-
-class colours(object):
-    def __init__(self):
-
-        self.orange = '#FF9933'
-        self.lightblue = '#66CCCC'
-        self.blue = '#0066CC'
-        self.pink = '#FF33CC'
-        self.turquoise = '#3399FF'
-        self.lightgreen = '#99CC99'
-        self.green = '#009933'
-        self.maroon = '#CC0066'
-        self.purple = '#9933FF'
-        self.red = '#CC0000'
-        self.lilac = '#CC99FF'

### 2016-11-11 code/tests/recover_kepler_injections.py

@@ -1,172 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from measure_GP_rotation import fit
-from Kepler_ACF import corr_run
-import h5py
-from plotstuff import params, colours
-reb = params()
-cols = colours()
-from gatspy.periodic import LombScargle
-import sys
-from simple_acf import simple_acf, make_plot
-from Kepler_ACF import corr_run
-from multiprocessing import Pool
-import glob
-from kepler_data import load_kepler_data
-
-plotpar = {'axes.labelsize': 22,
-           'font.size': 22,
-           'legend.fontsize': 22,
-           'xtick.labelsize': 22,
-           'ytick.labelsize': 22,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-def my_acf(id, x, y, yerr, fn, plot=False, amy=False):
-    """
-    takes id of the star, returns an array of period measurements and saves the
-    results.
-    (the files are saved in a directory that is a global variable).
-    """
-    if amy:
-        period, period_err = corr_run(x, y, yerr, id, fn, saveplot=plot)
-    else:
-        period, acf, lags = simple_acf(x, y)
-        np.savetxt("{0}/{1}_result.txt".format(fn, id, period))
-        if plot:
-            make_plot(acf, lags, id, fn)
-    return period
-
-def periodograms(id, x, y, yerr, fn, plot=False, savepgram=True):
-    """
-    takes id of the star, returns an array of period measurements and saves the
-    results.
-    (the files are saved in a directory that is a global variable).
-    """
-    # initialise with acf
-    try:
-        p_init, _ = np.genfromtxt("{0}/{1}_result.txt".format(fn, id))
-    except:
-        corr_run(x, y, yerr, id, fn, saveplot=False)
-        p_init, _ = np.genfromtxt("{0}/{1}_result.txt".format(fn, id))
-    print("acf period, err = ", p_init)
-
-    ps = np.linspace(p_init*.1, p_init*4, 1000)
-    model = LombScargle().fit(x, y, yerr)
-    pgram = model.periodogram(ps)
-
-    # find peaks
-    peaks = np.array([i for i in range(1, len(ps)-1) if pgram[i-1] <
[…]

### 2016-11-17 documents/ms.tex
adding all code from /GProtation/code/tests

@@
 lifetimes causing stellar flux variations to slowly shift in phase.
-A strictly periodic sinusoid is therefore not a representative generative
-model for the light curve of star with a rotational signal.
+A strictly periodic sinusoid is therefore not a representative model for the
+light curve of star with a rotational signal.
 Ideally, a physical model of the stellar surface would be conditioned on the
@@
 with a quasi-periodic covariance kernel function.
-By modelling the covariance matrix of the light curve with a GP, a highly
-flexible semi-parametric function, we avoid the necessity to choose a `best
-fitting' functional form, whilst sampling directly from the posterior
-probability distribution function of the periodic parameter and marginalising
-over the other kernel hyperparameters.
+By modelling the light curve with a GP, a highly flexible semi-parametric
+function, we avoid the necessity of choosing a `best fitting' functional form,
+whilst sampling directly from the posterior probability distribution function
+of the periodic parameter and marginalising over the other kernel
+hyperparameters.
 We used \nlightcurves\ simulated light curves with a range of rotation periods
@@
 The posterior probability distribution of the rotation period parameter was
-sampled using the affine invariant ensemble MCMC sampler {\tt emcee}
-\citep{Foreman-Mackey2013}, and the GP operations were performed using the
-{\tt george} python package \citep{George}.
+sampled using the affine invariant ensemble Markov Chain Monte Carlo (MCMC)
+sampler {\tt emcee} \citep{Foreman-Mackey2013}, and the GP operations were
+performed using the {\tt george} python package \citep{George}.
 Rotation periods inferred via this method are more precise and accurate than
@@
 conditioned on the data.
-A physically realistic, generative model would perfectly capture the complexity
-of shapes within stellar light curves as well as the quasi-periodic nature,
-allowing for extremely precise probabilistic period recovery.
+A physically realistic model would perfectly capture the complexity of shapes
+within stellar light curves as well as the quasi-periodic nature, allowing for
+extremely precise probabilistic period recovery.
 However, such physical models require many free parameters in order to
@@
 Considering that many stars have on the order of hundreds of spots, the number
-of free parameters quickly becomes unwieldy, especially if the posterior PDFs
-of these parameters are explored with MCMC.
+of free parameters quickly becomes unwieldy, especially if the posterior
+Probability Density Functions (PDFs) of these parameters are explored with
+MCMC.
 Simplified spot models, such as the one described in \citet{Lanza2014} where
@@
 to perfectly capture non-sinusoidal, QP behaviour.
-These requirements are perfectly fulfilled by a Gaussian process (GP) model.
+These requirements are fulfilled by a Gaussian process (GP) model.
 
@@
 representative uncertainty.
-And thirdly, to allow for an additional correlated noise model to be included
+Thirdly, to allow for an additional correlated noise model to be included
 during regression, the parameters of which could be marginalised over.
@@
 GPs are commonly used in the machine learning community and increasingly
-in other scientific fields, for example biology and geophysics (where GP
-regression is called `kriging').
+in other scientific fields, for example biology and geophysics.
 More recently, GPs have been used in the astronomy literature \citep[see
@@
 % star that rotates once every $\sim$ 15 days.
-This is the \kepler\ light curve of KID 1430163, a relatively active star that
+This is the \kepler\ light curve of KIC 1430163, a relatively active star that
 rotates once every $\sim$ 4 days.
@@
 Points that are close together in time are tightly correlated and those that
-are far apart are loosely correlated.
-It is the way in which the covariance varies with the separation between data
-points that is modelled when using GP regression.
-It is not the data but the {\it covariance structure} that is modelled.
+are widely separated in time are loosely correlated.
+The way in which the covariance varies with the separation between data points
+is modelled when using GP regression; it is not the data but the {\it
+covariance structure} that is modelled.
 This fact is what gives GPs their flexibility---they can model any time
@@
 % \ref{fig:GP_example}, shown in orange.
-The light curve of this star has been modelled with two different GP kernel
-functions in figure \ref{fig:GP_example}, shown in blue and pink.
-Both provide adequate fits to the data, however only the periodic kernel
-function, `QP' is a useful model because it has a periodic parameter.
-I return to this point shortly.
+In figure \ref{fig:GP_example} the light curve of KIC 1430163 has been
+modelled with a GP.
+% Both provide adequate fits to the data, however only the periodic kernel
+% function, `QP' is a useful model because it has a periodic parameter.
+% I return to this point shortly.
 
@@
 For example, the simplest and most commonly used kernel function, the `Squared
-Exponential' (SE) produces an adequate fit to the KID 1430163 light curve.
+Exponential' (SE) produces an adequate fit to the KIC 1430163 light curve.
 The SE kernel function is defined as,
@@
 k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2} -
-\frac{\sin^2(\frac{2\pi}{P})}{\Gamma^2} \right).
+\Gamma^2 \sin^2(\frac{2\pi}{P}) \right).
 \end{equation}
@@
 $P$ can be interpreted as the rotation period of the star and $\Gamma$
-controls the number of zero crossings per period.
+controls the amplitude of the sinusoidal term.
 This kernel function allows two data points that are separated in time by one
@@
 period can be weakly correlated.
-Both the QP and SE kernel functions were used to produce the blue and pink
-models shown in figure ~\ref{fig:GP_example}.
+Both the QP kernel function was used to produce the
+model shown in figure ~\ref{fig:GP_example}.
 
@@
 \caption[A light curve with a GP model.]
-{Light curve of KID 1430163, an active star with a rotation period of $\sim$ 4
+{Light curve of KIC 1430163, an active star with a rotation period of $\sim$ 4
 days.

### 2016-11-30 documents/ms.tex
running on a kepler target and making python 3 compatible

@@
 \newcommand{\uw}{3}
+\newcommand{\princeton}{4}
 \newcommand{\nlightcurves}{333}
@@
 
-\title{Inferring stellar rotation periods with Gaussian processes}
+\title{Inferring stellar rotation periods using Gaussian processes}
 
 \author{%
-   Ruth Angus\altaffilmark{\columbia}
-   Suzanne Aigrain\altaffilmark{\oxford}
+   Ruth Angus\altaffilmark{\columbia},
+   Suzanne Aigrain\altaffilmark{\oxford},
+   Timothy Morton\altaffilmark{\princeton}
    \& Daniel Foreman-Mackey\altaffilmark{\uw}
@@
 WA}
+\altaffiltext{\princeton}{Department of Astronomy, Princeton University}
 
@@
 For example, take the time-series in figure \ref{fig:GP_example}.
-% This is a \kepler\ light curve of Earth-like planet host Kepler-452: a G-type
-% star that rotates once every $\sim$ 15 days.
 This is the \kepler\ light curve of KIC 1430163, a relatively active star that
@@
 period can be weakly correlated.
-Both the QP kernel function was used to produce the
-model shown in figure ~\ref{fig:GP_example}.
+Both the QP kernel function was used to produce the model shown in figure
+~\ref{fig:GP_example}.
 
@@
 using MCMC.
-This latter approach comes at a cost: a GP model is expensive to compute once
-as it requires a matrix inversion and determinant evaluation, let alone the
-many thousands of times that is necessary to fully explore the posteriors of
-the parameters.
-However, fully exploring the posterior PDF of $P$, allows for an accurate
+This approach comes at a cost: a GP model is expensive to compute once as it
+requires a matrix inversion and determinant evaluation, let alone the many
+thousands of times that is necessary to fully explore the posteriors of the
+parameters.
+However, fully exploring the posterior PDF of $P$ allows for an accurate
 estimate of the uncertainty on the rotation period.
@@
 days.
-The blue solid line shows a fit to the data using
-a Gaussian process model with a QP covariance kernel function and the pink
-broken line shows a fit with a SE kernel function.
-Both kernel functions provide an adequate fit to the data, however we use the
-QP function because it has a useful parameter that corresponds to the star's
-rotation period.}
+The blue line shows a fit to the data using
+a Gaussian process model with a QP covariance kernel function.}
 \label{fig:GP_example}
@@
 a set of simulated light curves.
-We simulated \nlightcurves\ light curves using a spot model similar to that
-used
-% We used light curves simulated
-for the \citet{Aigrain2015} `hare and hounds' rotation period recovery
-experiment.
+We used \nlightcurves\ light curves generated for the \citet{Aigrain2015}
+`hare and hounds' rotation period recovery experiment.
 These light curves were simulated by placing dark, circular spots with slowly
@@
 % there are \nlightcurves.
-Unlike in their study we are not interested in recovering differential
-rotation in this work, so did not include it in our simulations.
+For the purpose of demonstrating that our method can recover rotation periods,
+we selected only the \nlightcurves\ out of the 1000 without differential
+rotation.
 We opted to use only solid-body rotators because differential rotation may
 produce some additional scatter in the rotation period measurements recovered.
-This code can also be adjusted to produce more realistic light curves by
-altering spot lifetimes.
-Stars with spot lifetimes that are long relative to their rotation periods
-will have highly periodic light curves.
-Those with short spot lifetimes will be more quasi-periodic.
-We fixed the mean spot lifetime at an arbitrary value of 30.5 days for all
-light curves in these initial tests.
-In future we plan to include both differential rotation and variable spot
-lifetimes in our light curve model.
-Light curves were simulated with a real \Kepler\ long-cadence time array, with
+In future we intend to test whether we can recover differential rotation using
+the GP method and will then use the full set of 1000 light curves.
+The distributions of input parameters used to generate these light curves are
+listed in \citet{Aigrain2015}.
+Light curves were simulated with a real \Kepler\ long-cadence time array:
 one data point every thirty minutes over a four year duration.
-The rotation periods of the simulations were randomly drawn from a log-uniform
-distribution between 0.5-60 days.
-Figure \ref{fig:noise-free_lc} shows an example of a simulated, noise-free
-light curve with a period of 4.2 days.
+90\% of the rotation periods of the simulations were randomly drawn from a
+log-uniform distribution between 10 and 50 days and 10\% from a log-uniform
+distribution between 1 and 10 days.
+The light curves were also generated with a range of stellar inclination
+angles, activity levels, spot lifetimes and more.
+Once simulated, the light curves were added to real \kepler\ light curves with
+no obvious astrophysical variability in order to preserve the noise properties
+of the time series.
+Figure \ref{fig:demo_lc} shows an example of a simulated, noise-free
+light curve with a period of 17.04 days.
 
-% The ranges and distributions of the physical stellar parameters used in the
-% simulated light curves are tabulated below, in table
-% \ref{tab:simulation_parameters}.
+The ranges and distributions of the physical stellar parameters used in the
+simulated light curves are tabulated below, in table
+\ref{tab:simulation_parameters}.
 
-% \begin{table*}
-% \begin{center}
-% \caption{Ranges and distributions of parameters used to simulate light curves
-% in \citet{aigrain2015}}
-% \begin{tabular}{lcc}
-% \hline\hline
-%     Parameter & Range & Distribution \\
-%     \hline
-%     Rotation period, $P_{rot}$ & 10 - 50 days (90\%) & log uniform \\
-%     & 1 - 10 days (10\%) & log uniform \\
-%     Activity cycle length & 1 - 10 years & log uniform \\
-%     Inclination & 0 - 90$^\circ$ & Uniform in $\sin^2i$ \\
-%     Decay timescale & (1 - 10) $\times P_{rot}$ & log uniform \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table*}
-% \label{tab:simulation_parameters}
+\begin{table*}
+\begin{center}
+\caption{Ranges and distributions of parameters used to simulate light curves
+in \citet{aigrain2015}}
+\begin{tabular}{lcc}
+\hline\hline
+    Parameter & Range & Distribution \\
+    \hline
+    Rotation period, $P_{rot}$ & 10 - 50 days (90\%) & log uniform \\
+    & 1 - 10 days (10\%) & log uniform \\
+    Activity cycle length & 1 - 10 years & log uniform \\
+    Inclination & 0 - 90$^\circ$ & Uniform in $\sin^2i$ \\
+    Decay timescale & (1 - 10) $\times P_{rot}$ & log uniform \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+\label{tab:simulation_parameters}
 
-% and figure \ref{fig:period_hist} shows the
-% distribution of rotation periods in the \citet{aigrain15} sample.
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/acf_compare_noise-free_hist.pdf}
-% \caption{A histogram of the rotation periods used to generate the simulated
-% light curves in \citet{aigrain15}.}
-% \label{fig:period_hist}
-% \end{center}
-% \end{figure}
+and figure \ref{fig:period_hist} shows the
+distribution of rotation periods in the \citet{aigrain15} sample.
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/period_hist.pdf}
+\caption{A histogram of the rotation periods used to generate the simulated
+light curves in \citet{aigrain15}.}
+\label{fig:period_hist}
+\end{center}
+\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/noise-free_lc.pdf}
-% \includegraphics[width=6in, clip=true]{figures/thesis_plot.pdf}
+\includegraphics[width=6in, clip=true]{figures/demo_lc.pdf}
 \caption[A simulated light curve.]
-{An example simulated, noise-free light curve. This `star' has a
-rotation period of 4.2 days.}
-\label{fig:noise-free_lc}
+{An example simulated light curve. This `star' has a rotation period of 20.8
+days.}
+\label{fig:demo_lc}
 \end{center}
@@
 
-Initial tests were conducted on noise-free light curves, in order to provide
-proof-of-concept.
 We attempted to recover the rotation periods of these \nlightcurves\ light
-curves
-% , both with and without noise, (\ie before {\it and} after being injected
-% into \kepler light curves)
-using three rotation period recovery methods: the ACF method, the LS
+curves using three rotation period recovery methods: the ACF method, the LS
 periodogram method and the GP method.
@@
 produce a light curve with two dips per rotation period.
-% By extension, if three active longitudes existed at 60$^\circ$ separations on
-% the stellar surface, the first peak in the ACF may be at one third of the true
-% rotation period and so on.
-An example ACF of the light curve in figure~\ref{fig:noise-free_lc} is shown
+An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
 in figure \ref{fig:ACF_example}.
 
-\begin{figure*}
+\begin{figure}
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/noise-free_acf.pdf}
-\caption[An ACF of a simulated light curve.]
-{The autocorrelation function of the simulated, noise-free light
-curve depicted in figure \ref{fig:noise-free_lc}.}
-\label{fig:ACF_example}
+\includegraphics[width=6in, clip=true]{figures/demo_ACF.pdf}
+\caption[ACF of a simulated light curve.]
+{An autocorrelation function of the simulated light curve shown in figure
+\ref{fig:demo_lc}.
+The vertical blue line shows the period measured using the ACF method (19.5
+days) and the pink dashed line shows the period that was used to simulate the
+light curve (20.8 days).}
+\label{fig:demo_lc}
 \end{center}
-\end{figure*}
+\end{figure}
 
@@
 rotation period measurements.
-Another advantage of the ACF method is that it is conveniently fast to
-implement.
+Another advantage of the ACF method is that it is fast to implement.
 
-We applied the ACF method to our sample of \nlightcurves\ noise-free,
-simulated light curves.
+We applied the ACF method to the sample of \nlightcurves\ simulated light
+curves.
 Periods measured using the ACF method are plotted against the original
@@
 true periods.
-We believe this is a feature of the peak position measurement method that is
-performed on the ACF.
-ACFs of stellar light curves can be roughly described as a cosine function
-super-imposed on top of a decaying exponential.
-In such a function the peak positions can be shifted towards the left--the
+This is a feature of the peak position measurement method that is performed on
+the ACF.
+ACFs of stellar light curves are periodic functions added to decaying
+exponentials.
+In such functions the peak positions can be shifted towards the left--the
 short period end--because the decaying exponential raises the left side of
 each peak more than the right.
-It is possible to model this effect of course, however the standard practice
-is to simply measure the peak position without taking it into account.
-We are still investigating the origin and implications of this effect.
+It is possible to model this effect, however the standard practice is to
+simply measure the peak position without taking it into account.
+Additionally, the presence of gaps in the data may be contributing to this
+effect.
+Although, in our ACF method, gaps are interpolated over, these gaps still
+affect the ACF by reducing its peak heights.
+This amplifies the effect described above, in which the period is
+underestimated due to the exponential decay in the ACF.
 % To demonstrate that this effect reproduces the underestimated periods seen in
@@
 \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/acf_compare_noise-free.pdf}
-\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
+\includegraphics[width=6in, clip=true]{figures/compare_acf.png}
 \caption[ACF results.]
-{Rotation periods measured using the ACF method as a function of the
-`true' period used to generate \nlightcurves\ simulated, noise-free light
-curves.}
-\label{fig:compare_noise_free}
+{The `true' rotation periods used to generate \nlightcurves\ simulated light
+curves vs the rotation periods measured using the ACF technique.}
+\label{fig:compare_acf}
 \end{center}
@@
 The resulting recovered rotation periods are plotted as a function of the true
-periods in figure \ref{fig:pgram_compare_noise_free}.
+periods in figure \ref{fig:pgram_compare}.
 These recovered rotation periods are in general more accurate than the ACF
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
+\includegraphics[width=6in, clip=true]{figures/compare_pgram.png}
 \caption[LS periodogram results.]
-{Rotation periods measured using the LS periodogram method as a
-function of the `true' rotation period used to generate \nlightcurves\
-simulated light curves.}
-\label{fig:pgram_compare_noise_free}
+{The `true' rotation periods used to generate \nlightcurves\ simulated light
+curves vs the rotation periods measured using a LS periodogram technique.}
+\label{fig:pgram_compare}
 \end{center}
@@
 decision is discussed further in section \textsection \ref{sec:discussion}).
-We use a uniform prior over rotation period with bounds described below and
-assert that the covariance decay timescale parameter, $l$ must be greater than
-the rotation period.
-This represents our assumption that the evolution timescale of active regions
-is greater than stellar rotation periods.
-This assumption may be more valid for late spectral types---hot stars have
-smaller active regions which are likely to be shorter-lived.
-For the remaining hyperparameters we use the following initial values and
-log-uniform prior distributions:
+We use Gausssian distributions in the natural logarithm of all the parameters
+of our GP model.
+The means and variances of the priors used were learned during the development
+of this method.
+In early stages of development we used very broad priors on all the parameters
+uniform in log-space between -20 and 20.
+We then examined the distributions of the model parameters for `sucessful' and
+`unsuccessful' rotation period recoveries, where a recovery was termed
+successful if it lay within 10\% of the injected period.
+For the successful cases, the parameter distributions were narrow Gaussians.
+The parameter distributions of the unsuccessful cases were broad Gaussians
+with similar means.
+We adopted the means and standard deviations of the narrow Gaussians
+describing the parameter distributions of the successful cases to describe our
+Gaussian priors.
+The means and standard deviations of these $\ln$ Gaussian priors are given in
+table \ref{tab:priors}.
+
+\begin{table*}
+\begin{center}
+\caption{Means and standard deviations of the Gaussian priors on the natural
+logarithms of the GP model parameters.}
+\begin{tabular}{lcc}
+Parameter & Mean & Standard Deviation \\
+    \hline
+    $A$ & -13 & 2.7 \\
+    $l$ & 6.2 & 1.5 \\
+    $\Gamma$ & -1.4 & 1.5 \\
+    Rotation period, $P_{rot}$ & ACF period & ACF period $\times 2$ \\
+    $\sigma$ & -17 & 5 \\
+\end{tabular}
+\end{center}
+\end{table*}
+\label{tab:priors}
+The parameter positions were also initialised at the means of these
+distributions during MCMC posterior exploration.
+
+In addition, we set the following bounds on the rotation period and the $l$
+parameter.
 \begin{eqnarray}
- 	&	A_{initial} = e^{-5}, \sim \exp(U[-20:20]) \\ \nonumber
- 	&	l_{initial} = e^{7}, \sim \exp(U[-20:20]), l<P \\ \nonumber
- 	&	\Gamma_{initial} = e^{0.6}, \sim \exp(U[-20:20]) \\ \nonumber
- 	&	\sigma_{initial} = e^{-16}, \sim \exp(U[-20:20]) \\ \nonumber
- 	&	P_{initial} = P_{ACF}, \sim U([1 - 0.4]P_{ACF}<P_{ACF}<[1 +
-0.4]P_{ACF}).
+ 	&	\ln(0.5) < \ln(P_{\mathrm{rot}}) < \ln(T/2) \\ \nonumber
+    &   0 < \ln(l) \\
 \end{eqnarray}
- \label{eq:initialisation}
+\label{eq:priors}
+Where $T$ is the total timespan of the light curve.
+We limit rotation periods to be greater than half a day because we are chiefly
+interested in rotation of FGK stars which are likely to be greater than this
+lower limit, however this restriction could be relaxed in principle.
+We also limit the rotation periods to be less than half of the total timespan
+of the data since this is the longest period we expect to reliable recover.
+We prevent $\ln(l)$ from falling beneath zero both because we expect that
+astrophysical signals in light curves should have longer correlation
+timescales than this and in practice we found that if $l$ is unrestricted, it
+can occasionally wander towards a small value, resulting in an incorrect
+rotation period estimate.
+
 $\sigma$ is an additional white noise term added to the diagonal elements of
@@
 uncertainties have been underestimated.
-If the errorbars reported on the data are too small, this parameter will be
-non-zero.
-In practice this parameter should always be non-zero when performing GP
-regression.
+If the uncertainties reported on the data are too small, this parameter will
+be non-zero.
+In practice however, this parameter should always be non-zero when performing
+GP regression.
 This is because the covariance matrix must be positive definite, however
 matrix inversion performed using most solvers is approximate, not exact,
-therefore slight deviations from positive definitism can arise.
+therefore slight deviations from positive definiteness can arise.
 Including a small amount of extra variance in the model allows enough
@@
 We subsampled the light curves in order to reduce computation time.
-The subsampling is altered for each light curve and depends, again, on the ACF
-period estimate.
-We found that retaining only 20 data points per rotation period, based on the
-ACF period estimate, significantly reduced computation time whilst maintaining
-performance.
+We randomly subsampled the data, retaining an average of four data points per
+day.
+This allowed us to recover periods down to half a day --- the minimum
+timescale of interest for this project, whilst reducing computational costs.
 The {\tt george} \citep{George} python package was used to implement the GP
@@
 sampler to explore the posterior PDFs of the model parameters.
-The rotation period was taken as the median value of the posterior PDF.
+The rotation period was taken as the sample with the maximum likelihood value.
 The resulting measured rotation periods are compared to the true rotation
-periods in figure \ref{fig:GP_compare_noise_free}.
-The root-mean-square (RMS) of residuals for the GP results is 0.26 days.
-The results of all three methods are plotted on the same axes in figure
-\ref{fig:compare_noise_free}.
-
-% The subsampling reduces computation time simply by contracting the data
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_gp.pdf}
-\caption[GP results.]
-{Measured vs true rotation period for \nlightcurves\ simulated light
-curves using the GP method.}
-\label{fig:GP_compare_noise_free}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare2.pdf}
-\caption[All results.]
-{Measured vs true rotation period for \nlightcurves\ simulated light
-curves from the three different methods described in the text.}
-\label{fig:compare_noise_free}
-\end{center}
-\end{figure*}
+periods in figure \ref{fig:compare_mcmc}.
 
@@
 % \end{figure*}
-
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/prediction.pdf}
-% \caption{A noisy simulated light curve with a rotation period of 17.04 days.
-% The blue curve shows the GP model with the best-fit parameters.}
-% \end{center}
-% \end{figure*}
-% \label{fig:compare_noise_free}
 

### 2016-11-30 code/data/download_NGC6819.py

@@ -1,9 +1,0 @@
-import numpy as np
-import kplr
-client = kplr.API()
-
-ids = np.genfromtxt("NGC6819_members.txt", dtype=str).T
-for id in ids:
-    print id
-    star = client.star(id)
-    star.get_light_curves(fetch=True, short_cadence=False)

### 2016-11-30 code/paper_plots.py

@@ -30,9 +30,10 @@ def acf_demo(x, y):
     plt.clf()
     plt.plot(lags, acf, "k")
     plt.axvline(period, color="CornFlowerBlue")
+    plt.axvline(20.8, color="LightCoral", ls="--")
     plt.xlabel("$\mathrm{Lags~(Days)}$")
     plt.ylabel("$\mathrm{Autocorrelation}$")
-    plt.xlim(0, 470)
+    plt.xlim(0, 200)  # 450
     plt.savefig(os.path.join(FIG_DIR, "demo_ACF.pdf"))
 
 

### 2016-11-30 gprotation/GProtation.py

@@ -92,8 +92,8 @@ def make_plot(sampler, xb, yb, yerrb, ID, RESULTS_DIR, trths, traces=False,
 
     nwalkers, nsteps, ndims = np.shape(sampler)
     flat = np.reshape(sampler, (nwalkers * nsteps, ndims))
-    mcmc_res = map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
-                      zip(*np.percentile(flat, [16, 50, 84], axis=0)))
+    mcmc_res = list(map(lambda v: (v[1], v[2]-v[1], v[1]-v[0]),
+                      zip(*np.percentile(flat, [16, 50, 84], axis=0))))
     med = np.concatenate([np.array(mcmc_res[i]) for i in
                           range(len(mcmc_res))])
     print("median values = ", med[::3])
@@ -105,7 +105,10 @@ def make_plot(sampler, xb, yb, yerrb, ID, RESULTS_DIR, trths, traces=False,
     r = np.concatenate((maxlike, med))
 
     # calculate autocorrelation times
-    acorr_t = emcee.autocorr.integrated_time(flat)
+    try:
+        acorr_t = emcee.autocorr.integrated_time(flat, c=1)
+    except:
+        acorr_t = emcee.autocorr.integrated_time(flat)
 
     # save data
     df = pd.DataFrame({"N": [ID], "A_max": [r[0]], "l_max": [r[1]],
@@ -135,14 +138,9 @@ def make_plot(sampler, xb, yb, yerrb, ID, RESULTS_DIR, trths, traces=False,
 
     if tri:
         DIR = "../code/simulations/kepler_diffrot_full/par/"
-        truths = pd.read_csv(os.path.join(DIR, "final_table.txt"),
-                             delimiter=" ")
-        true_p = np.log(truths.P_MIN.values[truths.N.values ==
-                                    int(filter(str.isdigit, ID))][0])
         print("Making triangle plot")
         fig = corner.corner(flat[:, :-1], labels=fig_labels,
-                            quantiles=[.16, .5, .84], show_titles=True,
-                            truths=trths)
+                            quantiles=[.16, .5, .84], show_titles=True)
         fig.savefig(os.path.join(RESULTS_DIR, "{0}_triangle".format(ID)))
         print(os.path.join("{0}_triangle.png".format(ID)))
 

### 2016-11-30 gprotation/Kepler_ACF.py

@@ -173,10 +173,10 @@ def corr_run(time, flux, flux_err, id, savedir, saveplot=True):
         pylab.plot(per_cent, var_arr_real, 'k.')
         pylab.axhline(amp_per[x], ls = '--', c = 'b')
         pylab.xlim(lc_tab.time.min(),lc_tab.time.max())
-	try:
-	    pylab.ylim(var_arr_real.min(), var_arr_real.max())
-	except ValueError:
-	    pass
+    try:
+        pylab.ylim(var_arr_real.min(), var_arr_real.max())
+    except ValueError:
+        pass
         ax = pylab.gca()
         pylab.text(0.415, -0.15, 'Time (days)', transform = ax.transAxes)
         pylab.text(0.415, -1.4, 'Period (days)', transform = ax.transAxes)

### 2016-11-30 gprotation/kepler_data.py

@@ -6,6 +6,7 @@ import glob
 from simple_acf import simple_acf
 from measure_GP_rotation import bin_data
 import h5py
+import os
 import time
 from GProtation import lnprob
 from multiprocessing import Pool
@@ -14,12 +15,13 @@ from george.kernels import ExpSquaredKernel, ExpSine2Kernel, \
         WhiteKernel
 
 
-def load_kepler_data(fnames):
+def load_kepler_data(KID, LC_DIR):
     """
     load and join quarters together.
     Takes a list of fits file names for a given star.
     Returns the concatenated arrays of time, flux and flux_err
     """
+    fnames = glob.glob(os.path.join(LC_DIR, "*fits"))
     hdulist = pyfits.open(fnames[0])
     t = hdulist[1].data
     time = t["TIME"]

### 2016-11-30 gprotation/simple_acf.py
removing trend from first 100 daus

@@ -37,12 +37,19 @@ def simple_acf(x, y):
     # just use the second bit (no reflection)
     acf_smooth, lags = acf_smooth[N:], lags[N:]
 
-    # cut it in half
+    # cut it in half (and reduce to 100 days)
     m = lags < max(lags)/2.
+    m = (lags < max(lags)/2.) * (lags < 100)
     acf_smooth, lags = acf_smooth[m], lags[m]
 
     # ditch the first point
     acf_smooth, lags = acf_smooth[1:], lags[1:]
+
+    # fit and subtract straight line
+    AT = np.vstack((lags, np.ones_like(lags))
+    ATA = np.dot(AT, AT.T)
+    m, b = np.linalg.solve(ATA, np.dot(AT, acf_smooth))
+    acf_smooth -= m*lags + b
 
     # find all the peaks
     peaks = np.array([i for i in range(1, len(lags)-1)

### 2016-11-30 gprotation/GProt.py
merging

@@ -5,7 +5,7 @@
 from __future__ import print_function
 import numpy as np
 from GProtation import make_plot, lnprob, Glnprob, Glnprob_split
-from Kepler_ACF import corr_run
+import Kepler_ACF
 import simple_acf as sa
 import h5py
 from gatspy.periodic import LombScargle

### 2016-11-30 code/simulations/kepler_diffrot_full/results/filter.py
deleting old files

@@ -1,240 +1,0 @@
-import numpy
-from scipy.signal import signaltools
-from scipy import interpolate
-import pylab
-import scipy
-
-def medsig(array):
-    med = numpy.median(array)
-    sig = 1.48 * numpy.median(abs(array - med))
-    return med, sig
-
-def filt1d(array, nmed, nlin, fill = False, circ = False):
-    """Nonlinear (median+boxcar) filter with edge reflection."""
-    nmed = 2 * int(nmed/2) + 1
-    N = len(array)
-    if N < (3*nmed): 
-        return array
-    # Check for NaNs
-    lnan = numpy.isnan(array)
-    nnan = sum(lnan)
-    if (nnan != 0):
-        lnotnan = numpy.where(lnan==0)
-        s = numpy.shape(lnotnan)
-        if (len(s)>1):
-            s = numpy.size(lnotnan)
-            lnotnan = numpy.reshape(lnotnan, s)
-            med = numpy.median(array[lnotnan])
-        # Fill in any data gaps by interpolating over them ...
-        work = numpy.zeros(N)
-        il = numpy.min(lnotnan)
-        ih = numpy.max(lnotnan)
-        xnew = numpy.arange(ih-il+1) + il
-        f = interpolate.interp1d(lnotnan, array[lnotnan])  
-        work[il:ih+1] = f(xnew)
-        # ... or extending slope of nearest data points if at the edges
-        if (il!=0):
-            slope = work[il+1] - work[il]
-            for i in range(il): work[i] = work[il] - slope*(il-i)
-        if (ih!=N-1):
-            slope = work[ih] - work[ih-1]
-            for i in range(N-ih-1)+ih+1: work[i] = work[ih] + slope*(i-ih)
-    else:
-        work = numpy.copy(array)
-    # Edge reflection
-    nr = min(20, nlin)
-    sz = max([nmed, nlin])
-    if sz >= (N-1): 
-        sz = N-2
-    if circ != False:
-        wl = work[N-sz:]
-        wr = work[:sz]
-    else:
-        wl = array[0:sz]
-        pivot = numpy.median(array[:nr])
-        wl = 2 * pivot - wl
-        wl = wl[::-1] # reverse array
-        wr = array[N-sz:N]
-        pivot = numpy.median(array[N-nr:N])
-        wr = 2 * pivot - wr
-        wr = wr[::-1] # reverse array
[…]

### 2016-11-30 code/simulations/kepler_diffrot_full/results/kepler_data.py

@@ -1,200 +1,0 @@
-from __future__ import print_function
-import numpy as np
-import matplotlib.pyplot as plt
-import pyfits
-import glob
-from measure_GP_rotation import bin_data
-import h5py
-import time
-from GProtation import lnprob
-from multiprocessing import Pool
-import george
-from george.kernels import ExpSquaredKernel, ExpSine2Kernel, \
-        WhiteKernel
-
-
-def load_kepler_data(fnames):
-    """
-    load and join quarters together.
-    Takes a list of fits file names for a given star.
-    Returns the concatenated arrays of time, flux and flux_err
-    """
-    hdulist = pyfits.open(fnames[0])
-    t = hdulist[1].data
-    time = t["TIME"]
-    flux = t["PDCSAP_FLUX"]
-    flux_err = t["PDCSAP_FLUX_ERR"]
-    q = t["SAP_QUALITY"]
-    m = np.isfinite(time) * np.isfinite(flux) * np.isfinite(flux_err) * \
-            (q == 0)
-    x = time[m]
-    med = np.median(flux[m])
-    y = flux[m]/med - 1
-    yerr = flux_err[m]/med
-    for fname in fnames[1:]:
-       hdulist = pyfits.open(fname)
-       t = hdulist[1].data
-       time = t["TIME"]
-       flux = t["PDCSAP_FLUX"]
-       flux_err = t["PDCSAP_FLUX_ERR"]
-       q = t["SAP_QUALITY"]
-       m = np.isfinite(time) * np.isfinite(flux) * np.isfinite(flux_err) * \
-               (q == 0)
-       x = np.concatenate((x, time[m]))
-       med = np.median(flux[m])
-       y = np.concatenate((y, flux[m]/med - 1))
-       yerr = np.concatenate((yerr, flux_err[m]/med))
-    return x, y, yerr
-
-
-def fit(x, y, yerr, id, p_init, plims, DIR, burnin=500, run=1500, npts=48,
-        nwalkers=32, cutoff=1000, sine_kernel=True, runMCMC=True, plot=False,
-        opt=False):
-    """
-    takes x, y, yerr and initial guesses and priors for period and does the
-    the GP MCMC.
-    id: the kepler id
-    p_init: period initial guess
-    plims: tuple, upper and lower limit for the prior
-    DIR: the directory to save the results
-    npts: number of points per bin (48 for one bin per day)
[…]

### 2016-11-30 code/simulations/kepler_diffrot_full/results/mpfit.py

@@ -1,1298 +1,0 @@
-import numpy
-
-"""mpfit: Non-linear least squares minimisation using the
-   Levenberg-Marquardt technique with optional fixed, limited or tied
-   parameters.
-
-   This module is intended primarily for fitting non-linear models to
-   data.  The user should supply a Python function which, given the
-   data and parameters, computes an array of weighted deviations
-   between model and data. The user can then use the function
-   mpfit.mpfit to minimise the sum of the squares of the deviates,
-   i.e. the chi-squared.
-
-   The function mpfit.mpfit takes as arguments the user defined
-   function and a numpy array containing inital guesses for the
-   parameters. The data, and any measurement uncertainties, are passed
-   to the user-defined function using the functkw keyword of
-   mpfit.mpfit. It returns an instance of the mpfit.results class,
-   which contains the best-fit values of the parameters as well as the
-   covariance matrix and formal parameter uncertainties.
-
-   Here is a quick example using a user function defined in this module:
-      import mpfit
-      import numpy
-      import pylab
-      N = 100
-      x = numpy.r_[0:10:N*1j]
-      p0 = numpy.array([5.7, 0.22])
-      noise = mpfit.F(x, p0) + numpy.random.randn(N) * 10
-      err = numpy.ones(N) * 10
-      pylab.errorbar(x, y, yerr = err)
-      p1 = numpy.array([6,0.2])
-      fa = {'x':x, 'y':y + noise, 'err':err}
-      m = mpfit(mpfit.example_func, p1, functkw = fa)
-      print 'parameters = ', m.params
-      print 'parameters = ', m.pcerror 
-# Note how the errors are underestimated because the two parameters
-# are strongly correlated!
-      pylab.plot(x, mpfit.F(x, m.params))
-      
-   The user function should be built according to a specific set of
-   rules, which are described below and illustrated in the example_func
-   function.
-
-   Simple constraints can be placed on parameter values by using the
-   fixed, limited & limits, and/or tied keywords of mpfit.mpfit, as
-   described in the documentation of function mpfit itself.
-   *** NB: in the present version, only the fixed keyword is
-       functional and tested. The limited / limits keywords don't work
-       properly and the tied keyword is untested. ***
-
-   See the documentation for the individual functions and classes for
-   more details, including some keywords not described here.
-
-   USER_DEFINED FUNCTION:
-
[…]

### 2016-11-30 code/simulations/kepler_diffrot_full/results/plotstuff.py

@@ -1,26 +1,0 @@
-import matplotlib.pyplot as plt
-
-def params():
-    plotpar = {'axes.labelsize': 18,
-               'text.fontsize': 10,
-               'legend.fontsize': 15,
-               'xtick.labelsize': 18,
-               'ytick.labelsize': 18,
-               'text.usetex': True}
-    plt.rcParams.update(plotpar)
-    return {'capsize':0, 'fmt':'k.', 'ecolor':'.8'}
-
-class colours(object):
-    def __init__(self):
-
-        self.orange = '#FF9933'
-        self.lightblue = '#66CCCC'
-        self.blue = '#0066CC'
-        self.pink = '#FF33CC'
-        self.turquoise = '#3399FF'
-        self.lightgreen = '#99CC99'
-        self.green = '#009933'
-        self.maroon = '#CC0066'
-        self.purple = '#9933FF'
-        self.red = '#CC0000'
-        self.lilac = '#CC99FF'

### 2016-11-30 code/simulations/kepler_diffrot_full/results/thesis_plot.py

@@ -1,31 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from plotstuff import params, colours
-reb = params()
-cols = colours()
-
-plotpar = {'axes.labelsize': 20,
-           'text.fontsize': 20,
-           'legend.fontsize': 20,
-           'xtick.labelsize': 18,
-           'ytick.labelsize': 18,
-           'text.usetex': True}
-plt.rcParams.update(plotpar)
-
-data = np.genfromtxt("../par/final_table.txt", skip_header=1).T
-m = data[13] == 0  # just the stars without diffrot
-ids = data[0][m]
-periods = data[-3][m]
-star = 1
-print("period = ", periods[star])
-id = str(int(ids[star])).zfill(4)
-x, y = np.genfromtxt("../noise_free/lightcurve_{0}.txt".format(id)).T
-y -= np.mean(y)
-
-plt.clf()
-plt.plot(x, y, "k")
-plt.xlabel("$\mathrm{Time~(days)}$")
-plt.ylabel("$\mathrm{Normalised~flux}$")
-plt.subplots_adjust(left=.2, bottom=.12)
-plt.xlim(min(x), max(x))
-plt.savefig("thesis_plot")

### 2016-11-30 ruth/colours.py
deleting files in ruth

@@ -1,15 +1,0 @@
-class plot_colours(object):
-
-    def __init__(self):
-
-        self.orange = '#FF9933'
-        self.lightblue = '#66CCCC'
-        self.blue = '#0066CC'
-        self.pink = '#FF33CC'
-        self.turquoise = '#3399FF'
-        self.lightgreen = '#99CC99'
-        self.green = '#009933'
-        self.maroon = '#CC0066'
-        self.purple = '#9933FF'
-        self.red = '#CC0000'
-        self.lilac = '#CC99FF'

### 2016-11-30 ruth/filter.py

@@ -1,240 +1,0 @@
-import numpy
-from scipy.signal import signaltools
-from scipy import interpolate
-import pylab
-import scipy
-
-def medsig(array):
-    med = numpy.median(array)
-    sig = 1.48 * numpy.median(abs(array - med))
-    return med, sig
-
-def filt1d(array, nmed, nlin, fill = False, circ = False):
-    """Nonlinear (median+boxcar) filter with edge reflection."""
-    nmed = 2 * int(nmed/2) + 1
-    N = len(array)
-    if N < (3*nmed):
-        return array
-    # Check for NaNs
-    lnan = numpy.isnan(array)
-    nnan = sum(lnan)
-    if (nnan != 0):
-        lnotnan = numpy.where(lnan==0)
-        s = numpy.shape(lnotnan)
-        if (len(s)>1):
-            s = numpy.size(lnotnan)
-            lnotnan = numpy.reshape(lnotnan, s)
-            med = numpy.median(array[lnotnan])
-        # Fill in any data gaps by interpolating over them ...
-        work = numpy.zeros(N)
-        il = numpy.min(lnotnan)
-        ih = numpy.max(lnotnan)
-        xnew = numpy.arange(ih-il+1) + il
-        f = interpolate.interp1d(lnotnan, array[lnotnan])
-        work[il:ih+1] = f(xnew)
-        # ... or extending slope of nearest data points if at the edges
-        if (il!=0):
-            slope = work[il+1] - work[il]
-            for i in range(il): work[i] = work[il] - slope*(il-i)
-        if (ih!=N-1):
-            slope = work[ih] - work[ih-1]
-            for i in range(N-ih-1)+ih+1: work[i] = work[ih] + slope*(i-ih)
-    else:
-        work = numpy.copy(array)
-    # Edge reflection
-    nr = min(20, nlin)
-    sz = max([nmed, nlin])
-    if sz >= (N-1):
-        sz = N-2
-    if circ != False:
-        wl = work[N-sz:]
-        wr = work[:sz]
-    else:
-        wl = array[0:sz]
-        pivot = numpy.median(array[:nr])
-        wl = 2 * pivot - wl
-        wl = wl[::-1] # reverse array
-        wr = array[N-sz:N]
-        pivot = numpy.median(array[N-nr:N])
-        wr = 2 * pivot - wr
-        wr = wr[::-1] # reverse array
[…]

### 2016-11-30 ruth/gls.py

@@ -1,322 +1,0 @@
-import scipy
-import scipy.linalg
-import pylab
-import mpfit
-#from planet import orbit
-
-mach = scipy.MachAr()
-small = 10 * mach.eps
-
-def sinefit(time, data, err = None, pmin = None, pmax = None, \
-                nper = 500, return_periodogram = False, doplot = False):
-    """Sine curve fit. Period by brute force, other pars linear.
-    per, amp, phase, dc = sinefit(x, y, [err, pmin, pmax, nper])"""
-    npts = len(time)
-    if pmin is None:
-        w = scipy.sort(time)
-        dt = w[1:] - w[:npts-1]
-        tstep = scipy.median(dt)
-        pmin = 2 * tstep
-    if pmax is None: pmax = (time.max() - time.min()) / 2.
-    # even sampling in log space
-    lpmin,lpmax = scipy.log10([pmin,pmax])
-    lpers = scipy.r_[lpmin:lpmax:nper*1j]
-    pers = 10.0**lpers
-    #print 'pmin vs pnewmin', pmin, min(pers)
-    #print 'pmax vs pnewmax', pmax, max(pers)
-
-    if err == None:
-        z = scipy.ones(npts)
-        err = scipy.zeros(npts)
-        mrk = '.'
-    else:
-        if len(err) is len(data):
-            z = 1.0 / err**2
-            mrk = '.'
-        else:
-            z = scipy.ones(npts)
-            err = scipy.zeros(npts)
-            mrk = '.'
-    sumwt = z.sum()
-    chi2_0 = scipy.sum((data-scipy.mean(data))**2*z)
-    p_w = scipy.zeros(nper)
-    a_w = scipy.zeros(nper)
-    p_max = 0.0
-
-    for i in scipy.arange(nper):
-        arg = 2 * scipy.pi * time / pers[i]
-        cosarg = scipy.cos(arg)
-        sinarg = scipy.sin(arg)
-        a = scipy.matrix([[scipy.sum(sinarg**2*z), scipy.sum(cosarg*sinarg*z), \
-                               scipy.sum(sinarg*z)], \
-                              [0, scipy.sum(cosarg**2*z), scipy.sum(cosarg*z)], \
-                              [0, 0, sumwt]])
[…]

### 2016-12-02 gprotation/comparison_plots.py
changing plot size

@@ -250,7 +250,7 @@ def pgram_plot(truths, DIR):
     plt.plot(np.log(xs), np.log(xs) + 2./3, "k--", alpha=.3, zorder=0)
     plt.scatter(np.log(true), np.log(pgram), c=np.log(amp), edgecolor="",
                 cmap="GnBu", vmin=min(np.log(amp)), vmax=max(np.log(amp)),
-                s=50, zorder=1)
+                s=15, zorder=1)
     cbar = plt.colorbar()
     cbar.ax.set_ylabel("$\ln\mathrm{(Amplitude)}$")
     plt.xlim(0, 4)
@@ -266,6 +266,6 @@ if __name__ == "__main__":
     DIR = "../code/simulations/kepler_diffrot_full/par/"
     truths = pd.read_csv(os.path.join(DIR, "final_table.txt"), delimiter=" ")
 
-    print("mcmc sigma rms = ", mcmc_plots(truths, "results_sigma"))
+#     print("mcmc sigma rms = ", mcmc_plots(truths, "results_sigma"))
     print("acf sigma rms = ", acf_plot(truths, "results_sigma"))
     print("pgram sigma rms = ", pgram_plot(truths, "results_sigma"))

### 2016-12-02 gprot/filter.py
added bandpass filtering

@@ -1,8 +1,85 @@
 import numpy as np
+from scipy.signal import butter, lfilter
+
 
 def sigma_clip(x, y, yerr, nsigma):
     med = np.median(y)
     std = (sum((med - y)**2)/float(len(y)))**.5
     m = np.abs(y - med) > (nsigma * std)
 
-    return x[~m], y[~m], yerr[~m]
+    return x[~m], y[~m], yerr[~m]    
+
+def butter_bandpass(lowcut, highcut, fs, order=5):
+    nyq = 0.5 * fs
+    low = lowcut / nyq
+    high = highcut / nyq
+    b, a = butter(order, [low, high], btype='band')
+    return b, a
+
+def butter_bandpass_filter(data, lowcut, highcut, fs, order=5):
+    b, a = butter_bandpass(lowcut, highcut, fs, order=order)
+    y = lfilter(b, a, data)
+    return y
+
+def fill_gaps(x, y, yerr, cadence=1766./86400, make_uniform=True):
+    # Guess cadence if not provided
+    if cadence is None:
+        cadence = np.median(np.diff(x))
+
+    new_x = x.copy()
+    new_y = y.copy()
+    new_yerr = yerr.copy()
+    
+    # Find data gaps.
+    dx = np.diff(x)
+    i_gaps = np.where(dx > 1.5*cadence)[0]
+
+    # Make gap filler arrays, using linear interpolation
+    x_gaps = []
+    y_gaps = []
+    yerr_gaps = []
+    for i in i_gaps:
+        x0, x1 = x[i:i+2]
+        y0, y1 = y[i:i+2]
+        xfill = np.arange(x0 + cadence, x1, cadence)
+        x_gaps.append(xfill)
+        y_gaps.append(y0 + (xfill - x0)*(y1 - y0)/(x1 - x0))
+        yerr_gaps.append(np.ones(len(xfill)) * yerr[i])
+        
+    # Insert gap fillers into gaps
+    shift = 1
+    i_new = []
+    for i, xg, yg, yerrg in zip(i_gaps, x_gaps, y_gaps, yerr_gaps):
+        ind = i + shift
+        new_x = np.insert(new_x, ind, xg)
+        new_y = np.insert(new_y, ind, yg)
+        new_yerr = np.insert(new_yerr, ind, yerrg)
+        n = len(xg)
+        i_new.append(np.arange(ind, ind+n))
+        shift += n
+    
+    if len(i_new) > 0:
+        i_new = np.concatenate(i_new)
[…]

### 2016-12-02 gprot/lc.py

@@ -5,10 +5,12 @@ import pandas as pd
 import matplotlib.pyplot as plt
 from cycler import cycler
 
+from scipy.interpolate import UnivariateSpline
+
 from collections import OrderedDict
 from pkg_resources import resource_filename
 
-from .filter import sigma_clip
+from .filter import sigma_clip, bandpass_filter
 from .plots import tableau20
 
 qtr_times = pd.read_table(resource_filename('gprot', 'data/qStartStop.txt'), 
@@ -42,6 +44,32 @@ class LightCurve(object):
     @property
     def df(self):
         return pd.DataFrame({'x':self.x, 'y':self.y, 'yerr':self.yerr})
+
+    def bandpass_filter(self, pmin=0.5, pmax=100, cadence=None, edge=2000):
+        """Applies a Butterworth bandpass filter to data
+
+        Replaces lightcurve data with new filtered, edge-cropped data.
+        """
+        if cadence is None:
+            try:
+                cadence = self.cadence
+            except AttributeError:
+                pass
+
+        x, y, yerr = bandpass_filter(self._x_full,
+                                     self._y_full,
+                                     self._yerr_full)
+        x = x[edge:-edge]
+        y = y[edge:-edge]
+        yerr = yerr[edge:-edge]
+
+        self._x = x.copy()
+        self._y = y.copy()
+        self._yerr = y.copy()
+
+        self._x_full = x.copy()
+        self._y_full = y.copy()
+        self._yerr_full = yerr.copy()
 
     def best_sublc(self, ndays, npoints=600, 
                     flat_order=3, **kwargs):
@@ -87,7 +115,7 @@ class LightCurve(object):
     def make_best_chunks(self, ndays=[800, 200, 50], seed=None, **kwargs):
         if not hasattr(ndays, '__iter__'):
             ndays = [ndays]
-            
+
         self._x_list = []
         self._y_list = []
         self._yerr_list = []

### 2016-12-02 gprotation/comparison_plots.py
better visibility

@@ -146,9 +146,9 @@ def mcmc_plots(truths, DIR):
 
     plt.errorbar(np.log(true), np.log(maxlike), yerr=[lnerrp, lnerrm],
                  fmt="k.", zorder=1, capsize=0, ecolor=".8", alpha=.5, ms=.1)
-    plt.scatter(np.log(true), np.log(maxlike), c=np.log(amp), edgecolor="",
+    plt.scatter(np.log(true), np.log(maxlike), c=np.log(amp), edgecolor="k",
                 cmap="GnBu_r", vmin=min(np.log(amp)), vmax=max(np.log(amp)),
-                s=15, zorder=2)
+                s=20, lw=.2, zorder=2)
 
     cbar = plt.colorbar()
     cbar.ax.set_ylabel("$\ln\mathrm{(Amplitude)}$")
@@ -165,9 +165,9 @@ def mcmc_plots(truths, DIR):
     plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
     plt.errorbar(np.log(true), np.log(maxlike), yerr=[lnerrp, lnerrm],
                  fmt="k.", zorder=1, capsize=0, ecolor=".8", alpha=.5, ms=.1)
-    plt.scatter(np.log(true), np.log(maxlike), c=mean_acorr, edgecolor="",
+    plt.scatter(np.log(true), np.log(maxlike), c=mean_acorr, edgecolor="k",
                 cmap="GnBu_r", vmin=min(mean_acorr), vmax=max(mean_acorr),
-                s=15, zorder=2)
+                s=20, lw=.2, zorder=2)
     cbar = plt.colorbar()
     cbar.ax.set_ylabel("$<\mathrm{(Autocorrelation~Time)}>$")
     plt.savefig(os.path.join(DIR, "compare_mcmc_convergence"))
@@ -216,9 +216,9 @@ def acf_plot(truths, DIR):
 
     plt.errorbar(np.log(true), np.log(acfs), yerr=(acf_errs/acfs), fmt="k.",
                  capsize=0, ecolor=".7", alpha=.4, ms=1, zorder=1)
-    plt.scatter(np.log(true), np.log(acfs), c=np.log(amp), edgecolor="",
+    plt.scatter(np.log(true), np.log(acfs), c=np.log(amp), edgecolor="k",
                 cmap="GnBu_r", vmin=min(np.log(amp)), vmax=max(np.log(amp)),
[…]

### 2016-12-02 documents/ms.tex
small change
editing the ACF sectino

@@
 days.
-The blue line shows a fit to the data using
-a Gaussian process model with a QP covariance kernel function.}
+The blue line shows a fit to the data using a Gaussian process model with a QP
+covariance kernel function.}
 \label{fig:GP_example}
@@
 simulated light curves are tabulated below, in table
-\ref{tab:simulation_parameters}.
+\ref{tab:simulation_parameters}, and figure \ref{fig:period_hist} shows the
+distribution of rotation periods in the \citet{Aigrain2015} sample.
 
@@
 
-and figure \ref{fig:period_hist} shows the
-distribution of rotation periods in the \citet{aigrain15} sample.
 \begin{figure}
@@
 \caption{A histogram of the rotation periods used to generate the simulated
-light curves in \citet{aigrain15}.}
+light curves in \citet{Aigrain2015}.}
 \label{fig:period_hist}
@@
 
-We calculated an ACF for each light curve using the method of
+We calculated an ACF for each light curve following the method of
 \citet{Mcquillan2013}.
@@
 convolving with a Gaussian with $\sigma=9$ days.
-A rotation period is estimated as the lag-time of the first peak in the ACF,
-unless the second peak is larger in which case {\it that} lag-time is
-interpreted as the true period.
-The second peak in an ACF can be larger than the first if there are two active
-regions at or near opposite longitudes on the surface of the star---this would
-produce a light curve with two dips per rotation period.
+A rotation period is estimated as the lag-time of the highest peak in the ACF,
+less than 100 days.
+This is not always the first peak: the second can be larger than the first if
+there are two active regions at or near opposite longitudes on the surface of
+the star---this would produce a light curve with two dips per rotation period.
+We restrict our search to periods less than 100 days because this is the
+maximum period that is realistically recoverable in \kepler\ light curves ---
+instrumental noise significantly distorts the signal on longer timescales.
 An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
-in figure \ref{fig:ACF_example}.
+in figure \ref{fig:demo_acf}.
 
@@
 \ref{fig:demo_lc}.
-The vertical blue line shows the period measured using the ACF method (19.5
+The vertical blue line shows the period measured using the ACF method (20.4
 days) and the pink dashed line shows the period that was used to simulate the
 light curve (20.8 days).}
-\label{fig:demo_lc}
+\label{fig:demo_acf}
 \end{center}
@@
 Another advantage of the ACF method is that it is fast to implement.
+A clear disadvantage however, as already mentioned above, are the poorly
+defined uncertainties for rotation periods estimated using an autocorrelation
+function.
+This is related to the fact that ACF period detection is a non-probabilistic
+method, another drawback of ACF periods.
 
@@
 rotation period values used to generate the light curves in figure
-\ref{fig:compare_noise_free}.
-% 73\% of the injected rotation periods were recovered with a value lying
-% within 10\% of the truth and 83\% within 20\%.
-A noteworthy feature of this figure is that many of the points fall below the
-$x=y$ line, \ie\ the recovered rotation periods are a little shorter than the
-true periods.
-This is a feature of the peak position measurement method that is performed on
-the ACF.
-ACFs of stellar light curves are periodic functions added to decaying
-exponentials.
-In such functions the peak positions can be shifted towards the left--the
-short period end--because the decaying exponential raises the left side of
-each peak more than the right.
-It is possible to model this effect, however the standard practice is to
-simply measure the peak position without taking it into account.
-Additionally, the presence of gaps in the data may be contributing to this
-effect.
-Although, in our ACF method, gaps are interpolated over, these gaps still
-affect the ACF by reducing its peak heights.
-This amplifies the effect described above, in which the period is
-underestimated due to the exponential decay in the ACF.
-% To demonstrate that this effect reproduces the underestimated periods seen in
-% figure \ref{fig:compare_noise_free}, we simulated 1000 ACFs by generating
-% cosine plus exponential functions with a range of periods, decay timescales
-% and relative amplitudes.
-% The measured position of the first peak in each simulated ACF is compared to
-% the true position in figure \ref{fig:exp_sine_test}.
-% The same underestimation of the peak position is apparent in this figure.
-% Dashed lines show harmonics of $2P_{\mathrm{rot}}$ and $1/2P_{\mathrm{rot}}$.
-% In several cases, twice the true rotation period is measured.
-% The RMS of residuals is 1.59 days.
-
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/exp_sine_test.pdf}
-% \caption[A flaw in the ACF method.]
-% {Recovered peak position as a function of the `true' period used to
-% generate simulated ACFs. In many cases the position is measured short-wards of
-% the truth, as also seen in figure \ref{fig:compare_noise_free}. The dashed
-% lines are at $x=2x$, $x=x$ and $x=\frac{1}{2}x$.}
-% \end{center}
-% \end{figure*}
-% \label{fig:exp_sine_test}
+\ref{fig:compare_acf}.
+The $2\times$ and $1/2\times$ harmonic lines are shown as dashed lines.
+The agreement between the injected and recovered rotation periods is good in
+general, however twice the true rotation period is often found.
+In addition, some rotation periods are drastically over- and underestimated.
 

### 2016-12-06 gprot/acf/__init__.py
basic acf

@@ -1,0 +1,1 @@
+from .acf import acf

### 2016-12-06 gprot/acf/acf.py

@@ -1,0 +1,12 @@
+import numpy as np
+import acor
+
+def acf(x, y, maxlag=100):
+    """Assumes regular sampling, zero-filled
+    """
+    cadence = np.median(np.diff(x))
+    maxlag_cad = maxlag/cadence
+    ac = acor.function(y, maxlag_cad)
+    lags = np.arange(len(ac))*cadence
+    return lags, ac
+

### 2016-12-06 gprot/filter.py

@@ -70,7 +70,7 @@ def fill_gaps(x, y, yerr, cadence=1766./86400, make_uniform=True):
     return new_x, new_y, new_yerr, i_new
 
 def bandpass_filter(x, y, yerr, pmin=0.5, pmax=100, cadence=1766./86400,
-                    edge=2000, order=3):
+                    edge=2000, order=3, zero_fill=False):
     x, y, yerr, i_new = fill_gaps(x, y, yerr)
 
     # Sampling and cutoff frequencies
@@ -80,6 +80,11 @@ def bandpass_filter(x, y, yerr, pmin=0.5, pmax=100, cadence=1766./86400,
 
     yfilt = butter_bandpass_filter(y, lowcut, highcut, fs, order=order) 
 
-    return (np.delete(x, i_new), 
-            np.delete(yfilt, i_new), 
-            np.delete(yerr, i_new))
+    if zero_fill:
+        if len(i_new) > 0:
+            yfilt[i_new] = 0
+        return x, yfilt, yerr
+    else:
+        return (np.delete(x, i_new), 
+                np.delete(yfilt, i_new), 
+                np.delete(yerr, i_new))

### 2016-12-06 gprot/lc.py

@@ -12,6 +12,7 @@ from pkg_resources import resource_filename
 
 from .filter import sigma_clip, bandpass_filter
 from .plots import tableau20
+from .acf import acf
 
 qtr_times = pd.read_table(resource_filename('gprot', 'data/qStartStop.txt'), 
                           delim_whitespace=True, index_col=0)
@@ -51,7 +52,8 @@ class LightCurve(object):
     def df(self):
         return pd.DataFrame({'x':self.x, 'y':self.y, 'yerr':self.yerr})
 
-    def bandpass_filter(self, pmin=0.5, pmax=100, cadence=None, edge=2000):
+    def bandpass_filter(self, pmin=0.5, pmax=100, cadence=None, edge=2000,
+                        zero_fill=False):
         """Applies a Butterworth bandpass filter to data
 
         Replaces lightcurve data with new filtered, edge-cropped data.
@@ -64,7 +66,7 @@ class LightCurve(object):
 
         x, y, yerr = bandpass_filter(self._x_full,
                                      self._y_full,
-                                     self._yerr_full)
+                                     self._yerr_full, zero_fill=zero_fill)
         x = x[edge:-edge]
         y = y[edge:-edge]
         yerr = yerr[edge:-edge]
@@ -83,6 +85,28 @@ class LightCurve(object):
 
         if self.sub is not None:
             self.subsample(self.sub)
+
+    def acf(self, maxlag=100, filter=True):
+        if filter:
+            x, y, yerr = bandpass_filter(self._x_full,
+                                         self._y_full,
+                                         self._yerr_full, zero_fill=True)
+        else:
+            x, y = self.x, self.y
+
+        lags, ac = acf(x, y)
+
+        return lags, ac
+
+    def plot_acf(self, truth=None, **kwargs):
+        lags, ac = self.acf(**kwargs)
+
+        fig, ax = plt.subplots(1,1)
+        ax.plot(lags, ac, 'k')
+        if truth is not None:
+            ax.axvline(truth, c='r', ls=':')
+
+        return fig
 
     def best_sublc(self, ndays, npoints=600, 
                     flat_order=3, **kwargs):

### 2016-12-06 gprot/plots.py

@@ -13,3 +13,28 @@ for i in range(len(tableau20)):
     r, g, b = tableau20[i]
     tableau20[i] = (r / 255., g / 255., b / 255.)    
 
+import os
+import matplotlib.pyplot as plt
+import numpy as np
+
+from emcee3.backends import HDFBackend
+
+from .model import GPRotModel, GPRotModel2
+
+def trace_plot(star, directory='mcmc_chains', kepler=False, thin=10):
+    filename = os.path.join(directory, '{}.h5'.format(star))
+    b = HDFBackend(filename)
+    coords = b.get_coords()
+    ndim = coords.shape[-1]
+    fig, axes = plt.subplots(ndim, 1, sharex=True, figsize=(8,8))
+    for i,ax in enumerate(axes):
+        ax.plot(coords[::thin, :, i], lw=1, alpha=0.2);
+        if i==ndim - 1:
+            from .aigrain import AigrainTruths
+            truth = AigrainTruths().df
+            ax.axhline(np.log(truth.ix[star, 'PEQ']), lw=1, color='r')        
+        ax.set_ylabel(GPRotModel.param_names[i])
+            
+    axes[0].set_title(star)
+    return fig    
+

### 2016-12-06 gprotation/GProt.py
getting rid of plotting convergence stuff - not compatible with emcee3

@@ -7,7 +7,6 @@ import numpy as np
 # from GProtation import make_plot, lnprob, Glnprob, Glnprob_split, \
 #         lnlike_split, Glnprior, MyModel
 from GProtation import make_plot, MyModel
-import Kepler_ACF
 import simple_acf as sa
 import h5py
 from gatspy.periodic import LombScargle
@@ -100,22 +99,23 @@ def mcmc_fit(x, y, yerr, p_init, p_max, id, RESULTS_DIR, truths, burnin=500,
                                      n_independent)
         mean_ind.append(ind_samp)
         mean_diff.append(diff)
-        print(conv)
+        print("Converged?", conv)
         if conv:
             break
 
-    col = "b"
-    if conv:
-        col = "r"
-    plt.clf()
-    plt.plot(autocorr_times, color=col)
-    plt.savefig(os.path.join(RESULTS_DIR, "{0}_acorr".format(id)))
-    plt.clf()
-    plt.plot(mean_ind, color=col)
-    plt.savefig(os.path.join(RESULTS_DIR, "{0}_ind".format(id)))
-    plt.clf()
-    plt.plot(mean_diff, color=col)
-    plt.savefig(os.path.join(RESULTS_DIR, "{0}_diff".format(id)))
+#     col = "b"
+#     if conv:
+#         col = "r"
+#     if autocorr_times:
+#         plt.clf()
+#         plt.plot(autocorr_times, color=col)
+#         plt.savefig(os.path.join(RESULTS_DIR, "{0}_acorr".format(id)))
+#         plt.clf()
+#         plt.plot(mean_ind, color=col)
+#         plt.savefig(os.path.join(RESULTS_DIR, "{0}_ind".format(id)))
+#         plt.clf()
+#         plt.plot(mean_diff, color=col)
+#         plt.savefig(os.path.join(RESULTS_DIR, "{0}_diff".format(id)))
     return
 
 
@@ -154,7 +154,7 @@ def evaluate_convergence(flat, mean_acorr, diff_threshold=0,
     try:
         acorr_t = emcee3.autocorr.integrated_time(flat, c=1)
     except emcee3.autocorr.AutocorrError:
-        return None, None, None, None
+        return converged, [], [], []
     mean_acorr.append(np.mean(acorr_t))
     mean_ind = len(flat) / np.mean(acorr_t)
     mean_diff = None

### 2016-12-06 gprotation/GProt.py
adding total timer

@@ -66,6 +66,7 @@ def mcmc_fit(x, y, yerr, p_init, p_max, id, RESULTS_DIR, truths, burnin=500,
     sampler = emcee3.Sampler(moves)
 
     print("burning in...")
+    total_start = time.time()
     ensemble = sampler.run(ensemble, burnin)
 
     flat = sampler.get_coords(flat=True)
@@ -102,6 +103,11 @@ def mcmc_fit(x, y, yerr, p_init, p_max, id, RESULTS_DIR, truths, burnin=500,
         print("Converged?", conv)
         if conv:
             break
+
+    total_end = time.time()
+    total_time = total_end - total_start
+    print("Total time taken = ", total_time/60., "minutes", total_time/3600.,
+          "hours")
 
 #     col = "b"
 #     if conv:

### 2016-12-06 gprotation/comparison_plots.py
adding labels to yaxes

@@ -142,7 +142,7 @@ def mcmc_plots(truths, DIR):
     plt.xlim(0, 4)
     plt.ylim(0, 6)
     plt.xlabel("$\ln(\mathrm{Injected~Period})$")
-    plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
+    plt.ylabel("$\ln(\mathrm{Recovered~Period,~GP~Method})$")
 
     plt.errorbar(np.log(true), np.log(maxlike), yerr=[lnerrp, lnerrm],
                  fmt="k.", zorder=1, capsize=0, ecolor=".8", alpha=.5, ms=.1)
@@ -224,7 +224,7 @@ def acf_plot(truths, DIR):
     plt.xlim(0, 4)
     plt.ylim(0, 6)
     plt.xlabel("$\ln(\mathrm{Injected~Period})$")
-    plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
+    plt.ylabel("$\ln(\mathrm{Recovered~Period~ACF~Method})$")
     plt.savefig(os.path.join(DIR, "compare_acf.pdf"))
     return (np.median((acfs - true)**2))**.5
 
@@ -257,7 +257,7 @@ def pgram_plot(truths, DIR):
     plt.ylim(0, 6)
     plt.xlabel("$\mathrm{Injected~Period~(Days)}$")
     plt.xlabel("$\ln(\mathrm{Injected~Period})$")
-    plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
+    plt.ylabel("$\ln(\mathrm{Recovered~Period~LS~Periodogram~method})$")
     plt.savefig(os.path.join(DIR, "compare_pgram.pdf"))
     return (np.median((pgram - true)**2))**.5
 

### 2016-12-06 gprotation/gprot_fit.py
adding default id value and some comments

@@ -13,7 +13,7 @@ from gatspy.periodic import LombScargle
 
 class fit(object):
 
-    def __init__(self, x, y, yerr, id, RESULTS_DIR="results", lc_size=200,
+    def __init__(self, x, y, yerr, id=0, RESULTS_DIR="results", lc_size=200,
                  ppd=4):
 
         self.x = x
@@ -67,6 +67,20 @@ class fit(object):
         return self.x[~m], self.y[~m], self.yerr[~m]
 
     def make_gaps(self, x, y, yerr, points_per_day):
+        """
+        Randomly subsample the light curve.
+
+        param x: (array)
+            The time array.
+        param y: (array)
+            The flux array.
+        param yerr: (array)
+            The flux uncertainty array.
+        param points_per_day: (int)
+            The mean number of data points to keep per day.
+
+        Returns subsampled x, y and yerr arrays.
+        """
         nkeep = points_per_day * (x[-1] - x[0])
         m = np.zeros(len(x), dtype=bool)
         l = np.random.choice(np.arange(len(x)), nkeep)

### 2016-12-06 gprotation/noise_run.py
adding noisy_sinusoid

@@ -13,11 +13,22 @@ def noise_fit(id, RESULTS_DIR):
     y = np.random.randn(len(x)) * 1e-5
     yerr = np.ones_like(y) * 1e-5
 
-    fit = gp.fit(x, y, yerr, id, RESULTS_DIR, lc_size=201)
+    fit = gp.fit(x, y, yerr, id, RESULTS_DIR)
     fit.gp_fit(burnin=1000, nwalkers=16, nruns=5, full_run=1000, nsets=1)
+
+def noisy_sinusoid(period, amp, id, RESULTS_DIR):
+
+    gap_days = .02043365
+    x = np.arange(0, 200, gap_days)
+    y = amp*np.sin(2*np.pi*(1./period)*x)
+    y += np.random.randn(len(x)) * 1e-5
+    yerr = np.ones_like(y) * 1e-5
+
+    fit = gp.fit(x, y, yerr, id, RESULTS_DIR)
+    fit.gp_fit(burnin=3000, nwalkers=16, nruns=5, full_run=1000, nsets=1)
 
 
 if __name__ == "__main__":
-    id = 0
     RESULTS_DIR = "noise"
-    noise_fit(id, RESULTS_DIR)
+#     noise_fit(0, RESULTS_DIR)
+    noisy_sinusoid(10, 1e-3, 1, RESULTS_DIR)

### 2016-12-06 documents/ms.tex
lots of changes

@@
 \newcommand{\princeton}{4}
-\newcommand{\nlightcurves}{333}
+\newcommand{\naigrain}{333}
+\newcommand{\nmcquillan}{100}
 \newcommand{\Kepler}{{\it Kepler}}
@@
 \newcommand{\numax}{$\nu_{\mathrm{max}}$}
+\newcommand{\acfRMS}{1.45}
+\newcommand{\pgramRMS}{1.10}
+\newcommand{\mcmcRMS}{0.87}
 
@@
 
-\altaffiltext{\columbia}{Simons Fellow, Columbia University, NY, NY,
-ruthangus@gmail.com}
+\altaffiltext{\columbia}{Simons Fellow, Department of Astronomy, Columbia
+University, NY, NY, RuthAngus@gmail.com}
 \altaffiltext{\oxford}{Subdepartment of Astrophysics, University of Oxford,
-OX1 3RH, UK}
+UK}
 \altaffiltext{\uw}{Department of Astronomy, University of Washington, Seattle,
 WA}
-\altaffiltext{\princeton}{Department of Astronomy, Princeton University}
+\altaffiltext{\princeton}{Department of Astronomy, Princeton University,
+Princeton, NJ}
 
@@
 function, we avoid the necessity of choosing a `best fitting' functional form,
-whilst sampling directly from the posterior probability distribution function
+whilst sampling directly from the posterior Probability Density Function (PDF)
 of the periodic parameter and marginalising over the other kernel
 hyperparameters.
-We used \nlightcurves\ simulated light curves with a range of rotation periods
+We used \naigrain\ simulated light curves with a range of rotation periods
 to test the GP model.
 We attempted to recover the rotation periods using three methods: our GP
-method, a sine-fitting periodogram method and an autocorrelation function
-method.
-The posterior probability distribution of the rotation period parameter was
-sampled using the affine invariant ensemble Markov Chain Monte Carlo (MCMC)
-sampler {\tt emcee} \citep{Foreman-Mackey2013}, and the GP operations were
-performed using the {\tt george} python package \citep{George}.
+method, a sine-fitting periodogram method and an AutoCorrelation Function
+(ACF) method.
+The posterior PDF of the rotation period parameter was sampled using the
+affine invariant ensemble Markov Chain Monte Carlo (MCMC) sampler {\tt emcee}
+\citep{Foreman-Mackey2013}, and the GP operations were performed using the
+{\tt george} python package \citep{George}.
 Rotation periods inferred via this method are more precise and accurate than
 the periodogram and ACF methods.
-Furthermore, the improvement is expected to be even more dramatic when applied
-to real, noisy {\it Kepler} light curves, since the GP method is well suited
-to modelling rotation signals and correlated noise simultaneously.
+Furthermore, rotation periods are inferred probabilistically, allowing them to
+be easily incorporated into follow-on studies, inferring stellar ages or
+characterising star-planet interactions for example.
+In addition, unlike the sine-fitting periodogram and ACF methods, the
+posterior PDF samples allow for proper characterisation of uncertainties.
+This method eliminates the need to define a heuristic detection criterion such
+as are often used in the literature: a light curve that does not contain a
+rotation signal will return the prior PDF.
+
+% Furthermore, the improvement is expected to be even more dramatic when applied
+% to real, noisy {\it Kepler} light curves, since the GP method is well suited
+% to modelling rotation signals and correlated noise simultaneously.
 
@@
 a set of simulated light curves.
-We used \nlightcurves\ light curves generated for the \citet{Aigrain2015}
+We used \naigrain\ light curves generated for the \citet{Aigrain2015}
 `hare and hounds' rotation period recovery experiment.
@@
 % those light curves simulated {\it without} differential rotation, of which
-% there are \nlightcurves.
+% there are \naigraincurves.
 For the purpose of demonstrating that our method can recover rotation periods,
-we selected only the \nlightcurves\ out of the 1000 without differential
+we selected only the \naigrain\ out of the 1000 without differential
 rotation.
@@
 \caption{Ranges and distributions of parameters used to simulate light curves
-in \citet{aigrain2015}}
+in \citet{Aigrain2015}}
 \begin{tabular}{lcc}
@@
 
-We attempted to recover the rotation periods of these \nlightcurves\ light
+We attempted to recover the rotation periods of these \naigrain\ light
 curves using three rotation period recovery methods: the ACF method, the LS
@@
 
-We applied the ACF method to the sample of \nlightcurves\ simulated light
+We applied the ACF method to the sample of \naigrain\ simulated light
 curves.
@@
 The $2\times$ and $1/2\times$ harmonic lines are shown as dashed lines.
+The RMS scatter about the $x=y$ line for rotation periods measured using the
+ACF method is \acfRMS.
 The agreement between the injected and recovered rotation periods is good in
@@
 In addition, some rotation periods are drastically over- and underestimated.
+Of the points clustered close to the $x=y$ line, more slightly below it than
+above it, \ie\ rotation periods are preferentially underestimated.
+This is a feature of the peak position measurement method that is performed on
+the ACF.
+ACFs of stellar light curves are, like the QP kernel function in equation
+\ref{eq:QP}, periodic functions added to decaying exponentials.
+In such functions the peak positions can be shifted towards the left, the
+short period end, because the decaying exponential raises the left side of
+each peak more than the right.
+It is possible to model this effect, however the standard practice is to
+simply measure the peak position without taking it into account.
+We adopt the standard practice approach here in order to provide a comparison
+between our new method and the current standard procedure.
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_acf.png}
+\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
 \caption[ACF results.]
-{The `true' rotation periods used to generate \nlightcurves\ simulated light
+{The `true' rotation periods used to generate \naigrain\ simulated light
 curves vs the rotation periods measured using the ACF technique.}
@@
 \citep{Lomb1976, Scargle1982} was computed over a grid of 10,000 periods,
-evenly spaced in frequency, between 1 and 100 days.
+evenly spaced in frequency, between 1 and 100 days\footnote{LS periodograms
+were calculated using the gatspy Python module:
+\url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
 The period of the highest peak in the periodogram was adopted as the rotation
@@
 results: they do not systematically over or under predict rotation period.
-They are however less precise than the ACF results---the 8.03 day RMS of
-residuals is five times larger.
+The RMS scatter for rotation periods measured using the LS periodogram
+technique is \pgramRMS.
 The longer-period measurements provide the dominant contribution to this RMS.
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_pgram.png}
+\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
 \caption[LS periodogram results.]
-{The `true' rotation periods used to generate \nlightcurves\ simulated light
+{The `true' rotation periods used to generate \naigrain\ simulated light
 curves vs the rotation periods measured using a LS periodogram technique.}
@@
     \hline
-    $A$ & -13 & 2.7 \\
-    $l$ & 6.2 & 1.5 \\
-    $\Gamma$ & -1.4 & 1.5 \\
-    Rotation period, $P_{rot}$ & ACF period & ACF period $\times 2$ \\
-    $\sigma$ & -17 & 5 \\
+    $\ln(A)$ & -13 & 2.7 \\
+    $\ln(l)$ & 6.2 & 1.5 \\
+    $\ln(\Gamma)$ & -1.4 & 1.5 \\
+    $\ln$(Rotation period, $P_{rot}$) & $\ln$(ACF period) &
+                2 $\times \ln$(ACF period) \\
+    $\ln\sigma$ & -17 & 5 \\
 \end{tabular}
@@
 distributions during MCMC posterior exploration.
-
 In addition, we set the following bounds on the rotation period and the $l$
@@
 We also limit the rotation periods to be less than half of the total timespan
-of the data since this is the longest period we expect to reliable recover.
+of the data since this is the longest period we expect to reliably recover.
 We prevent $\ln(l)$ from falling beneath zero both because we expect that
@@
 timescale of interest for this project, whilst reducing computational costs.
+In order to reduce computation time the light curves were split into small
+sections and the likelihood for each section calculated individually, then
+summed.
+The matrix operations scale as $N\ln(N)^2$ where $N$ is the number of data
+points in the light curve, however the likelihood calculation scales linearly
+with the number of light curve sections used.
+By adding together likelihoods of individial sections, the likelihood
+evaluation scales as $mn\ln(n)^2$, where $n$ is the number of data points per
+section and $m$ is the number of sections.
+This method therefore significantly reduces computation time.
+We split the simulated light curves into sections, 200 days long.
 The {\tt george} \citep{George} python package was used to implement the GP
@@
 the number of data points.
-We used {\tt emcee} \citep{Foreman-Mackey2013}, an affine invariant ensemble
-sampler to explore the posterior PDFs of the model parameters.
+We used {\tt emcee} \citep{Foreman-Mackey2013}, an affine invariant
+\citep{Goodman2010} ensemble sampler to explore the posterior PDFs of the
+model parameters.
 The rotation period was taken as the sample with the maximum likelihood value.
@@
 
-The marginal posterior distributions of the QP kernel hyperparameters for an
-example noise-free light curve  with a 14.4 day rotation period, are shown in
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
+\caption{The `true' rotation periods used to generate \naigrain\
+simulated light curves vs the rotation periods measured using the ACF
+technique.}
+\label{fig:compare_mcmc}
+\end{center}
+\end{figure*}
+
+The marginal posterior distributions of the QP kernel hyperparameters for the
+example simulated light curve with in figure \ref{fig:demo_lc}, are shown in
 figure \ref{fig:gp_posteriors}.
-The light curve itself with the best fit GP model is shown in figure
-\ref{fig:demo_lc_GP}.
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/demo_lc_GP.pdf}
-\caption{A simulated light curve with a rotation period of 14.4 days.
-The pink curve shows the GP model with the best-fit parameters.
-The marginal posteriors of the GP hyper-parameters used to produce this fit
-are shown in figure \ref{fig:gp_posteriors}.}
-\label{fig:demo_lc_GP}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/demo_triangle.pdf}
-\caption{Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
+\includegraphics[width=6in, clip=true]{figures/0002_triangle.png}
+\caption{(Place holder - to be updated.
+Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
 additional white noise term added to the diagonal elements of the covariance
 matrix. It is the fraction by which the observational uncertainties have been
-underestimated. If the errorbars reported on the data are too small, this
-parameter will be non-zero. Since this light curve was simulated without
-noise, this parameter is very close to zero.
-The true rotation period of this star is 14.4 days.}
+underestimated. If the uncertainties reported on the data are too small, this
+parameter will be non-zero.
+The true rotation period of this star is ... ($\exp(...)$) days.}
 \label{fig:gp_posteriors}
@@
 
-The noise-free light curves were injected into real \kepler\ light curves in
-order to test the performance of the GP method in a realistic case.
-Ideally, simulated light curves would have been injected into the \kepler\
-pixel-level data, then the same detrending method that is applied to the
-\kepler\ light curves would be applied to these light curves before attempting
-to measure rotation periods for these stars.
-However, since this work is designed to be more of a demonstration of the
-overall efficacy of the GP periodogram, rather than the development of a
-\Kepler\ specific method, this is beyond the scope of our demonstrations.
-
-Unfortunately, automating the GP method so that it can run on large numbers of
-real \kepler\ light curves is difficult.
-The added noise reduces the acceptance fraction within the MCMC chains,
-significantly increasing the required time for convergence.
-Despite this, the method works well on individual targets where chains can be
-run for a long time and subsampling can be reduced.
-We used a simulated, noisy light curve from \citet{Aigrain2015} that was
-generated by injecting a noise-free simulation into a real \kepler\ light
-curve in order to preserve realistic \kepler\ noise properties.
-
-\kepler\ light curves are big data and extracting the maximum amount of
-information from them requires either large numbers of CPU hours, or using
-work-arounds.
-One feature of \kepler\ light curves that works to our advantage is that they
+We applied our method to \nmcquillan\ stars with rotation periods previously
+measured by \citet{Mcquillan2014}.
+A feature of \kepler\ light curves that works to our advantage is that they
 are naturally broken up into smaller time units by quarterly (three month)
 breaks.
-Splitting the data set into quarters, rather than modelling the entire time
-series contiguously speeds up the computation time as inverting several
-small matrices is faster than inverting one large one.
+As explained above, splitting the data set into quarters, rather than
+modelling the entire time series contiguously speeds up the computation time
+as inverting several small matrices is faster than inverting one large one.
 The \kepler\ quarter divisions are natural places to split the time series
@@
 use global parameters only.
-Unfortunately, the application of this method to noisy test cases is still
-under development and cannot provide results here.
-We hope to develop this method further and apply it to real light curves in
-the near future.
 
-% The marginal posterior distributions of the QP kernel hyperparameters for an
-% example noisy light curve  with a ... day rotation period, are shown in
-% figure \ref{fig:gp_posteriors}.
-% The light curve itself with the best fit GP model is shown in figure
-% \ref{fig:demo_lc_GP}.
-
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/demo_lc_GP.pdf}
-% \caption{A simulated light curve with a rotation period of 14.4 days.
-% The pink curve shows the GP model with the best-fit parameters.
-% The marginal posteriors of the GP hyper-parameters used to produce this fit
-% are shown in figure \ref{fig:gp_posteriors}.}
-% \label{fig:demo_lc_GP}
-% \end{center}
-% \end{figure*}
-
-% \begin{figure*}
-% \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/demo_triangle.pdf}
-% \caption{Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
-% additional white noise term added to the diagonal elements of the covariance
-% matrix. It is the fraction by which the observational uncertainties have been
-% underestimated. If the errorbars reported on the data are too small, this
-% parameter will be non-zero. Since this light curve was simulated without
-% noise, this parameter is very close to zero.
-% The true rotation period of this star is 14.4 days.}
-% \label{fig:gp_posteriors}
-% \end{center}
-% \end{figure*}
+Our results show good agreement with the \citet{Mcquillan2014} measurements,
+as shown in ... %\ref{fig:mcquillan}.
 
@@
 which compromises accuracy.
+These shortcuts may only be necessary when working with a large number of
+stars (of order hundreds or thousands).
+If only interested in single or small numbers of targets it may be possible to
+reduce the extremity of these shortcuts or to avoid them altogether.
 The shortcuts that we employ here include:
@@
 independent of the ACF period unless MCMC chains are run for an infinitely
-long time.} \item{Subsampling.
+long time.}
+\item{Subsampling.
 Not only is some information always lost by subsampling, but the choice of
 subsampling frequency will also effect the results.
-We chose to retain only 20 data points per period --- based on the ACF
-period --- which, again means that our results are not independent of the ACF
-results.}
+We choose to retain an average of only 4 data points per day, based on the
+range of rotation periods we are interested in.}
 \item{Priors.
-We chose to use uniform, bounded priors to reduce parameter space and
-therefore convergence time.
-Unfortunately, if the true period lies outside the bounded region, it will
-never be found.}
+We choose to place a Gaussian prior in log-space over the rotation period
+parameter where the standard deviation is twice the log of the ACF rotation
+period.
+Unfortunately however, if the true period is far different to the ACF period,
+the MCMC chains will take a long time to reach the correct value.
+In practice we found that the introduction of a Gaussian prior improved
+results compared with a broad prior that was uniform in log-space.}
 \end{itemize}
-
-These shortcuts are only necessary when working with a large number of stars
-(of order hundreds or thousands).
-If only interested in single or small numbers of targets it may be possible to
-reduce the extremity of these shortcuts or to avoid them altogether.
-
-\subsection{The ACF method}
-The main advantage of the ACF method over the GP method is speed: it is
-extremely fast and therefore useful to get a quick estimate of a period.
-We have demonstrated that the ACF method may produce rotation periods that are
-slightly systematically biased towards lower rotation periods.
-This effect comes from the process of extracting the period from the ACF
-itself: typically superimposed onto a decaying exponential, that first peak in
-the ACF gets shifted towards faster periods.
-One may be able to get around this by modelling the ACF as a sum of a cosine
-and exponential function (a similar function to GP kernel function) but
-in practice ACFs tend to have a more complicated structure and it is not
-possible to model them with a simple function.
-This observation leads to the idea of modelling the ACF with a Gaussian
-processes, which then logically progresses to a using a Gaussian process with
-a Gaussian process kernel function.
-Although this may be an attractive idea, in practice a GP would not
-necessarily produce a positive semi-definite covariance kernel\footnote{One
-that can be inverted; a necessary regression operation.}.
-An alternative approach may be to correct for rotation period bias after the
-peak position has been measured by inferring a correction factor.
-Yet another approach could be to measure the mean separation between adjacent
-peaks in the ACF: the first peak should be most strongly affected by the
-exponential decay, the second peak less so, and so on.
 
@@
 as possible, or, alternatively to use a sampler that is designed to move
-around the parameter space much more quickly than {\tt emcee}; nested sampling
-for example.
+around the parameter space much more quickly than {\tt emcee}.
+% We implemented a version of our code using nested sampling but the results
+% were not as good as the {\tt emcee} method.
+% We believe this to be due to the initialisation --- there is no initialisation
+% in nested sampling, samples are drawn directly from the prior.
+% In a problem with a multimodal posterior PDF such as inferring periods,
+% initialising MCMC chains close to the maximum likelihood value may lead to a
+% more rapid convergence.
 We chose to use the ACF periods rather than the periodogram periods to
-initialise as, although there was some systematic bias present in these
-results, that bias was small and there were fewer large outliers.
+initialise the results since there were fewer large outliers.
 
@@
 the true rotation periods from the simulated light curves in this work.
-However, it is not clear whether the QP kernel function is the {\it best}
-function to use.
+However, it may not be the {\it best} function to use.
+It is, after all, an approximation to the shape of the covariance matrix of a
+real stellar light curve.
 There may be an alternative function which is better suited to capturing
@@
 
-\subsection{Future work}
-The next stages of improving this method are listed as follows:
+Not all \kepler\ light curves show evidence of stellar rotation.
+In some cases perhaps the star has few or no active regions, it is rotating
+pole-on, or it rotates so slowly that the \kepler\ data detrending pipeline
+removes any signal.
+In other cases there may be another source of variability present in the light
+curve, generating a false period detection.
+These sources may be physical: \eg\ binary star interactions, intra-pixel
+contamination from other astrophysical objects, pulsating variable stars,
+asteroseismic oscillations in giants and even stellar activity cycles.
+For some of these there is little we can do save attempt to eliminate these
+astrophysical false positives via alternative methods, \eg\ apply colour cuts
+to avoid giant contamination.
+For some however, like the variable stars, they may have distinctive
+hyperparameters that identify them, for example long coherence timescales.
+Testing this is beyond the scope of this paper but may be an interesting
+follow-up study.
+As well as astrophysical contamination, there may also be instrumental sources
+of contaminating variability: \eg\ temperature variations or pointing shifts
+of the \kepler\ spacecraft.
+These are unlikely to be periodic and, again, may produce unusual combinations
+of hyperparameters.
+This is something that we hope to test in future.
+In addition to this, there are several other aspects of the GP method we hope
+to test in future, listed as follows:.
 \begin{itemize}
-\item{To optimise the tradeoff between computational efficiency and accuracy.
-We have performed limited tests to explore the optimum subsampling strategy
-and prior bounds.
-These should be explored more thoroughly in future.
-For example, it would be useful to know how the precision and accuracy of the
-recovered rotation period varies as a function of subsampling frequency}
 \item{To perform model selection with different kernel functions. Again, these
@@
 be to use a sum of two kernel functions: one periodic and one non-periodic,
-each with its own amplitude parameter.
-If there is periodicity in the data, the amplitude of the periodic kernel will
-be larger than the amplitude of the non-periodic kernel and vice versa.}
+each with its own amplitude parameter.}
+\item{To attempt to detect differential rotation.
+We did not test our code on the light curves simulated with differential
+rotation in \citet{Aigrain2015} since we were only interested in recovering
+the most precise measurements of rotation period possible.
+In future we intend to investigate the possibility of recovering
+differential rotation by searching for close double peaks in the posterior
+PDFs of stars' rotation periods.}
 \item{To build in a noise model for \kepler\ data.
@@
 
-We simulated \nlightcurves\ noise-free, \kepler-like light curves using a spot
-model and attempted to recover the rotation periods used to generate them.
-Three methods were compared: a LS periodogram method, an ACF method and our GP
-method.
+We used the \naigrain\ simulated \kepler-like light curves for solid-body
+rotation from \citet{Aigrain2015} and attempted to recover the rotation
+periods used to generate them.
+Three methods were compared: a Lomb-Scargle periodogram method, an
+autocorrelation function method and a new Gaussian process method.
 The GP method produced the most precise and accurate rotation periods of the
@@
 This is because a GP is a semi-parametric model that is well suited to signals
-that are non-sinusoidal and quasi-periodic, because it does not require the
-assumption of regular time sampling and because a full posterior PDF of the
-rotation period parameter is produced.
-Not only this, unlike the other two methods, the GP method provides accurate
-rotation period uncertainties.
+that are non-sinusoidal and quasi-periodic.
+Not only this, unlike the other two methods, the GP method provides posterior
+PDF samples of rotation periods, allowing characterisation of uncertainties.
+Note that the posterior PDFs are often non-Gaussian, being bi- or multimodal.
+In these cases it is therefore misleading to report a 1$\sigma$ or 16th and
+84th percentiles as uncertainties.
+Rather, it is better to incorporate posterior PDF samples into analyses which
+require uncertainty propagation.
 
-The main drawback of this method is the computation time.
-We demonstrated that the GP method is capable of recovering signals from
-noise-free light curves, but were unable to automate the GP method for light
-curves with realistic noise properties.
-It may be that this method must remain a `boutique' method: ideal for small
-numbers of targets but not well suited to the entire \kepler\ catalogue.
-We continue to develop this method and remain hopeful that we can, eventually
-infer a probabilistic rotation period for thousands of \kepler\ stars.
+Because rotation periods are inferred probabilistically with the GP method,
+this allows them to be easily incorporated into follow-on studies; inferring
+stellar ages or characterising star-planet interactions for example.
+This the GP method also eliminates the need to define a heuristic detection
+criterion such as are often used in the literature: a light curve that does
+not contain a rotation signal will return the prior PDF.
+
+The main drawback of the GP method is the computation time.
 
 % acknowledgements
+This research was funded by the Simons Foundation and the Leverhulme Trust.
+Some of the data presented in this paper were obtained from the Mikulski
+Archive for Space Telescopes (MAST).
+STScI is operated by the Association of Universities for Research in
+Astronomy, Inc., under NASA contract NAS5-26555.
+Support for MAST for non-HST data is provided by the NASA Office of Space
+Science via grant NNX09AF08G and by other grants and contracts.
+This paper includes data collected by the Kepler mission. Funding for the
+Kepler mission is provided by the NASA Science Mission directorate.
 

### 2016-12-06 code/paper_plots.py

@@ -2,7 +2,6 @@ import numpy as np
 import matplotlib.pyplot as plt
 import os
 import pandas as pd
-import simple_acf as sa
 from acf import Kepler_ACF
 import simple_acf as sa
 
@@ -27,7 +26,7 @@ def acf_demo(x, y):
     yerr = np.ones_like(y) * 1e-5
     # period, err, lags, acf = Kepler_ACF.corr_run(x, y, yerr, 25)
     period, acf, lags, rvar = sa.simple_acf(x, y)
-
+    print("measured period = ", period)
     plt.clf()
     plt.plot(lags, acf, "k")
     plt.axvline(period, color="CornFlowerBlue")
@@ -54,6 +53,6 @@ if __name__ == "__main__":
     FIG_DIR = "/Users/ruthangus/projects/GProtation/documents/figures"
     x, y = np.genfromtxt(os.path.join(DATA_DIR,
                                       "final/lightcurve_0025.txt")).T
-    x, y = x[:10000], y[:10000]
+    x, y = x, y
     lc_demo(x, y)
     acf_demo(x, y)

### 2016-12-07 gprot/lc.py
added quality and tau to acf output

@@ -118,52 +118,58 @@ class LightCurve(object):
 
         return fig
 
-    def acf_prot(self, pmin=0.5, pmax=[10, 30, 100], delta=0.02, lookahead=30,
+    def acf_prot(self, pmin=0.5, pmax=100, delta=0.02, lookahead=30,
                  peak_to_trough=True, maxpeaks=1):
         """Returns best guess of prot from ACF, and height of peak
 
         Just pick first peak.
         """
-        if not hasattr(pmax, '__iter__'):
-            pmax = [pmax]
+        lags, ac = self.acf(pmin=pmin, pmax=pmax)
 
-        # Try different provided pmaxes; keep prot with best height.
-        best_prot = np.nan
-        best_height = -np.inf
-        for pm in pmax:
-            lags, ac = self.acf(pmin=pmin, pmax=pm)
+        # make sure lookahead isn't too long if pmax is small
+        lookahead = min(lookahead, pmax)
 
-            # make sure lookahead isn't too long if pmax is small
-            lookahead = min(lookahead, pm)
+        maxes, mins = peakdetect(ac, lags, delta=delta, lookahead=lookahead)
 
-            maxes, mins = peakdetect(ac, lags, delta=delta, lookahead=lookahead)
+        maxheight = -np.inf
+        pbest = np.nan
+        for i, ((xhi, yhi), (xlo, ylo)) in enumerate(zip(maxes, mins)):
+            if peak_to_trough:
+                # Calculate mean(peak-to-trough) height
+                h1 = yhi - ylo
+                try:
+                    h2 = yhi - mins[i+1][1]
+                    height = (h1+h2)/2.
+                except IndexError:
+                    height = h1
+            else:
+                height = yhi
+                print(i, height)
 
-            maxheight = -np.inf
-            pbest = np.nan
-            for i, ((xhi, yhi), (xlo, ylo)) in enumerate(zip(maxes, mins)):
-                if peak_to_trough:
-                    # Calculate mean(peak-to-trough) height
-                    h1 = yhi - ylo
-                    try:
-                        h2 = yhi - mins[i+1][1]
-                        height = (h1+h2)/2.
-                    except IndexError:
-                        height = h1
-                else:
-                    height = yhi
-                    print(i, height)
+            if height > maxheight:
+                pbest = xhi
+                maxheight = height
+            if i == maxpeaks-1:
+                break
 
-                if height > maxheight:
-                    pbest = xhi
[…]

### 2016-12-07 gprot/lc.py
corrected acf output

@@ -167,7 +167,7 @@ class LightCurve(object):
         fit = minimize(chisq, pbest)
 
         tau = fit.x[0]
-        quality = fit.fun / len(lags) / maxheight
+        quality =  1./ (fit.fun / len(lags) / maxheight) # bigger is better
 
         return pbest, maxheight, tau, quality
 

### 2016-12-07 documents/ms.tex

@@
 \newcommand{\nmcquillan}{100}
+\newcommand{\kepexample}{1430163}
+\newcommand{\kepexampleperiod}{4}
+\newcommand{\aigrainexampleperiod}{20.8}
 \newcommand{\Kepler}{{\it Kepler}}
@@
 Furthermore, rotation periods are inferred probabilistically, allowing them to
-be easily incorporated into follow-on studies, inferring stellar ages or
+be easily incorporated into follow-on studies; inferring stellar ages or
 characterising star-planet interactions for example.
@@
 Considering that many stars have on the order of hundreds of spots, the number
-of free parameters quickly becomes unwieldy, especially if the posterior
-Probability Density Functions (PDFs) of these parameters are explored with
-MCMC.
+of free parameters quickly becomes unwieldy, especially if the posterior PDFs
+of these parameters are explored with MCMC.
 Simplified spot models, such as the one described in \citet{Lanza2014} where
@@
 rotation period measurement.
-Many rotation periods in the literature have been inferred by measuring the
-position of the first peak in an ACF, however this approach can be dangerous.
-The exponential decay in correlation can shift the peak position short-wards
-of its true value, leading to an underestimate of the period.
-We return to this point in section \textsection \ref{sec:discussion}.
-% In order to avoid this, a period should be inferred by modelling the entire
-% ACF, not just measuring the position of the first peak.
+% Many rotation periods in the literature have been inferred by measuring the
+% position of the first peak in an ACF, however this approach can be dangerous.
+% The exponential decay in correlation can shift the peak position short-wards
+% of its true value, leading to an underestimate of the period.
+% We return to this point in section \textsection \ref{sec:discussion}.
 
@@
 Secondly, in order to infer {\it probabilistic} periods, i.e. to estimate the
-posterior PDF of the rotation period and thereby measure a realistic and
-representative uncertainty.
+posterior PDF of the rotation period and thereby produce a realistic and
+estimate of its uncertainty.
 Thirdly, to allow for an additional correlated noise model to be included
@@
 For example, take the time-series in figure \ref{fig:GP_example}.
-This is the \kepler\ light curve of KIC 1430163, a relatively active star that
+This is the \kepler\ light curve of KIC \kepexample, a relatively active star that
 rotates once every $\sim$ 4 days.
@@
 % \ref{fig:GP_example}, shown in orange.
-In figure \ref{fig:GP_example} the light curve of KIC 1430163 has been
+In figure \ref{fig:GP_example} the light curve of KIC \kepexample\ has been
 modelled with a GP.
@@
 For example, the simplest and most commonly used kernel function, the `Squared
-Exponential' (SE) produces an adequate fit to the KIC 1430163 light curve.
+Exponential' (SE) produces an adequate fit to the KIC \kepexample\ light
+curve.
 The SE kernel function is defined as,
@@
 % \citep[see \eg][]{Rajpaul2015}
-The SE kernel function may be an adequate model of the covariance in stellar
-light curves, but it is not a {\it useful} one for the problem of rotation
-period inference because it does not have a parameter that describes a period.
+The SE kernel function is not a good model of the covariance in stellar light
+curves, nor a {\it useful} one for the problem of rotation period inference
+because it does not capture periodic behaviour.
 In order to infer rotation periods it is necessary to use a periodic kernel
@@
 $P$ can be interpreted as the rotation period of the star and $\Gamma$
-controls the amplitude of the sinusoidal term.
+controls the amplitude of the $\sin^2$ term.
+If $\Gamma$ is very small, only points almost exactly one period away are
+tightly correlated and points that are slightly more or less than one period
+away are very loosely correlated.
+If $\Gamma$ is large, points separated by one period are tightly
+correlated and points separated by slightly more or less are still highly
+correlated although less so.
 This kernel function allows two data points that are separated in time by one
@@
 period can be weakly correlated.
-Both the QP kernel function was used to produce the model shown in figure
+We also use an extra parameter, $\sigma$ which is an additional white noise
+term added to the diagonal elements of the covariance matrix.
+It is the fraction by which the observational uncertainties have been
+underestimated --- if the uncertainties reported on the data are too small,
+this parameter will be non-zero.
+In practice however, this parameter should always be non-zero when performing
+GP regression.
+This is because the covariance matrix must be positive definite, however
+matrix inversion performed using most solvers is approximate, not exact,
+therefore slight deviations from positive definiteness can arise.
+Including a small amount of extra variance in the model allows enough
+flexibility that the matrix inversion algorithms do not run into numerical
+issues.
+
+This QP kernel function was used to produce the model shown in figure
 ~\ref{fig:GP_example}.
@@
 This approach comes at a cost: a GP model is expensive to compute once as it
-requires a matrix inversion and determinant evaluation, let alone the many
-thousands of times that is necessary to fully explore the posteriors of the
-parameters.
-However, fully exploring the posterior PDF of $P$ allows for an accurate
-estimate of the uncertainty on the rotation period.
+requires an inversion and determinant evaluation of the covariance matrix, let
+alone the many thousands of times that is necessary to fully explore the
+posteriors of the parameters.
+Fully exploring the posterior PDF of $P$ is important however, as it provides
+an accurate estimate of the uncertainty on the rotation period.
 
@@
 \caption[A light curve with a GP model.]
-{Light curve of KIC 1430163, an active star with a rotation period of $\sim$ 4
-days.
+{Light curve of KIC \kepexample, an active star with a rotation period of
+$\sim$ \kepexampleperiod\ days.
 The blue line shows a fit to the data using a Gaussian process model with a QP
@@
 
-\section{Rotation period injection and recovery}
+In \textsection \ref{sec:simulations} of this paper we describe the simulated
+light curves used to test rotation period inference methods.
+In \textsection \ref{sec:method} we describe the various methods used to
+recover the rotation periods: the periodogram, ACF and GP methods.
+In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data and
+the results of the paper are discussed in \textsection \ref{sec:discussion}.
+
+\section{Simulated light curves}
+\label{sec:simulations}
 
@@
 distribution between 1 and 10 days.
+A histogram of the \naigrain\ solid-body rotation periods is shown in figure
+\ref{fig:period_hist}.
 The light curves were also generated with a range of stellar inclination
@@
 of the time series.
-Figure \ref{fig:demo_lc} shows an example of a simulated, noise-free
-light curve with a period of 17.04 days.
+Figure \ref{fig:demo_lc} shows an example of a simulated light curve with a
+period of \aigrainexampleperiod days.
 
@@
 \includegraphics[width=6in, clip=true]{figures/period_hist.pdf}
-\caption{A histogram of the rotation periods used to generate the simulated
-light curves in \citet{Aigrain2015}.}
+\caption{A histogram of the rotation periods used to generate the \naigrain\
+simulated light curves in \citet{Aigrain2015}.}
 \label{fig:period_hist}
@@
 \caption[A simulated light curve.]
-{An example simulated light curve. This `star' has a rotation period of 20.8
-days.}
+{An example simulated light curve. This `star' has a rotation period of
+\aigrainexampleperiod\ days.}
 \label{fig:demo_lc}
@@
 
-We attempted to recover the rotation periods of these \naigrain\ light
-curves using three rotation period recovery methods: the ACF method, the LS
-periodogram method and the GP method.
+% We attempted to recover the rotation periods of these \naigrain\ light curves
+% using three methods: the ACF method; the LS periodogram method; and the GP
+% method, and compared their performances.
 
+\section{Rotation period recovery}
+\label{sec:method}
 \subsection{ACF}
@@
 \citet{Mcquillan2013}.
-In this method, an ACF is calculated for each light curve and smoothed by
-convolving with a Gaussian with $\sigma=9$ days.
+An ACF is calculated for each light curve and smoothed by convolving with a
+Gaussian with $\sigma=9$ days.
 A rotation period is estimated as the lag-time of the highest peak in the ACF,
@@
 there are two active regions at or near opposite longitudes on the surface of
-the star---this would produce a light curve with two dips per rotation period.
+the star---producing a light curve with two dips per rotation period.
 We restrict our search to periods less than 100 days because this is the
 maximum period that is realistically recoverable in \kepler\ light curves ---
-instrumental noise significantly distorts the signal on longer timescales.
+instrumental noise significantly distorts signals on longer timescales.
 An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
@@
 days) and the pink dashed line shows the period that was used to simulate the
-light curve (20.8 days).}
+light curve (\aigrainexampleperiod\ days).}
 \label{fig:demo_acf}
@@
 \ref{fig:compare_acf}.
-The $2\times$ and $1/2\times$ harmonic lines are shown as dashed lines.
+The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
 The RMS scatter about the $x=y$ line for rotation periods measured using the
@@
 In addition, some rotation periods are drastically over- and underestimated.
-Of the points clustered close to the $x=y$ line, more slightly below it than
-above it, \ie\ rotation periods are preferentially underestimated.
+Of the points clustered close to the $x=y$ line, more fall slightly below it
+than above it, \ie\ rotation periods are preferentially underestimated.
 This is a feature of the peak position measurement method that is performed on
 the ACF.
-ACFs of stellar light curves are, like the QP kernel function in equation
-\ref{eq:QP}, periodic functions added to decaying exponentials.
+ACFs of stellar light curves are similar to the QP kernel function in equation
+\ref{eq:QP}: periodic functions added to decaying exponentials.
 In such functions the peak positions can be shifted towards the left, the
@@
 We adopt the standard practice approach here in order to provide a comparison
-between our new method and the current standard procedure.
+between our new method and that used largely in the literature.
 
@@
 
-\subsection{The sine-fitting periodogram method}
+\subsection{LS periodogram}
 
-For each simulated light curve, a Lomb-Scargle (LS) periodogram
-\citep{Lomb1976, Scargle1982} was computed over a grid of 10,000 periods,
-evenly spaced in frequency, between 1 and 100 days\footnote{LS periodograms
-were calculated using the gatspy Python module:
+For each simulated light curve, a LS periodogram
+\footnote{LS periodograms were calculated using the gatspy Python module:
 \url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
+was computed over a grid of 10,000 periods, evenly spaced in frequency,
+between 1 and 100 days.
 The period of the highest peak in the periodogram was adopted as the rotation
@@
 periods in figure \ref{fig:pgram_compare}.
-These recovered rotation periods are in general more accurate than the ACF
-results: they do not systematically over or under predict rotation period.
 The RMS scatter for rotation periods measured using the LS periodogram
@@
 
-$\sigma$ is an additional white noise term added to the diagonal elements of
-the covariance matrix. It is the fraction by which the observational
-uncertainties have been underestimated.
-If the uncertainties reported on the data are too small, this parameter will
-be non-zero.
-In practice however, this parameter should always be non-zero when performing
-GP regression.
-This is because the covariance matrix must be positive definite, however
-matrix inversion performed using most solvers is approximate, not exact,
-therefore slight deviations from positive definiteness can arise.
-Including a small amount of extra variance in the model allows enough
-flexibility that the matrix inversion algorithms do not run into numerical
-issues.
-
 We subsampled the light curves in order to reduce computation time.
@@
 
-\subsection{Real \kepler\ data}
+\section{Real \kepler\ data}
+\label{sec:kepler}
 

### 2016-12-07 gprotation/kepler_run.py
running in series again

@@ -1,0 +1,59 @@
+# measure rotation periods for stars in the McQuillan catalogue
+from __future__ import print_function
+import numpy as np
+import subprocess
+import pandas as pd
+import os
+import gprot_fit as gp
+import kplr
+client = kplr.API()
+import kepler_data as kd
+import matplotlib.pyplot as plt
+from multiprocessing import Pool
+
+
+def run_gprot(stars):
+    for i, star in enumerate(stars):
+        print(int(star), i, "of", len(stars))
+        print("gprot-fit {0} --kepler -o -v".format(int(star)))
+        subprocess.call("gprot-fit {0} --kepler -o -v".format(int(star)),
+                        shell=True)
+
+
+def run_my_fit(i):
+    ruth1 = "/export/bbq1/angusr/GProtation/"
+    lc_path = "/home/angusr/.kplr/data/lightcurves/"
+    RESULTS_DIR = "/export/bbq1/angusr/GProtation/gprotation/kepler"
+
+    # McQuillan stars
+    s = pd.read_csv(os.path.join(ruth1, "code/data/Table_1_Periodic.txt"))
+    kid = s.KIC.values[i]
+
+    print(int(kid), i, "of", len(stars))
+    star = client.star(int(kid))
+    lcs = star.get_light_curves(fetch=True, short_cadence=False,
+                                clobber=False)
+    LC_DIR = os.path.join(lc_path, "{}".format(str(int(kid)).zfill(9)))
+    x, y, yerr = kd.load_kepler_data(LC_DIR)
+
+    emcee2 = False
+    if emcee2:
+        e2.gp_fit(x, y, yerr, "{}2".format(id), RESULTS_DIR)
+
+    fit = gp.fit(x, y, yerr, kid, RESULTS_DIR)
+    fit.gp_fit(burnin=1000, nwalkers=16, nruns=5, full_run=1000, nsets=2)
+#     fit.gp_fit(burnin=2, nwalkers=12, nruns=2, full_run=50, nsets=2)  # fast
+
+
+if __name__ == "__main__":
+    # NGC6891
+    ruth2 = "/export/bbq1/angusr/GProtation/"
+    stars = np.genfromtxt(os.path.join(ruth2,
+                          "code/data/NGC6819_members.txt")).T
+
+    for i in range(2):
+        run_my_fit(i)
+
+#     pool = Pool()
+#     pool.map(run_my_fit, range(2))
+#     pool.map(run_my_fit, range(100))

### 2016-12-07 gprotation/GProt.py
trying to print total time

@@ -110,8 +110,8 @@ def mcmc_fit(x, y, yerr, p_init, p_max, id, RESULTS_DIR, truths, burnin=500,
     print("Total time taken = ", total_time/60., "minutes", total_time/3600.,
           "hours")
 
-    with open(os.path.join(RESULTS_DIR, "{0}_time.txt".format(id))) as f:
-        f.write("{}".format(total_time))
+    with open(os.path.join(RESULTS_DIR, "{0}_time.txt".format(id)), "w") as f:
+        f.write(total_time)
 
 #     col = "b"
 #     if conv:

### 2016-12-07 gprotation/recover_suzannes.py
changing the run parameters

@@ -73,13 +73,13 @@ def gp_fit(x, y, yerr, sid, RESULTS_DIR):
 
     # fast settings
 #     burnin, nwalkers, nruns, full_run = 500, 12, 3, 100
-    burnin, nwalkers, nruns, full_run = 2, 12, 5, 50
+#     burnin, nwalkers, nruns, full_run = 2, 12, 5, 50
 #     xb[0], yb[0], yerrb[0] = xb[0][::100], yb[0][::100], yerrb[0][::100]
 
     trths = [None, None, None, None, None]
 #     trths = [None, None, None, None, np.log(truths.P_MIN.values[m][i])]
-#     mcmc_fit(xb[:2], yb[:2], yerrb[:2], p_init, p_max, sid, RESULTS_DIR,
-    mcmc_fit(xb[0], yb[0], yerrb[0], p_init, p_max, sid, RESULTS_DIR,
+#     mcmc_fit(xb[0], yb[0], yerrb[0], p_init, p_max, sid, RESULTS_DIR,
+    mcmc_fit(xb[:2], yb[:2], yerrb[:2], p_init, p_max, sid, RESULTS_DIR,
              truths=trths, burnin=burnin, nwalkers=nwalkers, nruns=nruns,
              full_run=full_run, diff_threshold=.5, n_independent=1000)
 

### 2016-12-07 documents/ms.tex
outline caveat, equation typo, anti-correlation description

@@
 
+At this point it is important to highlight the following caveat.
+Our method has been tuned to reproduce the rotation periods of the
+\citet{Aigrain2015} simulated light curves.
+These light curves were simulated using a very good but still imperfect model.
+The input parameters and their distributions were chosen to be as realistic as
+possible given current observations.
+Unfortunately however, there are limited data on some of these parameters for
+stars other than the Sun: for example the distrubution of spot lifetimes or
+activity cycle length.
+Ensuring that we can recover the rotation periods of these simulated light
+curves does not necessarily ensure that we can recover the rotation periods of
+real \kepler\ light curves with equal accuracy and precision.
+Even using priors on the QP kernel parameters that have been derived from the
+\citet{Aigrain2015} sample may be dangerous.
+This caveat should be taken into consideration when this method is applied to
+real data.
+
 \section{Discussion}
@@
 k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2}\right)
-\cos^2\left(\frac{2\pi}{P}\right)
+\cos\left(\frac{2\pi}{P}\right)
 \end{equation}
@@
 rotation period may not be well modelled.
+In other words, we do not want to force anti-correlation of points that are
+$\frac{1}{2}$ a period apart.
 It would be very worthwhile to test this assumption and this alternative

### 2016-12-20 documents/ms.tex
tdm rewrite of section 3.3

@@
 \usepackage{color}
+\usepackage{bm}
 
@@
 WA}
-\altaffiltext{\princeton}{Department of Astronomy, Princeton University,
+\altaffiltext{\princeton}{Department of Astrophysical Sciences, Princeton University,
 Princeton, NJ}
@@
 
-In order to recover the rotation periods of the simulated light curves using
-Gaussian processes, we sampled the posterior PDFs of the kernel
-hyperparameters described in equation \ref{eq:QP}.
-The likelihood function for a GP is similar to the simple Gaussian likelihood
-function that is used for optimisation problems where the uncertainties are
-Gaussian and uncorrelated.
-The latter can be written
+In order to recover the rotation period of a simulated light curve using
+Gaussian processes, we sample the following posterior PDF:
+\begin{equation}
+p({\bm \theta}\,|\,y) \propto \mathcal L(y\,|\,{\bm \theta}) p({\bm \theta}),
+\end{equation}
+\label{eq:posterior}
+where $y$ are the light curve data, $\bm \theta$ are the hyperparameters 
+of the kernel described in equation \ref{eq:QP}, $\mathcal L$ is the 
+GP likelihood function, and $p({\bm \theta})$ is the prior on the 
+hyperparameters.  The GP likelihood is similar to the simple Gaussian likelihood
+function used for optimisation problems where the uncertainties are
+Gaussian and uncorrelated. The latter can be written
 \begin{equation}
@@
 
-One could either maximise this likelihood function in order to find the
-best-fit hyper-parameters of the covariance kernel function or, as in our
-approach, sample from the posterior PDFs of the hyper-parameters.
-The advantage of the maximum-likelihood method is that the best-fit parameters
-will be found much faster, but the uncertainities on the rotation period
-will not be constrained.
-In order to measure accurate uncertainties, the posterior PDFs of the
-parameters must be explored.
-Since obtaining accurate uncertainties on rotation periods is one of our main
-motivations for this method, we use MCMC despite the added computational
-expense.
+Sampling the posterior described above presents several challenges:
+\begin{itemize}
+\item Evaluating $\mathcal L$ for an entire \Kepler\ lightcurve 
+($\sim$40,000 points) takes about $\sim$5\,s--- too computationally expensive 
+to perform inference on large numbers of light curves\footnote{All computational
+times cited in this section are based on evaluations on a 
+single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.  The matrix operations
+necessary to evaluate the GP likelihood (using the fast matrix solver 
+HODLR \citep{Ambikasaran2014}, implemented in the {\tt george} \citep{George} 
+python package) scale as $N\ln(N)^2$, where $N$ is
+the number of data points in the light curve.
 
-We use the ACF period as an initial guess for the rotation period (this
-decision is discussed further in section \textsection \ref{sec:discussion}).
-We use Gausssian distributions in the natural logarithm of all the parameters
-of our GP model.
-The means and variances of the priors used were learned during the development
-of this method.
-In early stages of development we used very broad priors on all the parameters
-uniform in log-space between -20 and 20.
-We then examined the distributions of the model parameters for `sucessful' and
-`unsuccessful' rotation period recoveries, where a recovery was termed
-successful if it lay within 10\% of the injected period.
-For the successful cases, the parameter distributions were narrow Gaussians.
-The parameter distributions of the unsuccessful cases were broad Gaussians
-with similar means.
-We adopted the means and standard deviations of the narrow Gaussians
-describing the parameter distributions of the successful cases to describe our
-Gaussian priors.
-The means and standard deviations of these $\ln$ Gaussian priors are given in
+\item The flexibility of this GP model allows for insidious kinds of 
+posterior multimodality and "over-fitting"--like behavior.  For example, 
+if $l$ is small,  the non-periodic factor in the covariance kernel may dominate, 
+allowing for a good fit to the data without requiring any 
+periodic covariance structure---even if
+clear periodic structure is present.  Additionally, if $\Gamma$ is too large,
+then [explain in more technical terms that "the GP gets way too wiggly and
+overfits the data without caring about the periodicity." ]
+
+\item In some cases, the posterior may also be multi-modal in period---that is, 
+there may be significant posterior density at half or twice the true period $P$.
+This is related to how the ACF method must guard against identifying periods 
+at half or twice the true period. 
+\end{itemize}
+
+We address the first of these difficulties in two distinct ways: subsampling
+the data and splitting the light curve into independent sections.  
+To subsample, we randomly select 1/30th of the points in the full light curve (an 
+average of $\sim$1.5 points per day).  This decreases the likelihood evaluation 
+time by a factor of about 50, down to about 100\,ms.
+We then split the light curve into equal-sized
+chunks containing approximately 300 points per section (corresponding to about
+200 days), and evaluate the log-likelihood as the sum of the log-likelihoods
+of the individual sections (all using the same parameters $\theta$).  This reduces
+computation time because the section-based likelihood evaluation scales as $mn\ln(n)^2$, where $n$ is the number of data points per section and $m$ is the number of sections.
+This method further reduces computation time for a typical light curve (subsampled
+by a factor of 30) by about a factor of two, to about 50\,ms.
+
+In order to manage the extreme flexibilty of the GP model and focus on 
+reliably retrieving the period parameter, we also impose priors
+on the non-period GP parameters that we 
+learned during the development of this method.  
+Beginning with very broad priors (uniform in log space between 
+-20 and 20) on all the parameters, we first explored what the 
+typical values of the non-period parameters tended to be when the 
+injected periods were successfully recovered.  We also experimented with
+constraining the allowed ranges of the parameters, after discovering 
+that some regions of parameter space (such as large values of $A$ and $\Gamma$
+and small values of $l$) tended to allow fits that ignored the desired
+periodicity.  The priors we used for this work are listed in
 table \ref{tab:priors}.  
+% We note that our default prior for period is 
+% log-uniform between the bounds, but that we were also able to 
+% further improve our results by using an ACF-inspired prior.
+
+To sample the posterior in a way that is sensitive to potential 
+multimodality, we use the \texttt{emcee3} MCMC sampler.
+We initialize 500 walkers with random samples from the prior 
+ and use a mix of different proposal distributions: 
+specifically, a 2:2:1 mix of KDE, differential evolution, and DESnooker moves. 
+[DFM: more words/citations here?]  
+We run fifty steps of the sampler at a time, checking for convergence after 
+each iteration, up to a maximum of fifty iterations.  We declare convergence if 
+the total effective chain length is at least 8$\times$ the maximum autocorrelation
+time.  When convergence is achieved, we drop the first two autocorrelation 
+lengths in the chain as a burn-in, and randomly choose 5000 samples as 
+representative of the posterior.  This fitting process takes several hours 
+for a typical simulated light curve.
+
+Figure \ref{fig:compare_mcmc} summarizes the results of this MCMC fitting procedure
+compared to the injected ``true'' stellar rotation periods.
 
@@
 \begin{center}
-\caption{Means and standard deviations of the Gaussian priors on the natural
-logarithms of the GP model parameters.}
+\caption{Priors and bounds on the natural logarithms of the GP model parameters.}
 \begin{tabular}{lcc}
-Parameter & Mean & Standard Deviation \\
+Parameter & Prior & Bounds\\
     \hline
-    $\ln(A)$ & -13 & 2.7 \\
-    $\ln(l)$ & 6.2 & 1.5 \\
-    $\ln(\Gamma)$ & -1.4 & 1.5 \\
-    $\ln$(Rotation period, $P_{rot}$) & $\ln$(ACF period) &
-                2 $\times \ln$(ACF period) \\
-    $\ln\sigma$ & -17 & 5 \\
+    $\ln A$ & $\mathcal N(-13, 5.7)$ & (-20, 0) \\
+    $\ln l$ & $\mathcal N(7.2, 1.2)$ & (2, 20) \\
+    $\ln \Gamma$ & $\mathcal N(-2.3, 1.4)$ & (-10, 3) \\
+    $\ln \sigma$ & $\mathcal N(-17, 5)$ & (-20, 0) \\
+    $\ln P $ & Uniform & ($\ln 0.5, \ln 100$) \\ 
 \end{tabular}
@@
 \label{tab:priors}
-The parameter positions were also initialised at the means of these
-distributions during MCMC posterior exploration.
-In addition, we set the following bounds on the rotation period and the $l$
-parameter.
-\begin{eqnarray}
- 	&	\ln(0.5) < \ln(P_{\mathrm{rot}}) < \ln(T/2) \\ \nonumber
-    &   0 < \ln(l) \\
-\end{eqnarray}
-\label{eq:priors}
-Where $T$ is the total timespan of the light curve.
-We limit rotation periods to be greater than half a day because we are chiefly
-interested in rotation of FGK stars which are likely to be greater than this
-lower limit, however this restriction could be relaxed in principle.
-We also limit the rotation periods to be less than half of the total timespan
-of the data since this is the longest period we expect to reliably recover.
-We prevent $\ln(l)$ from falling beneath zero both because we expect that
-astrophysical signals in light curves should have longer correlation
-timescales than this and in practice we found that if $l$ is unrestricted, it
-can occasionally wander towards a small value, resulting in an incorrect
-rotation period estimate.
-
-We subsampled the light curves in order to reduce computation time.
-We randomly subsampled the data, retaining an average of four data points per
-day.
-This allowed us to recover periods down to half a day --- the minimum
-timescale of interest for this project, whilst reducing computational costs.
-In order to reduce computation time the light curves were split into small
-sections and the likelihood for each section calculated individually, then
-summed.
-The matrix operations scale as $N\ln(N)^2$ where $N$ is the number of data
-points in the light curve, however the likelihood calculation scales linearly
-with the number of light curve sections used.
-By adding together likelihoods of individial sections, the likelihood
-evaluation scales as $mn\ln(n)^2$, where $n$ is the number of data points per
-section and $m$ is the number of sections.
-This method therefore significantly reduces computation time.
-We split the simulated light curves into sections, 200 days long.
-The {\tt george} \citep{George} python package was used to implement the GP
-model which uses the fast matrix solver, HODLR \citep{Ambikasaran2014}.
-Matrix operations performed with HODLR scale as $N \log ^2 (N)$, where $N$ is
-the number of data points.
-We used {\tt emcee} \citep{Foreman-Mackey2013}, an affine invariant
-\citep{Goodman2010} ensemble sampler to explore the posterior PDFs of the
-model parameters.
-The rotation period was taken as the sample with the maximum likelihood value.
-The resulting measured rotation periods are compared to the true rotation
-periods in figure \ref{fig:compare_mcmc}.
 

### 2016-12-20 gprot/model.py
took away restriction on l > p again (did not seem to do much, really)

@@ -127,8 +127,8 @@ class GPRotModel(object):
             return -np.inf
 
         # Don't let SE correlation length be shorter than P.
-        if theta[1] < theta[-1]:
-            return -np.inf
+        # if theta[1] < theta[-1]:
+        #     return -np.inf
 
         # if not (theta[1] > theta[4] and np.log(0.5) < theta[4]):
         #     return -np.inf

### 2016-12-20 documents/ms.tex

@@
 single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.  The matrix operations
-necessary to evaluate the GP likelihood (using the fast matrix solver 
+necessary to evaluate the GP likelihood\footnote{These operations use 
+the fast matrix solver 
 HODLR \citep{Ambikasaran2014}, implemented in the {\tt george} \citep{George} 
-python package) scale as $N\ln(N)^2$, where $N$ is
+python package.} scale as $N\ln(N)^2$, where $N$ is
 the number of data points in the light curve.
@@
 \item The flexibility of this GP model allows for insidious kinds of 
-posterior multimodality and "over-fitting"--like behavior.  For example, 
+posterior multimodality and ``over-fitting''--like behavior.  For example, 
 if $l$ is small,  the non-periodic factor in the covariance kernel may dominate, 
@@
 clear periodic structure is present.  Additionally, if $\Gamma$ is too large,
-then [explain in more technical terms that "the GP gets way too wiggly and
-overfits the data without caring about the periodicity." ]
+then [explain in more technical terms that ``the GP gets way too wiggly and
+overfits the data without caring about the periodicity.'' ]
 
@@
 
-We address the first of these difficulties in two distinct ways: subsampling
+We address the difficulty of computational speed in two distinct ways: subsampling
 the data and splitting the light curve into independent sections.  
@@
 periodicity.  The priors we used for this work are listed in
-table \ref{tab:priors}.  
+table \ref{tab:priors}.  We note that we do not have any particular
+quantitative justification for the use of these priors,
+except that they produce good results for this particular test set of 
+light curves.  We thus caution that other data sets might have different
+noise properties, potentially requiring modifications to these priors.
 % We note that our default prior for period is 
@@
 for a typical simulated light curve.
-
-Figure \ref{fig:compare_mcmc} summarizes the results of this MCMC fitting procedure
-compared to the injected ``true'' stellar rotation periods.
 
@@
 
+Figure \ref{fig:compare_mcmc} summarizes the results of this MCMC fitting procedure
+compared to the injected ``true'' stellar rotation periods. 
 The marginal posterior distributions of the QP kernel hyperparameters for the

### 2016-12-22 documents/ms.tex
adding a few words about wfirst and lsst

--- 2016-12-07 documents/ms.tex
@@
 \newcommand{\LSST}{{\it LSST}}
+\newcommand{\Wfirst}{{\it Wfirst}}
 \newcommand{\SDSS}{{\it SDSS}}
@@
 \newcommand{\numax}{$\nu_{\mathrm{max}}$}
-\newcommand{\acfRMS}{1.45}
-\newcommand{\pgramRMS}{1.10}
-\newcommand{\mcmcRMS}{0.87}
+\newcommand{\acfRMS}{9.27}
+\newcommand{\pgramRMS}{1.25}
+\newcommand{\mcmcRMS}{0.46}
 
@@
 rotation signal will return the prior PDF.
-
+Furthermore, the ACF method is not applicable to unevenly spaced data,
+therefore inappropriate for the ~$10^7$ stellar light curves expected to be
+produced by \LSST\ and \Wfirst.
 % Furthermore, the improvement is expected to be even more dramatic when applied
@@
 evenly-spaced\footnote{\citet{Edelson1988} describe a method for computing
-ACFs for unevenly-spaced data.} which is not the case with \Kepler\ light
-curves (although in many cases it can be approximated as uniformly sampled).
+ACFs for unevenly-spaced data.} which is not exactly the case with \Kepler\
+light curves (although in many cases it can be approximated as uniformly
+sampled) and not nearly the case for future large-scale survey instruments,
+\LSST\ and \Wfirst.
 An ACF is also an operation performed on the data, not a generative model of
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
+% \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_emcee3_13_12_2.png}
 \caption{The `true' rotation periods used to generate \naigrain\
-simulated light curves vs the rotation periods measured using the ACF
-technique.}
+simulated light curves vs the rotation periods measured using the GP
+technique.
+Since the posterior PDFs of rotation periods are often non-Gaussian,
+the points plotted here are the highest likelihood samples.
+The uncertainties are the 16th and 84th percentiles.}
 \label{fig:compare_mcmc}

### 2016-12-24 documents/ms.tex

--- 2016-12-20 documents/ms.tex
@@
 \usepackage{bm}
+\usepackage{enumerate}
 
@@
 
-In order to manage the extreme flexibilty of the GP model and focus on 
+In order to manage the great flexibilty of the GP model and focus on 
 reliably retrieving the period parameter, we also impose priors
@@
 and small values of $l$) tended to allow fits that ignored the desired
-periodicity.  The priors we used for this work are listed in
-table \ref{tab:priors}.  We note that we do not have any particular
+periodicity.  The priors and bounds we used for this work are listed in
+table \ref{tab:priors}.  We note that we do not have any specific
 quantitative justification for the use of these priors,
@@
 light curves.  We thus caution that other data sets might have different
-noise properties, potentially requiring modifications to these priors.
+noise properties, potentially requiring modifications to these priors
+on the non-period parameters.  For period, we tried two different methods:
+an uninformed (log-flat) prior between 0.5\,d and 100\,d, and an ACF-informed
+prior (details in Appendix.)
+
 % We note that our default prior for period is 
@@
 representative of the posterior.  This fitting process takes several hours 
-for a typical simulated light curve.
+for a typical simulated light curve, though in some cases it may take 12
+hours or longer to converge.
 
@@
     $\ln \sigma$ & $\mathcal N(-17, 5)$ & (-20, 0) \\
-    $\ln P $ & Uniform & ($\ln 0.5, \ln 100$) \\ 
+    $\ln P $ & Uniform / ACF-based & ($\ln 0.5, \ln 100$) \\ 
 \end{tabular}
@@
 
-Figure \ref{fig:compare_mcmc} summarizes the results of this MCMC fitting procedure
-compared to the injected ``true'' stellar rotation periods. 
+Figures \ref{fig:compare_mcmc_noprior} and \ref{fig:compare_mcmc_acfprior} 
+summarize the results of this MCMC fitting 
+procedure compared to the injected ``true'' stellar rotation periods, for the 
+uninformed and ACF-informed priors on period, respectively.
 The marginal posterior distributions of the QP kernel hyperparameters for the
@@
 
+\appendix
+
+\section{An ACF-informed Prior for Rotation Period}
+
+The autocorrelation function (ACF) has proven to be very useful
+for measuring the rotation periods of stars from time-series
+photometry.  
+However, as discussed in the body of this paper, the
+ACF-based method has several shortcomings, 
+most notably the inability to deliver uncertainties, but also 
+the necessity of several heuristic choices, 
+such as a timescale on which to smooth the ACF, 
+how to define a peak, whether the first or second peak 
+gets selected, and what constitutes a secure detection.
+While this paper presents a new MCMC-based method
+that avoids these shortcomings, 
+it seems prudent to still use information available from the ACF.
+This Appendix describes how we use the ACF to define a prior on period.
+
+Our strategy here is not to attempt to decide which single
+peak in the ACF best represents the true rotation period,
+but rather to identify several \emph{candidate} periods and define
+a weighting scheme in order to create a non-commital, though useful, 
+multimodal prior.  While this procedure does not avoid heuristic choices, 
+the fact that its goal is simply to create a \emph{prior} for a 
+probabilitistically justified inference scheme rather than
+to identify a single correct period softens the potential impact 
+of these choices.
+
+As another innovation beyond what ACF methods in the literature have 
+presented, we also bandpass filter the light curves (using a 5th order
+Butterworth filter, as implemented in \texttt{SciPy}) before calculating the 
+autocorrelation.  This suppresses power on timescales shorter than a chosen 
+minimum period $P_{\rm min}$ and longer than a chosen maximum $P_{\rm max}$,
+producing a cleaner autocorrelation signal than an unfiltered light curve.
+
+We use the following procedure to construct a prior for rotation period
+given a light curve:
+\begin{enumerate}
+\item{For each value of $P_i$, where $i = \{1, 3, 5, 10, 30, 50, 100\}$\,d,
+we apply a bandpass filter to the light curve using $P_{\rm min}=0.1$\,d
+and $P_{\rm max} = P_i$.  We then calculate the ACF of the filtered 
+light curve out to a maximum lag of $2P_i$ and smooth it with a boxcar 
+filter of width $P_i/10$.}
+
+\item{We identify the time lag corresponding to the 
+first peak of each of these ACFs, as well as the first peak's
+trough-to-peak height, creating a set of candidate periods
+$T_i$ and heights $h_i$.}
+
+\item{We assign a quality metric $Q_i$ to each of these candidate
+periods, as follows.  First, we model the ACF as a 
+damped oscillator with fixed period $T_i$:
+\begin{equation}
+y = A e^{-t/\tau} \cos{\frac{2\pi t}{T_i} },
+\end{equation}
+where and $t$ is the lag time, 
+and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
+least squares minimization procedure.  We then define the
+following heuristic quality metric:
+\begin{equation}
+\label{eq:quality}
+Q_i = \left(\frac{\tau_i}{T_i}\right) \left(\frac{N_i h_i}{R_i}\right),
+\end{equation}
+where $h_i$ is the height of the ACF peak at $T_i$, 
+$N_i$ is the length of the lag vector in the ACF (directly proportional
+to the maximum allowed period $P_i$), 
+and $R_i$ is the sum of squared residuals between the 
+damped oscillator model and the actual ACF data.  The idea behind this 
+quality metric to give a candidate period a higher score if
+
+    \begin{enumerate}[(a)]
+
+    \item{it has many regular sinusoidal peaks, such that the decay
+        time $\tau_i$ is long compared to the oscillation period $T_i$,}
+    \item{the ACF peak height is high, and}
+    \item{the damped oscillator model is a good fit (in a $\chi^2$
+      sense) to the ACF, with extra bonus for being 
+      a good fit over more points (larger $N_i$, or longer $P_i$).}
+
+    \end{enumerate}
+}
+
+\item{Given this set of candidate periods $T_i$ and quality metrics $Q_i$,
+we finally construct a multimodal prior on the $P$ parameter of the GP
+model as a weighted mixture of Gaussians:
+\begin{equation}
+\label{eq:mixture}
+p(\ln P) = \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) + 
+                                          0.05\mathcal N(\ln (T_i/2), 0.2) +
+                                          0.05\mathcal N(\ln (2 T_i), 0.2) \right)}
+                {\sum_i Q_i}.
+\end{equation}
+That is, in addition to taking the candidate periods themselves as mixture
+components, we also mix in twice and half each candidate period at a lower level, 
+which compensates for the possibility that the first peak in the ACF may actually 
+represent half or twice the actual rotation period.  The period width of 0.2 in 
+log space (corresponding to roughly 20\% uncertainty) is again a heuristic choice,
+balancing a healthy specificity with the desire to not have the results of 
+the inference overly determined by the ACF prior.
+}
+
+Incidentally, while we use the procedure described here to create a prior on 
+$P$ which we use while inferring the parameters of the quasi-periodic GP model,
+this same procedure may also be used in the service of a
+rotation-period estimating procedure all on its own, perhaps being even more 
+robust and accurate than the traditional ACF method.  We leave exploration of 
+this possibility to future work.
+
+\end{enumerate}
+
 \bibliographystyle{plainnat}

### 2017-01-16 documents/ms.tex
spacing changes
koi section edits tdm

@@
 \newcommand{\nmcquillan}{100}
+\newcommand{\nkois}{1114}
+\newcommand{\nkoimcq}{271}
+\newcommand{\nkoigood}{222}
+\newcommand{\nkoibad}{49}
 \newcommand{\kepexample}{1430163}
@@
 \label{sec:kepler}
+
+In order to test this rotation period inference method on real data,
+we applied it to a set of \nkois\ \Kepler\ Object of Interest (KOI)
+host stars, \nkoimcq\ of which had rotation periods previously measured by 
+the ACF method by \citet{Mcquillan2013}.  We use the pipeline-corrected flux 
+(\texttt{pdcsap_flux} column in the \Kepler\ light curve table), median-normalized
+and unit-subtracted.  As with the simulated light curves, we randomly subsample each 
+light curve by a factor of 30 and split it into segments of about 300 points
+for the purposes of evaluating the likelihood.  We also follow the same MCMC
+fitting procedure as with the simulated data.
+
+Figure \ref{fig:kois} compares the periods inferred with the GP method to the
+ACF-based periods from \citet{Mcquillan2013} for the \nkoimcq\ overlapping KOIs.
+This comparison shows generally very good agreement, 
+with the exception of a cluster of stars in the lower right of the figure
+where the GP method does not recover the ACF period.  
+However, rather than being primarily evidence of weakness of the GP method,
+this cluster of nominal failures actually reveals an important advantage
+for this method---the possibility of quantitatively motivated quality assessment.
 

### 2017-01-23 documents/ms.tex
small text changes

--- 2016-12-27 documents/ms.tex
@@
 \usepackage{bm}
-\usepackage{enumerate}
 
 \newcommand{\columbia}{1}
-\newcommand{\oxford}{2}
-\newcommand{\uw}{3}
-\newcommand{\princeton}{4}
+\newcommand{\oxford}{3j}
+\newcommand{\uw}{4}
+\newcommand{\princeton}{2}
 \newcommand{\naigrain}{333}
@@
    Ruth Angus\altaffilmark{\columbia},
+   Timothy Morton\altaffilmark{\princeton}
    Suzanne Aigrain\altaffilmark{\oxford},
-   Timothy Morton\altaffilmark{\princeton}
    \& Daniel Foreman-Mackey\altaffilmark{\uw}
@@
 WA}
-\altaffiltext{\princeton}{Department of Astrophysical Sciences, Princeton University,
-Princeton, NJ}
+\altaffiltext{\princeton}{Department of Astrophysical Sciences, Princeton
+University,  Princeton, NJ}
 
@@
 Furthermore, the ACF method is not applicable to unevenly spaced data,
-therefore inappropriate for the ~$10^7$ stellar light curves expected to be
-produced by \LSST\ and \Wfirst.
+therefore it is inappropriate for the ~$10^7$ stellar light curves expected
+to be produced by \LSST\ unlike the GP method.
 % Furthermore, the improvement is expected to be even more dramatic when applied
@@
 light curves (although in many cases it can be approximated as uniformly
-sampled) and not nearly the case for future large-scale survey instruments,
-\LSST\ and \Wfirst.
+sampled) and not nearly the case for the future large-scale survey instrument,
+\LSST.
 An ACF is also an operation performed on the data, not a generative model of
@@
 
-Sampling the posterior described above presents several challenges:
+Sampling the posterior PDF described above presents several challenges:
 \begin{itemize}
@@
 ($\sim$40,000 points) takes about $\sim$5\,s--- too computationally expensive
-to perform inference on large numbers of light curves\footnote{All computational
-times cited in this section are based on evaluations on a 
-single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.  The matrix operations
-necessary to evaluate the GP likelihood\footnote{These operations use 
-the fast matrix solver 
-HODLR \citep{Ambikasaran2014}, implemented in the {\tt george} \citep{George} 
-python package.} scale as $N\ln(N)^2$, where $N$ is
-the number of data points in the light curve.
+to perform inference on large numbers of light curves\footnote{All
+computational times cited in this section are based on evaluations on a
+single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.
+The matrix operations necessary to evaluate the GP likelihood\footnote{These
+operations use the fast matrix solver HODLR \citep{Ambikasaran2014},
+implemented in the {\tt george} \citep{George} python package.} scale as
+$N\ln(N)^2$, where $N$ is the number of data points in the light curve.
 
@@
 posterior multimodality and ``over-fitting''--like behavior.  For example,
-if $l$ is small,  the non-periodic factor in the covariance kernel may dominate, 
-allowing for a good fit to the data without requiring any 
+if $l$ is small,  the non-periodic factor in the covariance kernel may
+dominate, allowing for a good fit to the data without requiring any
 periodic covariance structure---even if
@@
 
-In order to manage the great flexibilty of the GP model and focus on 
+In order to manage the extreme flexibilty of the GP model and focus on
 reliably retrieving the period parameter, we also impose priors
@@
 and small values of $l$) tended to allow fits that ignored the desired
-periodicity.  The priors and bounds we used for this work are listed in
-table \ref{tab:priors}.  We note that we do not have any specific
+periodicity.  The priors we used for this work are listed in
+table \ref{tab:priors}.  We note that we do not have any particular
 quantitative justification for the use of these priors,
@@
 light curves.  We thus caution that other data sets might have different
-noise properties, potentially requiring modifications to these priors
-on the non-period parameters.  For period, we tried two different methods:
-an uninformed (log-flat) prior between 0.5\,d and 100\,d, and an ACF-informed
-prior (details in Appendix.)
-
+noise properties, potentially requiring modifications to these priors.
 % We note that our default prior for period is
@@
 representative of the posterior.  This fitting process takes several hours
-for a typical simulated light curve, though in some cases it may take 12
-hours or longer to converge.
+for a typical simulated light curve.
 
@@
     $\ln \sigma$ & $\mathcal N(-17, 5)$ & (-20, 0) \\
-    $\ln P $ & Uniform / ACF-based & ($\ln 0.5, \ln 100$) \\ 
+    $\ln P $ & Uniform & ($\ln 0.5, \ln 100$) \\
 \end{tabular}
@@
 % \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_emcee3_13_12_2.png}
+\includegraphics[width=6in, clip=true]{figures/comparison_emcee3_17_1.png}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 
-Figures \ref{fig:compare_mcmc_noprior} and \ref{fig:compare_mcmc_acfprior} 
-summarize the results of this MCMC fitting 
-procedure compared to the injected ``true'' stellar rotation periods, for the 
-uninformed and ACF-informed priors on period, respectively.
+Figure \ref{fig:compare_mcmc} summarizes the results of this MCMC fitting procedure
+compared to the injected ``true'' stellar rotation periods.
 The marginal posterior distributions of the QP kernel hyperparameters for the
@@
 Our results show good agreement with the \citet{Mcquillan2014} measurements,
-as shown in ... %\ref{fig:mcquillan}.
+as shown in \ref{fig:mcquillan}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/comparison_koi.pdf}
+\caption[Comparison with McQuillan results.]
+{A comparison of our GP rotation period measurements to those of
+\citet{Mcquillan2014}. The results show good agreement. In several cases we
+    measure a rotation period that is half the value reported by
+    \citet{Mcquillan2014}.}
+\label{fig:mcquillan}
+\end{center}
+\end{figure}
 
@@
 
-\appendix
-
-\section{An ACF-informed Prior for Rotation Period}
-
-The autocorrelation function (ACF) has proven to be very useful
-for measuring the rotation periods of stars from time-series
-photometry.  
-However, as discussed in the body of this paper, the
-ACF-based method has several shortcomings, 
-most notably the inability to deliver uncertainties, but also 
-the necessity of several heuristic choices, 
-such as a timescale on which to smooth the ACF, 
-how to define a peak, whether the first or second peak 
-gets selected, and what constitutes a secure detection.
-While this paper presents a new MCMC-based method
-that avoids these shortcomings, 
-it seems prudent to still use information available from the ACF.
-This Appendix describes how we use the ACF to define a prior on period.
-
-Our strategy here is not to attempt to decide which single
-peak in the ACF best represents the true rotation period,
-but rather to identify several \emph{candidate} periods and define
-a weighting scheme in order to create a non-commital, though useful, 
-multimodal prior.  While this procedure does not avoid heuristic choices, 
-the fact that its goal is simply to create a \emph{prior} for a 
-probabilitistically justified inference scheme rather than
-to identify a single correct period softens the potential impact 
-of these choices.
-
-As another innovation beyond what ACF methods in the literature have 
-presented, we also bandpass filter the light curves (using a 5th order
-Butterworth filter, as implemented in \texttt{SciPy}) before calculating the 
-autocorrelation.  This suppresses power on timescales shorter than a chosen 
-minimum period $P_{\rm min}$ and longer than a chosen maximum $P_{\rm max}$,
-producing a cleaner autocorrelation signal than an unfiltered light curve.
-
-We use the following procedure to construct a prior for rotation period
-given a light curve:
-\begin{enumerate}
-\item{For each value of $P_i$, where $i = \{1, 3, 5, 10, 30, 50, 100\}$\,d,
-we apply a bandpass filter to the light curve using $P_{\rm min}=0.1$\,d
-and $P_{\rm max} = P_i$.  We then calculate the ACF of the filtered 
-light curve out to a maximum lag of $2P_i$ and smooth it with a boxcar 
-filter of width $P_i/10$.}
-
-\item{We identify the time lag corresponding to the 
-first peak of each of these ACFs, as well as the first peak's
-trough-to-peak height, creating a set of candidate periods
-$T_i$ and heights $h_i$.}
-
-\item{We assign a quality metric $Q_i$ to each of these candidate
-periods, as follows.  First, we model the ACF as a 
-damped oscillator with fixed period $T_i$:
-\begin{equation}
-y = A e^{-t/\tau} \cos{\frac{2\pi t}{T_i} },
-\end{equation}
-where and $t$ is the lag time, 
-and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
-least squares minimization procedure.  We then define the
-following heuristic quality metric:
-\begin{equation}
-\label{eq:quality}
-Q_i = \left(\frac{\tau_i}{T_i}\right) \left(\frac{N_i h_i}{R_i}\right),
-\end{equation}
-where $h_i$ is the height of the ACF peak at $T_i$, 
-$N_i$ is the length of the lag vector in the ACF (directly proportional
-to the maximum allowed period $P_i$), 
-and $R_i$ is the sum of squared residuals between the 
-damped oscillator model and the actual ACF data.  The idea behind this 
-quality metric to give a candidate period a higher score if
-
-    \begin{enumerate}[(a)]
-
-    \item{it has many regular sinusoidal peaks, such that the decay
-        time $\tau_i$ is long compared to the oscillation period $T_i$,}
-    \item{the ACF peak height is high, and}
-    \item{the damped oscillator model is a good fit (in a $\chi^2$
-      sense) to the ACF, with extra bonus for being 
-      a good fit over more points (larger $N_i$, or longer $P_i$).}
-
-    \end{enumerate}
-}
-
-\item{Given this set of candidate periods $T_i$ and quality metrics $Q_i$,
-we finally construct a multimodal prior on the $P$ parameter of the GP
-model as a weighted mixture of Gaussians:
-\begin{equation}
-\label{eq:mixture}
-p(\ln P) = \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) + 
-                                          0.05\mathcal N(\ln (T_i/2), 0.2) +
-                                          0.05\mathcal N(\ln (2 T_i), 0.2) \right)}
-                {\sum_i Q_i}.
-\end{equation}
-That is, in addition to taking the candidate periods themselves as mixture
-components, we also mix in twice and half each candidate period at a lower level, 
-which compensates for the possibility that the first peak in the ACF may actually 
-represent half or twice the actual rotation period.  The period width of 0.2 in 
-log space (corresponding to roughly 20\% uncertainty) is again a heuristic choice,
-balancing a healthy specificity with the desire to not have the results of 
-the inference overly determined by the ACF prior.
-}
-
-Incidentally, while we use the procedure described here to create a prior on 
-$P$ which we use while inferring the parameters of the quasi-periodic GP model,
-this same procedure may also be used in the service of a
-rotation-period estimating procedure all on its own, perhaps being even more 
-robust and accurate than the traditional ACF method.  We leave exploration of 
-this possibility to future work.
-
-\end{enumerate}
-
 \bibliographystyle{plainnat}

### 2017-01-23 documents/ms.tex

--- 2017-01-16 documents/ms.tex
@@
 \newcommand{\columbia}{1}
-\newcommand{\oxford}{2}
-\newcommand{\uw}{3}
-\newcommand{\princeton}{4}
+\newcommand{\oxford}{3j}
+\newcommand{\uw}{4}
+\newcommand{\princeton}{2}
 \newcommand{\naigrain}{333}
@@
 prior (details in Appendix.)
-
 % We note that our default prior for period is
@@
 representative of the posterior.  This fitting process takes several hours 
-for a typical simulated light curve, though in some cases it may take 12
+for a typical simulated light curve, though in some cases it can take 12
 hours or longer to converge.

### 2017-01-23 documents/ms.tex
new koi section

@@
 \newcommand{\nmcquillan}{100}
-\newcommand{\nkois}{1114}
-\newcommand{\nkoimcq}{271}
-\newcommand{\nkoigood}{222}
-\newcommand{\nkoibad}{49}
+\newcommand{\nkois}{1102}
+\newcommand{\nkoimcq}{275}
+% \newcommand{\nkoigood}{222}
+% \newcommand{\nkoibad}{49}
 \newcommand{\kepexample}{1430163}
@@
 for the purposes of evaluating the likelihood.  We also follow the same MCMC
-fitting procedure as with the simulated data.
-
-Figure \ref{fig:kois} compares the periods inferred with the GP method to the
+fitting procedure as with the simulated data, using the ACF-based prior as before.
+
+Initially, we also use the same priors on the hyperparameters for the KOIs 
+as for the simulated light curves (Table \ref{tab:priors}).  
+However, we found that $\ln l$ and $\ln \Gamma$ tended toward slightly different
+values than the simulations.  We also found that the allowed hyperparameter
+range that we used for the simulations was too large for the KOI population, 
+as maybe $\sim$15\% of the fits tended toward corners in the hyperparameter
+space, resulting in poor period measurements.  As a result, after this initial
+test, we subsequently adjusted the priors and re-fit all the KOIs.  The final 
+priors and parameter ranges that we used are in Table \ref{tab:koipriors}.
+
+\begin{table*}
+\begin{center}
+\caption{Priors and bounds on the natural logarithms of the GP model parameters, 
+        for \Kepler\ light curves}
+\begin{tabular}{lcc}
+Parameter & Prior & Bounds\\
+    \hline
+    $\ln A$ & $\mathcal N(-13, 5.7)$ & (-20, 0) \\
+    $\ln l$ & $\mathcal N(5.0, 1.2)$ & (2, 8) \\
+    $\ln \Gamma$ & $\mathcal N(1.9, 1.4)$ & (0, 3) \\
+    $\ln \sigma$ & $\mathcal N(-17, 5)$ & (-20, 0) \\
+    $\ln P $ & Uniform / ACF-based & ($\ln 0.5, \ln 100$) \\ 
+\end{tabular}
+\end{center}
+\end{table*}
+\label{tab:koipriors}
+
+Figure \ref{fig:mcquillan} compares the periods inferred with the GP method to the
 ACF-based periods from \citet{Mcquillan2013} for the \nkoimcq\ overlapping KOIs.
-This comparison shows generally very good agreement, 
-with the exception of a cluster of stars in the lower right of the figure
-where the GP method does not recover the ACF period.  
-However, rather than being primarily evidence of weakness of the GP method,
-this cluster of nominal failures actually reveals an important advantage
-for this method---the possibility of quantitatively motivated quality assessment.
-
-We applied our method to \nmcquillan\ stars with rotation periods previously
-measured by \citet{Mcquillan2014}.
-A feature of \kepler\ light curves that works to our advantage is that they
-are naturally broken up into smaller time units by quarterly (three month)
-breaks.
-As explained above, splitting the data set into quarters, rather than
-modelling the entire time series contiguously speeds up the computation time
-as inverting several small matrices is faster than inverting one large one.
-The \kepler\ quarter divisions are natural places to split the time series
-because the spacecraft rotates by ninety degrees every quarter (three months),
-placing each star on a new CCD module.
-Pixel response functions and background flux differs from pixel-to-pixel and
-module-to-module, so noise properties of \kepler\ light curves change every
-quarter.
-Additionally, changes in the spacecraft's orientation and position during
-quarterly re-pointings temporarily affect the temperature of the CCD,
-producing systematic features in the light curves at the start of some
-quarters.
-We model each quarter separately but the parameters of the GP kernel function
-are global, {\it i.e.} we do not use a separate period parameter for each
-quarter---there is just one period parameter for an entire light curve.
-It would be possible to model the time series with a mixture of some global
-parameters and some quarter-specific parameters, for example one might expect
-that the amplitude of covariance, $A$ or white noise level, $\sigma$ to vary
-on a quarterly basis.
-However, since there are seventeen quarters this would lead to thirty-seven
-parameters, and in the interest of minimising computation time (adding
-parameters leads to longer MCMC burn in and convergence time), we choose to
-use global parameters only.
-
-Our results show good agreement with the \citet{Mcquillan2014} measurements,
-as shown in \ref{fig:mcquillan}.
+This comparison shows generally very good agreement, with only a few exceptions, 
+demonstrating that this method works not only on simulated data, but also on 
+real data---with the caveat that for any particular data set, some care is needed 
+regarding the setting the priors and ranges for the GP hyperparameters.
 
@@
 
-At this point it is important to highlight the following caveat.
-Our method has been tuned to reproduce the rotation periods of the
-\citet{Aigrain2015} simulated light curves.
-These light curves were simulated using a very good but still imperfect model.
-The input parameters and their distributions were chosen to be as realistic as
-possible given current observations.
-Unfortunately however, there are limited data on some of these parameters for
-stars other than the Sun: for example the distrubution of spot lifetimes or
-activity cycle length.
-Ensuring that we can recover the rotation periods of these simulated light
-curves does not necessarily ensure that we can recover the rotation periods of
-real \kepler\ light curves with equal accuracy and precision.
-Even using priors on the QP kernel parameters that have been derived from the
-\citet{Aigrain2015} sample may be dangerous.
-This caveat should be taken into consideration when this method is applied to
-real data.
+% At this point it is important to highlight the following caveat.
+% Our method has been tuned to reproduce the rotation periods of the
+% \citet{Aigrain2015} simulated light curves.
+% These light curves were simulated using a very good but still imperfect model.
+% The input parameters and their distributions were chosen to be as realistic as
+% possible given current observations.
+% Unfortunately however, there are limited data on some of these parameters for
+% stars other than the Sun: for example the distrubution of spot lifetimes or
+% activity cycle length.
+% Ensuring that we can recover the rotation periods of these simulated light
+% curves does not necessarily ensure that we can recover the rotation periods of
+% real \kepler\ light curves with equal accuracy and precision.
+% Even using priors on the QP kernel parameters that have been derived from the
+% \citet{Aigrain2015} sample may be dangerous.
+% This caveat should be taken into consideration when this method is applied to
+% real data.
 

### 2017-01-25 documents/ms.tex
fixed latex typo with underscore
transits masked

@@
 the ACF method by \citet{Mcquillan2013}.  We use the pipeline-corrected flux 
-(\texttt{pdcsap_flux} column in the \Kepler\ light curve table), median-normalized
-and unit-subtracted.  As with the simulated light curves, we randomly subsample each 
+(\texttt{pdcsap\_flux} column in the \Kepler\ light curve table), median-normalized
+and unit-subtracted, and mask out all known transiting planet candidate signals.  
+As with the simulated light curves, we randomly subsample each 
 light curve by a factor of 30 and split it into segments of about 300 points

### 2017-01-26 documents/ms.tex
updating plots and adding words about kois

--- 2017-01-23 documents/ms.tex
@@
 \usepackage{bm}
-\usepackage{enumerate}
 
 \newcommand{\columbia}{1}
-\newcommand{\oxford}{3j}
+\newcommand{\oxford}{3}
 \newcommand{\uw}{4}
@@
 \newcommand{\nmcquillan}{100}
-\newcommand{\nkois}{1114}
-\newcommand{\nkoimcq}{271}
-\newcommand{\nkoigood}{222}
-\newcommand{\nkoibad}{49}
 \newcommand{\kepexample}{1430163}
@@
    Suzanne Aigrain\altaffilmark{\oxford},
-   \& Daniel Foreman-Mackey\altaffilmark{\uw}
+   Daniel Foreman-Mackey\altaffilmark{\uw}
+   Suzanne Aigrain\altaffilmark{\oxford},
+   \& Vinesh Rajpaul\altaffilmark{\oxford}
 }
@@
 {The `true' rotation periods used to generate \naigrain\ simulated light
-curves vs the rotation periods measured using a LS periodogram technique.}
+curves vs the rotation periods measured using a LS periodogram technique.
+In many cases a large peak at a long period was present in the periodogram,
+    producing a significant over-estimate of the period.
+    }
 \label{fig:pgram_compare}
@@
 light curves.  We thus caution that other data sets might have different
-noise properties, potentially requiring modifications to these priors
-on the non-period parameters.  For period, we tried two different methods:
-an uninformed (log-flat) prior between 0.5\,d and 100\,d, and an ACF-informed
-prior (details in Appendix.)
+noise properties, potentially requiring modifications to these priors.
 % We note that our default prior for period is
@@
 representative of the posterior.  This fitting process takes several hours
-for a typical simulated light curve, though in some cases it can take 12
-hours or longer to converge.
+for a typical simulated light curve.
 
@@
 \label{sec:kepler}
-
-In order to test this rotation period inference method on real data,
-we applied it to a set of \nkois\ \Kepler\ Object of Interest (KOI)
-host stars, \nkoimcq\ of which had rotation periods previously measured by 
-the ACF method by \citet{Mcquillan2013}.  We use the pipeline-corrected flux 
-(\texttt{pdcsap_flux} column in the \Kepler\ light curve table), median-normalized
-and unit-subtracted.  As with the simulated light curves, we randomly subsample each 
-light curve by a factor of 30 and split it into segments of about 300 points
-for the purposes of evaluating the likelihood.  We also follow the same MCMC
-fitting procedure as with the simulated data.
-
-Figure \ref{fig:kois} compares the periods inferred with the GP method to the
-ACF-based periods from \citet{Mcquillan2013} for the \nkoimcq\ overlapping KOIs.
-This comparison shows generally very good agreement, 
-with the exception of a cluster of stars in the lower right of the figure
-where the GP method does not recover the ACF period.  
-However, rather than being primarily evidence of weakness of the GP method,
-this cluster of nominal failures actually reveals an important advantage
-for this method---the possibility of quantitatively motivated quality assessment.
 
@@
 as shown in \ref{fig:mcquillan}.
+The uncertainties published in the \citet{Mcquillan2014} catalogue were
+calculated from the scatter in the positions of peaks in the ACF that are
+separated by integer multiples of the period.
+We show a comparison of our results to those of the \kepler\ Objects of
+Interest (KOIS) \citet{Mcquillan2014} in order to demonstrate that our
+technique works on real \kepler\ light curves, not just simulated ones.
+These are perhaps not the ideal test subjects because many KOI light curves
+have deep stellar occulatations and planetary transits which can look
+like non-Gaussian noise.
+Since Gaussian processes require that the data are Gaussian, this could have
+detremental effects on the results.
+Despite this however, our results are, in most cases, similar to those of
+\citet{Mcquillan2014} indicating that our method still works well even in the
+presence of transits and eclipses.
+In many cases where there is a large discrepancy between our results and the
+\citet{Mcquillan2014} results, either the variability is low amplitude or
+frequent, deep eclipses/transits are present in the data.
+In most other cases we expect our results to be the more accurate ones and the
+discrepancy may be due to quasi-periodicity in the light curve.
 
@@
 {A comparison of our GP rotation period measurements to those of
-\citet{Mcquillan2014}. The results show good agreement. In several cases we
-    measure a rotation period that is half the value reported by
-    \citet{Mcquillan2014}.}
+\citet{Mcquillan2014}.
+The data points are coloured by the range of variability measured by
+    \citet{Mcquillan2014}, defined as the interval between the 5th and 95th
+    percentiles of normalized flux per period bin in millimagnitudes.
+    The results show good agreement.}
 \label{fig:mcquillan}
@@
 autocorrelation function method and a new Gaussian process method.
-The GP method produced the most precise and accurate rotation periods of the
-three techniques.
-This is because a GP is a semi-parametric model that is well suited to signals
-that are non-sinusoidal and quasi-periodic.
+We demonstrate that the GP method produced the most precise and accurate
+rotation periods of the three techniques.
+In addition, we compared the rotation periods of \kepler\ objects of
+interest measured using the GP method to those measured previously by
+\citet{Mcquillan2014}.
+The good agreement between the two sets of results demonstrates that the GP
+method works well on real \kepler\ data and is, in most cases robust to the
+presents of transits and eclipses in the light curves.
+
+Although rotation period uncertainties calculated using the GP method are
+better estimates than those calculated using the ACF method, they are still
+imperfect estimates of the intrinsic uncertainties.
+This is because the posterior PDF of the rotation period parameter is, in many
+cases, multi-modal.
+If it were possible to run an MCMC sampler for an infinite number of steps, we
+would the entire posterior PDF would eventually be sampled, producing an
+accurate uncertainty measurement.
+Practically speaking however, these imperfect estimates may be the best
+currently achievable.
+
+A GP is a semi-parametric model that is well suited to signals that are
+non-sinusoidal and quasi-periodic.
 Not only this, unlike the other two methods, the GP method provides posterior
@@
 Note that the posterior PDFs are often non-Gaussian, being bi- or multimodal.
-In these cases it is therefore misleading to report a 1$\sigma$ or 16th and
+In these cases it is therefore misleading to report median values or 16th and
 84th percentiles as uncertainties.
@@
 require uncertainty propagation.
-
 Because rotation periods are inferred probabilistically with the GP method,
@@
 stellar ages or characterising star-planet interactions for example.
-This the GP method also eliminates the need to define a heuristic detection
-criterion such as are often used in the literature: a light curve that does
-not contain a rotation signal will return the prior PDF.
-
-The main drawback of the GP method is the computation time.
+
+The main drawback of the GP method is the computational cost.
 

### 2017-01-26 code/plot_koi.py

@@ -35,7 +35,10 @@ def load_and_plot(data, DATA_DIR, RESULTS_DIR):
     xerr = data.P_rot_err.values
     lnxerr = xerr/x
     amp = data.R_var.values
-    l = recovered > 0
+    # remove binaries.
+    l = (recovered > 0) * (data.koi_id.values != 197) * \
+        (data.koi_id.values != 279)
+
 
     plt.clf()
     xs = np.log(np.linspace(0, 100, 100))
@@ -50,11 +53,28 @@ def load_and_plot(data, DATA_DIR, RESULTS_DIR):
                 edgecolor=".5", cmap="GnBu_r", vmin=min(np.log(amp[l])),
                 vmax=max(np.log(amp[l])), s=20, zorder=2, lw=.2)
 
+    import kplr
+    import kepler_data as kd
+    client = kplr.API()
+    m = 20 < (np.abs(x[l] - recovered[l])/x[l] * 100)
+    for i, koi in enumerate(data.koi_id.values):
+        print(data.koi_id.values[l][m][i], x[l][m][i], recovered[l][m][i])
+        star = client.koi("{}.01".format(str(data.koi_id.values[l][m][i])))
+        # star.get_light_curves(fetch=True, shortcadence=False)
+        print(star.kepid)
+        kid = str(int(star.kepid)).zfill(9)
+        fname = "/Users/ruthangus/.kplr/data/lightcurves/{}".format(kid)
+        x, y, yerr = kd.load_kepler_data(fname)
+        plt.clf()
+        plt.plot(x, y, "k.")
+        plt.xlim(150, 300)
+        plt.savefig("{}".format(data.koi_id.values[l][m][i]))
+
     plt.ylim(0, 4)
     plt.xlim(0, 4)
     cbar = plt.colorbar()
     cbar.ax.set_ylabel("$\ln(\mathrm{R}_{\mathrm{var}})$")
-    plt.xlabel("$\ln(\mathrm{McQuillan~{\it et~al.}~(2014)~Period})$")
+    plt.xlabel("$\ln(\mathrm{McQuillan~{\it et~al.,}~2013~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
     plt.savefig("comparison_koi")
     plt.savefig("../documents/figures/comparison_koi.pdf")

### 2017-01-26 documents/ms.tex

--- 2017-01-25 documents/ms.tex
@@
 \newcommand{\columbia}{1}
-\newcommand{\oxford}{3j}
+\newcommand{\oxford}{3}
 \newcommand{\uw}{4}
@@
 regarding the setting the priors and ranges for the GP hyperparameters.
+>>>>>>> 265508362e5e7a01a39bdd616c754c276059397c
 

### 2017-01-27 documents/ms.tex
adding figures
minor edits

@@
 However, in the case where there is evidence for correlated
-`noise'\footnote{In our case the `noise' is actually the model!}, as in the
-case of \Kepler\ light curves, those off-diagonal elements are non-zero.
-With GP regression, a covariance matrix generated by the kernel function,
-${\bf K}$ replaces ${\bf C}$ in the above equation.
-Incidentally, this approach is the reverse of the regression techniques
+`noise'\footnote{In our case the `noise' is actually the model!  Incidentally, this approach is the reverse of the regression techniques
 usually employed by astronomers.
@@
 Here, the parameters describing the correlated noise are what we are
-interested in and our mean model is simply a straight line at $y=0$.
+interested in and our mean model is simply a straight line at $y=0$.}, as in the
+case of \Kepler\ light curves, those off-diagonal elements are non-zero.
+With GP regression, a covariance matrix generated by the kernel function,
+${\bf K}$ replaces ${\bf C}$ in the above equation.
 
@@
 200 days), and evaluate the log-likelihood as the sum of the log-likelihoods
-of the individual sections (all using the same parameters $\theta$).  This reduces
+of the individual sections (all using the same parameters ${\bm \theta}$).  This reduces
 computation time because the section-based likelihood evaluation scales as $mn\ln(n)^2$, where $n$ is the number of data points per section and $m$ is the number of sections.
@@
 noise properties, potentially requiring modifications to these priors
-on the non-period parameters.  For period, we tried two different methods:
+on the non-period parameters (e.g., see Section \ref{sec:kepler}).  
+For period, we tried two different methods:
 an uninformed (log-flat) prior between 0.5\,d and 100\,d, and an ACF-informed
@@
 % \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_emcee3_17_1.png}
+\includegraphics[width=6in, clip=true]{figures/comparison_emcee3_24_0.png}
 \caption{The `true' rotation periods used to generate \naigrain\

### 2017-01-28 documents/ms.tex
some words

--- 2017-01-26 documents/ms.tex
@@
 {The `true' rotation periods used to generate \naigrain\ simulated light
-curves vs the rotation periods measured using the ACF technique.}
+curves vs the rotation periods measured using the ACF technique.
+    Several light curves have over-estimated rotation periods and some
+    are drastically underestimated.
+    In some cases the recovered period is twice the true period used to
+    simulate the light curve.}
 \label{fig:compare_acf}
@@
 
-To sample the posterior in a way that is sensitive to potential
-multimodality, we use the \texttt{emcee3} MCMC sampler.
-We initialize 500 walkers with random samples from the prior
- and use a mix of different proposal distributions:
-specifically, a 2:2:1 mix of KDE, differential evolution, and DESnooker moves.
+To sample the posterior in a way that is sensitive to potential multimodality,
+we use the \texttt{emcee3} MCMC sampler.
+We initialize 500 walkers with random samples from the prior and use a mix of
+different proposal distributions: specifically, a 2:2:1 mix of KDE,
+differential evolution, and DESnooker moves.
 [DFM: more words/citations here?]
 We run fifty steps of the sampler at a time, checking for convergence after
-each iteration, up to a maximum of fifty iterations.  We declare convergence if
-the total effective chain length is at least 8$\times$ the maximum autocorrelation
-time.  When convergence is achieved, we drop the first two autocorrelation
-lengths in the chain as a burn-in, and randomly choose 5000 samples as
-representative of the posterior.  This fitting process takes several hours
-for a typical simulated light curve, though in some cases it can take 12
-hours or longer to converge.
+each iteration, up to a maximum of fifty iterations.
+We declare convergence if the total effective chain length is at least
+8$\times$ the maximum autocorrelation time.
+When convergence is achieved, we drop the first two autocorrelation lengths in
+the chain as a burn-in, and randomly choose 5000 samples as representative of
+the posterior.
+This fitting process takes several hours for a typical simulated light curve,
+though in some cases it can take 12 hours or longer to converge.
 
@@
 \begin{center}
-\caption{Priors and bounds on the natural logarithms of the GP model parameters.}
+\caption{Priors and bounds on the natural logarithms of the GP model
+    parameters.}
 \begin{tabular}{lcc}
@@
 the points plotted here are the highest likelihood samples.
-The uncertainties are the 16th and 84th percentiles.}
+The uncertainties are the 16th and 84th percentiles.
+In many cases, the uncertainties are under-estimated.
+    }
 \label{fig:compare_mcmc}

### 2017-01-31 documents/ms.tex
small changes to text

@@
 
-\item The flexibility of this GP model allows for insidious kinds of
-posterior multimodality and ``over-fitting''--like behavior.  For example,
-if $l$ is small,  the non-periodic factor in the covariance kernel may
-dominate, allowing for a good fit to the data without requiring any
-periodic covariance structure---even if
-clear periodic structure is present.  Additionally, if $\Gamma$ is too large,
-then [explain in more technical terms that ``the GP gets way too wiggly and
-overfits the data without caring about the periodicity.'' ]
+\item The flexibility of this GP model allows for posterior multimodality and
+``over-fitting''--like behavior.
+For example, if $l$ is small,  the non-periodic factor in the covariance
+        kernel may dominate, allowing for a good fit to the data without
+        requiring any periodic covariance structure---even if clear periodic
+        structure is present.
+Additionally, if $\Gamma$ is too large, the GP model becomes extremely
+        flexible and can fit the data without varying the period.
 
@@
 there may be significant posterior density at half or twice the true period $P$.
-This is related to how the ACF method must guard against identifying periods
-at half or twice the true period.
 \end{itemize}
 
-We address the difficulty of computational speed in two distinct ways: subsampling
-the data and splitting the light curve into independent sections.
-To subsample, we randomly select 1/30th of the points in the full light curve (an
-average of $\sim$1.5 points per day).  This decreases the likelihood evaluation
-time by a factor of about 50, down to about 100\,ms.
-We then split the light curve into equal-sized
-chunks containing approximately 300 points per section (corresponding to about
-200 days), and evaluate the log-likelihood as the sum of the log-likelihoods
-of the individual sections (all using the same parameters ${\bm \theta}$).  This reduces
-computation time because the section-based likelihood evaluation scales as $mn\ln(n)^2$, where $n$ is the number of data points per section and $m$ is the number of sections.
-This method further reduces computation time for a typical light curve (subsampled
-by a factor of 30) by about a factor of two, to about 50\,ms.
-
-In order to manage the great flexibilty of the GP model and focus on
-reliably retrieving the period parameter, we also impose priors
-on the non-period GP parameters that we
-learned during the development of this method.
-Beginning with very broad priors (uniform in log space between
--20 and 20) on all the parameters, we first explored what the
-typical values of the non-period parameters tended to be when the
-injected periods were successfully recovered.  We also experimented with
-constraining the allowed ranges of the parameters, after discovering
-that some regions of parameter space (such as large values of $A$ and $\Gamma$
-and small values of $l$) tended to allow fits that ignored the desired
-periodicity.  The priors and bounds we used for this work are listed in
-table \ref{tab:priors}.  We note that we do not have any specific
-quantitative justification for the use of these priors,
-except that they produce good results for this particular test set of
-light curves.  We thus caution that other data sets might have different
-noise properties, potentially requiring modifications to these priors
-on the non-period parameters (e.g., see Section \ref{sec:kepler}).  
-For period, we tried two different methods:
-an uninformed (log-flat) prior between 0.5\,d and 100\,d, and an ACF-informed
-prior (details in Appendix.)
-% We note that our default prior for period is
-% log-uniform between the bounds, but that we were also able to
-% further improve our results by using an ACF-inspired prior.
+We address the difficulty of computational speed in two distinct ways:
+subsampling the data and splitting the light curve into independent sections.
+To subsample, we randomly select 1/30th of the points in the full light curve
+(an average of $\sim$1.5 points per day).
+This decreases the likelihood evaluation time by a factor of about 50, down to
+about 100\,ms.
+We then split the light curve into equal-sized chunks containing approximately
+300 points per section (corresponding to about 200 days), and evaluate the
+log-likelihood as the sum of the log-likelihoods of the individual sections
+(all using the same parameters ${\bm \theta}$).
+This reduces computation time because the section-based likelihood evaluation
+scales as $mn\ln(n)^2$, where $n$ is the number of data points per section and
+$m$ is the number of sections.
+This method further reduces computation time for a typical light curve
+(subsampled by a factor of 30) by about a factor of two, to about 50\,ms.
+
+In order to manage the flexibilty of the GP model and focus on reliably
+retrieving the period parameter, we also impose priors on the non-period GP
+parameters, learned during the development of this method.
+Beginning with very broad priors (uniform in log space between -20 and 20) on
+all the parameters, we first explored the typical values of the non-period
+parameters when the injected periods were successfully recovered.
+We also experimented with constraining the allowed ranges of the parameters
+after discovering that some regions of parameter space (such as large values
+of $A$ and $\Gamma$ and small values of $l$) tended to allow fits that ignored
+the desired periodicity.
+The priors and bounds we used for this work are listed in table
+\ref{tab:priors}.
+We note that we do not have any specific quantitative justification for the
+use of these priors, except that they produce good results for this particular
+test set of light curves.
+We thus caution that other data sets might have different noise properties,
+potentially requiring modifications to these priors on the non-period
+parameters (e.g., see Section \ref{sec:kepler}).
+For the period prior, we tried two different methods: an uninformed (log-flat)
+prior between 0.5\,d and 100\,d, and an ACF-informed prior (details in
+Appendix.)
+% We note that our default prior for period is log-uniform between the bounds,
+% but that we were also able to further improve our results by using an
+% ACF-inspired prior.
 

### 2017-01-31 gprotation/GProt.py

@@ -10,7 +10,8 @@ import os
 import time
 import h5py
 
-import emcee3
+import emcee
+# import emcee3
 from gatspy.periodic import LombScargle
 
 from GProtation import make_plot, MyModel

### 2017-01-31 gprotation/GProtation.py

@@ -9,7 +9,8 @@ import os
 
 import george
 from george.kernels import ExpSine2Kernel, ExpSquaredKernel, WhiteKernel
-import emcee3
+# import emcee3
+import emcee
 import corner
 
 

### 2017-01-31 gprotation/recover_suzannes.py

@@ -82,10 +82,10 @@ def gp_fit(x, y, yerr, sid, RESULTS_DIR):
 
 def recover(i):
 
-    RESULTS_DIR = "results_emcee3_move"
-    DATA_DIR = "../code/simulations/kepler_diffrot_full/final"
+    RESULTS_DIR = "acf_results"
+    DATA_DIR = "kepler_diffrot_full/final"
 
-    DIR = "../code/simulations/kepler_diffrot_full/par/"
+    DIR = "kepler_diffrot_full/par/"
     truths = pd.read_csv(os.path.join(DIR, "final_table.txt"), delimiter=" ")
     m = truths.DELTA_OMEGA.values == 0
 
@@ -96,12 +96,12 @@ def recover(i):
     yerr = np.ones_like(y) * 1e-5
 
     fit = gp.fit(x, y, yerr, sid, RESULTS_DIR)
-    fit.gp_fit(burnin=1000, nwalkers=20, nruns=10, full_run=1000, nsets=2)
+    # fit.gp_fit(burnin=1000, nwalkers=20, nruns=10, full_run=1000, nsets=2)
 
 
 if __name__ == "__main__":
 
-    DIR = "../code/simulations/kepler_diffrot_full/par/"
+    DIR = "kepler_diffrot_full/par/"
     truths = pd.read_csv(os.path.join(DIR, "final_table.txt"), delimiter=" ")
     m = truths.DELTA_OMEGA.values == 0
 

### 2017-01-31 gprotation/simple_acf.py

@@ -56,6 +56,10 @@ def simple_acf(x, y):
     m, b = np.linalg.solve(ATA, np.dot(AT, acf_smooth))
     acf_smooth -= m*lags + b
 
+    # Cut off the first half a day
+    m = lags > .5
+    acf_smooth, lags = acf_smooth[m], lags[m]
+
     # find all the peaks
     peaks = np.array([i for i in range(1, len(lags)-1)
                      if acf_smooth[i-1] < acf_smooth[i] and

### 2017-02-01 documents/ms.tex
updating discussion, part of the conclusion and corner plot

@@
 \newcommand{\numax}{$\nu_{\mathrm{max}}$}
-\newcommand{\acfRMS}{9.27}
-\newcommand{\pgramRMS}{1.25}
+\newcommand{\acfRMS}{0.60}
+\newcommand{\pgramRMS}{0.98}
 \newcommand{\mcmcRMS}{0.46}
@@
 
-\title{Inferring stellar rotation periods using Gaussian processes}
+\title{Inferring probabilistic stellar rotation periods using Gaussian
+processes}
 
@@
 hyperparameters.
-We used \naigrain\ simulated light curves with a range of rotation periods
-to test the GP model.
-We attempted to recover the rotation periods using three methods: our GP
-method, a sine-fitting periodogram method and an AutoCorrelation Function
-(ACF) method.
-The posterior PDF of the rotation period parameter was sampled using the
-affine invariant ensemble Markov Chain Monte Carlo (MCMC) sampler {\tt emcee}
+We use \naigrain\ simulated light curves with a range of rotation periods to
+    test the GP model and compare our results with a sine-fitting periodogram
+    method and an AutoCorrelation Function (ACF) method.
+The posterior PDFs of the parameters were sampled using the affine invariant
+    ensemble Markov Chain Monte Carlo (MCMC) sampler {\tt emcee3}
     \citep{Foreman-Mackey2013}, and the GP operations were performed using the
     {\tt george} python package \citep{George}.
-Rotation periods inferred via this method are more precise and accurate than
-the periodogram and ACF methods.
-Furthermore, rotation periods are inferred probabilistically, allowing them to
-be easily incorporated into follow-on studies; inferring stellar ages or
-characterising star-planet interactions for example.
-In addition, unlike the sine-fitting periodogram and ACF methods, the
-posterior PDF samples allow for proper characterisation of uncertainties.
-This method eliminates the need to define a heuristic detection criterion such
-as are often used in the literature: a light curve that does not contain a
-rotation signal will return the prior PDF.
-Furthermore, the ACF method is not applicable to unevenly spaced data,
-therefore it is inappropriate for the ~$10^7$ stellar light curves expected
-to be produced by \LSST\ unlike the GP method.
+We demonstrate that rotation periods inferred via this method are more
+    accurate than the periodogram and ACF methods.
+Furthermore, the rotation periods are inferred probabilistically, allowing
+    them to be easily incorporated into follow-on studies; inferring stellar
+    ages or characterising star-planet interactions for example.
+% Thehe ACF method is not applicable to unevenly spaced data,
+% therefore it is inappropriate for the ~$10^7$ stellar light curves expected
+% to be produced by \LSST\ unlike the GP method.
 % Furthermore, the improvement is expected to be even more dramatic when applied
@@
 The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
-The RMS scatter about the $x=y$ line for rotation periods measured using the
-ACF method is \acfRMS.
+The Median Absolute Deviation (MAD) of the ACF-recovered periods is \acfRMS\
+in log-space (1.8 days).
 The agreement between the injected and recovered rotation periods is good in
@@
 periods in figure \ref{fig:pgram_compare}.
-The RMS scatter for rotation periods measured using the LS periodogram
-technique is \pgramRMS.
-The longer-period measurements provide the dominant contribution to this RMS.
+The MAD for rotation periods measured using the LS periodogram technique is
+\pgramRMS\ in log-space (2.7 days).
 % We believe that the unsuccessful recoveries occurring preferentially for
@@
 prior between 0.5\,d and 100\,d, and an ACF-informed prior (details in
-Appendix.)
+Appendix).
 % We note that our default prior for period is log-uniform between the bounds,
@@
 In many cases, the uncertainties are under-estimated.
+The ACF-informed prior on rotation period used to generate these results is
+    described in the appendix.
+The MAD of these results is 0.38 in log-space, \ie\ 1.46 days.
     }
-\label{fig:compare_mcmc}
+\label{fig:compare_mcmc_acfprior}
 \end{center}
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/comparison_noprior.pdf}
+\caption{The `true' rotation periods used to generate \naigrain\
+simulated light curves vs the rotation periods measured using the GP
+technique with no .
+Since the posterior PDFs of rotation periods are often non-Gaussian,
+the points plotted here are the highest likelihood samples.
+The uncertainties are the 16th and 84th percentiles.
+In many cases, the uncertainties are under-estimated.
+An uninformative prior, flat in the natural log of the rotation period between
+    0.5 and 100 days was used to generate these results.
+The MAD of these results is 0.48 in log-space, \ie\ 1.61 days.
+    }
+\label{fig:compare_mcmc_noprior}
+\end{center}
+\end{figure*}
+
 Figures \ref{fig:compare_mcmc_noprior} and \ref{fig:compare_mcmc_acfprior}
-summarize the results of this MCMC fitting
-procedure compared to the injected ``true'' stellar rotation periods, for the
-uninformed and ACF-informed priors on period, respectively.
+summarize the results of this MCMC fitting procedure compared to the injected
+`true' stellar rotation periods, for the uninformed and ACF-informed priors on
+period, respectively.
+Both the informative and uninformative prior versions of the GP method perform
+better than the ACF and periodogram methods, with MADs of 0.38 and 0.48,
+respectively.
 The marginal posterior distributions of the QP kernel hyperparameters for the
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/0002_triangle.png}
-\caption{(Place holder - to be updated.
-Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
+\includegraphics[width=6in, clip=true]{figures/25.png}
+\caption{Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
 additional white noise term added to the diagonal elements of the covariance
@@
 parameter will be non-zero.
-The true rotation period of this star is ... ($\exp(...)$) days.}
+    The green line in the period panel shows the injected period.}
 \label{fig:gp_posteriors}
@@
 regarding the setting the priors and ranges for the GP hyperparameters.
->>>>>>> 265508362e5e7a01a39bdd616c754c276059397c
 
@@
 
-The main drawback of the GP method is computation time.
-Because GPs are so expensive to compute, it is necessary to come up with
-shortcuts in order to perform inference on \kepler\ light curves, each of
-which compromises accuracy.
-These shortcuts may only be necessary when working with a large number of
-stars (of order hundreds or thousands).
-If only interested in single or small numbers of targets it may be possible to
-reduce the extremity of these shortcuts or to avoid them altogether.
-The shortcuts that we employ here include:
-\begin{itemize}
-\item{Initialisation.
-The choice of initial parameter values makes a difference to computation time:
-the closer they are to the maximum likelihood parameters the shorter the MCMC
-burn-in time.
-Using the ACF period to initialise is not ideal as results will not be fully
-independent of the ACF period unless MCMC chains are run for an infinitely
-long time.}
-\item{Subsampling.
-Not only is some information always lost by subsampling, but the choice of
-subsampling frequency will also effect the results.
-We choose to retain an average of only 4 data points per day, based on the
-range of rotation periods we are interested in.}
-\item{Priors.
-We choose to place a Gaussian prior in log-space over the rotation period
-parameter where the standard deviation is twice the log of the ACF rotation
-period.
-Unfortunately however, if the true period is far different to the ACF period,
-the MCMC chains will take a long time to reach the correct value.
-In practice we found that the introduction of a Gaussian prior improved
-results compared with a broad prior that was uniform in log-space.}
-\end{itemize}
-
-\subsection{Initialisation}
-Using the ACF to initialise the MCMC chains is not ideal because, of course,
-one becomes reliant on the assumption that the ACF period is close to the true
-period.
-Of course, if you had infinite CPU time, this would not be a problem as you
-would eventually sample the entire posterior PDF of the period parameter,
-however in practice this is likely to be an issue.
-The only way to get around this problem is to run the MCMC chains for as long
-as possible, or, alternatively to use a sampler that is designed to move
-around the parameter space much more quickly than {\tt emcee}.
-% We implemented a version of our code using nested sampling but the results
-% were not as good as the {\tt emcee} method.
-% We believe this to be due to the initialisation --- there is no initialisation
-% in nested sampling, samples are drawn directly from the prior.
-% In a problem with a multimodal posterior PDF such as inferring periods,
-% initialising MCMC chains close to the maximum likelihood value may lead to a
-% more rapid convergence.
-We chose to use the ACF periods rather than the periodogram periods to
-initialise the results since there were fewer large outliers.
-
-\subsection{Kernel function choice and interpretation of hyper-parameters}
+The uncertainties produced by the GP method are more representative than those
+produced by the ACF and periodogram methods because the posterior PDF of the
+period is explored using MCMC.
+Whilst they are an improvement, unfortunately these uncertainties are still
+underestimated in the majority cases.
+Of the rotation periods recovered from the simulated light curves, only 25\%
+of the measured periods lay within 1$\sigma$ of the true period.
+50\% lay within 2$\sigma$ and 66\% within 3$\sigma$.
+The largest outlier was 114$\sigma$ away from the true value.
+In the majority of these cases the uncertainties are underestimated due to the
+multi-modal nature of the period posterior PDF.
+Because the posteriors are multi-modal they are difficult to sample and,
+without running the MCMC for infinitely long, the samples will not necessarily
+always cover the full range of periods with significant power.
+The underestimated uncertainties can also be attributed to the model choice.
+Although a GP model is more appropriate than a sinusoid, it is still an
+imperfect, effective model.
+A perfect physical model of the star would produce better results with more
+representative uncertainties.
+
 The QP kernel function represents a simplistic effective model of a stellar
@@
 rotation periods of the three techniques.
+As expected, it provides a large improvement over the periodogram method and a
+moderate one over the ACF method.
+The periodogram method performs poorly because a sinusoid is not a perfect
+model choice for stellar light curves --- they do not always display
+sinusoidal behaviour.
+On the other hand, the ACF method does not rely on any model choice.
+
 In addition, we compared the rotation periods of \kepler\ objects of
@@
 The good agreement between the two sets of results demonstrates that the GP
-method works well on real \kepler\ data and is, in most cases robust to the
-presents of transits and eclipses in the light curves.
-
-Although rotation period uncertainties calculated using the GP method are
-better estimates than those calculated using the ACF method, they are still
-imperfect estimates of the intrinsic uncertainties.
+method works well on real \kepler\ data.
+
+The GP method provides uncertainties because it produces posterior PDF samples
+but they are still imperfect estimates of the intrinsic uncertainties.
 This is because the posterior PDF of the rotation period parameter is, in many
 cases, multi-modal.
-If it were possible to run an MCMC sampler for an infinite number of steps, we
-would the entire posterior PDF would eventually be sampled, producing an
-accurate uncertainty measurement.
+If it were possible to run an MCMC sampler for an infinite number of steps,
+the entire posterior PDF would eventually be sampled, producing an accurate
+uncertainty measurement.
 Practically speaking however, these imperfect estimates may be the best
@@
 PDF samples of rotation periods, allowing characterisation of uncertainties.
-Note that the posterior PDFs are often non-Gaussian, being bi- or multimodal.
-In these cases it is therefore misleading to report median values or 16th and
-84th percentiles as uncertainties.
+The posterior PDFs are often non-Gaussian, being bi- or multimodal.
 Rather, it is better to incorporate posterior PDF samples into analyses which
@@
 stellar ages or characterising star-planet interactions for example.
-
-The main drawback of the GP method is the computational cost.
 

### 2017-02-01 code/plot_koi.py

@@ -36,9 +36,8 @@ def load_and_plot(data, DATA_DIR, RESULTS_DIR):
     lnxerr = xerr/x
     amp = data.R_var.values
     # remove binaries.
-    l = (recovered > 0) * (data.koi_id.values != 197) * \
+    l = np.isfinite(x) * (recovered > 0) * (data.koi_id.values != 197) * \
         (data.koi_id.values != 279)
-
 
     plt.clf()
     xs = np.log(np.linspace(0, 100, 100))
@@ -53,23 +52,6 @@ def load_and_plot(data, DATA_DIR, RESULTS_DIR):
                 edgecolor=".5", cmap="GnBu_r", vmin=min(np.log(amp[l])),
                 vmax=max(np.log(amp[l])), s=20, zorder=2, lw=.2)
 
-    import kplr
-    import kepler_data as kd
-    client = kplr.API()
-    m = 20 < (np.abs(x[l] - recovered[l])/x[l] * 100)
-    for i, koi in enumerate(data.koi_id.values):
-        print(data.koi_id.values[l][m][i], x[l][m][i], recovered[l][m][i])
-        star = client.koi("{}.01".format(str(data.koi_id.values[l][m][i])))
-        # star.get_light_curves(fetch=True, shortcadence=False)
-        print(star.kepid)
-        kid = str(int(star.kepid)).zfill(9)
-        fname = "/Users/ruthangus/.kplr/data/lightcurves/{}".format(kid)
-        x, y, yerr = kd.load_kepler_data(fname)
-        plt.clf()
-        plt.plot(x, y, "k.")
-        plt.xlim(150, 300)
-        plt.savefig("{}".format(data.koi_id.values[l][m][i]))
-
     plt.ylim(0, 4)
     plt.xlim(0, 4)
     cbar = plt.colorbar()
@@ -77,9 +59,40 @@ def load_and_plot(data, DATA_DIR, RESULTS_DIR):
     plt.xlabel("$\ln(\mathrm{McQuillan~{\it et~al.,}~2013~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
     plt.savefig("comparison_koi")
-    plt.savefig("../documents/figures/comparison_koi.pdf")
+    # plt.savefig("../documents/figures/comparison_koi.pdf")
 
-    print((np.median((x[l] - recovered[l])**2))**.5)
[…]

### 2017-02-01 gprotation/GProt.py

@@ -16,6 +16,7 @@ from gatspy.periodic import LombScargle
 
 from GProtation import make_plot, MyModel
 import simple_acf as sa
+import acf
 
 DATA_DIR = "data/"
 RESULTS_DIR = "results/"
@@ -177,7 +178,7 @@ def evaluate_convergence(flat, mean_acorr, diff_threshold=0,
     return converged, mean_acorr, mean_ind, mean_diff
 
 
-def calc_p_init(x, y, yerr, id, RESULTS_DIR, clobber=False):
+def calc_p_init(x, y, yerr, id, RESULTS_DIR, clobber=True):
     fname = os.path.join(RESULTS_DIR, "{0}_acf_pgram_results.txt".format(id))
     if not clobber and os.path.exists(fname):
         print("Previous ACF pgram result found")
@@ -191,6 +192,9 @@ def calc_p_init(x, y, yerr, id, RESULTS_DIR, clobber=False):
         print("Calculating ACF")
         acf_period, acf, lags, rvar = sa.simple_acf(x, y)
         err = .1 * acf_period
+        # acf_period, err, lags, acf = acf.corr_run(x, y, np.ones_like(y)*1e-5,
+                                                  # id, RESULTS_DIR)
+
         plt.clf()
         plt.plot(lags, acf)
         plt.axvline(acf_period, color="r")

### 2017-02-01 gprotation/comparison_plots.py

@@ -15,7 +15,8 @@ plotpar = {'axes.labelsize': 18,
            'text.usetex': True}
 plt.rcParams.update(plotpar)
 
-DATA_DIR = "../code/simulations/kepler_diffrot_full/final"
+DATA_DIR = "Users/ruthangus/projects/GProtation/" \
+    "code/kepler_diffrot_full/final"
 
 def calc_p_init(x, y, yerr, id, RESULTS_DIR):
     fname = os.path.join(RESULTS_DIR, "{0}_acf_pgram_results.txt".format(id))
@@ -232,7 +233,7 @@ def acf_plot(truths, DIR):
     plt.xlabel("$\ln(\mathrm{Injected~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period~ACF~Method})$")
     plt.savefig(os.path.join(DIR, "compare_acf.pdf"))
-    return (np.median((acfs - true)**2))**.5
+    return np.median(np.abs(acfs - true))
 
 def pgram_plot(truths, DIR):
     """
@@ -266,23 +267,16 @@ def pgram_plot(truths, DIR):
     plt.xlabel("$\ln(\mathrm{Injected~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period~LS~Periodogram~method})$")
     plt.savefig(os.path.join(DIR, "compare_pgram.pdf"))
-    return (np.median((pgram - true)**2))**.5
+    return np.median(np.abs(pgram - true))
 
 if __name__ == "__main__":
 
-    DIR = "../code/simulations/kepler_diffrot_full/par/"
+    DIR = "/Users/ruthangus/projects/GProtation/code/kepler_diffrot_full/par/"
     truths = pd.read_csv(os.path.join(DIR, "final_table.txt"), delimiter=" ")
 
     # remove 17 for now
 #     m = truths.N.values != 17
 #     truths = truths.iloc[m]
 
-#     print("mcmc sigma rms = ", mcmc_plots(truths, "results_emcee3"))
-    print("acf sigma rms = ", acf_plot(truths, "results_emcee3_move"))
-    print("pgram sigma rms = ", pgram_plot(truths, "results_emcee3_move"))
-#     print("mcmc sigma rms = ", mcmc_plots(truths, "results_emcee2"))
-#     print("acf sigma rms = ", acf_plot(truths, "results_emcee2"))
-#     print("pgram sigma rms = ", pgram_plot(truths, "results_emcee2"))
[…]

### 2017-02-01 gprotation/recover_suzannes.py

@@ -105,5 +105,7 @@ if __name__ == "__main__":
     truths = pd.read_csv(os.path.join(DIR, "final_table.txt"), delimiter=" ")
     m = truths.DELTA_OMEGA.values == 0
 
-    pool = Pool()
-    results = pool.map(recover, range(len(truths.N.values[m])))
+    for i in range(len(truths.N.values[m])):
+        recover(i)
+    # pool = Pool()
+    # results = pool.map(recover, range(len(truths.N.values[m])))

### 2017-02-01 gprotation/simple_acf.py

@@ -50,11 +50,11 @@ def simple_acf(x, y):
     # ditch the first point
     acf_smooth, lags = acf_smooth[1:], lags[1:]
 
-    # fit and subtract straight line
-    AT = np.vstack((lags, np.ones_like(lags)))
-    ATA = np.dot(AT, AT.T)
-    m, b = np.linalg.solve(ATA, np.dot(AT, acf_smooth))
-    acf_smooth -= m*lags + b
+    # # fit and subtract straight line
+    # AT = np.vstack((lags, np.ones_like(lags)))
+    # ATA = np.dot(AT, AT.T)
+    # m, b = np.linalg.solve(ATA, np.dot(AT, acf_smooth))
+    # acf_smooth -= m*lags + b
 
     # Cut off the first half a day
     m = lags > .5

### 2017-02-01 documents/ms.tex
tidying up a bit, adding more words to the conclusion

@@
 
-We calculated an ACF for each light curve following the method of
+We measured an ACF-based period for each light curve following the method of
 \citet{Mcquillan2013}.
@@
 The agreement between the injected and recovered rotation periods is good in
-general, however twice the true rotation period is often found.
-In addition, some rotation periods are drastically over- and underestimated.
+general, however some rotation periods are drastically over- and
+underestimated.
 Of the points clustered close to the $x=y$ line, more fall slightly below it
@@
 the ACF.
-ACFs of stellar light curves are similar to the QP kernel function in equation
-\ref{eq:QP}: periodic functions added to decaying exponentials.
+ACFs of stellar light curves have similar functional forms to the QP kernel
+function in equation \ref{eq:QP}: periodic functions added to decaying
+exponentials.
 In such functions the peak positions can be shifted towards the left, the
@@
     Several light curves have over-estimated rotation periods and some
-    are drastically underestimated.
-    In some cases the recovered period is twice the true period used to
-    simulate the light curve.}
+    are drastically underestimated.}
 \label{fig:compare_acf}
@@
 \pgramRMS\ in log-space (2.7 days).
-% We believe that the unsuccessful recoveries occurring preferentially for
-% shorter rotation periods are produced by additional, longer timescale
-% variations in some of the light curves.
-% Specifically, by fluctuations produced by changes in the total spot coverage
-% on the stellar surface.
-% These fluctuations can occasionally be larger in magnitude than those produced
-% by the rotational signal itself and are present in the majority of outlying
-% light curves that occupy the top left of figure
-% \ref{fig:pgram_compare_noise-free}.
+There is a large number of drastically overestimated periods recovered using
+the periodogram.
+Excess power at long periods can be generated by long-term trends that are not
+related to rotation.
+This problem becomes exacerbated when instrumental noise is also present in
+light curves.
 
@@
 
-Figures \ref{fig:compare_mcmc_noprior} and \ref{fig:compare_mcmc_acfprior}
+Figures \ref{fig:compare_mcmc_acfprior} and \ref{fig:compare_mcmc_noprior}
 summarize the results of this MCMC fitting procedure compared to the injected
@@
 Both the informative and uninformative prior versions of the GP method perform
-better than the ACF and periodogram methods, with MADs of 0.38 and 0.48,
-respectively.
+better than the ACF and periodogram methods, with respective MADs of 0.38 and
+0.48, compared to \acfRMS\ (ACF) and \pgramRMS\ (periodogram).
 The marginal posterior distributions of the QP kernel hyperparameters for the
-example simulated light curve with in figure \ref{fig:demo_lc}, are shown in
+example simulated light curve in figure \ref{fig:demo_lc}, are shown in
 figure \ref{fig:gp_posteriors}.
@@
 \includegraphics[width=6in, clip=true]{figures/25.png}
-\caption{Marginal posterior PDFs of the QP GP model parameters. $\sigma$ is an
+\caption{Marginal posterior PDFs of the QP GP model parameters, fit to the
+    light curve in figure \ref{fig:demo_lc}.
+$\sigma$ is an
 additional white noise term added to the diagonal elements of the covariance
@@
 
-% At this point it is important to highlight the following caveat.
-% Our method has been tuned to reproduce the rotation periods of the
-% \citet{Aigrain2015} simulated light curves.
-% These light curves were simulated using a very good but still imperfect model.
-% The input parameters and their distributions were chosen to be as realistic as
-% possible given current observations.
-% Unfortunately however, there are limited data on some of these parameters for
-% stars other than the Sun: for example the distrubution of spot lifetimes or
-% activity cycle length.
-% Ensuring that we can recover the rotation periods of these simulated light
-% curves does not necessarily ensure that we can recover the rotation periods of
-% real \kepler\ light curves with equal accuracy and precision.
-% Even using priors on the QP kernel parameters that have been derived from the
-% \citet{Aigrain2015} sample may be dangerous.
-% This caveat should be taken into consideration when this method is applied to
-% real data.
-
 \section{Discussion}
@@
 Of the rotation periods recovered from the simulated light curves, only 25\%
-of the measured periods lay within 1$\sigma$ of the true period.
-50\% lay within 2$\sigma$ and 66\% within 3$\sigma$.
-The largest outlier was 114$\sigma$ away from the true value.
+of the measured periods lie within 1 $\sigma$ of the true period.
+50\% lie within 2 $\sigma$ and 66\% within 3 $\sigma$.
+The largest outlier is 114 $\sigma$ away from the true value.
 In the majority of these cases the uncertainties are underestimated due to the
-multi-modal nature of the period posterior PDF.
-Because the posteriors are multi-modal they are difficult to sample and,
-without running the MCMC for infinitely long, the samples will not necessarily
-always cover the full range of periods with significant power.
+multi-modal nature of the period posterior PDFs which makes them difficult to
+sample.
 The underestimated uncertainties can also be attributed to the model choice.
@@
 parameter of interest.
-It may in fact be even {\it more} suited to modelling stellar
-light curves as it describes a Gaussian in frequency space.
-It is easy to imagine a differentially rotating star with a period that is a
-Gaussian in frequency space: the mean frequency would be the frequency at the
-most active latitude, at or near which spots spend the majority of their time
-and the tails would be occupied by spots that drift near the equator or poles.
-The main difference between this cosine and the QP function is that the cosine
-function allows negative covariances and the QP function does not.
+% It may in fact be even {\it more} suited to modelling stellar
+% light curves as it describes a Gaussian in frequency space.
+% It is easy to imagine a differentially rotating star with a period that is a
+% Gaussian in frequency space: the mean frequency would be the frequency at the
+% most active latitude, at or near which spots spend the majority of their time
+% and the tails would be occupied by spots that drift near the equator or poles.
+The main practical difference between this cosine and the QP function is that
+the cosine function allows negative covariances and the QP function does not.
 Is is realistic to allow negative covariances?
@@
 
-Clearly, stellar rotation periods are well represented by the $P$ parameter in
-our QP kernel function, as evidenced by the its impressive ability to recover
-the true rotation periods from the simulated light curves in this work.
-However, it may not be the {\it best} function to use.
-It is, after all, an approximation to the shape of the covariance matrix of a
-real stellar light curve.
-There may be an alternative function which is better suited to capturing
-stellar variability and is able to recover periods even more precisely than
-the QP kernel.
-There may also be an alternative function which is more physically motivated,
-that captures not just the rotation period but also (for example) the spot
-lifetime or differential surface rotation.
-This is beyond the scope of this work but we hope that these questions will be
-answered in the near future.
+% Clearly, stellar rotation periods are well represented by the $P$ parameter in
+% our QP kernel function, as evidenced by the its impressive ability to recover
+% the true rotation periods from the simulated light curves in this work.
+% However, it may not be the {\it best} function to use.
+% It is, after all, an approximation to the shape of the covariance matrix of a
+% real stellar light curve.
+% There may be an alternative function which is better suited to capturing
+% stellar variability and is able to recover periods even more precisely than
+% the QP kernel.
+% There may also be an alternative function which is more physically motivated,
+% that captures not just the rotation period but also (for example) the spot
+% lifetime or differential surface rotation.
+% This is beyond the scope of this work but we hope that these questions will be
+% answered in the near future.
 
@@
 This idea is being explored by another member of my research group.}
-\item{To develop a detection criterion to assess whether a rotation period was
-measured.
-Another general problem in rotation period inference is deciding whether a
-real rotation period was measured at all.
-Detection thresholds are usually set using the amplitudes of peaks in LS
-periodograms or ACFs in order to eliminate contaminants.
-Using a GP model it may be possible to do this via model selection: \ie\ are
-the data better described by a periodic model, or a non-periodic model?
-There is more than one way to implement such a model comparison, one way could
-be to use a sum of two kernel functions: one periodic and one non-periodic,
-each with its own amplitude parameter.}
+% \item{To develop a detection criterion to assess whether a rotation period was
+% measured.
+% Another general problem in rotation period inference is deciding whether a
+% real rotation period was measured at all.
+% Detection thresholds are usually set using the amplitudes of peaks in LS
+% periodograms or ACFs in order to eliminate contaminants.
+% Using a GP model it may be possible to do this via model selection: \ie\ are
+% the data better described by a periodic model, or a non-periodic model?
+% There is more than one way to implement such a model comparison, one way could
+% be to use a sum of two kernel functions: one periodic and one non-periodic,
+% each with its own amplitude parameter.}
 \item{To attempt to detect differential rotation.
@@
 
-\section{Conclusions}
+\section{Conclusion}
 
@@
 moderate one over the ACF method.
-The periodogram method performs poorly because a sinusoid is not a perfect
-model choice for stellar light curves --- they do not always display
-sinusoidal behaviour.
-On the other hand, the ACF method does not rely on any model choice.
-
+We find that the ACF method often slightly under-predicts the rotation period
+due to a subtlety in the standard peak detection algorithm.
+% The periodogram method performs poorly because a sinusoid is not a perfect
+% model choice for stellar light curves --- they do not always display
+% sinusoidal behaviour.
+% On the other hand, the ACF method does not rely on any model choice and
 In addition, we compared the rotation periods of \kepler\ objects of
@@
 
-The GP method provides uncertainties because it produces posterior PDF samples
-but they are still imperfect estimates of the intrinsic uncertainties.
+The GP method provides approximate uncertainties because it produces posterior
+PDF samples but they are still imperfect estimates of the intrinsic
+uncertainties.
 This is because the posterior PDF of the rotation period parameter is, in many
@@
 uncertainty measurement.
-Practically speaking however, these imperfect estimates may be the best
-currently achievable.
-
-A GP is a semi-parametric model that is well suited to signals that are
-non-sinusoidal and quasi-periodic.
-Not only this, unlike the other two methods, the GP method provides posterior
-PDF samples of rotation periods, allowing characterisation of uncertainties.
-The posterior PDFs are often non-Gaussian, being bi- or multimodal.
-Rather, it is better to incorporate posterior PDF samples into analyses which
-require uncertainty propagation.
-Because rotation periods are inferred probabilistically with the GP method,
-this allows them to be easily incorporated into follow-on studies; inferring
-stellar ages or characterising star-planet interactions for example.
+We find that only a quarter of the reported uncertainties are representative.
+
+The main aim of this work was to develop a rotation period inference method
+that is probabilistic.
+Probabilistic periods methods are necessary for hierarchical Bayesian
+inference and for performing population analysis.
+Using this method to generate a catalogue of probabilistic rotation periods
+could, for example, reveal the period distribution of (and therefore
+potentially the age distribution) of stars in the Milky Way.
+There is still some work to be done on the development of this model such that
+the uncertainty estimates are truly representative, but this new method
+provides a improvement upon previous methods in that regard.
 

### 2017-02-01 gprotation/comparison_plots.py
making histograms of residuals

@@ -233,7 +233,8 @@ def acf_plot(truths, DIR):
     plt.xlabel("$\ln(\mathrm{Injected~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period~ACF~Method})$")
     plt.savefig(os.path.join(DIR, "compare_acf.pdf"))
-    return np.median(np.abs(acfs - true))
+
+    return np.median(np.abs(acfs - true)), np.log(true) - np.log(acfs)
 
 def pgram_plot(truths, DIR):
     """
@@ -267,7 +268,9 @@ def pgram_plot(truths, DIR):
     plt.xlabel("$\ln(\mathrm{Injected~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period~LS~Periodogram~method})$")
     plt.savefig(os.path.join(DIR, "compare_pgram.pdf"))
-    return np.median(np.abs(pgram - true))
+
+    return np.median(np.abs(pgram - true)), np.log(true) - np.log(pgram)
+
 
 if __name__ == "__main__":
 
@@ -278,5 +281,19 @@ if __name__ == "__main__":
 #     m = truths.N.values != 17
 #     truths = truths.iloc[m]
 
-    print("acf sigma rms = ", acf_plot(truths, "acf_results"))
-    print("pgram sigma rms = ", pgram_plot(truths, "acf_results"))
+    acf_MAD, acf_resids = acf_plot(truths, "acf_results")
+    pgram_MAD, pgram_resids = pgram_plot(truths, "acf_results")
+    print("acf MAD = ", acf_MAD)
+    print("pgram MAD = ", pgram_MAD)
+    m1 = (-.5 < acf_resids) * (acf_resids < .5)
+    m2 = (-.5 < pgram_resids) * (pgram_resids < .5)
+    plt.clf()
+    plt.hist(pgram_resids[m2], 100, normed=True, alpha=.5, color="r",
+             label="pgram")
+    plt.hist(acf_resids[m1], 100, normed=True, alpha=.5, color="b",
+             label="ACF")
+    print(np.median(np.abs(acf_resids)))
+    print(np.median(np.abs(pgram_resids)))
+    plt.legend()
+    plt.xlim(-1, 1)
+    plt.savefig("acf_pgram_resids")

### 2017-02-02 documents/ms.tex
tdm abstract edits

@@
 quasi-periodic---spots are not static on the stellar surface and have finite
-lifetimes causing stellar flux variations to slowly shift in phase.
-A strictly periodic sinusoid is therefore not a representative model for the
-light curve of star with a rotational signal.
-Ideally, a physical model of the stellar surface would be conditioned on the
-data, however the parameters of such models can be highly degenerate
-\citep[\eg][]{Russell1906, Jeffers2009}.
-Instead, we use an appropriate {\it effective} model: a Gaussian Process (GP)
+lifetimes, causing stellar flux variations to slowly shift in phase.
+A strictly periodic sinusoid therefore cannot accurately model a 
+rotationally modulated stellar light curve.
+Physical models of the stellar surface also have many drawbacks preventing effective
+inference, such as having highly degenerate or high-dimensional parameter spaces
+\citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
+In this work, we introduce an appropriate {\it effective} model: a Gaussian Process (GP)
 with a quasi-periodic covariance kernel function.
-By modelling the light curve with a GP, a highly flexible semi-parametric
-function, we avoid the necessity of choosing a `best fitting' functional form,
-whilst sampling directly from the posterior Probability Density Function (PDF)
-of the periodic parameter and marginalising over the other kernel
-hyperparameters.
-We use \naigrain\ simulated light curves with a range of rotation periods to
-    test the GP model and compare our results with a sine-fitting periodogram
-    method and an AutoCorrelation Function (ACF) method.
-The posterior PDFs of the parameters were sampled using the affine invariant
-    ensemble Markov Chain Monte Carlo (MCMC) sampler {\tt emcee3}
-    \citep{Foreman-Mackey2013}, and the GP operations were performed using the
-    {\tt george} python package \citep{George}.
-We demonstrate that rotation periods inferred via this method are more
-    accurate than the periodogram and ACF methods.
-Furthermore, the rotation periods are inferred probabilistically, allowing
-    them to be easily incorporated into follow-on studies; inferring stellar
-    ages or characterising star-planet interactions for example.
+Using this highly flexible model, 
+we are able to sample from the posterior Probability Density Function
+(PDF) of the periodic parameter, marginalising over the other kernel hyperparameters
+using a Markov Chain Monte Carlo approach.
+To test this method's effectiveness, we infer rotation periods from 
+\naigrain\ simulated light curves, demonstrating that our results 
+    are more accurate than both a sine-fitting periodogram and an 
+    AutoCorrelation Function (ACF) method.
+We also demonstrate that it works effectively on real data, by 
+inferring rotation periods for \nkois\ \Kepler\ stars.
+Because this method delivers posteriors PDFs, 
+it will enable other studies involving stellar rotation, 
+such as inferring stellar
+ages or stellar obliquities in exoplanet systems, or characterising star-planet interactions.
 % Thehe ACF method is not applicable to unevenly spaced data,

### 2017-02-02 documents/ms.tex
removed references in abstract

@@
 Physical models of the stellar surface also have many drawbacks preventing effective
-inference, such as having highly degenerate or high-dimensional parameter spaces
-\citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
+inference, such as having highly degenerate or high-dimensional parameter spaces.
+% \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
 In this work, we introduce an appropriate {\it effective} model: a Gaussian Process (GP)

### 2017-02-02 documents/ms.tex
fixing #5

@@
 To sample the posterior in a way that is sensitive to potential multimodality,
-we use the \texttt{emcee3} MCMC sampler.
-We initialize 500 walkers with random samples from the prior and use a mix of
-different proposal distributions: specifically, a 2:2:1 mix of KDE,
-differential evolution, and DESnooker moves.
-[DFM: more words/citations here?]
+we use the \texttt{emcee3}\footnote{\url{https://github.com/dfm/emcee3}} MCMC
+sampler.
+\texttt{emcee3} is the successor to the \texttt{emcee} project
+\citep{Foreman-Mackey2013} that includes a suite of ensemble MCMC proposals
+that can be combined to efficiently sample more distributions than the stretch
+move \citep{Goodman2010} in \texttt{emcee}.
+For this project, we use a weighted mixture of three proposals.
+First, we include a proposal based on the \texttt{kombine}
+package\footnote{\url{https://github.com/bfarr/kombine}} (Farr \& Farr, in
+prep.) where a kernel density estimate (KDE) of the density represented by the
+complementary ensemble is used as the proposal for the other walkers.
+The other two proposals are a ``Differential Evolution (DE) MCMC'' proposal
+\citep{terBraak2006, Nelson2014} and the ``snooker'' extension of DE
+\citep{terBraak2008}.
+
+We initialize 500 walkers with random samples from the prior and use a
+weighted mixture of the KDE, DE, and snooker proposals with weights of 0.4,
+0.4, and 0.2 respectively.
 We run fifty steps of the sampler at a time, checking for convergence after

### 2017-02-02 code/plot_tim.py
using MAP instead of median

@@ -1,0 +1,127 @@
+import numpy as np
+import os
+import pandas as pd
+import matplotlib.pyplot as plt
+
+plotpar = {'axes.labelsize': 18,
+           'font.size': 10,
+           'legend.fontsize': 15,
+           'xtick.labelsize': 18,
+           'ytick.labelsize': 18,
+           'text.usetex': True}
+plt.rcParams.update(plotpar)
+
+prior = True
+
+if prior:
+    RESULTS_DIR = "results_24_01"
+else:
+    RESULTS_DIR = "results_noprior"
+truths = pd.read_csv("final_table.txt", delimiter=" ")
+
+# remove differential rotators and take just the first 100
+m = truths.DELTA_OMEGA.values == 0
+truths = truths.iloc[m]
+# ml = np.arange(100)
+# truths = truths.iloc[ml]
+
+recovered = np.zeros(len(truths.N.values))
+errp, errm = [np.zeros(len(truths.N.values)) for i in range(2)]
+lnerrp, lnerrm = [np.zeros(len(truths.N.values)) for i in range(2)]
+for i, id in enumerate(truths.N.values):
+    fn = os.path.join(RESULTS_DIR, "{}.h5".format(id))
+    if os.path.exists(fn):
+        # store = pd.HDFStore(fn)
+        # print(store)
+        df = pd.read_hdf(fn, key="samples")
+        phist, bins = np.histogram(df.ln_period.values, 100)
+        ln_p = bins[phist == max(phist)][0]
+        # ln_p = np.median(df.ln_period.values)
+        recovered[i] = np.exp(ln_p)
+        lnerrp[i] = np.percentile(df.ln_period.values, 84) - ln_p
+        lnerrm[i] = ln_p - np.percentile(df.ln_period.values, 16)
+        errp[i] = np.exp(lnerrp[i]/ln_p)
+        errm[i] = np.exp(lnerrm[i]/ln_p)
+
+x = .5 * (truths.P_MIN.values + truths.P_MAX.values)
+amp = truths.AMP.values
+l = recovered > 0
+
+plt.clf()
+xs = np.log(np.linspace(0, 55, 100))
+plt.plot(xs, xs, "-", color=".7", zorder=0)
+plt.plot(xs, xs + 2./3, "--", color=".7", zorder=0)
[…]

### 2017-02-02 documents/ms.tex

--- 2017-02-02 documents/ms.tex
@@
 rotate in and out of view.
-The non-sinusoidal quality is caused by the complicated surface spot patterns
-and the quasi-periodicity is produced both by the finite lifetimes of these
-active regions and the presence of differential rotation on the stellar
-surface.
-A strictly periodic sinusoid is therefore not a good model for stellar light
-curves.
-In an ideal world, a physical model of the stellar surface would be
-conditioned on the data.
-A physically realistic model would perfectly capture the complexity of shapes
+Complicated surface spot patterns produce the non-sinusoidal quality,
+and the finite lifetimes of these active regions and the presence of 
+differential rotation on the stellar surface produce quasi-periodicity.
+A strictly periodic sinusoid therefore does not well describe 
+rotationally modulated stellar light curves.
+A physically realistic model of the stellar surface 
+would perfectly capture the complexity of shapes
 within stellar light curves as well as the quasi-periodic nature, allowing for
-extremely precise probabilistic period recovery.
+extremely precise probabilistic period recovery when conditioned on the data.
 However, such physical models require many free parameters in order to
 accurately represent a stellar surface and some of these parameters are
-extremely degenerate \citep[\eg][]{Russell1906, Jeffers2009}.
+extremely degenerate \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
 In addition to global stellar parameters such as inclination and rotation
@@
 latitude, size, temperature and lifetime.
-Considering that many stars have on the order of hundreds of spots, the number
-of free parameters quickly becomes unwieldy, especially if the posterior PDFs
-of these parameters are explored with MCMC.
+Considering that many stars likely have hundreds of spots, the number
+of free parameters quickly becomes unwieldy, especially to explore
+their posterior PDFs with MCMC.
 Simplified spot models, such as the one described in \citet{Lanza2014} where
-only two spots are modelled, have produced successful results, however these
-simplified models sacrifice some precision due to lack of model flexibility.
-Instead of using a physical model for stellar light curves, we choose to use
-an {\it effective} model: one which captures the behaviour but is not
-physically motivated, although the parameters of this model may be {\it
-interpreted} as physical ones.
-An ideal effective model for the light curve of a spotted, rotating star is
-one with a small number of non-degenerate parameters that is flexible enough
-to perfectly capture non-sinusoidal, QP behaviour.
-These requirements are fulfilled by a Gaussian process (GP) model.
+only two spots are modelled, have produced successful results; however, such 
+relatively inflexible models sacrifice precision.
 
@@
 peaks in a Lomb-Scargle \citep{Lomb1976, Scargle1982} (LS) periodogram
-\citep[e.g.][]{Reinhold2013}, Auto-Correlation Functions (ACFs)
+\citep[e.g.][]{Reinhold2013}, Auto-Correlation Functions (ACFs),
 \citep{Mcquillan2013} and wavelet transforms \citep{Garcia2014}.
 The precision of the LS periodogram and wavelet methods are limited by the
-suitability of the model choice.
-A sinusoid is used in the case of the LS periodogram and the wavelet method
-relies on a choice of mother wavelet that is assumed to describe the data over
-a range of transpositions \citep[see, \eg][]{Carter2010}.
+suitability of the model choice: a sinusoid for the LS periodogram, 
+and a choice of mother wavelet, assumed to desribe the data
+over a range of transpositions \citep[see, \eg][]{Carter2010}, 
+for the wavelet method.
 In contrast, the ACF method is much better suited to signals that are
 non-sinusoidal.
-In fact, it doesn't matter what shape the signal is: as long as it is
-approximately periodic the ACF will display peaks located at the rotation
-period.
-A drawback of the ACF method however, is that it requires data to be
+In fact, as long as it is approximately periodic the ACF will 
+display a peak at the rotation period, no matter the signal shape.
+A drawback of the ACF method, however, is that it requires data to be
 evenly-spaced\footnote{\citet{Edelson1988} describe a method for computing
-ACFs for unevenly-spaced data.} which is not exactly the case with \Kepler\
+ACFs for unevenly-spaced data.}, which is not exactly the case with \Kepler\
 light curves (although in many cases it can be approximated as uniformly
-sampled) and not nearly the case for the future large-scale survey instrument,
+sampled) and not nearly the case for the future large-scale survey instrument
 \LSST.
 An ACF is also an operation performed on the data, not a generative model of
-the data and is therefore not inherently probabilistic.
-For this reason it is very difficult to estimate the uncertainty on an ACF
+the data, and so not inherently probabilistic.
+Therefore, there is no justified way to estimate the uncertainty on an ACF
 rotation period measurement.
@@
 
-The motivation for developing a GP-based method for rotation period inference
-is, firstly to measure more accurate and precise rotation periods using a
-better-suited generative model than a sinusoid for the reasons explained
-above.
-Secondly, in order to infer {\it probabilistic} periods, i.e. to estimate the
-posterior PDF of the rotation period and thereby produce a realistic and
-estimate of its uncertainty.
-Thirdly, to allow for an additional correlated noise model to be included
-during regression, the parameters of which could be marginalised over.
-% And fourthly to provide some way of determining whether a periodic model is
-% supported by the data over a purely stochastic one.
+In this work, we introduce an {\it effective} model for rotationally
+modulated stellar light curves
+which captures the salient behaviour but is not
+physically motivated---although some parameters may indeed be 
+{\it interpreted} as physical ones.
+An ideal effective model should have a small number of non-degenerate parameters 
+and be flexible enough to perfectly capture non-sinusoidal, QP behaviour.
+A Gaussian process (GP) model fulfills these requirements. We use this GP 
+model as the generative model at the backbone of a method to probabilistically 
+infer accurate and precise stellar rotation periods.  This enables us to 
+estimate the posterior PDF of the rotation period, thereby producing a 
+justified estimate of its uncertainty while simultaneously marginalizing over
+parameters which model correlated noise.
 
@@
 Rajpaul2016, Aigrain2016}.
-% FIXME: CITATIONS Gibson, Aigrain, Evans, Rajpaul, Waldman(?), Cosmology,
-% geology, biology, Rasmussen & Williams, etc.
 They are useful in regression problems involving any stochastic process,
@@
 multi-variate Gaussian.
-If the probability of obtaining a data-set is a Gaussian in $N$ dimensions,
-where $N$ is the number of data points, that data-set can be described as, and
-with, a Gaussian process.
+If the probability of obtaining a dataset is a Gaussian in $N$ dimensions,
+where $N$ is the number of data points, a GP can describe that dataset. 
 An in-depth introduction to Gaussian processes is provided by
@@
 
-GP models parameterise the covariance between data points and a kernel
-function provides the form of covariance matrix parameterisation.
-For example, take the time-series in figure \ref{fig:GP_example}.
-This is the \kepler\ light curve of KIC \kepexample, a relatively active star that
-rotates once every $\sim$ 4 days.
-The stochastic variability in this time-series is typical of \kepler\ FGK
-stars.
+GP models parameterise the covariance between data points by means of a kernel
+function that defines the form of the covariance matrix.
+For example, take the time-series in Figure \ref{fig:GP_example}: 
+the \kepler\ light curve of KIC \kepexample, a relatively active star that
+rotates once every $\sim$ 4 days, with stochastic variability typical of 
+\kepler\ FGK stars.
 Clearly, data points in this light curve are correlated.
-Points that are close together in time are tightly correlated and those that
-are widely separated in time are loosely correlated.
-The way in which the covariance varies with the separation between data points
-is modelled when using GP regression; it is not the data but the {\it
-covariance structure} that is modelled.
-This fact is what gives GPs their flexibility---they can model any time
+Points close together in time are tightly correlated and points more
+widely separated are loosely correlated.
+A GP models the way in which this correlation varies with the separation between 
+data points; that is, it models not the data directly but rather the {\it 
+covariance structure}.
+This gives GPs their flexibility---they can model any time
 series with a similar covariance structure.
@@
 structure of a light curve, whereas a much more complicated one may be
-required to model the time-series itself.
-% The light curve of \Kepler-452 has been modelled with a GP in figure
-% \ref{fig:GP_example}, shown in orange.
-In figure \ref{fig:GP_example} the light curve of KIC \kepexample\ has been
-modelled with a GP.
+required to model the time series itself.
+Figure \ref{fig:GP_example} demonstrates how the light curve of KIC \kepexample\ 
+may be modelled with a GP.
 % Both provide adequate fits to the data, however only the periodic kernel
@@
 
-A range of kernel functions could be used to describe stellar variability.
+A range of kernel functions can describe stellar variability.
 For example, the simplest and most commonly used kernel function, the `Squared
-Exponential' (SE) produces an adequate fit to the KIC \kepexample\ light
-curve.
-The SE kernel function is defined as,
+Exponential' (SE), defined as follows, 
+produces an adequate fit to the KIC \kepexample\ light curve:
 \begin{equation}
@@
 Here $A$ is the amplitude of covariance, $l$ is the length scale of
-exponential decay and $x_i-x_j$ is the separation between data points.
-The SE kernel function has the advantage of being very simple---it has just
-two parameters, a covariance amplitude and length scale: $A$ and $l$.
-If $l$ is large two data points far apart in $x$ will be tightly correlated,
+exponential decay, and $x_i-x_j$ is the separation between data points.
+The SE kernel function has the advantage of being very simple, with
+just two parameters, $A$ and $l$.
+If $l$ is large, two data points far apart in $x$ will be tightly correlated,
 and if small they will be loosely correlated.
@@
 % \citep[see \eg][]{Rajpaul2015}
-The SE kernel function is not a good model of the covariance in stellar light
-curves, nor a {\it useful} one for the problem of rotation period inference
-because it does not capture periodic behaviour.
-In order to infer rotation periods it is necessary to use a periodic kernel
-function.
-For this reason, we use the `Quasi-Periodic' kernel function.
-The QP kernel function is defined as
+
+However, The SE kernel function is not a good model of the covariance 
+in stellar light curves, nor a {\it useful} one for the problem of 
+rotation period inference because it does not capture periodic behaviour.
+In order to infer rotation periods it is thus necessary to use a periodic kernel
+function.  For this reason, we use the `Quasi-Periodic' kernel, defined as
 \begin{equation}
@@
 This kernel function allows two data points that are separated in time by one
-rotation period to be tightly correlated, while points separated by half a
-period can be weakly correlated.
-We also use an extra parameter, $\sigma$ which is an additional white noise
+rotation period to be tightly correlated, while also allowing 
+points separated by half a period to be weakly correlated.
+We also use an extra parameter $\sigma$ which is an additional white noise
 term added to the diagonal elements of the covariance matrix.
@@
 GP regression.
-This is because the covariance matrix must be positive definite, however
+This is because the covariance matrix must be positive definite; however,
 matrix inversion performed using most solvers is approximate, not exact,
-therefore slight deviations from positive definiteness can arise.
+so slight deviations from positive definiteness can arise.
 Including a small amount of extra variance in the model allows enough
 flexibility that the matrix inversion algorithms do not run into numerical
-issues.
-
-This QP kernel function was used to produce the model shown in figure
-~\ref{fig:GP_example}.
-
-In order to infer a stellar rotation period from a light curve, we fit a GP
-model with a QP kernel function to the data.
-The likelihood of the model, conditioned on the data could then be maximised
-in order to find the maximum likelihood value for $P$.
-In this study however, the full posterior PDFs of the parameters are explored
-using MCMC.
-This approach comes at a cost: a GP model is expensive to compute once as it
-requires an inversion and determinant evaluation of the covariance matrix, let
-alone the many thousands of times that is necessary to fully explore the
-posteriors of the parameters.
-Fully exploring the posterior PDF of $P$ is important however, as it provides
-an accurate estimate of the uncertainty on the rotation period.
+issues.  We use this QP kernel function to produce the GP model 
+that fits the \Kepler\ light curve in in Figure \ref{fig:GP_example}.
+
+To infer a stellar rotation period from a light curve, we fit this QP-kernel
+GP model to the data.  As with any model-fitting exercise, the likelihood
+of the model could be maximized to find the maximum-likelihood value for $P$.
+In this study, however, we explore the full posterior PDFs using MCMC.  While
+this approach comes at a computational cost, such posterior exploration 
+importantly provides a justified estimate of the uncertainty on rotation period.
 
@@
 In \textsection \ref{sec:simulations} of this paper we describe the simulated
-light curves used to test rotation period inference methods.
-In \textsection \ref{sec:method} we describe the various methods used to
-recover the rotation periods: the periodogram, ACF and GP methods.
-In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data and
-the results of the paper are discussed in \textsection \ref{sec:discussion}.
+light curves used to test the three rotation period recovery methods---periodogram,
+ACF, and GP---that we describe in detail in \textsection \ref{sec:method}.
+In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data, and
+we discuss the results in \textsection \ref{sec:discussion}.
 

### 2017-02-02 documents/ms.tex
minor edits to abstract/intro

--- 2017-02-02 documents/ms.tex
@@
 \begin{abstract}
-The light curves of spotted, rotating stars are often non-sinusoidal and
-quasi-periodic---spots are not static on the stellar surface and have finite
+The light curves of spotted, rotating stars often vary in a non-sinusoidal and
+quasi-periodic fashion---spots move on the stellar surface and have finite
 lifetimes, causing stellar flux variations to slowly shift in phase.
@@
 Physical models of the stellar surface also have many drawbacks preventing effective
-inference, such as having highly degenerate or high-dimensional parameter spaces.
+inference, such as highly degenerate or high-dimensional parameter spaces.
 % \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
@@
 To test this method's effectiveness, we infer rotation periods from 
-\naigrain\ simulated light curves, demonstrating that our results 
+\naigrain\ simulated stellar light curves, demonstrating that our results 
     are more accurate than both a sine-fitting periodogram and an 
@@
 
-The light curves of spotted, rotating stars are often non-sinusoidal and
-Quasi-Periodic (QP).
-These stars vary in brightness due to active regions on their surfaces which
+The brightness of a spotted, rotating star often varies in a non-sinusoidal and
+Quasi-Periodic (QP) manner, due to active regions on its surface which
 rotate in and out of view.
-Complicated surface spot patterns produce the non-sinusoidal quality,
-and the finite lifetimes of these active regions and the presence of 
+Complicated surface spot patterns produce non-sinusoidal variations,
+and the finite lifetimes of these active regions and
 differential rotation on the stellar surface produce quasi-periodicity.
@@
 A physically realistic model of the stellar surface 
-would perfectly capture the complexity of shapes
+would ideally perfectly capture the complexity of shapes
 within stellar light curves as well as the quasi-periodic nature, allowing for
@@
 However, such physical models require many free parameters in order to
-accurately represent a stellar surface and some of these parameters are
+accurately represent a stellar surface, and some of these parameters are
 extremely degenerate \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
@@
 latitude, size, temperature and lifetime.
-Considering that many stars likely have hundreds of spots, the number
-of free parameters quickly becomes unwieldy, especially to explore
-their posterior PDFs with MCMC.
+Considering that stars may have hundreds of spots, the number
+of free parameters in such a model quickly becomes unwieldy, especially to explore
+its posterior probability distribution function (PDF).
 Simplified spot models, such as the one described in \citet{Lanza2014} where
@@
 
-The standard methods used for measuring rotation periods include detecting
+Standard methods to measure rotation periods include detecting
 peaks in a Lomb-Scargle \citep{Lomb1976, Scargle1982} (LS) periodogram
@@
 \citep{Mcquillan2013} and wavelet transforms \citep{Garcia2014}.
-The precision of the LS periodogram and wavelet methods are limited by the
+The precisions of the LS periodogram and wavelet methods are limited by the
 suitability of the model choice: a sinusoid for the LS periodogram, 
@@
 non-sinusoidal.
-In fact, as long as it is approximately periodic the ACF will 
-display a peak at the rotation period, no matter the signal shape.
+In fact, as long as the signal is approximately periodic the ACF will 
+display a peak at the rotation period, no matter its precise shape.
 A drawback of the ACF method, however, is that it requires data to be
@@
 light curves (although in many cases it can be approximated as uniformly
-sampled) and not nearly the case for the future large-scale survey instrument
-\LSST.
-An ACF is also an operation performed on the data, not a generative model of
-the data, and so not inherently probabilistic.
+sampled) and not nearly the case for most ground-based measurements, such 
+as the future Large Synoptic Survey Telescope (\LSST).
+An ACF is also an operation performed on the data rather than a generative 
+model of the data, and so is not inherently probabilistic.
 Therefore, there is no justified way to estimate the uncertainty on an ACF
@@
 An ideal effective model should have a small number of non-degenerate parameters 
-and be flexible enough to perfectly capture non-sinusoidal, QP behaviour.
-A Gaussian process (GP) model fulfills these requirements. We use this GP 
-model as the generative model at the backbone of a method to probabilistically 
+and be flexible enough to perfectly capture non-sinusoidal and QP behaviour.
+A Gaussian process (GP) model fulfills these requirements. We thus use a GP 
+as the generative model at the core of a method to probabilistically 
 infer accurate and precise stellar rotation periods.  This enables us to 
@@
 GPs are commonly used in the machine learning community and increasingly
-in other scientific fields, for example biology and geophysics.
+in other scientific fields such as biology and geophysics.
 More recently, GPs have been used in the astronomy literature \citep[see
@@
 function that defines the form of the covariance matrix.
-For example, take the time-series in Figure \ref{fig:GP_example}: 
+As a qualitatively demonstration, we present the time-series 
+in Figure \ref{fig:GP_example}: 
 the \kepler\ light curve of KIC \kepexample, a relatively active star that
@@
 Clearly, data points in this light curve are correlated.
-Points close together in time are tightly correlated and points more
+Points close together in time are tightly correlated, and points more
 widely separated are loosely correlated.
@@
 covariance structure}.
-This gives GPs their flexibility---they can model any time
+This lends GPs their flexibility---they can model any time
 series with a similar covariance structure.
 In addition, a very simple function can usually capture the covariance
-structure of a light curve, whereas a much more complicated one may be
-required to model the time series itself.
-Figure \ref{fig:GP_example} demonstrates how the light curve of KIC \kepexample\ 
-may be modelled with a GP.
+structure of a light curve, whereas modelling the time series itself
+might require much more complexity.
+Figure \ref{fig:GP_example} demonstrates how a GP model fits the light curve of 
+KIC \kepexample.
 % Both provide adequate fits to the data, however only the periodic kernel
@@
 For example, the simplest and most commonly used kernel function, the `Squared
-Exponential' (SE), defined as follows, 
-produces an adequate fit to the KIC \kepexample\ light curve:
+Exponential' (SE), defined as follows, adequately fits the 
+KIC \kepexample\ light curve:
 \begin{equation}
@@
 Another property of the SE kernel function is that it produces functions that
-are infinitely differentiable.
-It is therefore possible to model a data set and its derivatives
-simultaneously.
+are infinitely differentiable, making it possible to model a data set 
+and its derivatives simultaneously.
 % \citep[see \eg][]{Rajpaul2015}
 
-However, The SE kernel function is not a good model of the covariance 
-in stellar light curves, nor a {\it useful} one for the problem of 
+However, The SE kernel function does not well describe the covariance 
+in stellar light curves, nor is it {\it useful} for the problem of 
 rotation period inference because it does not capture periodic behaviour.
-In order to infer rotation periods it is thus necessary to use a periodic kernel
+Inferring rotation periods thus requires a periodic kernel
 function.  For this reason, we use the `Quasi-Periodic' kernel, defined as
@@
 that describes the periodic covariance structure.
-$P$ can be interpreted as the rotation period of the star and $\Gamma$
+$P$ can be interpreted as the rotation period of the star, and $\Gamma$
 controls the amplitude of the $\sin^2$ term.
@@
 If $\Gamma$ is large, points separated by one period are tightly
-correlated and points separated by slightly more or less are still highly
-correlated although less so.
+correlated, and points separated by slightly more or less are still highly
+correlated, although less so.
 This kernel function allows two data points that are separated in time by one
@@
 term added to the diagonal elements of the covariance matrix.
-It is the fraction by which the observational uncertainties have been
+This represents the fraction by which the observational uncertainties have been
 underestimated --- if the uncertainties reported on the data are too small,
-this parameter will be non-zero.
-In practice however, this parameter should always be non-zero when performing
-GP regression.
-This is because the covariance matrix must be positive definite; however,
-matrix inversion performed using most solvers is approximate, not exact,
-so slight deviations from positive definiteness can arise.
-Including a small amount of extra variance in the model allows enough
-flexibility that the matrix inversion algorithms do not run into numerical
+it will be non-zero.
+In practice, this parameter should always be non-zero when performing
+GP regression, because the covariance matrix must be positive definite and  
+slight deviations from positive definiteness can arise due to most 
+solvers using approximate, rather than exact, matrix inversion.
+Including a small amount of extra variance in the model ensures enough
+flexibility for matrix inversion algorithms to not run into such numerical
 issues.  We use this QP kernel function to produce the GP model 
@@
 
-To infer a stellar rotation period from a light curve, we fit this QP-kernel
+To infer a stellar rotation period $P$ from a light curve, we fit this QP-kernel
 GP model to the data.  As with any model-fitting exercise, the likelihood
 of the model could be maximized to find the maximum-likelihood value for $P$.
-In this study, however, we explore the full posterior PDFs using MCMC.  While
+In this study, however, we explore the full posterior PDFs using a Markov
+Chain Monte Carlo (MCMC) procedure.  While
 this approach comes at a computational cost, such posterior exploration 
-importantly provides a justified estimate of the uncertainty on rotation period.
+importantly provides a justified uncertainty estimate.
 

### 2017-02-02 documents/ms.tex
fixing double Suzanne in author list

--- 2017-02-02 documents/ms.tex
@@
    Ruth Angus\altaffilmark{\columbia},
-   Timothy Morton\altaffilmark{\princeton}
+   Timothy Morton\altaffilmark{\princeton},
    Suzanne Aigrain\altaffilmark{\oxford},
    Daniel Foreman-Mackey\altaffilmark{\uw}
-   Suzanne Aigrain\altaffilmark{\oxford},
    \& Vinesh Rajpaul\altaffilmark{\oxford}
@@
 rotate in and out of view.
-Complicated surface spot patterns produce the non-sinusoidal quality,
-and the finite lifetimes of these active regions and the presence of 
-differential rotation on the stellar surface produce quasi-periodicity.
-A strictly periodic sinusoid therefore does not well describe 
-rotationally modulated stellar light curves.
-A physically realistic model of the stellar surface 
-would perfectly capture the complexity of shapes
+The non-sinusoidal quality is caused by the complicated surface spot patterns
+and the quasi-periodicity is produced both by the finite lifetimes of these
+active regions and the presence of differential rotation on the stellar
+surface.
+A strictly periodic sinusoid is therefore not a good model for stellar light
+curves.
+In an ideal world, a physical model of the stellar surface would be
+conditioned on the data.
+A physically realistic model would perfectly capture the complexity of shapes
 within stellar light curves as well as the quasi-periodic nature, allowing for
-extremely precise probabilistic period recovery when conditioned on the data.
+extremely precise probabilistic period recovery.
 However, such physical models require many free parameters in order to
 accurately represent a stellar surface and some of these parameters are
-extremely degenerate \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
+extremely degenerate \citep[\eg][]{Russell1906, Jeffers2009}.
 In addition to global stellar parameters such as inclination and rotation
@@
 latitude, size, temperature and lifetime.
-Considering that many stars likely have hundreds of spots, the number
-of free parameters quickly becomes unwieldy, especially to explore
-their posterior PDFs with MCMC.
+Considering that many stars have on the order of hundreds of spots, the number
+of free parameters quickly becomes unwieldy, especially if the posterior PDFs
+of these parameters are explored with MCMC.
 Simplified spot models, such as the one described in \citet{Lanza2014} where
-only two spots are modelled, have produced successful results; however, such 
-relatively inflexible models sacrifice precision.
+only two spots are modelled, have produced successful results, however these
+simplified models sacrifice some precision due to lack of model flexibility.
+Instead of using a physical model for stellar light curves, we choose to use
+an {\it effective} model: one which captures the behaviour but is not
+physically motivated, although the parameters of this model may be {\it
+interpreted} as physical ones.
+An ideal effective model for the light curve of a spotted, rotating star is
+one with a small number of non-degenerate parameters that is flexible enough
+to perfectly capture non-sinusoidal, QP behaviour.
+These requirements are fulfilled by a Gaussian process (GP) model.
 
@@
 peaks in a Lomb-Scargle \citep{Lomb1976, Scargle1982} (LS) periodogram
-\citep[e.g.][]{Reinhold2013}, Auto-Correlation Functions (ACFs),
+\citep[e.g.][]{Reinhold2013}, Auto-Correlation Functions (ACFs)
 \citep{Mcquillan2013} and wavelet transforms \citep{Garcia2014}.
 The precision of the LS periodogram and wavelet methods are limited by the
-suitability of the model choice: a sinusoid for the LS periodogram, 
-and a choice of mother wavelet, assumed to desribe the data
-over a range of transpositions \citep[see, \eg][]{Carter2010}, 
-for the wavelet method.
+suitability of the model choice.
+A sinusoid is used in the case of the LS periodogram and the wavelet method
+relies on a choice of mother wavelet that is assumed to describe the data over
+a range of transpositions \citep[see, \eg][]{Carter2010}.
 In contrast, the ACF method is much better suited to signals that are
 non-sinusoidal.
-In fact, as long as it is approximately periodic the ACF will 
-display a peak at the rotation period, no matter the signal shape.
-A drawback of the ACF method, however, is that it requires data to be
+In fact, it doesn't matter what shape the signal is: as long as it is
+approximately periodic the ACF will display peaks located at the rotation
+period.
+A drawback of the ACF method however, is that it requires data to be
 evenly-spaced\footnote{\citet{Edelson1988} describe a method for computing
-ACFs for unevenly-spaced data.}, which is not exactly the case with \Kepler\
+ACFs for unevenly-spaced data.} which is not exactly the case with \Kepler\
 light curves (although in many cases it can be approximated as uniformly
-sampled) and not nearly the case for the future large-scale survey instrument
+sampled) and not nearly the case for the future large-scale survey instrument,
 \LSST.
 An ACF is also an operation performed on the data, not a generative model of
-the data, and so not inherently probabilistic.
-Therefore, there is no justified way to estimate the uncertainty on an ACF
+the data and is therefore not inherently probabilistic.
+For this reason it is very difficult to estimate the uncertainty on an ACF
 rotation period measurement.
@@
 
-In this work, we introduce an {\it effective} model for rotationally
-modulated stellar light curves
-which captures the salient behaviour but is not
-physically motivated---although some parameters may indeed be 
-{\it interpreted} as physical ones.
-An ideal effective model should have a small number of non-degenerate parameters 
-and be flexible enough to perfectly capture non-sinusoidal, QP behaviour.
-A Gaussian process (GP) model fulfills these requirements. We use this GP 
-model as the generative model at the backbone of a method to probabilistically 
-infer accurate and precise stellar rotation periods.  This enables us to 
-estimate the posterior PDF of the rotation period, thereby producing a 
-justified estimate of its uncertainty while simultaneously marginalizing over
-parameters which model correlated noise.
+The motivation for developing a GP-based method for rotation period inference
+is, firstly to measure more accurate and precise rotation periods using a
+better-suited generative model than a sinusoid for the reasons explained
+above.
+Secondly, in order to infer {\it probabilistic} periods, i.e. to estimate the
+posterior PDF of the rotation period and thereby produce a realistic and
+estimate of its uncertainty.
+Thirdly, to allow for an additional correlated noise model to be included
+during regression, the parameters of which could be marginalised over.
+% And fourthly to provide some way of determining whether a periodic model is
+% supported by the data over a purely stochastic one.
 
@@
 Rajpaul2016, Aigrain2016}.
+% FIXME: CITATIONS Gibson, Aigrain, Evans, Rajpaul, Waldman(?), Cosmology,
+% geology, biology, Rasmussen & Williams, etc.
 They are useful in regression problems involving any stochastic process,
@@
 multi-variate Gaussian.
-If the probability of obtaining a dataset is a Gaussian in $N$ dimensions,
-where $N$ is the number of data points, a GP can describe that dataset. 
+If the probability of obtaining a data-set is a Gaussian in $N$ dimensions,
+where $N$ is the number of data points, that data-set can be described as, and
+with, a Gaussian process.
 An in-depth introduction to Gaussian processes is provided by
@@
 
-GP models parameterise the covariance between data points by means of a kernel
-function that defines the form of the covariance matrix.
-For example, take the time-series in Figure \ref{fig:GP_example}: 
-the \kepler\ light curve of KIC \kepexample, a relatively active star that
-rotates once every $\sim$ 4 days, with stochastic variability typical of 
-\kepler\ FGK stars.
+GP models parameterise the covariance between data points and a kernel
+function provides the form of covariance matrix parameterisation.
+For example, take the time-series in figure \ref{fig:GP_example}.
+This is the \kepler\ light curve of KIC \kepexample, a relatively active star that
+rotates once every $\sim$ 4 days.
+The stochastic variability in this time-series is typical of \kepler\ FGK
+stars.
 Clearly, data points in this light curve are correlated.
-Points close together in time are tightly correlated and points more
-widely separated are loosely correlated.
-A GP models the way in which this correlation varies with the separation between 
-data points; that is, it models not the data directly but rather the {\it 
-covariance structure}.
-This gives GPs their flexibility---they can model any time
+Points that are close together in time are tightly correlated and those that
+are widely separated in time are loosely correlated.
+The way in which the covariance varies with the separation between data points
+is modelled when using GP regression; it is not the data but the {\it
+covariance structure} that is modelled.
+This fact is what gives GPs their flexibility---they can model any time
 series with a similar covariance structure.
@@
 structure of a light curve, whereas a much more complicated one may be
-required to model the time series itself.
-Figure \ref{fig:GP_example} demonstrates how the light curve of KIC \kepexample\ 
-may be modelled with a GP.
+required to model the time-series itself.
+% The light curve of \Kepler-452 has been modelled with a GP in figure
+% \ref{fig:GP_example}, shown in orange.
+In figure \ref{fig:GP_example} the light curve of KIC \kepexample\ has been
+modelled with a GP.
 % Both provide adequate fits to the data, however only the periodic kernel
@@
 
-A range of kernel functions can describe stellar variability.
+A range of kernel functions could be used to describe stellar variability.
 For example, the simplest and most commonly used kernel function, the `Squared
-Exponential' (SE), defined as follows, 
-produces an adequate fit to the KIC \kepexample\ light curve:
+Exponential' (SE) produces an adequate fit to the KIC \kepexample\ light
+curve.
+The SE kernel function is defined as,
 \begin{equation}
@@
 Here $A$ is the amplitude of covariance, $l$ is the length scale of
-exponential decay, and $x_i-x_j$ is the separation between data points.
-The SE kernel function has the advantage of being very simple, with
-just two parameters, $A$ and $l$.
-If $l$ is large, two data points far apart in $x$ will be tightly correlated,
+exponential decay and $x_i-x_j$ is the separation between data points.
+The SE kernel function has the advantage of being very simple---it has just
+two parameters, a covariance amplitude and length scale: $A$ and $l$.
+If $l$ is large two data points far apart in $x$ will be tightly correlated,
 and if small they will be loosely correlated.
@@
 % \citep[see \eg][]{Rajpaul2015}
-
-However, The SE kernel function is not a good model of the covariance 
-in stellar light curves, nor a {\it useful} one for the problem of 
-rotation period inference because it does not capture periodic behaviour.
-In order to infer rotation periods it is thus necessary to use a periodic kernel
-function.  For this reason, we use the `Quasi-Periodic' kernel, defined as
+The SE kernel function is not a good model of the covariance in stellar light
+curves, nor a {\it useful} one for the problem of rotation period inference
+because it does not capture periodic behaviour.
+In order to infer rotation periods it is necessary to use a periodic kernel
+function.
+For this reason, we use the `Quasi-Periodic' kernel function.
+The QP kernel function is defined as
 \begin{equation}
@@
 This kernel function allows two data points that are separated in time by one
-rotation period to be tightly correlated, while also allowing 
-points separated by half a period to be weakly correlated.
-We also use an extra parameter $\sigma$ which is an additional white noise
+rotation period to be tightly correlated, while points separated by half a
+period can be weakly correlated.
+We also use an extra parameter, $\sigma$ which is an additional white noise
 term added to the diagonal elements of the covariance matrix.
@@
 GP regression.
-This is because the covariance matrix must be positive definite; however,
+This is because the covariance matrix must be positive definite, however
 matrix inversion performed using most solvers is approximate, not exact,
-so slight deviations from positive definiteness can arise.
+therefore slight deviations from positive definiteness can arise.
 Including a small amount of extra variance in the model allows enough
 flexibility that the matrix inversion algorithms do not run into numerical
-issues.  We use this QP kernel function to produce the GP model 
-that fits the \Kepler\ light curve in in Figure \ref{fig:GP_example}.
-
-To infer a stellar rotation period from a light curve, we fit this QP-kernel
-GP model to the data.  As with any model-fitting exercise, the likelihood
-of the model could be maximized to find the maximum-likelihood value for $P$.
-In this study, however, we explore the full posterior PDFs using MCMC.  While
-this approach comes at a computational cost, such posterior exploration 
-importantly provides a justified estimate of the uncertainty on rotation period.
+issues.
+
+This QP kernel function was used to produce the model shown in figure
+~\ref{fig:GP_example}.
+
+In order to infer a stellar rotation period from a light curve, we fit a GP
+model with a QP kernel function to the data.
+The likelihood of the model, conditioned on the data could then be maximised
+in order to find the maximum likelihood value for $P$.
+In this study however, the full posterior PDFs of the parameters are explored
+using MCMC.
+This approach comes at a cost: a GP model is expensive to compute once as it
+requires an inversion and determinant evaluation of the covariance matrix, let
+alone the many thousands of times that is necessary to fully explore the
+posteriors of the parameters.
+Fully exploring the posterior PDF of $P$ is important however, as it provides
+an accurate estimate of the uncertainty on the rotation period.
 
@@
 In \textsection \ref{sec:simulations} of this paper we describe the simulated
-light curves used to test the three rotation period recovery methods---periodogram,
-ACF, and GP---that we describe in detail in \textsection \ref{sec:method}.
-In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data, and
-we discuss the results in \textsection \ref{sec:discussion}.
+light curves used to test rotation period inference methods.
+In \textsection \ref{sec:method} we describe the various methods used to
+recover the rotation periods: the periodogram, ACF and GP methods.
+In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data and
+the results of the paper are discussed in \textsection \ref{sec:discussion}.
 

### 2017-02-02 documents/ms.tex
a few additions to the manuscript

@@
 \newcommand{\oxford}{3}
-\newcommand{\uw}{4}
+\newcommand{\sagan}{4}
+\newcommand{\uw}{5}
 \newcommand{\princeton}{2}
@@
 
+\newcommand{\dfmcomment}[1]{{\color{red}#1}}
+
 \begin{document}
@@
    Suzanne Aigrain\altaffilmark{\oxford},
-   Daniel Foreman-Mackey\altaffilmark{\uw}
+   Daniel Foreman-Mackey\altaffilmark{\uw,\sagan}
    \& Vinesh Rajpaul\altaffilmark{\oxford}
@@
 University, NY, NY, RuthAngus@gmail.com}
+\altaffiltext{\princeton}{Department of Astrophysical Sciences, Princeton
+University,  Princeton, NJ}
 \altaffiltext{\oxford}{Subdepartment of Astrophysics, University of Oxford,
 UK}
+\altaffiltext{\sagan}{Sagan Fellow}
 \altaffiltext{\uw}{Department of Astronomy, University of Washington, Seattle,
 WA}
-\altaffiltext{\princeton}{Department of Astrophysical Sciences, Princeton
-University,  Princeton, NJ}
 
@@
 ages or stellar obliquities in exoplanet systems, or characterising star-planet interactions.
+\dfmcomment{Why do posteriors ``enable'' other studies? Hierarchical
+    inference? This could include more of an argument here.}
 % Thehe ACF method is not applicable to unevenly spaced data,
@@
 model of the data, and so is not inherently probabilistic.
-Therefore, there is no justified way to estimate the uncertainty on an ACF
-rotation period measurement.
+This means that the effects of the observational uncertainties cannot be
+formally propagated to constraints on the rotation period.
 % Many rotation periods in the literature have been inferred by measuring the
@@
 in other scientific fields such as biology and geophysics.
-More recently, GPs have been used in the astronomy literature \citep[see
-\eg][]{Gibson2012, Haywood2014, Haywood2015, Evans2015, Rajpaul2015,
-Rajpaul2016, Aigrain2016}.
+More recently, GPs have been used in the astronomy \dfmcomment{(this should
+probably say something more specific because GPs have been used in other parts
+of astronomy for decades)} literature \citep[see
+\eg][]{Gibson2012, Haywood2014, Dawson2014, Barclay2015, Haywood2015,
+Evans2015, Rajpaul2015, Rajpaul2016, Aigrain2016}.
 They are useful in regression problems involving any stochastic process,
@@
 
-GP models parameterise the covariance between data points by means of a kernel
-function that defines the form of the covariance matrix.
+GP models parameterise the covariance between data points by means of a
+parametric kernel function that models the autocorrelation behaviour of the
+signal and defines the form of the covariance matrix.
 As a qualitatively demonstration, we present the time-series
@@
 function.  For this reason, we use the `Quasi-Periodic' kernel, defined as
+\citep{Rasmussen2005}
 \begin{equation}
-k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2} -
-\Gamma^2 \sin^2(\frac{2\pi}{P}) \right).
+k_{i,j} = A \exp \left[-\frac{(x_i - x_j)^2}{2l^2} -
+\Gamma^2 \sin^2\left(\frac{2\pi}{P}\right) \right].
 \end{equation}
@@
 term added to the diagonal elements of the covariance matrix.
-This represents the fraction by which the observational uncertainties have been
-underestimated --- if the uncertainties reported on the data are too small,
-it will be non-zero.
-In practice, this parameter should always be non-zero when performing
-GP regression, because the covariance matrix must be positive definite and  
-slight deviations from positive definiteness can arise due to most 
-solvers using approximate, rather than exact, matrix inversion.
-Including a small amount of extra variance in the model ensures enough
-flexibility for matrix inversion algorithms to not run into such numerical
-issues.  We use this QP kernel function to produce the GP model 
+This can be interpreted to represent underestimation of observational
+uncertainties~---~if the uncertainties reported on the data are too small, it
+will be non-zero~---~or it can capture any remaining ``jitter'' or residuals
+that are not captured by the effective GP model.
+%In practice, this parameter should always be non-zero when performing
+%GP regression, because the covariance matrix must be positive definite and
+%slight deviations from positive definiteness can arise due to most
+%solvers using approximate, rather than exact, matrix inversion.
+%Including a small amount of extra variance in the model ensures enough
+%flexibility for matrix inversion algorithms to not run into such numerical
+%issues.
+We use this QP kernel function to produce the GP model
 that fits the \Kepler\ light curve in in Figure \ref{fig:GP_example}.
@@
 parameter will be non-zero.
-    The green line in the period panel shows the injected period.}
+    The green line in the period panel shows the injected period.
+    This figure was made using \texttt{corner.py} \citep{Corner}.}
 \label{fig:gp_posteriors}

### 2017-02-02 documents/ms.tex
minor edit

--- 2017-02-02 documents/ms.tex
@@
    Ruth Angus\altaffilmark{\columbia},
-   Timothy Morton\altaffilmark{\princeton},
+   Timothy Morton\altaffilmark{\princeton}
    Suzanne Aigrain\altaffilmark{\oxford},
    Daniel Foreman-Mackey\altaffilmark{\uw}
+   Suzanne Aigrain\altaffilmark{\oxford},
    \& Vinesh Rajpaul\altaffilmark{\oxford}
@@
 
-In order to test the GP method we attempted to the measure rotation periods of
-a set of simulated light curves.
-We used \naigrain\ light curves generated for the \citet{Aigrain2015}
+In order to test our rotation period recovery method and compare it to
+other standard procedures, we use a set of simulated light curves: namely,
+\naigrain\ light curves generated for the \citet{Aigrain2015}
 `hare and hounds' rotation period recovery experiment.
-These light curves were simulated by placing dark, circular spots with slowly
-evolving size, on the surface of bright, rotating spheres, ignoring
+These light curves result from placing dark, circular spots with slowly
+evolving size on the surface of bright, rotating spheres, ignoring
 limb-darkening effects.
-\citet{Aigrain2015} simulated one thousand light curves in order to test the
+\citet{Aigrain2015} simulated one thousand light curves to test the
 ability of participating teams to recover both the stellar rotation periods
-and the rotational shear: the amplitude of surface differential rotation.
+and the rotational shear (the amplitude of surface differential rotation).
 % Since we are not interested in recovering differential rotation we only used

### 2017-02-03 documents/ms.tex
rearranging method section and correcting Dan's comments

--- 2017-02-02 documents/ms.tex
@@
 % \newcommand{\nkoibad}{49}
-\newcommand{\kepexample}{1430163}
-\newcommand{\kepexampleperiod}{4}
+\newcommand{\kepexample}{5809890}
+\newcommand{\kepexampleperiod}{30.5}
 \newcommand{\aigrainexampleperiod}{20.8}
@@
 inferring rotation periods for \nkois\ \Kepler\ stars.
-Because this method delivers posteriors PDFs,
-it will enable other studies involving stellar rotation,
-such as inferring stellar
-ages or stellar obliquities in exoplanet systems, or characterising star-planet interactions.
-\dfmcomment{Why do posteriors ``enable'' other studies? Hierarchical
-    inference? This could include more of an argument here.}
+Because this method delivers posteriors PDFs it will enable hierarchical
+    studies involving stellar rotation, particularly those involving
+    population modelling, such as inferring stellar ages, obliquities in
+    exoplanet systems, or characterising star-planet interactions.
 % Thehe ACF method is not applicable to unevenly spaced data,
@@
 GPs are commonly used in the machine learning community and increasingly
-in other scientific fields such as biology and geophysics.
-More recently, GPs have been used in the astronomy \dfmcomment{(this should
-probably say something more specific because GPs have been used in other parts
-of astronomy for decades)} literature \citep[see
-\eg][]{Gibson2012, Haywood2014, Dawson2014, Barclay2015, Haywood2015,
-Evans2015, Rajpaul2015, Rajpaul2016, Aigrain2016}.
+in other scientific fields such as biology, geophysics and cosmology.
+More recently, GPs have been used in stellar and exoplanet fields to capture
+stellar variability or instrumental systematics \citep[see \eg][]{Gibson2012,
+Haywood2014, Dawson2014, Barclay2015, Haywood2015, Evans2015, Rajpaul2015,
+Rajpaul2016, Aigrain2016}.
 They are useful in regression problems involving any stochastic process,
@@
 Inferring rotation periods thus requires a periodic kernel
-function.  For this reason, we use the `Quasi-Periodic' kernel, defined as
-\citep{Rasmussen2005}
+function.  For this reason, we use the `Quasi-Periodic' kernel.
+This kernel is used by \citet{Rasmussen2005} to capture the variable behaviour
+of CO$_2$ emissions from the Mauna Loa volcano in Hawai'i.
+It is defined as
 \begin{equation}
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/001430163.pdf}
+\includegraphics[width=6in, clip=true]{figures/koi_lc_demo.pdf}
 \caption[A light curve with a GP model.]
@@
 
+This paper is laid out as follows.
 In \textsection \ref{sec:simulations} of this paper we describe the simulated
-light curves used to test the three rotation period recovery methods---periodogram,
-ACF, and GP---that we describe in detail in \textsection \ref{sec:method}.
-In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data, and
-we discuss the results in \textsection \ref{sec:discussion}.
+light curves used to test the three rotation period recovery
+methods (ACF, LS periodogram and GP).
+The GP method is described in \textsection \ref{sec:method} and the ACF and LS
+periodogram methods are described in \textsection \ref{sec:comparison}.
+In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data,
+and the results are discussed in \textsection \ref{sec:discussion}.
 
@@
 
-\section{Rotation period recovery}
+\section{The GP method}
 \label{sec:method}
-\subsection{ACF}
-
-We measured an ACF-based period for each light curve following the method of
-\citet{Mcquillan2013}.
-An ACF is calculated for each light curve and smoothed by convolving with a
-Gaussian with $\sigma=9$ days.
-A rotation period is estimated as the lag-time of the highest peak in the ACF,
-less than 100 days.
-This is not always the first peak: the second can be larger than the first if
-there are two active regions at or near opposite longitudes on the surface of
-the star---producing a light curve with two dips per rotation period.
-We restrict our search to periods less than 100 days because this is the
-maximum period that is realistically recoverable in \kepler\ light curves ---
-instrumental noise significantly distorts signals on longer timescales.
-An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
-in figure \ref{fig:demo_acf}.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/demo_ACF.pdf}
-\caption[ACF of a simulated light curve.]
-{An autocorrelation function of the simulated light curve shown in figure
-\ref{fig:demo_lc}.
-The vertical blue line shows the period measured using the ACF method (20.4
-days) and the pink dashed line shows the period that was used to simulate the
-light curve (\aigrainexampleperiod\ days).}
-\label{fig:demo_acf}
-\end{center}
-\end{figure}
-
-The ACF method has proven to be extremely useful for measuring rotation
-periods.
-The catalogue of rotation periods of \Kepler\ stars provided in
-\citet{Mcquillan2013} has been widely used by the community and has provided
-ground-breaking results for stellar and exoplanetary science.
-The results of the ACF method as tested in \citet{Aigrain2015} were positive
-(see, for example their figure 8) as it produced a large number of accurate
-rotation period measurements.
-Another advantage of the ACF method is that it is fast to implement.
-A clear disadvantage however, as already mentioned above, are the poorly
-defined uncertainties for rotation periods estimated using an autocorrelation
-function.
-This is related to the fact that ACF period detection is a non-probabilistic
-method, another drawback of ACF periods.
-
-We applied the ACF method to the sample of \naigrain\ simulated light
-curves.
-Periods measured using the ACF method are plotted against the original
-rotation period values used to generate the light curves in figure
-\ref{fig:compare_acf}.
-The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
-The Median Absolute Deviation (MAD) of the ACF-recovered periods is \acfRMS\
-in log-space (1.8 days).
-The agreement between the injected and recovered rotation periods is good in
-general, however some rotation periods are drastically over- and
-underestimated.
-Of the points clustered close to the $x=y$ line, more fall slightly below it
-than above it, \ie\ rotation periods are preferentially underestimated.
-This is a feature of the peak position measurement method that is performed on
-the ACF.
-ACFs of stellar light curves have similar functional forms to the QP kernel
-function in equation \ref{eq:QP}: periodic functions added to decaying
-exponentials.
-In such functions the peak positions can be shifted towards the left, the
-short period end, because the decaying exponential raises the left side of
-each peak more than the right.
-It is possible to model this effect, however the standard practice is to
-simply measure the peak position without taking it into account.
-We adopt the standard practice approach here in order to provide a comparison
-between our new method and that used largely in the literature.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
-\caption[ACF results.]
-{The `true' rotation periods used to generate \naigrain\ simulated light
-curves vs the rotation periods measured using the ACF technique.
-    Several light curves have over-estimated rotation periods and some
-    are drastically underestimated.}
-\label{fig:compare_acf}
-\end{center}
-\end{figure*}
-
-\subsection{LS periodogram}
-
-For each simulated light curve, a LS periodogram
-\footnote{LS periodograms were calculated using the gatspy Python module:
-\url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
-was computed over a grid of 10,000 periods, evenly spaced in frequency,
-between 1 and 100 days.
-The period of the highest peak in the periodogram was adopted as the rotation
-period.
-The resulting recovered rotation periods are plotted as a function of the true
-periods in figure \ref{fig:pgram_compare}.
-The MAD for rotation periods measured using the LS periodogram technique is
-\pgramRMS\ in log-space (2.7 days).
-There is a large number of drastically overestimated periods recovered using
-the periodogram.
-Excess power at long periods can be generated by long-term trends that are not
-related to rotation.
-This problem becomes exacerbated when instrumental noise is also present in
-light curves.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
-\caption[LS periodogram results.]
-{The `true' rotation periods used to generate \naigrain\ simulated light
-curves vs the rotation periods measured using a LS periodogram technique.
-In many cases a large peak at a long period was present in the periodogram,
-    producing a significant over-estimate of the period.
-    }
-\label{fig:pgram_compare}
-\end{center}
-\end{figure*}
-
-\subsection{The GP method}
 
@@
 % \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_emcee3_24_0.png}
+\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_02_03.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_noprior.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_noprior_02_03.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 
+\section{Comparison with standard methods}
+\label{sec:comparison}
+\subsection{ACF}
+
+We measured an ACF-based period for each light curve following the method of
+\citet{Mcquillan2013}.
+An ACF is calculated for each light curve and smoothed by convolving with a
+Gaussian with $\sigma=9$ days.
+A rotation period is estimated as the lag-time of the highest peak in the ACF,
+less than 100 days.
+This is not always the first peak: the second can be larger than the first if
+there are two active regions at or near opposite longitudes on the surface of
+the star---producing a light curve with two dips per rotation period.
+We restrict our search to periods less than 100 days because this is the
+maximum period that is realistically recoverable in \kepler\ light curves ---
+instrumental noise significantly distorts signals on longer timescales.
+An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
+in figure \ref{fig:demo_acf}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/demo_ACF.pdf}
+\caption[ACF of a simulated light curve.]
+{An autocorrelation function of the simulated light curve shown in figure
+\ref{fig:demo_lc}.
+The vertical blue line shows the period measured using the ACF method (20.4
+days) and the pink dashed line shows the period that was used to simulate the
+light curve (\aigrainexampleperiod\ days).}
+\label{fig:demo_acf}
+\end{center}
+\end{figure}
+
+The ACF method has proven to be extremely useful for measuring rotation
+periods.
+The catalogue of rotation periods of \Kepler\ stars provided in
+\citet{Mcquillan2013} has been widely used by the community and has provided
+ground-breaking results for stellar and exoplanetary science.
+The results of the ACF method as tested in \citet{Aigrain2015} were positive
+(see, for example their figure 8) as it produced a large number of accurate
+rotation period measurements.
+Another advantage of the ACF method is that it is fast to implement.
+A clear disadvantage however, as already mentioned above, are the poorly
+defined uncertainties for rotation periods estimated using an autocorrelation
+function.
+This is related to the fact that ACF period detection is a non-probabilistic
+method, another drawback of ACF periods.
+
+We applied the ACF method to the sample of \naigrain\ simulated light
+curves.
+Periods measured using the ACF method are plotted against the original
+rotation period values used to generate the light curves in figure
+\ref{fig:compare_acf}.
+The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
+The Median Absolute Deviation (MAD) of the ACF-recovered periods is \acfRMS\
+in log-space (1.8 days).
+The agreement between the injected and recovered rotation periods is good in
+general, however some rotation periods are drastically over- and
+underestimated.
+Of the points clustered close to the $x=y$ line, more fall slightly below it
+than above it, \ie\ rotation periods are preferentially underestimated.
+This is a feature of the peak position measurement method that is performed on
+the ACF.
+ACFs of stellar light curves have similar functional forms to the QP kernel
+function in equation \ref{eq:QP}: periodic functions added to decaying
+exponentials.
+In such functions the peak positions can be shifted towards the left, the
+short period end, because the decaying exponential raises the left side of
+each peak more than the right.
+It is possible to model this effect, however the standard practice is to
+simply measure the peak position without taking it into account.
+We adopt the standard practice approach here in order to provide a comparison
+between our new method and that used largely in the literature.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
+\caption[ACF results.]
+{The `true' rotation periods used to generate \naigrain\ simulated light
+curves vs the rotation periods measured using the ACF technique.
+    Several light curves have over-estimated rotation periods and some
+    are drastically underestimated.}
+\label{fig:compare_acf}
+\end{center}
+\end{figure*}
+
+\subsection{LS periodogram}
+
+For each simulated light curve, a LS periodogram
+\footnote{LS periodograms were calculated using the gatspy Python module:
+\url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
+was computed over a grid of 10,000 periods, evenly spaced in frequency,
+between 1 and 100 days.
+The period of the highest peak in the periodogram was adopted as the rotation
+period.
+The resulting recovered rotation periods are plotted as a function of the true
+periods in figure \ref{fig:pgram_compare}.
+The MAD for rotation periods measured using the LS periodogram technique is
+\pgramRMS\ in log-space (2.7 days).
+There is a large number of drastically overestimated periods recovered using
+the periodogram.
+Excess power at long periods can be generated by long-term trends that are not
+related to rotation.
+This problem becomes exacerbated when instrumental noise is also present in
+light curves.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
+\caption[LS periodogram results.]
+{The `true' rotation periods used to generate \naigrain\ simulated light
+curves vs the rotation periods measured using a LS periodogram technique.
+In many cases a large peak at a long period was present in the periodogram,
+    producing a significant over-estimate of the period.
+    }
+\label{fig:pgram_compare}
+\end{center}
+\end{figure*}
+
 \section{Real \kepler\ data}
@@
 
-Figure \ref{fig:mcquillan} compares the periods inferred with the GP method to the
-ACF-based periods from \citet{Mcquillan2013} for the \nkoimcq\ overlapping KOIs.
-This comparison shows generally very good agreement, with only a few exceptions,
-demonstrating that this method works not only on simulated data, but also on
-real data---with the caveat that for any particular data set, some care is needed
-regarding the setting the priors and ranges for the GP hyperparameters.
+Figure \ref{fig:mcquillan} compares the periods inferred with the GP method to
+the ACF-based periods from \citet{Mcquillan2013} for the \nkoimcq\ overlapping
+KOIs.
+This comparison shows generally very good agreement, with only a few
+exceptions, demonstrating that this method works not only on simulated data,
+but also on real data---with the caveat that for any particular data set, some
+care is needed regarding the setting the priors and ranges for the GP
+hyperparameters.
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_koi.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_koi_02_03.pdf}
 \caption[Comparison with McQuillan results.]

### 2017-02-03 code/plot_koi.py
updating plots to reflect final results

@@ -16,7 +16,7 @@ plotpar = {'axes.labelsize': 18,
 plt.rcParams.update(plotpar)
 
 
-def load_and_plot(data, DATA_DIR, RESULTS_DIR):
+def load_and_plot(data):
 
     recovered, errp, errm, lnerrp, lnerrm = [np.zeros(len(data.koi_id.values))
                                              for i in range(5)]
@@ -59,7 +59,7 @@ def load_and_plot(data, DATA_DIR, RESULTS_DIR):
     plt.xlabel("$\ln(\mathrm{McQuillan~{\it et~al.,}~2013~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
     plt.savefig("comparison_koi")
-    # plt.savefig("../documents/figures/comparison_koi.pdf")
+    plt.savefig(os.path.join(FIG_DIR, "comparison_koi_02_03.pdf"))
 
     print("MAD = ", np.median(np.abs(np.log(x[l]) - np.log(recovered[l]))))
 
@@ -96,7 +96,9 @@ def load_and_plot(data, DATA_DIR, RESULTS_DIR):
 
 
 if __name__ == "__main__":
-    RESULTS_DIR = "koi_results_01_23"
-    DATA_DIR = "data"
+    path = "/Users/ruthangus/projects/GProtation"
+    RESULTS_DIR = os.path.join(path, "code/koi_results_02_03")
+    DATA_DIR = os.path.join(path, "code/data")
+    FIG_DIR = os.path.join(path, "documents/figures")
     data = pd.read_csv(os.path.join(DATA_DIR, "Table_1.txt"), skiprows=19)
-    load_and_plot(data, DATA_DIR, RESULTS_DIR)
+    load_and_plot(data)

### 2017-02-03 code/plot_tim.py

@@ -1,5 +1,7 @@
-import numpy as np
+
 import os
+
+import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
 
@@ -11,117 +13,134 @@ plotpar = {'axes.labelsize': 18,
            'text.usetex': True}
 plt.rcParams.update(plotpar)
 
-prior = True
 
-if prior:
-    RESULTS_DIR = "results_24_01"
-else:
-    RESULTS_DIR = "results_noprior"
-truths = pd.read_csv("final_table.txt", delimiter=" ")
+def plot(prior, nbins):
+    """
+    Make the two GP comparison plots for the paper --- one with an ACF prior
+    (prior = True) and one without (prior = False).
+    params:
+    ------
+    prior: (bool)
+        If True the ACF prior results are plotted, if False the no prior
+        results are plotted.
+    nbins: (int)
+        The number of bins to use to find the MAP.
+    """
 
-# remove differential rotators and take just the first 100
-m = truths.DELTA_OMEGA.values == 0
-truths = truths.iloc[m]
-# ml = np.arange(100)
-# truths = truths.iloc[ml]
+    if prior:
+        RESULTS_DIR = "results_01_24"
+    else:
+        RESULTS_DIR = "results_noprior_02_03"
+    truths = pd.read_csv("final_table.txt", delimiter=" ")
 
-recovered = np.zeros(len(truths.N.values))
-errp, errm = [np.zeros(len(truths.N.values)) for i in range(2)]
-lnerrp, lnerrm = [np.zeros(len(truths.N.values)) for i in range(2)]
-for i, id in enumerate(truths.N.values):
-    fn = os.path.join(RESULTS_DIR, "{}.h5".format(id))
-    if os.path.exists(fn):
-        # store = pd.HDFStore(fn)
-        # print(store)
-        df = pd.read_hdf(fn, key="samples")
-        phist, bins = np.histogram(df.ln_period.values, 100)
-        ln_p = bins[phist == max(phist)][0]
-        # ln_p = np.median(df.ln_period.values)
-        recovered[i] = np.exp(ln_p)
-        lnerrp[i] = np.percentile(df.ln_period.values, 84) - ln_p
[…]

### 2017-02-03 code/plot_koi.py
small changes

@@ -72,6 +72,13 @@ def load_and_plot(data):
     print("median_err = ", median_err)
     print(np.median(lnxerr[np.isfinite(lnxerr)]), "mcquillan err")
 
+
+    print("MAD = ", np.median(np.abs(x[l] - recovered[l])))
+    print("MAD (log) = ", np.median(np.abs(np.log(x[l]) - np
+                                           .log(recovered[l]))))
+    print("MAD relative % = ", np.median((np.abs(x[l] -
+                                                 recovered[l]))/x[l])*100)
+
 #     errs = .5*(lnerrp[l] + lnerrm[l])
 #     plt.clf()
 #     plt.hist(errs, 100)

### 2017-02-03 code/plot_tim.py

@@ -102,14 +102,21 @@ def plot(prior, nbins):
     else:
         plt.savefig("gp_hist_noprior.pdf")
 
-    print("percentiles:")
-    print(np.percentile(resids, 16), np.percentile(resids, 84))
-    print("diff = ",
-          np.abs(np.percentile(resids, 16) - np.percentile(resids, 84)))
-    print("median_err = ", median_err)
+    print(len(x[l]), "stars")
+    # print("percentiles:")
+    # print(np.percentile(resids, 16), np.percentile(resids, 84))
+    # print("diff = ",
+    #       np.abs(np.percentile(resids, 16) - np.percentile(resids, 84)))
+    # print("median_err = ", median_err)
 
-    print("MAD = ", (np.median((x[l] - recovered[l])**2)**.5))
-    print("MAD = ", (np.median((np.log(x[l]) - np.log(recovered[l]))**2)**.5))
+    print("MAD = ", np.median(np.abs(x[l] - recovered[l])))
+    print("MAD (log) = ", np.median(np.abs(np.log(x[l]) -
+                                           np.log(recovered[l]))))
+    print("MAD relative % = ", np.median((np.abs(x[l] -
+                                                 recovered[l]))/x[l])*100)
+    # print("MAD relative (log) = ", np.median(np.abs(np.log(x[l]) -
+    #                                            np.log(recovered[l]))
+    #                                          /np.log(x[l]))*100j)
 
     errs = .5*(lnerrp[l] + lnerrm[l])
     plt.clf()
@@ -142,5 +149,7 @@ def plot(prior, nbins):
 
 if __name__ == "__main__":
     FIG_DIR = "/Users/ruthangus/projects/GProtation/documents/figures"
+    print("ACF prior")
     plot(True, 100)  # with ACF prior
+    print("\n", "No prior")
     plot(False, 100)  # without ACF prior

### 2017-02-03 documents/ms.tex
adding MAD/MRD table

@@
 \newcommand{\princeton}{2}
-\newcommand{\naigrain}{333}
-\newcommand{\nmcquillan}{100}
-\newcommand{\nkois}{1102}
-\newcommand{\nkoimcq}{275}
-% \newcommand{\nkoigood}{222}
-% \newcommand{\nkoibad}{49}
-\newcommand{\kepexample}{5809890}
-\newcommand{\kepexampleperiod}{30.5}
-\newcommand{\aigrainexampleperiod}{20.8}
+
 \newcommand{\Kepler}{{\it Kepler}}
@@
 \newcommand{\numax}{$\nu_{\mathrm{max}}$}
-\newcommand{\acfRMS}{0.60}
-\newcommand{\pgramRMS}{0.98}
-\newcommand{\mcmcRMS}{0.46}
+
+\newcommand{\naigrain}{333}
+\newcommand{\nmcquillan}{100}
+\newcommand{\nkois}{1102}
+\newcommand{\nkoimcq}{275}
+\newcommand{\kepexample}{5809890}
+\newcommand{\kepexampleperiod}{30.5}
+\newcommand{\aigrainexampleperiod}{20.8}
+
+\newcommand{\lnacfMAD}{0.04}
+\newcommand{\acfMAD}{0.60}
+\newcommand{\percentacfMAD}{3.65}
+\newcommand{\lnpgramMAD}{0.04}
+\newcommand{\pgramMAD}{0.98}
+\newcommand{\percentpgramMAD}{4.32}
+\newcommand{\lngpMADnp}{0.02}
+\newcommand{\gpMADnp}{0.48}
+\newcommand{\percentgpMADnp}{2.45}
+\newcommand{\lngpMAD}{0.02}
+\newcommand{\gpMAD}{0.38}
+\newcommand{\percentgpMAD}{1.85}
 
@@
 rotationally modulated stellar light curve.
-Physical models of the stellar surface also have many drawbacks preventing effective
-inference, such as highly degenerate or high-dimensional parameter spaces.
+Physical models of the stellar surface also have many drawbacks preventing
+    effective inference, such as highly degenerate or high-dimensional
+    parameter spaces.
 % \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
-In this work, we introduce an appropriate {\it effective} model: a Gaussian Process (GP)
-with a quasi-periodic covariance kernel function.
-Using this highly flexible model,
-we are able to sample from the posterior Probability Density Function
-(PDF) of the periodic parameter, marginalising over the other kernel hyperparameters
-using a Markov Chain Monte Carlo approach.
+In this work, we introduce an appropriate {\it effective} model: a Gaussian
+    Process (GP) with a quasi-periodic covariance kernel function.
+Using this highly flexible model, we are able to sample from the posterior
+    Probability Density Function (PDF) of the periodic parameter,
+    marginalising over the other kernel hyperparameters using a Markov Chain
+    Monte Carlo approach.
 To test this method's effectiveness, we infer rotation periods from
@@
 period, respectively.
-Both the informative and uninformative prior versions of the GP method perform
-better than the ACF and periodogram methods, with respective MADs of 0.38 and
-0.48, compared to \acfRMS\ (ACF) and \pgramRMS\ (periodogram).
+In order to assess the performance of the GP and other period recovery
+methods, we compute Median Absolute Deviations (MADs) of the results, relative
+to the input periods.
+We also compute the Median Relative Deviation (MRD), as a percentage.
+These metrics are presented for the three different methods tested in this
+paper in table \ref{tab:MADs}.
+The informative and uninformative prior versions of the GP method have MRDs of
+\percentgpMAD\ \% and \percentgpMADnp \% respectively.
 The marginal posterior distributions of the QP kernel hyperparameters for the
@@
 The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
-The Median Absolute Deviation (MAD) of the ACF-recovered periods is \acfRMS\
-in log-space (1.8 days).
+The MRD of the ACF-recovered periods is \percentacfMAD\ \% (see table
+\ref{tab:MADs} for a side-by-side comparison with the other methods).
 The agreement between the injected and recovered rotation periods is good in
@@
 periods in figure \ref{fig:pgram_compare}.
-The MAD for rotation periods measured using the LS periodogram technique is
-\pgramRMS\ in log-space (2.7 days).
+The MRD of the pgram-recovered periods is \percentpgramMAD\ \% (see table
+\ref{tab:MADs} for a side-by-side comparison with the other methods).
 There is a large number of drastically overestimated periods recovered using
@@
 \end{figure*}
+
+\begin{table*}
+\begin{center}
+    \caption{Median absolute (MAD) and median relative (MRD) deviations for
+    the ACF, LS periodogram and GP period recovery methods.}
+\begin{tabular}{lccc}
+Method & MAD & MRD \\
+    \hline
+    ACF & \acfMAD\ days & \percentacfMAD\ \% \\
+    LS periodogram & \pgramMAD\ days & \percentpgramMAD\ \% \\
+    GP (no prior) & \gpMADnp\ days & \percentgpMADnp\ \% \\
+    GP (acf prior) & \gpMAD\ days & \percentgpMAD\ \% \\
+\end{tabular}
+\end{center}
+\end{table*}
+\label{tab:MADs}
 

### 2017-02-03 documents/ms.tex
editing the conclusion

@@
 
-We used the \naigrain\ simulated \kepler-like light curves for solid-body
-rotation from \citet{Aigrain2015} and attempted to recover the rotation
-periods used to generate them.
-Three methods were compared: a Lomb-Scargle periodogram method, an
-autocorrelation function method and a new Gaussian process method.
-We demonstrate that the GP method produced the most precise and accurate
-rotation periods of the three techniques.
-As expected, it provides a large improvement over the periodogram method and a
-moderate one over the ACF method.
-We find that the ACF method often slightly under-predicts the rotation period
-due to a subtlety in the standard peak detection algorithm.
-% The periodogram method performs poorly because a sinusoid is not a perfect
-% model choice for stellar light curves --- they do not always display
-% sinusoidal behaviour.
-% On the other hand, the ACF method does not rely on any model choice and
+We attempted to recover the rotation periods of \naigrain\ simulated
+\kepler-like light curves for solid-body rotators in \citet{Aigrain2015} using
+three different methods: a new Gaussian process method, a Lomb-Scargle
+periodogram method and an autocorrelation function method.
+We demonstrate that the GP method produces the most accurate rotation periods
+of the three techniques, providing a large improvement over the LS periodogram
+method and a moderate one over the ACF method.
+We also find that the standard version of the ACF method, most commonly
+implemented in the literature, often slightly under-predicts the rotation
+period due to a subtle feature of its peak detection algorithm.
 In addition, we compared the rotation periods of \kepler\ objects of
@@
 
-The GP method provides approximate uncertainties because it produces posterior
-PDF samples but they are still imperfect estimates of the intrinsic
-uncertainties.
-This is because the posterior PDF of the rotation period parameter is, in many
-cases, multi-modal.
+Unlike the ACF and LS periodogram methods, the GP method provides posterior
+PDF samples which can be used to estimate rotation periods uncertainties.
+Although an improvement on competing methods, these uncertainties are still
+approximate because is many cases the posterior PDFs of the parameters are
+multi-modal.
+In addition, although the GP model used here is clearly a good one, it is
+still only an effective model, not an accurate physical model.
+It can only capture the posterior PDF for the periodic component of the
+covariance matrix, not the rotation period of a physical star.
 If it were possible to run an MCMC sampler for an infinite number of steps,
-the entire posterior PDF would eventually be sampled, producing an accurate
-uncertainty measurement.
-We find that only a quarter of the reported uncertainties are representative.
+the entire posterior PDF would eventually be sampled, and a more accurate
+uncertainty would result.
+Unfortunately, due to realistic computational resources and the limitations of
+MCMC sampling techniques for multi-modal posteriors, we are not always able to
+map out the entire posterior PDFs.
+We find that only a quarter of the reported uncertainties are accurate.
 
@@
 that is probabilistic.
-Probabilistic periods methods are necessary for hierarchical Bayesian
-inference and for performing population analysis.
-Using this method to generate a catalogue of probabilistic rotation periods
-could, for example, reveal the period distribution of (and therefore
-potentially the age distribution) of stars in the Milky Way.
-There is still some work to be done on the development of this model such that
-the uncertainty estimates are truly representative, but this new method
-provides a improvement upon previous methods in that regard.
+Probabilistic periods are necessary for hierarchical Bayesian inference,
+particularly when performing population analysis.
+This new GP method is capable of generating a catalogue of probabilistic
+rotation periods which could, for example, reveal the period distribution of
+(and therefore potentially the age distribution) of stars in the Milky Way.
+There is still room for improvement of this model since that the uncertainty
+estimates are not yet truly representative and it is still only an `effective'
+model, not a physical one, but it is probabilistic, it provides more accurate
+periods and, we argue, is the best method for rotation period inference
+currently available.
 

### 2017-02-03 documents/ms.tex
small edits

@@
 }
-
 \altaffiltext{\columbia}{Simons Fellow, Department of Astronomy, Columbia
@@
 UK}
+\altaffiltext{\uw}{Department of Astronomy, University of
+Washington, Seattle, WA}
 \altaffiltext{\sagan}{Sagan Fellow}
-\altaffiltext{\uw}{Department of Astronomy, University of Washington, Seattle,
-WA}
 
@@
     Monte Carlo approach.
-To test this method's effectiveness, we infer rotation periods from
+To test the effectiveness of this method, we infer rotation periods from
 \naigrain\ simulated stellar light curves, demonstrating that our results
@@
 differential rotation on the stellar surface produce quasi-periodicity.
-A strictly periodic sinusoid therefore does not well describe
-rotationally modulated stellar light curves.
+A strictly periodic sinusoid is therefore not necessarily a good model choice
+for these time-series.
 A physically realistic model of the stellar surface
-would ideally perfectly capture the complexity of shapes
+would, ideally,  perfectly capture the complexity of shapes
 within stellar light curves as well as the quasi-periodic nature, allowing for
@@
 latitude, size, temperature and lifetime.
-Considering that stars may have hundreds of spots, the number
-of free parameters in such a model quickly becomes unwieldy, especially to explore
-its posterior probability distribution function (PDF).
+Considering that stars may have hundreds of spots, the number of free
+parameters in such a model quickly becomes unwieldy, especially to explore its
+posterior Probability Density Function (PDF).
 Simplified spot models, such as the one described in \citet{Lanza2014} where
@@
 In fact, as long as the signal is approximately periodic the ACF will
-display a peak at the rotation period, no matter its precise shape.
+display a peak at the rotation period, no matter its shape.
 A drawback of the ACF method, however, is that it requires data to be
@@
 in other scientific fields such as biology, geophysics and cosmology.
-More recently, GPs have been used in stellar and exoplanet fields to capture
-stellar variability or instrumental systematics \citep[see \eg][]{Gibson2012,
-Haywood2014, Dawson2014, Barclay2015, Haywood2015, Evans2015, Rajpaul2015,
-Rajpaul2016, Aigrain2016}.
+More recently, GPs have been used in the stellar and exoplanet fields within
+astronomy, to capture stellar variability or instrumental systematics
+\citep[see \eg][]{Gibson2012, Haywood2014, Dawson2014, Barclay2015,
+Haywood2015, Evans2015, Rajpaul2015, Rajpaul2016, Aigrain2016}.
 They are useful in regression problems involving any stochastic process,
@@
 where $N$ is the number of data points, a GP can describe that dataset.
-An in-depth introduction to Gaussian processes is provided by
+An in-depth introduction to Gaussian processes is provided in
 \citet{Rasmussen2005}.
@@
 signal and defines the form of the covariance matrix.
-As a qualitatively demonstration, we present the time-series
-in Figure \ref{fig:GP_example}:
-the \kepler\ light curve of KIC \kepexample, a relatively active star that
-rotates once every $\sim$ 4 days, with stochastic variability typical of
-\kepler\ FGK stars.
+As a qualitative demonstration, we present the time-series in Figure
+\ref{fig:GP_example}: the \kepler\ light curve of KIC \kepexample.
+This is a relatively active star that rotates once every $\sim$
+\kepexampleperiod\ days, with stochastic variability typical of \kepler\ FGK
+stars.
 Clearly, data points in this light curve are correlated.
@@
 widely separated are loosely correlated.
-A GP models the way in which this correlation varies with the separation between
-data points; that is, it models not the data directly but rather the {\it
-covariance structure}.
+A GP models this variation in correlation with the separation between data
+points; that is, it models the {\it covariance structure} rather than the data
+directly.
 This lends GPs their flexibility---they can model any time
@@
 For example, the simplest and most commonly used kernel function, the `Squared
-Exponential' (SE), defined as follows, adequately fits the
+Exponential' (SE), defined as follows, could adequately fit the
 KIC \kepexample\ light curve:
@@
 % \citep[see \eg][]{Rajpaul2015}
-
 However, The SE kernel function does not well describe the covariance
@@
 Inferring rotation periods thus requires a periodic kernel
-function.  For this reason, we use the `Quasi-Periodic' kernel.
-This kernel is used by \citet{Rasmussen2005} to capture the variable behaviour
-of CO$_2$ emissions from the Mauna Loa volcano in Hawai'i.
-It is defined as
+function.
+For this reason, we use the `Quasi-Periodic' kernel.
+\citet{Rasmussen2005} model QP variability in CO$_2$ concentration on the
+summit of the Mauna Loa volcano in Hawaii \citep[data from][]{Keeling2004}
+using a kernel which is the product of a periodic and a SE kernel: the QP
+kernel.
+This kernel is defined as
 \begin{equation}
@@
 This paper is laid out as follows.
-In \textsection \ref{sec:simulations} of this paper we describe the simulated
-light curves used to test the three rotation period recovery
-methods (ACF, LS periodogram and GP).
+In \textsection \ref{sec:simulations} we describe the simulated light curves
+used to test the three rotation period recovery methods (ACF, LS periodogram
+and GP).
 The GP method is described in \textsection \ref{sec:method} and the ACF and LS
@@
 
-In order to test the GP method we attempted to the measure rotation periods of
+In order to test the GP method we attempted to measure the rotation periods of
 a set of simulated light curves.
@@
 The distributions of input parameters used to generate these light curves are
-listed in \citet{Aigrain2015}.
+listed in table \ref{tab:simulation_parameters}.
 Light curves were simulated with a real \Kepler\ long-cadence time array:
@@
 angles, activity levels, spot lifetimes and more.
-Once simulated, the light curves were added to real \kepler\ light curves with
-no obvious astrophysical variability in order to preserve the noise properties
-of the time series.
+Once simulated, the light curves were added to real, quiet \kepler\ light
+curves --- with no obvious astrophysical variability --- in order to preserve
+the noise properties of the time series.
 Figure \ref{fig:demo_lc} shows an example of a simulated light curve with a
-period of \aigrainexampleperiod days.
-
-The ranges and distributions of the physical stellar parameters used in the
-simulated light curves are tabulated below, in table
-\ref{tab:simulation_parameters}, and figure \ref{fig:period_hist} shows the
-distribution of rotation periods in the \citet{Aigrain2015} sample.
+period of \aigrainexampleperiod\ days.
 
@@
 \begin{equation}
-p({\bm \theta}\,|\,y) \propto \mathcal L(y\,|\,{\bm \theta}) p({\bm \theta}),
+p({\bm \theta}\,|\,y) \propto p(y\,|\,{\bm \theta}) p({\bm \theta}),
 \end{equation}
@@
 where $y$ are the light curve data, $\bm \theta$ are the hyperparameters
-of the kernel described in equation \ref{eq:QP}, $\mathcal L$ is the
-GP likelihood function, and $p({\bm \theta})$ is the prior on the
-hyperparameters.  The GP likelihood is similar to the simple Gaussian likelihood
-function used for optimisation problems where the uncertainties are
-Gaussian and uncorrelated. The latter can be written
+of the kernel described in equation \ref{eq:QP}, $p(y\,|\,{\bm \theta}) =
+\mathcal{L}$, is the GP likelihood function, and $p({\bm \theta})$ is the prior
+ on the hyperparameters.
+The GP likelihood is similar to the simple Gaussian likelihood function used
+for optimisation problems where the uncertainties are Gaussian and
+uncorrelated.
+The latter can be written
 \begin{equation}
@@
 The informative and uninformative prior versions of the GP method have MRDs of
-\percentgpMAD\ \% and \percentgpMADnp \% respectively.
+\percentgpMAD \% and \percentgpMADnp \% respectively.
 The marginal posterior distributions of the QP kernel hyperparameters for the
@@
     light curve in figure \ref{fig:demo_lc}.
-$\sigma$ is an
-additional white noise term added to the diagonal elements of the covariance
-matrix. It is the fraction by which the observational uncertainties have been
-underestimated. If the uncertainties reported on the data are too small, this
-parameter will be non-zero.
     The green line in the period panel shows the injected period.
@@
 
-\section{Comparison with standard methods}
+\section{Comparison with literature methods}
 \label{sec:comparison}
@@
 periods in figure \ref{fig:pgram_compare}.
-The MRD of the pgram-recovered periods is \percentpgramMAD\ \% (see table
+The MRD of the pgram-recovered periods is \percentpgramMAD\% (see table
 \ref{tab:MADs} for a side-by-side comparison with the other methods).
@@
     \hline
-    ACF & \acfMAD\ days & \percentacfMAD\ \% \\
-    LS periodogram & \pgramMAD\ days & \percentpgramMAD\ \% \\
-    GP (no prior) & \gpMADnp\ days & \percentgpMADnp\ \% \\
-    GP (acf prior) & \gpMAD\ days & \percentgpMAD\ \% \\
+    ACF & \acfMAD\ days & \percentacfMAD \% \\
+    LS periodogram & \pgramMAD\ days & \percentpgramMAD \% \\
+    GP (no prior) & \gpMADnp\ days & \percentgpMADnp \% \\
+    GP (acf prior) & \gpMAD\ days & \percentgpMAD \% \\
 \end{tabular}
@@
 In the majority of these cases the uncertainties are underestimated due to the
-multi-modal nature of the period posterior PDFs which makes them difficult to
+multi-modal nature of the period posterior PDFs, which makes them difficult to
 sample.
@@
 imperfect, effective model.
-A perfect physical model of the star would produce better results with more
-representative uncertainties.
+A perfect physical model of the star would still produce better results with
+more representative uncertainties.
 
@@
 This is something that we hope to test in future.
-In addition to this, there are several other aspects of the GP method we hope
-to test in future, listed as follows:.
+In addition to this, there are several other aspects of the GP method we are
+continuing to develop, listed as follows:.
 \begin{itemize}
@@
 We attempted to recover the rotation periods of \naigrain\ simulated
-\kepler-like light curves for solid-body rotators in \citet{Aigrain2015} using
+\kepler-like light curves for solid-body rotators \citep{Aigrain2015} using
 three different methods: a new Gaussian process method, a Lomb-Scargle
@@
 Although an improvement on competing methods, these uncertainties are still
-approximate because is many cases the posterior PDFs of the parameters are
+approximate because in many cases the posterior PDFs of the parameters are
 multi-modal.
-In addition, although the GP model used here is clearly a good one, it is
-still only an effective model, not an accurate physical model.
-It can only capture the posterior PDF for the periodic component of the
-covariance matrix, not the rotation period of a physical star.
 If it were possible to run an MCMC sampler for an infinite number of steps,
@@
 map out the entire posterior PDFs.
+In addition, although the GP model used here is clearly a good one, it is
+still only an effective model, not an accurate physical model.
+It can only capture the posterior PDF of the periodic component of the
+covariance matrix, not the actual rotation period of a physical star.
 We find that only a quarter of the reported uncertainties are accurate.
 
-The main aim of this work was to develop a rotation period inference method
-that is probabilistic.
+The main aim of this work is to develop a probabilistic rotation period
+inference method.
 Probabilistic periods are necessary for hierarchical Bayesian inference,
@@
 This new GP method is capable of generating a catalogue of probabilistic
-rotation periods which could, for example, reveal the period distribution of
+rotation periods which could, for example, reveal the period distribution
 (and therefore potentially the age distribution) of stars in the Milky Way.
-There is still room for improvement of this model since that the uncertainty
-estimates are not yet truly representative and it is still only an `effective'
-model, not a physical one, but it is probabilistic, it provides more accurate
-periods and, we argue, is the best method for rotation period inference
-currently available.
+There is still room for improvement since the uncertainty estimates are not
+yet truly representative and it is still only an `effective' model, not a
+physical one.
+However, it is probabilistic and provides more accurate periods than
+alternative methods.
+We therefore argue that the GP method is the best method for rotation period
+inference currently available.
 

### 2017-02-03 documents/ms.tex
prelim middle sections/reorganization

--- 2017-02-03 documents/ms.tex
@@
 \newcommand{\princeton}{2}
-
+\newcommand{\naigrain}{333}
+\newcommand{\nmcquillan}{100}
+\newcommand{\nkois}{1102}
+\newcommand{\nkoimcq}{275}
+% \newcommand{\nkoigood}{222}
+% \newcommand{\nkoibad}{49}
+\newcommand{\kepexample}{1430163}
+\newcommand{\kepexampleperiod}{4}
+\newcommand{\aigrainexampleperiod}{20.8}
 \newcommand{\Kepler}{{\it Kepler}}
@@
 \newcommand{\numax}{$\nu_{\mathrm{max}}$}
-
-\newcommand{\naigrain}{333}
-\newcommand{\nmcquillan}{100}
-\newcommand{\nkois}{1102}
-\newcommand{\nkoimcq}{275}
-\newcommand{\kepexample}{5809890}
-\newcommand{\kepexampleperiod}{30.5}
-\newcommand{\aigrainexampleperiod}{20.8}
-
-\newcommand{\lnacfMAD}{0.04}
-\newcommand{\acfMAD}{0.60}
-\newcommand{\percentacfMAD}{3.65}
-\newcommand{\lnpgramMAD}{0.04}
-\newcommand{\pgramMAD}{0.98}
-\newcommand{\percentpgramMAD}{4.32}
-\newcommand{\lngpMADnp}{0.02}
-\newcommand{\gpMADnp}{0.48}
-\newcommand{\percentgpMADnp}{2.45}
-\newcommand{\lngpMAD}{0.02}
-\newcommand{\gpMAD}{0.38}
-\newcommand{\percentgpMAD}{1.85}
+\newcommand{\acfRMS}{0.60}
+\newcommand{\pgramRMS}{0.98}
+\newcommand{\mcmcRMS}{0.46}
 
 \newcommand{\dfmcomment}[1]{{\color{red}#1}}
+\newcommand{\tdmcomment}[1]{{\color{blue}#1}}
 
@@
 rotationally modulated stellar light curve.
-Physical models of the stellar surface also have many drawbacks preventing
-    effective inference, such as highly degenerate or high-dimensional
-    parameter spaces.
+Physical models of the stellar surface also have many drawbacks preventing effective
+inference, such as highly degenerate or high-dimensional parameter spaces.
 % \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
-In this work, we introduce an appropriate {\it effective} model: a Gaussian
-    Process (GP) with a quasi-periodic covariance kernel function.
-Using this highly flexible model, we are able to sample from the posterior
-    Probability Density Function (PDF) of the periodic parameter,
-    marginalising over the other kernel hyperparameters using a Markov Chain
-    Monte Carlo approach.
+In this work, we introduce an appropriate {\it effective} model: a Gaussian Process (GP)
+with a quasi-periodic covariance kernel function.
+Using this highly flexible model,
+we are able to sample from the posterior Probability Density Function
+(PDF) of the periodic parameter, marginalising over the other kernel hyperparameters
+using a Markov Chain Monte Carlo approach.
 To test this method's effectiveness, we infer rotation periods from
@@
 inferring rotation periods for \nkois\ \Kepler\ stars.
-Because this method delivers posteriors PDFs it will enable hierarchical
-    studies involving stellar rotation, particularly those involving
-    population modelling, such as inferring stellar ages, obliquities in
-    exoplanet systems, or characterising star-planet interactions.
+Because this method delivers posteriors PDFs,
+it will enable other studies involving stellar rotation,
+such as inferring stellar
+ages or stellar obliquities in exoplanet systems, or characterising star-planet interactions.
+\dfmcomment{Why do posteriors ``enable'' other studies? Hierarchical
+    inference? This could include more of an argument here.}
 % Thehe ACF method is not applicable to unevenly spaced data,
@@
 GPs are commonly used in the machine learning community and increasingly
-in other scientific fields such as biology, geophysics and cosmology.
-More recently, GPs have been used in stellar and exoplanet fields to capture
-stellar variability or instrumental systematics \citep[see \eg][]{Gibson2012,
-Haywood2014, Dawson2014, Barclay2015, Haywood2015, Evans2015, Rajpaul2015,
-Rajpaul2016, Aigrain2016}.
+in other scientific fields such as biology and geophysics.
+More recently, GPs have been used in the astronomy \dfmcomment{(this should
+probably say something more specific because GPs have been used in other parts
+of astronomy for decades)} literature \citep[see
+\eg][]{Gibson2012, Haywood2014, Dawson2014, Barclay2015, Haywood2015,
+Evans2015, Rajpaul2015, Rajpaul2016, Aigrain2016}.
 They are useful in regression problems involving any stochastic process,
@@
 Inferring rotation periods thus requires a periodic kernel
-function.  For this reason, we use the `Quasi-Periodic' kernel.
-This kernel is used by \citet{Rasmussen2005} to capture the variable behaviour
-of CO$_2$ emissions from the Mauna Loa volcano in Hawai'i.
-It is defined as
+function.  For this reason, we use the `Quasi-Periodic' kernel, defined as
+\citep{Rasmussen2005}
 \begin{equation}
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/koi_lc_demo.pdf}
+\includegraphics[width=6in, clip=true]{figures/001430163.pdf}
 \caption[A light curve with a GP model.]
@@
 
-This paper is laid out as follows.
 In \textsection \ref{sec:simulations} of this paper we describe the simulated
-light curves used to test the three rotation period recovery
-methods (ACF, LS periodogram and GP).
-The GP method is described in \textsection \ref{sec:method} and the ACF and LS
-periodogram methods are described in \textsection \ref{sec:comparison}.
-In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data,
-and the results are discussed in \textsection \ref{sec:discussion}.
+light curves used to test the three rotation period recovery methods---periodogram,
+ACF, and GP---that we describe in detail in \textsection \ref{sec:method}.
+In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data, and
+we discuss the results in \textsection \ref{sec:discussion}.
+
+\section{GP Rotation Period Inference}
+
+In order to recover a stellar rotation period from a light curve using a
+Gaussian processes, we sample the following posterior PDF:
+\begin{equation}
+p({\bm \theta}\,|\,y) \propto \mathcal L(y\,|\,{\bm \theta}) p({\bm \theta}),
+\end{equation}
+\label{eq:posterior}
+where $y$ are the light curve flux data, $\bm \theta$ are the hyperparameters
+of the kernel described in Equation \ref{eq:QP}, $\mathcal L$ is the
+GP likelihood function, and $p({\bm \theta})$ is the prior on the
+hyperparameters.  The GP likelihood is similar to the simple Gaussian likelihood
+function used for optimisation problems where the uncertainties are
+Gaussian and uncorrelated. The latter can be written
+\begin{equation}
+\ln \mathcal{L} = -\frac{1}{2}\sum_{n=1}^N\frac{(y_n-\mu)^2}{\sigma_n^2}
+    - \frac{N}{2}\ln(2\pi\sigma_n^2),
+\end{equation}
+\label{eq:chi2}
+where $y_n$ are the data, $\mu$ is the mean model and $\sigma_n$ are the
+Gaussian uncertainites on the data.
+The equivalent equation in matrix notation is
+\begin{equation}
+\ln \mathcal{L} = -\frac{1}{2}\bf{r}^T\bf{C}^{-1}\bf{r}-\ln|\bf{C}|
+    + \mathrm{constant},
+\end{equation}
+\label{eq:lhf1}
+where $\bf{r}$ is the vector of residuals and $\bf{C}$ is the covariance
+matrix,
+\begin{eqnarray}
+    \mathbf{C} &=& \left (\begin{array}{cccc}
+    \sigma^2_1 & \sigma_{2, 1} & \cdots & \sigma_{N, 1} \\
+    \sigma_{1, 2} & \sigma^2_2 & \cdots & \sigma_{N, 2} \\
+    && \vdots & \\
+    \sigma_{1, N} & \sigma_{2, N} & \cdots & \sigma^2_N
+\end{array}\right )
+\end{eqnarray}
+In the case where the uncertainties are uncorrelated, the noise is `white',
+(which is a frequent assumption made by astronomers and is sometimes
+justified) and the off-diagonal elements of the covariance matrix are zero.
+However, in the case where there is evidence for correlated
+`noise'\footnote{In our case the `noise' is actually the model!  Incidentally, this approach is the reverse of the regression techniques
+usually employed by astronomers.
+In most problems in astronomy one tries to infer the parameters that describe
+the mean model and, if correlated noise is present, to marginalise over that
+noise.
+Here, the parameters describing the correlated noise are what we are
+interested in and our mean model is simply a straight line at $y=0$.}, as in the
+case of \Kepler\ light curves, those off-diagonal elements are non-zero.
+With GP regression, a covariance matrix generated by the kernel function
+${\bf K}$ replaces ${\bf C}$ in the above equation.
+
+Sampling the posterior PDF described above presents several challenges:
+\begin{itemize}
+\item Evaluating $\mathcal L$ for a large number of points
+can be computationally expensive.  For example, evaluating
+$\mathcal L$ for an entire \Kepler\ lightcurve
+($\sim$40,000 points) takes about $\sim$5\,s--- too slow
+to perform inference on large numbers of light curves\footnote{All
+computational times cited in this section are based on evaluations on a
+single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.
+The matrix operations necessary to evaluate the GP likelihood\footnote{These
+operations use the fast matrix solver HODLR \citep{Ambikasaran2014},
+implemented in the {\tt george} \citep{George} python package.} scale as
+$N\ln(N)^2$, where $N$ is the number of data points in the light curve.
+
+\item The flexibility of this GP model allows for posterior multimodality and
+``over-fitting''--like behavior.
+For example, if $l$ is small,  the non-periodic factor in the covariance
+        kernel may dominate, allowing for a good fit to the data without
+        requiring any periodic covariance structure---even if clear periodic
+        structure is present.
+Additionally, if $\Gamma$ is too large, the GP model becomes extremely
+        flexible and can fit the data without varying the period.
+
+\item In some cases, the posterior may also be multi-modal in period---that is,
+there may be significant posterior density at half or twice the true period $P$.
+\end{itemize}
+
+We address the difficulty of computational speed in two distinct ways:
+subsampling the data and splitting the light curve into independent sections.
+To subsample \Kepler\ data, for example, 
+we randomly select 1/30th of the points 
+in the full light curve (an average of $\sim$1.5 points per day).
+This decreases the likelihood evaluation time by a factor of about 50, down to
+about 100\,ms.
+We then split the light curve into equal-sized chunks containing approximately
+300 points per section (corresponding to about 200 days), and evaluate the
+log-likelihood as the sum of the log-likelihoods of the individual sections
+(all using the same parameters ${\bm \theta}$).
+This reduces computation time because the section-based likelihood evaluation
+scales as $mn\ln(n)^2$, where $n$ is the number of data points per section and
+$m$ is the number of sections.
+This method further reduces computation time for a typical light curve
+(subsampled by a factor of 30) by about a factor of two, to about 50\,ms.
+
+Managing the flexibilty of the GP model to focus on reliably
+retrieving the period parameter requires imposing priors on the non-period
+GP parameters.
+
+Additionally,
+
+In order to manage the flexibilty of the GP model and focus on reliably
+retrieving the period parameter, we also impose priors on the non-period GP
+parameters, learned during the development of this method.
+Beginning with very broad priors (uniform in log space between -20 and 20) on
+all the parameters, we first explored the typical values of the non-period
+parameters when the injected periods were successfully recovered.
+We also experimented with constraining the allowed ranges of the parameters
+after discovering that some regions of parameter space (such as large values
+of $A$ and $\Gamma$ and small values of $l$) tended to allow fits that ignored
+the desired periodicity.
+The priors and bounds we used for this work are listed in table
+\ref{tab:priors}.
+We note that we do not have any specific quantitative justification for the
+use of these priors, except that they produce good results for this particular
+test set of light curves.
+We thus caution that other data sets might have different noise properties,
+potentially requiring modifications to these priors on the non-period
+parameters (e.g., see Section \ref{sec:kepler}).
+For the period prior, we tried two different methods: an uninformed (log-flat)
+prior between 0.5\,d and 100\,d, and an ACF-informed prior (details in
+Appendix).
 
@@
 
-In order to test the GP method we attempted to the measure rotation periods of
-a set of simulated light curves.
-We used \naigrain\ light curves generated for the \citet{Aigrain2015}
+In order to test our new rotation period recovery method and compare it to
+other standard procedures, we use a set of simulated light curves: namely,
+\naigrain\ light curves generated for the \citet{Aigrain2015}
 `hare and hounds' rotation period recovery experiment.
-These light curves were simulated by placing dark, circular spots with slowly
-evolving size, on the surface of bright, rotating spheres, ignoring
+These light curves result from placing dark, circular spots with slowly
+evolving size on the surface of bright, rotating spheres, ignoring
 limb-darkening effects.
-\citet{Aigrain2015} simulated one thousand light curves in order to test the
+\citet{Aigrain2015} simulated one thousand such light curves to test the
 ability of participating teams to recover both the stellar rotation periods
-and the rotational shear: the amplitude of surface differential rotation.
+and the rotational shear (the amplitude of surface differential rotation).
 % Since we are not interested in recovering differential rotation we only used
@@
 % there are \naigraincurves.
-For the purpose of demonstrating that our method can recover rotation periods,
-we selected only the \naigrain\ out of the 1000 without differential
-rotation.
-We opted to use only solid-body rotators because differential rotation may
-produce some additional scatter in the rotation period measurements recovered.
-In future we intend to test whether we can recover differential rotation using
-the GP method and will then use the full set of 1000 light curves.
-The distributions of input parameters used to generate these light curves are
-listed in \citet{Aigrain2015}.
-Light curves were simulated with a real \Kepler\ long-cadence time array:
+However, in this work, in order to focus on demonstrating reliable period recovery, 
+we select only the \naigrain\ light curves without differential rotation, 
+as differential rotation may produce additional scatter in the 
+measured rotation periods.
+In future we intend to test to what extent we can recover 
+differential rotation using the GP method and will then use the 
+full set of 1000 light curves.
+
+Each of these light curve simulations uses a real \Kepler\ long-cadence time array:
 one data point every thirty minutes over a four year duration.
-90\% of the rotation periods of the simulations were randomly drawn from a
-log-uniform distribution between 10 and 50 days and 10\% from a log-uniform
+90\% of the rotation periods of the simulations come from a
+log-uniform distribution between 10 and 50 days, and 10\% from a log-uniform
 distribution between 1 and 10 days.
-A histogram of the \naigrain\ solid-body rotation periods is shown in figure
-\ref{fig:period_hist}.
-The light curves were also generated with a range of stellar inclination
-angles, activity levels, spot lifetimes and more.
-Once simulated, the light curves were added to real \kepler\ light curves with
-no obvious astrophysical variability in order to preserve the noise properties
-of the time series.
+Figure \ref{fig:period_hist} shows the distribution of solid-body rotation periods.
+The simulations also have a range of stellar inclination
+angles, activity levels, spot lifetimes and more 
+(see Table \ref{tab:simulation_parameters}).
+In order to preserve \kepler\ noise properties, 
+\citet{Aigrain2015} add real \kepler\ light curves with no obvious astrophysical
+variability to the theoretical rotational modulated light curves. 
 Figure \ref{fig:demo_lc} shows an example of a simulated light curve with a
 period of \aigrainexampleperiod days.
-
-The ranges and distributions of the physical stellar parameters used in the
-simulated light curves are tabulated below, in table
-\ref{tab:simulation_parameters}, and figure \ref{fig:period_hist} shows the
-distribution of rotation periods in the \citet{Aigrain2015} sample.
 
@@
 
-\section{The GP method}
+\section{Rotation period recovery}
 \label{sec:method}
+\subsection{ACF}
+
+We measured an ACF-based period for each light curve following the method of
+\citet{Mcquillan2013}.
+An ACF is calculated for each light curve and smoothed by convolving with a
+Gaussian with $\sigma=9$ days.
+A rotation period is estimated as the lag-time of the highest peak in the ACF,
+less than 100 days.
+This is not always the first peak: the second can be larger than the first if
+there are two active regions at or near opposite longitudes on the surface of
+the star---producing a light curve with two dips per rotation period.
+We restrict our search to periods less than 100 days because this is the
+maximum period that is realistically recoverable in \kepler\ light curves ---
+instrumental noise significantly distorts signals on longer timescales.
+An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
+in figure \ref{fig:demo_acf}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/demo_ACF.pdf}
+\caption[ACF of a simulated light curve.]
+{An autocorrelation function of the simulated light curve shown in figure
+\ref{fig:demo_lc}.
+The vertical blue line shows the period measured using the ACF method (20.4
+days) and the pink dashed line shows the period that was used to simulate the
+light curve (\aigrainexampleperiod\ days).}
+\label{fig:demo_acf}
+\end{center}
+\end{figure}
+
+The ACF method has proven to be extremely useful for measuring rotation
+periods.
+The catalogue of rotation periods of \Kepler\ stars provided in
+\citet{Mcquillan2013} has been widely used by the community and has provided
+ground-breaking results for stellar and exoplanetary science.
+The results of the ACF method as tested in \citet{Aigrain2015} were positive
+(see, for example their figure 8) as it produced a large number of accurate
+rotation period measurements.
+Another advantage of the ACF method is that it is fast to implement.
+A clear disadvantage however, as already mentioned above, are the poorly
+defined uncertainties for rotation periods estimated using an autocorrelation
+function.
+This is related to the fact that ACF period detection is a non-probabilistic
+method, another drawback of ACF periods.
+
+We applied the ACF method to the sample of \naigrain\ simulated light
+curves.
+Periods measured using the ACF method are plotted against the original
+rotation period values used to generate the light curves in figure
+\ref{fig:compare_acf}.
+The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
+The Median Absolute Deviation (MAD) of the ACF-recovered periods is \acfRMS\
+in log-space (1.8 days).
+The agreement between the injected and recovered rotation periods is good in
+general, however some rotation periods are drastically over- and
+underestimated.
+Of the points clustered close to the $x=y$ line, more fall slightly below it
+than above it, \ie\ rotation periods are preferentially underestimated.
+This is a feature of the peak position measurement method that is performed on
+the ACF.
+ACFs of stellar light curves have similar functional forms to the QP kernel
+function in equation \ref{eq:QP}: periodic functions added to decaying
+exponentials.
+In such functions the peak positions can be shifted towards the left, the
+short period end, because the decaying exponential raises the left side of
+each peak more than the right.
+It is possible to model this effect, however the standard practice is to
+simply measure the peak position without taking it into account.
+We adopt the standard practice approach here in order to provide a comparison
+between our new method and that used largely in the literature.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
+\caption[ACF results.]
+{The `true' rotation periods used to generate \naigrain\ simulated light
+curves vs the rotation periods measured using the ACF technique.
+    Several light curves have over-estimated rotation periods and some
+    are drastically underestimated.}
+\label{fig:compare_acf}
+\end{center}
+\end{figure*}
+
+\subsection{LS periodogram}
+
+For each simulated light curve, a LS periodogram
+\footnote{LS periodograms were calculated using the gatspy Python module:
+\url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
+was computed over a grid of 10,000 periods, evenly spaced in frequency,
+between 1 and 100 days.
+The period of the highest peak in the periodogram was adopted as the rotation
+period.
+The resulting recovered rotation periods are plotted as a function of the true
+periods in figure \ref{fig:pgram_compare}.
+The MAD for rotation periods measured using the LS periodogram technique is
+\pgramRMS\ in log-space (2.7 days).
+There is a large number of drastically overestimated periods recovered using
+the periodogram.
+Excess power at long periods can be generated by long-term trends that are not
+related to rotation.
+This problem becomes exacerbated when instrumental noise is also present in
+light curves.
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
+\caption[LS periodogram results.]
+{The `true' rotation periods used to generate \naigrain\ simulated light
+curves vs the rotation periods measured using a LS periodogram technique.
+In many cases a large peak at a long period was present in the periodogram,
+    producing a significant over-estimate of the period.
+    }
+\label{fig:pgram_compare}
+\end{center}
+\end{figure*}
+
+\subsection{The GP method}
 
@@
 % \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_02_03.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_emcee3_24_0.png}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_noprior_02_03.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_noprior.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 period, respectively.
-In order to assess the performance of the GP and other period recovery
-methods, we compute Median Absolute Deviations (MADs) of the results, relative
-to the input periods.
-We also compute the Median Relative Deviation (MRD), as a percentage.
-These metrics are presented for the three different methods tested in this
-paper in table \ref{tab:MADs}.
-The informative and uninformative prior versions of the GP method have MRDs of
-\percentgpMAD\ \% and \percentgpMADnp \% respectively.
+Both the informative and uninformative prior versions of the GP method perform
+better than the ACF and periodogram methods, with respective MADs of 0.38 and
+0.48, compared to \acfRMS\ (ACF) and \pgramRMS\ (periodogram).
 The marginal posterior distributions of the QP kernel hyperparameters for the
@@
 \end{figure*}
-
-\section{Comparison with standard methods}
-\label{sec:comparison}
-\subsection{ACF}
-
-We measured an ACF-based period for each light curve following the method of
-\citet{Mcquillan2013}.
-An ACF is calculated for each light curve and smoothed by convolving with a
-Gaussian with $\sigma=9$ days.
-A rotation period is estimated as the lag-time of the highest peak in the ACF,
-less than 100 days.
-This is not always the first peak: the second can be larger than the first if
-there are two active regions at or near opposite longitudes on the surface of
-the star---producing a light curve with two dips per rotation period.
-We restrict our search to periods less than 100 days because this is the
-maximum period that is realistically recoverable in \kepler\ light curves ---
-instrumental noise significantly distorts signals on longer timescales.
-An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
-in figure \ref{fig:demo_acf}.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/demo_ACF.pdf}
-\caption[ACF of a simulated light curve.]
-{An autocorrelation function of the simulated light curve shown in figure
-\ref{fig:demo_lc}.
-The vertical blue line shows the period measured using the ACF method (20.4
-days) and the pink dashed line shows the period that was used to simulate the
-light curve (\aigrainexampleperiod\ days).}
-\label{fig:demo_acf}
-\end{center}
-\end{figure}
-
-The ACF method has proven to be extremely useful for measuring rotation
-periods.
-The catalogue of rotation periods of \Kepler\ stars provided in
-\citet{Mcquillan2013} has been widely used by the community and has provided
-ground-breaking results for stellar and exoplanetary science.
-The results of the ACF method as tested in \citet{Aigrain2015} were positive
-(see, for example their figure 8) as it produced a large number of accurate
-rotation period measurements.
-Another advantage of the ACF method is that it is fast to implement.
-A clear disadvantage however, as already mentioned above, are the poorly
-defined uncertainties for rotation periods estimated using an autocorrelation
-function.
-This is related to the fact that ACF period detection is a non-probabilistic
-method, another drawback of ACF periods.
-
-We applied the ACF method to the sample of \naigrain\ simulated light
-curves.
-Periods measured using the ACF method are plotted against the original
-rotation period values used to generate the light curves in figure
-\ref{fig:compare_acf}.
-The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
-The MRD of the ACF-recovered periods is \percentacfMAD\ \% (see table
-\ref{tab:MADs} for a side-by-side comparison with the other methods).
-The agreement between the injected and recovered rotation periods is good in
-general, however some rotation periods are drastically over- and
-underestimated.
-Of the points clustered close to the $x=y$ line, more fall slightly below it
-than above it, \ie\ rotation periods are preferentially underestimated.
-This is a feature of the peak position measurement method that is performed on
-the ACF.
-ACFs of stellar light curves have similar functional forms to the QP kernel
-function in equation \ref{eq:QP}: periodic functions added to decaying
-exponentials.
-In such functions the peak positions can be shifted towards the left, the
-short period end, because the decaying exponential raises the left side of
-each peak more than the right.
-It is possible to model this effect, however the standard practice is to
-simply measure the peak position without taking it into account.
-We adopt the standard practice approach here in order to provide a comparison
-between our new method and that used largely in the literature.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
-\caption[ACF results.]
-{The `true' rotation periods used to generate \naigrain\ simulated light
-curves vs the rotation periods measured using the ACF technique.
-    Several light curves have over-estimated rotation periods and some
-    are drastically underestimated.}
-\label{fig:compare_acf}
-\end{center}
-\end{figure*}
-
-\subsection{LS periodogram}
-
-For each simulated light curve, a LS periodogram
-\footnote{LS periodograms were calculated using the gatspy Python module:
-\url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
-was computed over a grid of 10,000 periods, evenly spaced in frequency,
-between 1 and 100 days.
-The period of the highest peak in the periodogram was adopted as the rotation
-period.
-The resulting recovered rotation periods are plotted as a function of the true
-periods in figure \ref{fig:pgram_compare}.
-The MRD of the pgram-recovered periods is \percentpgramMAD\ \% (see table
-\ref{tab:MADs} for a side-by-side comparison with the other methods).
-There is a large number of drastically overestimated periods recovered using
-the periodogram.
-Excess power at long periods can be generated by long-term trends that are not
-related to rotation.
-This problem becomes exacerbated when instrumental noise is also present in
-light curves.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
-\caption[LS periodogram results.]
-{The `true' rotation periods used to generate \naigrain\ simulated light
-curves vs the rotation periods measured using a LS periodogram technique.
-In many cases a large peak at a long period was present in the periodogram,
-    producing a significant over-estimate of the period.
-    }
-\label{fig:pgram_compare}
-\end{center}
-\end{figure*}
-
-\begin{table*}
-\begin{center}
-    \caption{Median absolute (MAD) and median relative (MRD) deviations for
-    the ACF, LS periodogram and GP period recovery methods.}
-\begin{tabular}{lccc}
-Method & MAD & MRD \\
-    \hline
-    ACF & \acfMAD\ days & \percentacfMAD\ \% \\
-    LS periodogram & \pgramMAD\ days & \percentpgramMAD\ \% \\
-    GP (no prior) & \gpMADnp\ days & \percentgpMADnp\ \% \\
-    GP (acf prior) & \gpMAD\ days & \percentgpMAD\ \% \\
-\end{tabular}
-\end{center}
-\end{table*}
-\label{tab:MADs}
 
@@
 
-Figure \ref{fig:mcquillan} compares the periods inferred with the GP method to
-the ACF-based periods from \citet{Mcquillan2013} for the \nkoimcq\ overlapping
-KOIs.
-This comparison shows generally very good agreement, with only a few
-exceptions, demonstrating that this method works not only on simulated data,
-but also on real data---with the caveat that for any particular data set, some
-care is needed regarding the setting the priors and ranges for the GP
-hyperparameters.
+Figure \ref{fig:mcquillan} compares the periods inferred with the GP method to the
+ACF-based periods from \citet{Mcquillan2013} for the \nkoimcq\ overlapping KOIs.
+This comparison shows generally very good agreement, with only a few exceptions,
+demonstrating that this method works not only on simulated data, but also on
+real data---with the caveat that for any particular data set, some care is needed
+regarding the setting the priors and ranges for the GP hyperparameters.
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_koi_02_03.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_koi.pdf}
 \caption[Comparison with McQuillan results.]

### 2017-02-03 documents/ms.tex

--- 2017-02-03 documents/ms.tex
@@
 \newcommand{\dfmcomment}[1]{{\color{red}#1}}
+\newcommand{\tdmcomment}[1]{{\color{blue}#1}}
 
@@
 
+\section{GP Rotation Period Inference}
+
+In order to recover a stellar rotation period from a light curve using a
+Gaussian processes, we sample the following posterior PDF:
+\begin{equation}
+p({\bm \theta}\,|\,y) \propto \mathcal L(y\,|\,{\bm \theta}) p({\bm \theta}),
+\end{equation}
+\label{eq:posterior}
+where $y$ are the light curve flux data, $\bm \theta$ are the hyperparameters
+of the kernel described in Equation \ref{eq:QP}, $\mathcal L$ is the
+GP likelihood function, and $p({\bm \theta})$ is the prior on the
+hyperparameters.  The GP likelihood is similar to the simple Gaussian likelihood
+function used for optimisation problems where the uncertainties are
+Gaussian and uncorrelated. The latter can be written
+\begin{equation}
+\ln \mathcal{L} = -\frac{1}{2}\sum_{n=1}^N\frac{(y_n-\mu)^2}{\sigma_n^2}
+    - \frac{N}{2}\ln(2\pi\sigma_n^2),
+\end{equation}
+\label{eq:chi2}
+where $y_n$ are the data, $\mu$ is the mean model and $\sigma_n$ are the
+Gaussian uncertainites on the data.
+The equivalent equation in matrix notation is
+\begin{equation}
+\ln \mathcal{L} = -\frac{1}{2}\bf{r}^T\bf{C}^{-1}\bf{r}-\ln|\bf{C}|
+    + \mathrm{constant},
+\end{equation}
+\label{eq:lhf1}
+where $\bf{r}$ is the vector of residuals and $\bf{C}$ is the covariance
+matrix,
+\begin{eqnarray}
+    \mathbf{C} &=& \left (\begin{array}{cccc}
+    \sigma^2_1 & \sigma_{2, 1} & \cdots & \sigma_{N, 1} \\
+    \sigma_{1, 2} & \sigma^2_2 & \cdots & \sigma_{N, 2} \\
+    && \vdots & \\
+    \sigma_{1, N} & \sigma_{2, N} & \cdots & \sigma^2_N
+\end{array}\right )
+\end{eqnarray}
+In the case where the uncertainties are uncorrelated, the noise is `white',
+(which is a frequent assumption made by astronomers and is sometimes
+justified) and the off-diagonal elements of the covariance matrix are zero.
+However, in the case where there is evidence for correlated
+`noise'\footnote{In our case the `noise' is actually the model!  Incidentally, this approach is the reverse of the regression techniques
+usually employed by astronomers.
+In most problems in astronomy one tries to infer the parameters that describe
+the mean model and, if correlated noise is present, to marginalise over that
+noise.
+Here, the parameters describing the correlated noise are what we are
+interested in and our mean model is simply a straight line at $y=0$.}, as in the
+case of \Kepler\ light curves, those off-diagonal elements are non-zero.
+With GP regression, a covariance matrix generated by the kernel function
+${\bf K}$ replaces ${\bf C}$ in the above equation.
+
+Sampling the posterior PDF described above presents several challenges:
+\begin{itemize}
+\item Evaluating $\mathcal L$ for a large number of points
+can be computationally expensive.  For example, evaluating
+$\mathcal L$ for an entire \Kepler\ lightcurve
+($\sim$40,000 points) takes about $\sim$5\,s--- too slow
+to perform inference on large numbers of light curves\footnote{All
+computational times cited in this section are based on evaluations on a
+single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.
+The matrix operations necessary to evaluate the GP likelihood\footnote{These
+operations use the fast matrix solver HODLR \citep{Ambikasaran2014},
+implemented in the {\tt george} \citep{George} python package.} scale as
+$N\ln(N)^2$, where $N$ is the number of data points in the light curve.
+
+\item The flexibility of this GP model allows for posterior multimodality and
+``over-fitting''--like behavior.
+For example, if $l$ is small,  the non-periodic factor in the covariance
+        kernel may dominate, allowing for a good fit to the data without
+        requiring any periodic covariance structure---even if clear periodic
+        structure is present.
+Additionally, if $\Gamma$ is too large, the GP model becomes extremely
+        flexible and can fit the data without varying the period.
+
+\item In some cases, the posterior may also be multi-modal in period---that is,
+there may be significant posterior density at half or twice the true period $P$.
+\end{itemize}
+
+We address the difficulty of computational speed in two distinct ways:
+subsampling the data and splitting the light curve into independent sections.
+To subsample \Kepler\ data, for example, 
+we randomly select 1/30th of the points 
+in the full light curve (an average of $\sim$1.5 points per day).
+This decreases the likelihood evaluation time by a factor of about 50, down to
+about 100\,ms.
+We then split the light curve into equal-sized chunks containing approximately
+300 points per section (corresponding to about 200 days), and evaluate the
+log-likelihood as the sum of the log-likelihoods of the individual sections
+(all using the same parameters ${\bm \theta}$).
+This reduces computation time because the section-based likelihood evaluation
+scales as $mn\ln(n)^2$, where $n$ is the number of data points per section and
+$m$ is the number of sections.
+This method further reduces computation time for a typical light curve
+(subsampled by a factor of 30) by about a factor of two, to about 50\,ms.
+
+Managing the flexibilty of the GP model to focus on reliably
+retrieving the period parameter requires imposing priors on the non-period
+GP parameters.
+
+Additionally,
+
+In order to manage the flexibilty of the GP model and focus on reliably
+retrieving the period parameter, we also impose priors on the non-period GP
+parameters, learned during the development of this method.
+Beginning with very broad priors (uniform in log space between -20 and 20) on
+all the parameters, we first explored the typical values of the non-period
+parameters when the injected periods were successfully recovered.
+We also experimented with constraining the allowed ranges of the parameters
+after discovering that some regions of parameter space (such as large values
+of $A$ and $\Gamma$ and small values of $l$) tended to allow fits that ignored
+the desired periodicity.
+The priors and bounds we used for this work are listed in table
+\ref{tab:priors}.
+We note that we do not have any specific quantitative justification for the
+use of these priors, except that they produce good results for this particular
+test set of light curves.
+We thus caution that other data sets might have different noise properties,
+potentially requiring modifications to these priors on the non-period
+parameters (e.g., see Section \ref{sec:kepler}).
+For the period prior, we tried two different methods: an uninformed (log-flat)
+prior between 0.5\,d and 100\,d, and an ACF-informed prior (details in
+Appendix).
+
 \section{Simulated light curves}
@@
 
-In order to test the GP method we attempted to measure the rotation periods of
-a set of simulated light curves.
-We used \naigrain\ light curves generated for the \citet{Aigrain2015}
+In order to test our new rotation period recovery method and compare it to
+other standard procedures, we use a set of simulated light curves: namely,
+\naigrain\ light curves generated for the \citet{Aigrain2015}
 `hare and hounds' rotation period recovery experiment.
-These light curves were simulated by placing dark, circular spots with slowly
-evolving size, on the surface of bright, rotating spheres, ignoring
+These light curves result from placing dark, circular spots with slowly
+evolving size on the surface of bright, rotating spheres, ignoring
 limb-darkening effects.
-\citet{Aigrain2015} simulated one thousand light curves in order to test the
+\citet{Aigrain2015} simulated one thousand such light curves to test the
 ability of participating teams to recover both the stellar rotation periods
-and the rotational shear: the amplitude of surface differential rotation.
+and the rotational shear (the amplitude of surface differential rotation).
 % Since we are not interested in recovering differential rotation we only used
@@
 % there are \naigraincurves.
-For the purpose of demonstrating that our method can recover rotation periods,
-we selected only the \naigrain\ out of the 1000 without differential
-rotation.
-We opted to use only solid-body rotators because differential rotation may
-produce some additional scatter in the rotation period measurements recovered.
-In future we intend to test whether we can recover differential rotation using
-the GP method and will then use the full set of 1000 light curves.
-The distributions of input parameters used to generate these light curves are
-listed in table \ref{tab:simulation_parameters}.
-Light curves were simulated with a real \Kepler\ long-cadence time array:
+However, in this work, in order to focus on demonstrating reliable period recovery, 
+we select only the \naigrain\ light curves without differential rotation, 
+as differential rotation may produce additional scatter in the 
+measured rotation periods.
+In future we intend to test to what extent we can recover 
+differential rotation using the GP method and will then use the 
+full set of 1000 light curves.
+
+Each of these light curve simulations uses a real \Kepler\ long-cadence time array:
 one data point every thirty minutes over a four year duration.
-90\% of the rotation periods of the simulations were randomly drawn from a
-log-uniform distribution between 10 and 50 days and 10\% from a log-uniform
+90\% of the rotation periods of the simulations come from a
+log-uniform distribution between 10 and 50 days, and 10\% from a log-uniform
 distribution between 1 and 10 days.
-A histogram of the \naigrain\ solid-body rotation periods is shown in figure
-\ref{fig:period_hist}.
-The light curves were also generated with a range of stellar inclination
-angles, activity levels, spot lifetimes and more.
-Once simulated, the light curves were added to real, quiet \kepler\ light
-curves --- with no obvious astrophysical variability --- in order to preserve
-the noise properties of the time series.
+Figure \ref{fig:period_hist} shows the distribution of solid-body rotation periods.
+The simulations also have a range of stellar inclination
+angles, activity levels, spot lifetimes and more 
+(see Table \ref{tab:simulation_parameters}).
+In order to preserve \kepler\ noise properties, 
+\citet{Aigrain2015} add real \kepler\ light curves with no obvious astrophysical
+variability to the theoretical rotational modulated light curves. 
 Figure \ref{fig:demo_lc} shows an example of a simulated light curve with a
-period of \aigrainexampleperiod\ days.
+period of \aigrainexampleperiod days.
 

### 2017-02-05 documents/ms.tex
reorganized/edited sections 2 and 3

@@
 \begin{equation}
+\label{eq:SE}
 k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2} \right).
 \end{equation}
-\label{eq:SE}
 Here $A$ is the amplitude of covariance, $l$ is the length scale of
@@
 \begin{equation}
+\label{eq:QP}
 k_{i,j} = A \exp \left[-\frac{(x_i - x_j)^2}{2l^2} -
@@
 \end{equation}
-\label{eq:QP}
 It is the product of the SE kernel function, which describes the overall
@@
 \section{GP Rotation Period Inference}
+\label{sec:method}
 
 In order to recover a stellar rotation period from a light curve using a
-Gaussian processes, we sample the following posterior PDF:
+quasi-periodic Gaussian process (QP-GP), we sample the following posterior PDF:
 \begin{equation}
+\label{eq:posterior}
 p({\bm \theta}\,|\,y) \propto \mathcal L(y\,|\,{\bm \theta}) p({\bm \theta}),
 \end{equation}
-\label{eq:posterior}
 where $y$ are the light curve flux data, $\bm \theta$ are the hyperparameters
 of the kernel described in Equation \ref{eq:QP}, $\mathcal L$ is the
-GP likelihood function, and $p({\bm \theta})$ is the prior on the
-hyperparameters.  The GP likelihood is similar to the simple Gaussian likelihood
+QP-GP likelihood function, and $p({\bm \theta})$ is the prior on the
+hyperparameters.  Sampling this posterior presents several challenges:
+\begin{itemize}
+    \item The likelihood evaulation is computationally expensive;
+    \item The GP model is very flexible, sometimes at the expense of 
+    reliable recovery of the period parameter; and
+    \item The posterior may often be multimodal.
+\end{itemize}
+This Section discusses how we address these challenges through 
+the implementation details of the likelihood (Section \ref{sec:GP_lhood}), priors 
+(Section \ref{sec:GP_prior}), and sampling method (Section \ref{sec:sampling}).
+
+\subsection{Likelihood}
+\label{sec:GP_lhood}
+
+The GP likelihood is similar to the simple Gaussian likelihood
 function used for optimisation problems where the uncertainties are
@@
 With GP regression, a covariance matrix generated by the kernel function
-${\bf K}$ replaces ${\bf C}$ in the above equation.
-
-Sampling the posterior PDF described above presents several challenges:
-\begin{itemize}
-\item Evaluating $\mathcal L$ for a large number of points
-can be computationally expensive.  For example, evaluating
+${\bf K}$ replaces ${\bf C}$ in the above equation (for our purposes, the QP 
+kernel of Equation \ref{eq:QP}).
+
+Evaluating this likeihood for a large number of points can be computationally
+expensive.  For example, evaluating
 $\mathcal L$ for an entire \Kepler\ lightcurve
@@
 
-\item The flexibility of this GP model allows for posterior multimodality and
-``over-fitting''--like behavior.
-For example, if $l$ is small,  the non-periodic factor in the covariance
-        kernel may dominate, allowing for a good fit to the data without
-        requiring any periodic covariance structure---even if clear periodic
-        structure is present.
-Additionally, if $\Gamma$ is too large, the GP model becomes extremely
-        flexible and can fit the data without varying the period.
-
-\item In some cases, the posterior may also be multi-modal in period---that is,
-there may be significant posterior density at half or twice the true period $P$.
-\end{itemize}
-
-We address the difficulty of computational speed in two distinct ways:
+We accelerate the likelihood calculation using two complementary strategies:
 subsampling the data and splitting the light curve into independent sections.
@@
 
-Managing the flexibilty of the GP model to focus on reliably
-retrieving the period parameter requires imposing priors on the non-period
-GP parameters.
-
-Additionally,
-
-In order to manage the flexibilty of the GP model and focus on reliably
-retrieving the period parameter, we also impose priors on the non-period GP
-parameters, learned during the development of this method.
-Beginning with very broad priors (uniform in log space between -20 and 20) on
-all the parameters, we first explored the typical values of the non-period
-parameters when the injected periods were successfully recovered.
-We also experimented with constraining the allowed ranges of the parameters
-after discovering that some regions of parameter space (such as large values
-of $A$ and $\Gamma$ and small values of $l$) tended to allow fits that ignored
-the desired periodicity.
-The priors and bounds we used for this work are listed in table
-\ref{tab:priors}.
-We note that we do not have any specific quantitative justification for the
-use of these priors, except that they produce good results for this particular
-test set of light curves.
-We thus caution that other data sets might have different noise properties,
-potentially requiring modifications to these priors on the non-period
-parameters (e.g., see Section \ref{sec:kepler}).
-For the period prior, we tried two different methods: an uninformed (log-flat)
-prior between 0.5\,d and 100\,d, and an ACF-informed prior (details in
-Appendix).
-
-\section{Simulated light curves}
+\subsection{Priors}
+\label{sec:GP_prior}
+
+\subsubsection{Non-period Hyperparameters}
+\label{sec:nonperiod_prior}
+
+The flexibility of this GP model allows for posterior multimodality and 
+``over-fitting''--like behavior. 
+For example, if $l$ is small, the non-periodic factor in the covariance
+kernel may dominate, allowing for a good fit to the data without
+requiring any periodic covariance structure---even if clear periodic
+structure is present.
+Additionally, if $\Gamma$ is too large, the GP model becomes extremely
+flexible and can fit the data without varying the period.
+
+Managing this flexibility to focus on reliably
+retrieving the period parameter requires imposing priors 
+on the non-period GP parameters.
+While there are no strictly {\it a priori} reasons to believe
+that any of the hyperparameters should take values in any specific range,
+we have found in this work that reasonable priors for any particular
+dataset can be determined by experimentation.  
+We note this is a somewhat subjective art, with a goal of 
+not allowing the GP to become so flexible that it does not require
+periodicity to fit the data even in the presence of clear periodicity.
+In particular, we find it necessary to avoid large values of 
+$A$ and $\Gamma$, and small values of $l$; though the exact details of what 
+works well may differ among datasets (see \ref{sec:} and 
+\ref{sec:} for two different examples of hyperparameter priors).
+
+\subsubsection{Period}
+\label{sec:period_prior}
+
+While using an uninformed prior on period (e.g., log-flat) often suffices,
+we also develop a method to construct an informed period prior, based on the 
+autocorrelation function (ACF). 
+The ACF has proven to be very useful for measuring stellar rotation periods 
+\citep{Mcquillan2012, Mcquillan2013, Mcquillan2014}; however, the
+method has several shortcomings,
+most notably the inability to deliver uncertainties, but also
+the necessity of several heuristic choices,
+such as a timescale on which to smooth the ACF,
+how to define a peak, whether the first or second peak
+gets selected, and what constitutes a secure detection.
+While this paper presents a rotation period inference method
+that avoids these shortcomings,
+it seems prudent to still use information available from the ACF.
+We thus use the ACF to define a prior on period,
+which can help the posterior sampling converge on the true period.
+
+We do not attempt to decide which single
+peak in the ACF best represents the true rotation period,
+but rather we identify several \emph{candidate} periods and define
+a weighting scheme in order to create a non-commital, though useful,
+multimodal prior.  While this procedure does not avoid heuristic choices,
+the fact that its goal is simply to create a \emph{prior} for a
+probabilitistically justified inference scheme rather than
+to identify a single correct period softens the potential impact
+of these choices.
+
+As another innovation beyond what ACF methods in the literature have
+presented, we also bandpass filter the light curves (using a 5th order
+Butterworth filter, as implemented in \texttt{SciPy}) before calculating the
+autocorrelation.  This suppresses power on timescales shorter than a chosen
+minimum period $P_{\rm min}$ and longer than a chosen maximum $P_{\rm max}$,
+producing a cleaner autocorrelation signal than an unfiltered light curve.
+
+We use the following procedure to construct a prior for rotation period
+given a light curve:
+
+\begin{enumerate}
+\item{For each value of $P_i$, where $i = \{1, 3, 5, 10, 30, 50, 100\}$\,d,
+we apply a bandpass filter to the light curve using $P_{\rm min}=0.1$\,d
+and $P_{\rm max} = P_i$.  We then calculate the ACF of the filtered
+light curve out to a maximum lag of $2P_i$ and smooth it with a boxcar
+filter of width $P_i/10$.}
+
+\item{We identify the time lag corresponding to the
+first peak of each of these ACFs, as well as the first peak's
+trough-to-peak height, creating a set of candidate periods
+$T_i$ and heights $h_i$.}
+
+\item{We assign a quality metric $Q_i$ to each of these candidate
+periods, as follows.  First, we model the ACF as a
+damped oscillator with fixed period $T_i$:
+\begin{equation}
+y = A e^{-t/\tau} \cos{\frac{2\pi t}{T_i} },
+\end{equation}
+where and $t$ is the lag time,
+and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
+least squares minimization procedure.  We then define the
+following heuristic quality metric:
+\begin{equation}
+\label{eq:quality}
+Q_i = \left(\frac{\tau_i}{T_i}\right) \left(\frac{N_i h_i}{R_i}\right),
+\end{equation}
+where $h_i$ is the height of the ACF peak at $T_i$,
+$N_i$ is the length of the lag vector in the ACF (directly proportional
+to the maximum allowed period $P_i$),
+and $R_i$ is the sum of squared residuals between the
+damped oscillator model and the actual ACF data.  The idea behind this
+quality metric to give a candidate period a higher score if
+
+    \begin{enumerate}[(a)]
+
+    \item{it has many regular sinusoidal peaks, such that the decay
+        time $\tau_i$ is long compared to the oscillation period $T_i$,}
+    \item{the ACF peak height is high, and}
+    \item{the damped oscillator model is a good fit (in a $\chi^2$
+      sense) to the ACF, with extra bonus for being
+      a good fit over more points (larger $N_i$, or longer $P_i$).}
+
+    \end{enumerate}
+}
+
+\item{Given this set of candidate periods $T_i$ and quality metrics $Q_i$,
+we finally construct a multimodal prior on the $P$ parameter of the GP
+model as a weighted mixture of Gaussians:
+\begin{equation}
+\label{eq:mixture}
+p(\ln P) = \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) +
+                                          0.05\mathcal N(\ln (T_i/2), 0.2) +
+                                          0.05\mathcal N(\ln (2 T_i), 0.2) \right)}
+                {\sum_i Q_i}.
+\end{equation}
+That is, in addition to taking the candidate periods themselves as mixture
+components, we also mix in twice and half each candidate period at a lower level,
+which compensates for the possibility that the first peak in the ACF may actually
+represent half or twice the actual rotation period.  The period width of 0.2 in
+log space (corresponding to roughly 20\% uncertainty) is again a heuristic choice,
+balancing a healthy specificity with the desire to not have the results of
+the inference overly determined by the ACF prior.
+}
+\end{enumerate}
+
+Incidentally, while we use the procedure described here to create a prior on
+$P$ which we use while inferring the parameters of the quasi-periodic GP model,
+this same procedure may also be used in the service of a
+rotation-period estimating procedure all on its own, perhaps being even more
+robust and accurate than the traditional ACF method.  We leave exploration of
+this possibility to future work.
+
+\subsection{Sampling}
+\label{sec:sampling}
+
+To sample the posterior in a way that is sensitive to potential multimodality,
+we use the \texttt{emcee3}\footnote{\url{https://github.com/dfm/emcee3}} MCMC
+sampler.
+\texttt{emcee3} is the successor to the \texttt{emcee} project
+\citep{Foreman-Mackey2013} that includes a suite of ensemble MCMC proposals
+that can be combined to efficiently sample more distributions than the stretch
+move \citep{Goodman2010} in \texttt{emcee}.
+For this project, we use a weighted mixture of three proposals.
+First, we include a proposal based on the \texttt{kombine}
+package\footnote{\url{https://github.com/bfarr/kombine}} (Farr \& Farr, in
+prep.) where a kernel density estimate (KDE) of the density represented by the
+complementary ensemble is used as the proposal for the other walkers.
+The other two proposals are a ``Differential Evolution (DE) MCMC'' proposal
+\citep{terBraak2006, Nelson2014} and the ``snooker'' extension of DE
+\citep{terBraak2008}.
+
+We initialize 500 walkers with random samples from the prior and use a
+weighted mixture of the KDE, DE, and snooker proposals with weights of 0.4,
+0.4, and 0.2 respectively.
+We run fifty steps of the sampler at a time, checking for convergence after
+each iteration, up to a maximum of fifty iterations.
+We declare convergence if the total effective chain length is at least
+8$\times$ the maximum autocorrelation time.
+When convergence is achieved, we drop the first two autocorrelation lengths in
+the chain as a burn-in, and randomly choose 5000 samples as representative of
+the posterior.
+This fitting process takes several hours for a typical simulated light curve,
+though in some cases it can take 12 hours or longer to converge.
+
+\section{Performance and Comparison to Literature: Simulated Data}
+\label{sec:perf_and_comp}
+
+In order to benchmark this new rotation period recovery method, we apply 
+it to a set of simulated light curves
+and compare to the performance of established literature methods.
+Section \ref{sec:simulations} describes the simulated data we use; 
+Section \ref{sec:performance} demonstrates the performance of the 
+QP-GP method; and Section \ref{sec:comparison} compares to the performance
+of the Lomb-Scargle periodogram and autocorrelation function methods.
+
+\subsection{Simulated light curves}
 \label{sec:simulations}
 
-In order to test our new rotation period recovery method and compare it to
-other standard procedures, we use a set of simulated light curves: namely,
-\naigrain\ light curves generated for the \citet{Aigrain2015}
-`hare and hounds' rotation period recovery experiment.
+We take our test data set from the \citet{Aigrain2015} `hare and hounds' 
+rotation period recovery experiment.
 These light curves result from placing dark, circular spots with slowly
@@
 
-% We attempted to recover the rotation periods of these \naigrain\ light curves
-% using three methods: the ACF method; the LS periodogram method; and the GP
-% method, and compared their performances.
-
-\section{The GP method}
-\label{sec:method}
-
-In order to recover the rotation period of a simulated light curve using
-Gaussian processes, we sample the following posterior PDF:
-\begin{equation}
-p({\bm \theta}\,|\,y) \propto p(y\,|\,{\bm \theta}) p({\bm \theta}),
-\end{equation}
-\label{eq:posterior}
-where $y$ are the light curve data, $\bm \theta$ are the hyperparameters
-of the kernel described in equation \ref{eq:QP}, $p(y\,|\,{\bm \theta}) =
-\mathcal{L}$, is the GP likelihood function, and $p({\bm \theta})$ is the prior
- on the hyperparameters.
-The GP likelihood is similar to the simple Gaussian likelihood function used
-for optimisation problems where the uncertainties are Gaussian and
-uncorrelated.
-The latter can be written
-\begin{equation}
-\ln \mathcal{L} = -\frac{1}{2}\sum_{n=1}^N\frac{(y_n-\mu)^2}{\sigma_n^2}
-    - \frac{N}{2}\ln(2\pi\sigma_n^2),
-\end{equation}
-\label{eq:chi2}
-where $y_n$ are the data, $\mu$ is the mean model and $\sigma_n$ are the
-Gaussian uncertainites on the data.
-The equivalent equation in matrix notation is
-\begin{equation}
-\ln \mathcal{L} = -\frac{1}{2}\bf{r}^T\bf{C}^{-1}\bf{r}-\ln|\bf{C}|
-    + \mathrm{constant},
-\end{equation}
-\label{eq:lhf1}
-where $\bf{r}$ is the vector of residuals and $\bf{C}$ is the covariance
-matrix,
-\begin{eqnarray}
-	\mathbf{C} &=& \left (\begin{array}{cccc}
-	\sigma^2_1 & \sigma_{2, 1} & \cdots & \sigma_{N, 1} \\
-	\sigma_{1, 2} & \sigma^2_2 & \cdots & \sigma_{N, 2} \\
-    && \vdots & \\
-	\sigma_{1, N} & \sigma_{2, N} & \cdots & \sigma^2_N
-\end{array}\right )
-\end{eqnarray}
-In the case where the uncertainties are uncorrelated, the noise is `white',
-(which is a frequent assumption made by astronomers and is sometimes
-justified) and the off-diagonal elements of the covariance matrix are zero.
-However, in the case where there is evidence for correlated
-`noise'\footnote{In our case the `noise' is actually the model!  Incidentally, this approach is the reverse of the regression techniques
-usually employed by astronomers.
-In most problems in astronomy one tries to infer the parameters that describe
-the mean model and, if correlated noise is present, to marginalise over that
-noise.
-Here, the parameters describing the correlated noise are what we are
-interested in and our mean model is simply a straight line at $y=0$.}, as in the
-case of \Kepler\ light curves, those off-diagonal elements are non-zero.
-With GP regression, a covariance matrix generated by the kernel function,
-${\bf K}$ replaces ${\bf C}$ in the above equation.
-
-Sampling the posterior PDF described above presents several challenges:
-\begin{itemize}
-\item Evaluating $\mathcal L$ for an entire \Kepler\ lightcurve
-($\sim$40,000 points) takes about $\sim$5\,s--- too computationally expensive
-to perform inference on large numbers of light curves\footnote{All
-computational times cited in this section are based on evaluations on a
-single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.
-The matrix operations necessary to evaluate the GP likelihood\footnote{These
-operations use the fast matrix solver HODLR \citep{Ambikasaran2014},
-implemented in the {\tt george} \citep{George} python package.} scale as
-$N\ln(N)^2$, where $N$ is the number of data points in the light curve.
-
-\item The flexibility of this GP model allows for posterior multimodality and
-``over-fitting''--like behavior.
-For example, if $l$ is small,  the non-periodic factor in the covariance
-        kernel may dominate, allowing for a good fit to the data without
-        requiring any periodic covariance structure---even if clear periodic
-        structure is present.
-Additionally, if $\Gamma$ is too large, the GP model becomes extremely
-        flexible and can fit the data without varying the period.
-
-\item In some cases, the posterior may also be multi-modal in period---that is,
-there may be significant posterior density at half or twice the true period $P$.
-\end{itemize}
-
-We address the difficulty of computational speed in two distinct ways:
-subsampling the data and splitting the light curve into independent sections.
-To subsample, we randomly select 1/30th of the points in the full light curve
-(an average of $\sim$1.5 points per day).
-This decreases the likelihood evaluation time by a factor of about 50, down to
-about 100\,ms.
-We then split the light curve into equal-sized chunks containing approximately
-300 points per section (corresponding to about 200 days), and evaluate the
-log-likelihood as the sum of the log-likelihoods of the individual sections
-(all using the same parameters ${\bm \theta}$).
-This reduces computation time because the section-based likelihood evaluation
-scales as $mn\ln(n)^2$, where $n$ is the number of data points per section and
-$m$ is the number of sections.
-This method further reduces computation time for a typical light curve
-(subsampled by a factor of 30) by about a factor of two, to about 50\,ms.
-
-In order to manage the flexibilty of the GP model and focus on reliably
-retrieving the period parameter, we also impose priors on the non-period GP
-parameters, learned during the development of this method.
-Beginning with very broad priors (uniform in log space between -20 and 20) on
-all the parameters, we first explored the typical values of the non-period
-parameters when the injected periods were successfully recovered.
+\subsection{Method Performance}
+\label{sec:performance}
+
+We apply the QP-GP inference method described in Section \ref{sec:method} 
+to each of these 333 simulated light curves.  As discussed in Section
+\ref{sec:nonperiod_prior}, reliable inference
+requires defining a useful set of priors on the non-period
+hyperparameters.
+For this simulation dataset, we determined these by first running 
+the method using very broad priors on all the non-period parameters 
+(log-flat between -20 and 20)
+and then inspecting the distribution of their posteriors for those cases 
+that successfully recovered the true period.  
 We also experimented with constraining the allowed ranges of the parameters
@@
 the desired periodicity.
-The priors and bounds we used for this work are listed in table
-\ref{tab:priors}.
+We list the final priors and bounds this process led us to adopt
+in Table \ref{tab:priors}.
 We note that we do not have any specific quantitative justification for the
@@
 We thus caution that other data sets might have different noise properties,
-potentially requiring modifications to these priors on the non-period
-parameters (e.g., see Section \ref{sec:kepler}).
+potentially requiring modifications to these priors 
+(e.g., see Section \ref{sec:kepler}).
 For the period prior, we tried two different methods: an uninformed (log-flat)
-prior between 0.5\,d and 100\,d, and an ACF-informed prior (details in
-Appendix).
-% We note that our default prior for period is log-uniform between the bounds,
-% but that we were also able to further improve our results by using an
-% ACF-inspired prior.
-
-To sample the posterior in a way that is sensitive to potential multimodality,
-we use the \texttt{emcee3}\footnote{\url{https://github.com/dfm/emcee3}} MCMC
-sampler.
-\texttt{emcee3} is the successor to the \texttt{emcee} project
-\citep{Foreman-Mackey2013} that includes a suite of ensemble MCMC proposals
-that can be combined to efficiently sample more distributions than the stretch
-move \citep{Goodman2010} in \texttt{emcee}.
-For this project, we use a weighted mixture of three proposals.
-First, we include a proposal based on the \texttt{kombine}
-package\footnote{\url{https://github.com/bfarr/kombine}} (Farr \& Farr, in
-prep.) where a kernel density estimate (KDE) of the density represented by the
-complementary ensemble is used as the proposal for the other walkers.
-The other two proposals are a ``Differential Evolution (DE) MCMC'' proposal
-\citep{terBraak2006, Nelson2014} and the ``snooker'' extension of DE
-\citep{terBraak2008}.
-
-We initialize 500 walkers with random samples from the prior and use a
-weighted mixture of the KDE, DE, and snooker proposals with weights of 0.4,
-0.4, and 0.2 respectively.
-We run fifty steps of the sampler at a time, checking for convergence after
-each iteration, up to a maximum of fifty iterations.
-We declare convergence if the total effective chain length is at least
-8$\times$ the maximum autocorrelation time.
-When convergence is achieved, we drop the first two autocorrelation lengths in
-the chain as a burn-in, and randomly choose 5000 samples as representative of
-the posterior.
-This fitting process takes several hours for a typical simulated light curve,
-though in some cases it can take 12 hours or longer to converge.
+prior between 0.5\,d and 100\,d, and an ACF-informed prior 
+(Section \ref{sec:period_prior}).
+
+Figures \ref{fig:compare_mcmc_acfprior} and \ref{fig:compare_mcmc_noprior}
+summarize the results of this MCMC fitting procedure compared to the injected
+`true' stellar rotation periods, for the uninformed and ACF-informed priors on
+period, respectively.
+In order to assess the performance of the GP and other period recovery
+methods, we compute Median Absolute Deviations (MADs) of the results, relative
+to the input periods.
+We also compute the Median Relative Deviation (MRD), as a percentage.
+These metrics are presented for the three different methods tested in this
+paper in table \ref{tab:MADs}.
+The informative and uninformative prior versions of the GP method have MRDs of
+\percentgpMAD \% and \percentgpMADnp \% respectively.
+The marginal posterior distributions of the QP kernel hyperparameters for the
+example simulated light curve in figure \ref{fig:demo_lc}, are shown in
+figure \ref{fig:gp_posteriors}.
 
@@
 
-Figures \ref{fig:compare_mcmc_acfprior} and \ref{fig:compare_mcmc_noprior}
-summarize the results of this MCMC fitting procedure compared to the injected
-`true' stellar rotation periods, for the uninformed and ACF-informed priors on
-period, respectively.
-In order to assess the performance of the GP and other period recovery
-methods, we compute Median Absolute Deviations (MADs) of the results, relative
-to the input periods.
-We also compute the Median Relative Deviation (MRD), as a percentage.
-These metrics are presented for the three different methods tested in this
-paper in table \ref{tab:MADs}.
-The informative and uninformative prior versions of the GP method have MRDs of
-\percentgpMAD \% and \percentgpMADnp \% respectively.
-The marginal posterior distributions of the QP kernel hyperparameters for the
-example simulated light curve in figure \ref{fig:demo_lc}, are shown in
-figure \ref{fig:gp_posteriors}.
-
 \begin{figure*}
@@
 
-\section{Comparison with literature methods}
+\subsection{Comparison with literature methods}
 \label{sec:comparison}
-\subsection{ACF}
+\subsubsection{ACF}
 
@@
 
-\subsection{LS periodogram}
+\subsubsection{LS periodogram}
 
@@
 
-\appendix
-
-\section{An ACF-informed Prior for Rotation Period}
-
-The autocorrelation function (ACF) has proven to be very useful
-for measuring the rotation periods of stars from time-series
-photometry.
-However, as discussed in the body of this paper, the
-ACF-based method has several shortcomings,
-most notably the inability to deliver uncertainties, but also
-the necessity of several heuristic choices,
-such as a timescale on which to smooth the ACF,
-how to define a peak, whether the first or second peak
-gets selected, and what constitutes a secure detection.
-While this paper presents a new MCMC-based method
-that avoids these shortcomings,
-it seems prudent to still use information available from the ACF.
-This Appendix describes how we use the ACF to define a prior on period.
-
-Our strategy here is not to attempt to decide which single
-peak in the ACF best represents the true rotation period,
-but rather to identify several \emph{candidate} periods and define
-a weighting scheme in order to create a non-commital, though useful,
-multimodal prior.  While this procedure does not avoid heuristic choices,
-the fact that its goal is simply to create a \emph{prior} for a
-probabilitistically justified inference scheme rather than
-to identify a single correct period softens the potential impact
-of these choices.
-
-As another innovation beyond what ACF methods in the literature have
-presented, we also bandpass filter the light curves (using a 5th order
-Butterworth filter, as implemented in \texttt{SciPy}) before calculating the
-autocorrelation.  This suppresses power on timescales shorter than a chosen
-minimum period $P_{\rm min}$ and longer than a chosen maximum $P_{\rm max}$,
-producing a cleaner autocorrelation signal than an unfiltered light curve.
-
-We use the following procedure to construct a prior for rotation period
-given a light curve:
-\begin{enumerate}
-\item{For each value of $P_i$, where $i = \{1, 3, 5, 10, 30, 50, 100\}$\,d,
-we apply a bandpass filter to the light curve using $P_{\rm min}=0.1$\,d
-and $P_{\rm max} = P_i$.  We then calculate the ACF of the filtered
-light curve out to a maximum lag of $2P_i$ and smooth it with a boxcar
-filter of width $P_i/10$.}
-
-\item{We identify the time lag corresponding to the
-first peak of each of these ACFs, as well as the first peak's
-trough-to-peak height, creating a set of candidate periods
-$T_i$ and heights $h_i$.}
-
-\item{We assign a quality metric $Q_i$ to each of these candidate
-periods, as follows.  First, we model the ACF as a
-damped oscillator with fixed period $T_i$:
-\begin{equation}
-y = A e^{-t/\tau} \cos{\frac{2\pi t}{T_i} },
-\end{equation}
-where and $t$ is the lag time,
-and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
-least squares minimization procedure.  We then define the
-following heuristic quality metric:
-\begin{equation}
-\label{eq:quality}
-Q_i = \left(\frac{\tau_i}{T_i}\right) \left(\frac{N_i h_i}{R_i}\right),
-\end{equation}
-where $h_i$ is the height of the ACF peak at $T_i$,
-$N_i$ is the length of the lag vector in the ACF (directly proportional
-to the maximum allowed period $P_i$),
-and $R_i$ is the sum of squared residuals between the
-damped oscillator model and the actual ACF data.  The idea behind this
-quality metric to give a candidate period a higher score if
-
-    \begin{enumerate}[(a)]
-
-    \item{it has many regular sinusoidal peaks, such that the decay
-        time $\tau_i$ is long compared to the oscillation period $T_i$,}
-    \item{the ACF peak height is high, and}
-    \item{the damped oscillator model is a good fit (in a $\chi^2$
-      sense) to the ACF, with extra bonus for being
-      a good fit over more points (larger $N_i$, or longer $P_i$).}
-
-    \end{enumerate}
-}
-
-\item{Given this set of candidate periods $T_i$ and quality metrics $Q_i$,
-we finally construct a multimodal prior on the $P$ parameter of the GP
-model as a weighted mixture of Gaussians:
-\begin{equation}
-\label{eq:mixture}
-p(\ln P) = \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) +
-                                          0.05\mathcal N(\ln (T_i/2), 0.2) +
-                                          0.05\mathcal N(\ln (2 T_i), 0.2) \right)}
-                {\sum_i Q_i}.
-\end{equation}
-That is, in addition to taking the candidate periods themselves as mixture
-components, we also mix in twice and half each candidate period at a lower level,
-which compensates for the possibility that the first peak in the ACF may actually
-represent half or twice the actual rotation period.  The period width of 0.2 in
-log space (corresponding to roughly 20\% uncertainty) is again a heuristic choice,
-balancing a healthy specificity with the desire to not have the results of
-the inference overly determined by the ACF prior.
-}
-
-Incidentally, while we use the procedure described here to create a prior on
-$P$ which we use while inferring the parameters of the quasi-periodic GP model,
-this same procedure may also be used in the service of a
-rotation-period estimating procedure all on its own, perhaps being even more
-robust and accurate than the traditional ACF method.  We leave exploration of
-this possibility to future work.
-
-\end{enumerate}
+% \appendix
+
+% \section{An ACF-informed Prior for Rotation Period}
+
+% The autocorrelation function (ACF) has proven to be very useful
+% for measuring the rotation periods of stars from time-series
+% photometry.
+% However, as discussed in the body of this paper, the
+% ACF-based method has several shortcomings,
+% most notably the inability to deliver uncertainties, but also
+% the necessity of several heuristic choices,
+% such as a timescale on which to smooth the ACF,
+% how to define a peak, whether the first or second peak
+% gets selected, and what constitutes a secure detection.
+% While this paper presents a new MCMC-based method
+% that avoids these shortcomings,
+% it seems prudent to still use information available from the ACF.
+% This Appendix describes how we use the ACF to define a prior on period.
+
+% Our strategy here is not to attempt to decide which single
+% peak in the ACF best represents the true rotation period,
+% but rather to identify several \emph{candidate} periods and define
+% a weighting scheme in order to create a non-commital, though useful,
+% multimodal prior.  While this procedure does not avoid heuristic choices,
+% the fact that its goal is simply to create a \emph{prior} for a
+% probabilitistically justified inference scheme rather than
+% to identify a single correct period softens the potential impact
+% of these choices.
+
+% As another innovation beyond what ACF methods in the literature have
+% presented, we also bandpass filter the light curves (using a 5th order
+% Butterworth filter, as implemented in \texttt{SciPy}) before calculating the
+% autocorrelation.  This suppresses power on timescales shorter than a chosen
+% minimum period $P_{\rm min}$ and longer than a chosen maximum $P_{\rm max}$,
+% producing a cleaner autocorrelation signal than an unfiltered light curve.
+
+% We use the following procedure to construct a prior for rotation period
+% given a light curve:
+% \begin{enumerate}
+% \item{For each value of $P_i$, where $i = \{1, 3, 5, 10, 30, 50, 100\}$\,d,
+% we apply a bandpass filter to the light curve using $P_{\rm min}=0.1$\,d
+% and $P_{\rm max} = P_i$.  We then calculate the ACF of the filtered
+% light curve out to a maximum lag of $2P_i$ and smooth it with a boxcar
+% filter of width $P_i/10$.}
+
+% \item{We identify the time lag corresponding to the
+% first peak of each of these ACFs, as well as the first peak's
+% trough-to-peak height, creating a set of candidate periods
+% $T_i$ and heights $h_i$.}
+
+% \item{We assign a quality metric $Q_i$ to each of these candidate
+% periods, as follows.  First, we model the ACF as a
+% damped oscillator with fixed period $T_i$:
+% \begin{equation}
+% y = A e^{-t/\tau} \cos{\frac{2\pi t}{T_i} },
+% \end{equation}
+% where and $t$ is the lag time,
+% and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
+% least squares minimization procedure.  We then define the
+% following heuristic quality metric:
+% \begin{equation}
+% \label{eq:quality}
+% Q_i = \left(\frac{\tau_i}{T_i}\right) \left(\frac{N_i h_i}{R_i}\right),
+% \end{equation}
+% where $h_i$ is the height of the ACF peak at $T_i$,
+% $N_i$ is the length of the lag vector in the ACF (directly proportional
+% to the maximum allowed period $P_i$),
+% and $R_i$ is the sum of squared residuals between the
+% damped oscillator model and the actual ACF data.  The idea behind this
+% quality metric to give a candidate period a higher score if
+
+%     \begin{enumerate}[(a)]
+
+%     \item{it has many regular sinusoidal peaks, such that the decay
+%         time $\tau_i$ is long compared to the oscillation period $T_i$,}
+%     \item{the ACF peak height is high, and}
+%     \item{the damped oscillator model is a good fit (in a $\chi^2$
+%       sense) to the ACF, with extra bonus for being
+%       a good fit over more points (larger $N_i$, or longer $P_i$).}
+
+%     \end{enumerate}
+% }
+
+% \item{Given this set of candidate periods $T_i$ and quality metrics $Q_i$,
+% we finally construct a multimodal prior on the $P$ parameter of the GP
+% model as a weighted mixture of Gaussians:
+% \begin{equation}
+% \label{eq:mixture}
+% p(\ln P) = \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) +
+%                                           0.05\mathcal N(\ln (T_i/2), 0.2) +
+%                                           0.05\mathcal N(\ln (2 T_i), 0.2) \right)}
+%                 {\sum_i Q_i}.
+% \end{equation}
+% That is, in addition to taking the candidate periods themselves as mixture
+% components, we also mix in twice and half each candidate period at a lower level,
+% which compensates for the possibility that the first peak in the ACF may actually
+% represent half or twice the actual rotation period.  The period width of 0.2 in
+% log space (corresponding to roughly 20\% uncertainty) is again a heuristic choice,
+% balancing a healthy specificity with the desire to not have the results of
+% the inference overly determined by the ACF prior.
+% }
+
+% Incidentally, while we use the procedure described here to create a prior on
+% $P$ which we use while inferring the parameters of the quasi-periodic GP model,
+% this same procedure may also be used in the service of a
+% rotation-period estimating procedure all on its own, perhaps being even more
+% robust and accurate than the traditional ACF method.  We leave exploration of
+% this possibility to future work.
 

### 2017-02-05 documents/ms.tex
acf section edits

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 Figures \ref{fig:compare_mcmc_acfprior} and \ref{fig:compare_mcmc_noprior}
-summarize the results of this MCMC fitting procedure compared to the injected
+summarize our results compared to the injected
 `true' stellar rotation periods, for the uninformed and ACF-informed priors on
 period, respectively.
-In order to assess the performance of the GP and other period recovery
+To assess the performance of the QP-GP and other period recovery
 methods, we compute Median Absolute Deviations (MADs) of the results, relative
@@
 
-We measured an ACF-based period for each light curve following the method of
+We measure an ACF-based period for each light curve, 
+following the method of
 \citet{Mcquillan2013}.
-An ACF is calculated for each light curve and smoothed by convolving with a
-Gaussian with $\sigma=9$ days.
-A rotation period is estimated as the lag-time of the highest peak in the ACF,
-less than 100 days.
-This is not always the first peak: the second can be larger than the first if
-there are two active regions at or near opposite longitudes on the surface of
-the star---producing a light curve with two dips per rotation period.
-We restrict our search to periods less than 100 days because this is the
-maximum period that is realistically recoverable in \kepler\ light curves ---
-instrumental noise significantly distorts signals on longer timescales.
-An example ACF of the light curve in figure~\ref{fig:demo_lc} is shown
-in figure \ref{fig:demo_acf}.
+After calculating the ACF for a light curve, we smooth it by convolving with a
+Gaussian filter ($\sigma=9$\,d), and select the rotation period as the
+lag-time of the highest peak in the ACF less than 100\,d.
+This is not always the first peak---the second can be larger than the first if
+two active regions are at or near opposite longitudes on the surface of
+the star, producing a light curve with two dips per rotation period.
+We restrict our search to periods less than 100\,d because instrumental
+noise and light-curve systematics significantly distort \kepler\ light 
+curves on longer timescales, rendering longer periods realistically unrecoverable.
+Figure \ref{fig:demo_lc} contains an example ACF of the light curve in 
+Figure \ref{fig:demo_acf}.
 
@@
 
-The ACF method has proven to be extremely useful for measuring rotation
-periods.
+The ACF method has proven extremely useful for measuring rotation periods.
 The catalogue of rotation periods of \Kepler\ stars provided in
@@
 ground-breaking results for stellar and exoplanetary science.
-The results of the ACF method as tested in \citet{Aigrain2015} were positive
-(see, for example their figure 8) as it produced a large number of accurate
-rotation period measurements.
-Another advantage of the ACF method is that it is fast to implement.
-A clear disadvantage however, as already mentioned above, are the poorly
-defined uncertainties for rotation periods estimated using an autocorrelation
-function.
-This is related to the fact that ACF period detection is a non-probabilistic
-method, another drawback of ACF periods.
-
-We applied the ACF method to the sample of \naigrain\ simulated light
+The method also performed well in the \citet{Aigrain2015}
+recovery experiment,
+producing a large number of accurate rotation period measurements
+(see, e.g., their Figure 8).  
+Another advantage is its fast implementation speed.
+However, because the ACF method is non-probabilistic, ACF-estimated
+rotation period uncertainties are poorly defined---a clear 
+disadvantage.
+
+We apply the ACF method to the sample of \naigrain\ simulated light
 curves.
-Periods measured using the ACF method are plotted against the original
-rotation period values used to generate the light curves in figure
-\ref{fig:compare_acf}.
-The $2n$ and $\frac{1}{2}n$ harmonic lines are shown as dashed lines.
+Figure \ref{fig:compare_acf} shows ACF-measured versus true rotation
+periods, with the $2n$ and $\frac{1}{2}n$ harmonic lines as dashed lines.
 The MRD of the ACF-recovered periods is \percentacfMAD\ \% (see table
 \ref{tab:MADs} for a side-by-side comparison with the other methods).
-The agreement between the injected and recovered rotation periods is good in
-general, however some rotation periods are drastically over- and
-underestimated.
-Of the points clustered close to the $x=y$ line, more fall slightly below it
-than above it, \ie\ rotation periods are preferentially underestimated.
-This is a feature of the peak position measurement method that is performed on
-the ACF.
+The injected and recovered rotation periods generally agree well,
+though with a few drastic over- or underestimates.
+Additionally, of the points clustered close to the 1:1 line, 
+more fall slightly below than above it, \ie\ 
+rotation periods are preferentially underestimated.
+This stems from how peak positions in the ACF are measured.
 ACFs of stellar light curves have similar functional forms to the QP kernel
-function in equation \ref{eq:QP}: periodic functions added to decaying
+function in Equation \ref{eq:QP}: periodic functions added to decaying
 exponentials.
-In such functions the peak positions can be shifted towards the left, the
-short period end, because the decaying exponential raises the left side of
+In such functions the peak positions can be shifted towards the left 
+(towards shorter periods), because the decaying exponential raises the left side of
 each peak more than the right.
-It is possible to model this effect, however the standard practice is to
-simply measure the peak position without taking it into account.
-We adopt the standard practice approach here in order to provide a comparison
-between our new method and that used largely in the literature.
+It is possible to model this effect; however, standard practice 
+simply measures the peak position without taking it into account.
+We adopt this standard practice here to faithfully compare
+our new method to that used in the literature.
 

### 2017-02-05 documents/ms.tex
ls section edits

@@
 
-For each simulated light curve, a LS periodogram
+For each simulated light curve, we compute a LS periodogram
 \footnote{LS periodograms were calculated using the gatspy Python module:
 \url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
-was computed over a grid of 10,000 periods, evenly spaced in frequency,
+over a grid of 10,000 periods, evenly spaced in frequency,
 between 1 and 100 days.
-The period of the highest peak in the periodogram was adopted as the rotation
+We adopt the period of the highest peak in the periodogram as the rotation
 period.
-The resulting recovered rotation periods are plotted as a function of the true
-periods in figure \ref{fig:pgram_compare}.
-The MRD of the pgram-recovered periods is \percentpgramMAD\% (see table
+Figure \ref{fig:pgram_compare} shows the resulting recovered rotation periods
+as a function of true period.
+The MRD of the periodogram-recovered periods is \percentpgramMAD\% (see table
 \ref{tab:MADs} for a side-by-side comparison with the other methods).
-There is a large number of drastically overestimated periods recovered using
-the periodogram.
-Excess power at long periods can be generated by long-term trends that are not
-related to rotation.
-This problem becomes exacerbated when instrumental noise is also present in
-light curves.
+This method drastically overestimates many rotation periods, 
+because long-term trends unrelated to rotation can cause excess power
+at long periods---a problem further exacerbated by instrumental noise. 
+\tdmcomment{I don't quite understand this---what do we mean here by 
+``instrumental noise'' as distinct from long-term trends?}
 

### 2017-02-05 documents/ms.tex
first-pass tdm edits through section 4

@@
 \citep[e.g.][]{Reinhold2013}, Auto-Correlation Functions (ACFs),
-\citep{Mcquillan2013} and wavelet transforms \citep{Garcia2014}.
+\citep{Mcquillan13b} and wavelet transforms \citep{Garcia2014}.
 The precisions of the LS periodogram and wavelet methods are limited by the
@@
 The ACF has proven to be very useful for measuring stellar rotation periods 
-\citep{Mcquillan2012, Mcquillan2013, Mcquillan2014}; however, the
+\citep{Mcquillan2012, Mcquillan13b, Mcquillan2014}; however, the
 method has several shortcomings,
@@
 \subsubsection{ACF}
+\label{sec:acf}
 
@@
 \subsubsection{LS periodogram}
+\label{sec:ls}
 
@@
 
-In order to test this rotation period inference method on real data,
-we applied it to a set of \nkois\ \Kepler\ Object of Interest (KOI)
+In order to test our rotation period inference method on real data,
+we apply it to a set of \nkois\ \Kepler\ Object of Interest (KOI)
 host stars, \nkoimcq\ of which had rotation periods previously measured by
@@
 space, resulting in poor period measurements.  As a result, after this initial
-test, we subsequently adjusted the priors and re-fit all the KOIs.  The final
-priors and parameter ranges that we used are in Table \ref{tab:koipriors}.
+test, we subsequently adjusted the priors and re-fit all the KOIs.  
+The final priors and parameter ranges that we used are in Table \ref{tab:koipriors}.
 
@@
     $\ln \sigma$ & $\mathcal N(-17, 5)$ & (-20, 0) \\
-    $\ln P $ & Uniform / ACF-based & ($\ln 0.5, \ln 100$) \\
+    $\ln P $ & ACF-based & ($\ln 0.5, \ln 100$) \\
 \end{tabular}
@@
 hyperparameters.
+Notably, below $\ln P$ of about 2.5, the GP method recovers periods 
+systematically slightly larger than \citet{Mcquillan2013}---a likely consequence
+of correcting for the ACF peak measurement bias discussed in Section \ref{sec:acf}.
 
@@
 {A comparison of our GP rotation period measurements to those of
-\citet{Mcquillan2014}.
+\citet{Mcquillan2013}.
 The data points are coloured by the range of variability measured by
-    \citet{Mcquillan2014}, defined as the interval between the 5th and 95th
+    \citet{Mcquillan2013}, defined as the interval between the 5th and 95th
     percentiles of normalized flux per period bin in millimagnitudes.

### 2017-02-05 documents/ms.tex
a few comments

@@
 50\% lie within 2 $\sigma$ and 66\% within 3 $\sigma$.
+\tdmcomment{What are these numbers for the uninformed-prior experiment?}
 The largest outlier is 114 $\sigma$ away from the true value.
@@
 multi-modal nature of the period posterior PDFs, which makes them difficult to
-sample.
+sample. \tdmcomment{Is there any evidence of this?  Doesn't sound convincing to me.}
 The underestimated uncertainties can also be attributed to the model choice.
@@
 interest measured using the GP method to those measured previously by
-\citet{Mcquillan2014}.
+\citet{Mcquillan2013}.
 The good agreement between the two sets of results demonstrates that the GP
@@
 map out the entire posterior PDFs.
+\tdmcomment{I don't think this is the problem;
+I think the fits that we judge to have converged have actually converged.}
+
 In addition, although the GP model used here is clearly a good one, it is
@@
 This research was funded by the Simons Foundation and the Leverhulme Trust.
+TDM is supported by NASA grant NNX14AE11G, and acknowledges the 
+hospitality of the Institute for Advanced Study, 
+where part of this work was completed.
 Some of the data presented in this paper were obtained from the Mikulski

### 2017-02-06 documents/ms.tex

@@
 multi-modal nature of the period posterior PDFs, which makes them difficult to
-sample. \tdmcomment{Is there any evidence of this?  Doesn't sound convincing to me.}
+sample. \tdmcomment{Is there any evidence of this?  Doesn't sound convincing to me.
+I would think the multi-modality would only be relevant if the posterior is 
+closer to the 1/2x or 2x period, but this isn't most of them.}
 The underestimated uncertainties can also be attributed to the model choice.
@@
 A perfect physical model of the star would still produce better results with
-more representative uncertainties.
+more representative uncertainties. \tdmcomment{Not sure saying ``a perfect
+model would give better results'' is worth saying...}
 

### 2017-02-12 documents/ms.tex
As for previous one

@@
 \newcommand{\tdmcomment}[1]{{\color{blue}#1}}
+\newcommand{\vrcomment}[1]{{\color{magenta}#1}}
 
@@
 and the finite lifetimes of these active regions and
-differential rotation on the stellar surface produce quasi-periodicity.
+differential rotation on the stellar surface produce quasi-periodicity \vrcomment{\citep{dumusque2012}}.
 A strictly periodic sinusoid is therefore not necessarily a good model choice
@@
 suitability of the model choice: a sinusoid for the LS periodogram,
-and a choice of mother wavelet, assumed to desribe the data
+and a choice of mother wavelet, assumed to \vrcomment{describe [VR: spelling]} the data
 over a range of transpositions \citep[see, \eg][]{Carter2010},
@@
 A range of kernel functions can describe stellar variability.
-For example, the simplest and most commonly used kernel function, the `Squared
+For example, the simplest and most commonly used kernel function \vrcomment{[VR: it's true that it's the most commonly used kernel, but not true that it's the `simplest' -- a constant or linear kernel is simpler; an exponential is as `simple' in terms of number of hyper-parameters; etc.
+There are of course other motivations for using SE kernel; see e.g.
+Rasmussen \& Williams pg.\ 14, although these details are probably not relevant to this paper.
+Maybe it's the simplest stationary kernel function that admits infinitely-differentiable function draws, although I stand to be corrected on that.]}, the `Squared
 Exponential' (SE), defined as follows, could adequately fit the
@@
 \end{equation}
-Here $A$ is the amplitude of covariance, $l$ is the length scale of
+Here $A \vrcomment{>0}$ is the amplitude of covariance \vrcomment{[VR: either say $A>0$ or use $A^2$ in the definition of this kernel; need to ensure positive values so QP covariance matrix has nonnegative eigenvalues i.e.\ is positive semi-definite.]}, $l$ is the length scale of
 exponential decay, and $x_i-x_j$ is the separation between data points.
@@
 k_{i,j} = A \exp \left[-\frac{(x_i - x_j)^2}{2l^2} -
-\Gamma^2 \sin^2\left(\frac{2\pi}{P}\right) \right].
+\Gamma^2 \sin^2\left(\frac{\vrcomment{\pi(x_i - x_j)}}{P}\right) \right].
 \end{equation}
-It is the product of the SE kernel function, which describes the overall
+\vrcomment{[VR: there was a term missing from the above definition, viz.\ $(x_i - x_j)$ inside the $\sin()$.
+Without it, the $\sin()$ term reduces to a constant.
+Also, I changed $2\pi$ to $\pi$: if you define it with $2\pi$ then the period will differ from the physical period by a factor of $2$, given the squared term modifying the sine wave: ${\sin ^2}(\frac{\theta }{2}) = \frac{1}{2} - \frac{1}{2}\cos \theta $.
+This can be checked by plotting some covariance kernels with large $l$ and noting that peaks appear with expecting spacing $P$ only if you remove the factor of $2$ that originally appeared in the definition.]}   It is the product of the SE kernel function, which describes the overall
 covariance decay, and an exponentiated, squared, sinusoidal kernel function
@@
 controls the amplitude of the $\sin^2$ term.
-If $\Gamma$ is very small, only points almost exactly one period away are
+If $\Gamma$ is very \vrcomment{large}, only points almost exactly one period away are
 tightly correlated and points that are slightly more or less than one period
 away are very loosely correlated.
-If $\Gamma$ is large, points separated by one period are tightly
+If $\Gamma$ is \vrcomment{small}, points separated by one period are tightly
 correlated, and points separated by slightly more or less are still highly
 correlated, although less so.
+\vrcomment{[VR: behaviour for large vs.\ small values of $\Gamma$ was described backwards; i.e.\ correct explanation is that large $\Gamma$ leads to only points $\sim$~one period apart being tightly correlated, with weak correlations for other separations.
+Another way of expressing this would be to say that large values of $\Gamma$ lead to periodic variations with increasingly complex harmonic content.
+Or in non-technical terms: more bumps and wiggles per function repetition.]}
 This kernel function allows two data points that are separated in time by one
@@
 term added to the diagonal elements of the covariance matrix.
+\vrcomment{[VR: I think it might be clearer if you just included this white noise term directly in the definition of the kernel, especially for people who might not be familiar with GPs and for whom it won't necessarily be obvious how the white noise term connects to the rest of the kernel.
+I.e.\ I'd suggest you simply defined the kernel as below.]}
+\begin{equation}
+\label{eq:QP2}
+\vrcomment{k_{i,j} = A \exp \left[-\frac{(x_i - x_j)^2}{2l^2} -
+\Gamma^2 \sin^2\left(\frac{\pi\vrcomment{(x_i - x_j)}}{P}\right) \right] + \sigma^2\delta_{ij}.}
+\end{equation}
 This can be interpreted to represent underestimation of observational
@@
 and the results are discussed in \textsection \ref{sec:discussion}.
+\vrcomment{[VR: the order doesn't make sense.
+As things stand, you've laid it out as \textsection3.1,\textsection2, then  \textsection3.3.
+Presumably should be \textsection2,\textsection3.1, \textsection3.3 -- and \textsection3.2 is missing].}
 
@@
 \begin{itemize}
-    \item The likelihood evaulation is computationally expensive;
+    \item The likelihood \vrcomment{evaluation [VR: spelling]} is computationally expensive;
     \item The GP model is very flexible, sometimes at the expense of 
@@
 where $y_n$ are the data, $\mu$ is the mean model and $\sigma_n$ are the
-Gaussian uncertainites on the data.
+Gaussian \vrcomment{uncertainties [VR: spelling]} on the data. \vrcomment{VR: the above expression is incorrect. It should read:
+\begin{equation}
+\ln \mathcal{L} =  - \frac{1}{2}\sum\limits_{n = 1}^N {\frac{{{{({y_n} - \mu )}^2}}}{{\sigma _n^2}} - } \frac{1}{2}\sum\limits_{n = 1}^N {\ln (2\pi \sigma _n^2)},
+\end{equation}
+or equivalently
+\begin{equation}
+\ln \mathcal{L} =  - \frac{1}{2}\sum\limits_{n = 1}^N {\left\{ {\frac{{{{({y_n} - \mu )}^2}}}{{\sigma _n^2}} + \ln (2\pi \sigma _n^2)} \right\}}.
+\end{equation}
+The problem with the expression as written above is that either the second term must be included in the summation, assuming $\sigma_n\neq\sigma$, in which case the factor of $N$ should not appear in the second term; or if $\sigma_n=\sigma$~$\forall n$ (which now no longer represents the general case), then the factor of $N$ is correct, but the subscript $n$ should be removed from $\sigma_n^2$.}
+
 The equivalent equation in matrix notation is
@@
 \label{eq:lhf1}
-where $\bf{r}$ is the vector of residuals and $\bf{C}$ is the covariance
+\vrcomment{(VR: why not just write $\frac{N}{2}\log{2\pi}$ instead of `constant'?)}~where $\bf{r}$ is the vector of residuals and $\bf{C}$ is the covariance
 matrix,
@@
 
-Evaluating this likeihood for a large number of points can be computationally
+Evaluating this \vrcomment{likelihood [VR: spelling]} for a large number of points can be computationally
 expensive.  For example, evaluating
@@
 
+\vrcomment{VR: in my opinion, this is the weakest part of the paper.
+I don't agree that there are no \emph{a priori} reasons to believe that any of the hyperparameters should take values in any specific range, and the phrase `somewhat subjective art' seems likely to set alarm bells ringing for referees/readers.
+Going through the hyperparameters in turn, the following things come to mind.
+
+\begin{itemize}
+\item Amplitude, $A$: their is a closed form solution for the maximum-likelihood value of $A$ (solve ${\partial _A}\mathcal{L} = 0$).
+Steve Roberts alluded to this in his 2012 tutorial paper and I know I've worked through the mathematics myself -- offhand I can't remember the solution, but it's a fairly straightforward looking thing involving the precision matrix like $A\sim{{\mathbf{y}}^{\textrm{T}}}{{\mathbf{K}}^{ - 1}}{\mathbf{y}}$ with some scaling constant.
+I'll see if I can dig up my workings and/or find a reference.
+I think this means that if you're anyway going to use a flat prior on $A$, your MAP solution will end up corresponding to the ML solution.
+In which case you could speed things up a good deal by just setting $A$ to its optimal value \emph{ab initio}, and not bothering to explore the full posterior PDF.
+More importantly though, if for some reason the MAP value you use doesn't correspond more or less to the ML value (e.g.\ if something went wrong with the MCMC convergence) -- in other words, if you end up using a non-optimal value for $A$ -- I've found you can sometimes get rather bizarre behaviour.
+For example, when I fit data with an obvious period, sometimes I end up recovering a very wrong period if $A$ is also very wrong e.g.\ much too large.
+Such problems can be avoided quite straightforwardly by constraining $A$ to sensible values.
+\item Harmonic complexity, $\Gamma$: for large values of $l$ ($l\gg P$), $\Gamma$ can be related to the typical number of stationary/non-stationary points of inflection, per period, found in functions drawn from the corresponding GP.
+This behaviour will be independent of the exact value of $l$.
+For `intermediate' values of $l$ ($l\gtrsim P$) this can still be done but the behaviour will be sensitive to the value of $l$.
+And small values of $l$ e.g.\ $l<P$ are probably not relevant -- see below.
+If useful I could make a plot or two illustrating this.
+Anyway, the point is that depending on the processes giving rise to observed stellar signals, you may not expect to require certain types of functions to model observed signals.
+In typical simulations e.g.\ of arbitrary numbers of spots on rotating stars, I think you're unlikely to end up with more than say three or four turning points per period in the disk-integrated photometric signal (as opposed to say $20$; of course $2$ turning points is the minimum for periodic functions).
+Such simulations, crude though they may be, can be used to come up with physically-motivated priors for $\Gamma$.
+\item Evolution time-scale, $l$: it's sensible to enforce something like $l>P$ if we desire functions that have at least some fairly clear periodicity, as intuitively $l<P$ would mean the functions are allowed to evolve significantly over time scales shorter than one period.
+So a function $f(t+P)$ would look signficantly different to $f(t)$, in which case any claims of periodicity become dubious.
+Actually, you can come with a more rigorous criterion (which also takes into account dependencies on $\Gamma$) than this -- see below.
+\item Suzanne and I have shown that provided
+\begin{equation}
+\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} < \frac{{4\pi }}{3},
+\end{equation}
+the QP covariance function will have at least one inflection point other than at $|x_i-x_j|=0$ (the derivation is straightforward though a bit long).
+So $\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} = \frac{{4\pi }}{3}$ corresponds to a critical point at which secondary maxima disappear, and the behaviour of the kernel becomes dominated by the SE term rather than the periodic term.
+For $\Gamma\sim\mathcal{O}(1)$, this translates into $l\gtrsim P/2$.
+In practice satisfying this criterion doesn't absolutely guarantee that function draws will have clear evidence of periodicity, but it does give you a quantitative boundary between obviously periodic behaviour and not-so-obviously periodic behaviour (if you're trying to fit rotation periods then obviously you don't want to use an SE-like kernel to do so) and provides some justification for a more stringent criterion such as $l>P$, or $l>2P$ (ensure that functions evolve significantly over time-scales longer than 2 periods), etc.
+\end{itemize}
+The upshot of all of this is that I don't think constraining hyperparameters needs to be some sort of dark art or ad hockery, and it'd do the underlying method a disservice by suggesting as much.
+As I see it, there are at least two ways to proceed here.
+\begin{enumerate}
+\item Provide as much detail on constraining the hyperparameters in this paper as we possibly can, including quantitative justification, illustrations of qualitatively-different behaviours for different regions of the hyperparameter space, etc.
+Disadvantages: if we do this it'd be necessary to re-run your fits with the new priors; and the paper would be come a good deal longer, with the focus possibly shifted away from the essential idea of using a GP to infer rotation periods.
+\item Alternatively, at least allude to some of these approaches to constraining hyperparameters (instead of saying that it's a subjective art) but defer a focused discussion on how to do that to a short, separate paper/research note.
+As I've already done much of the work for such a study, I'd be happy to take the lead on that at some point later this year.
+Advantage here would be your paper remains more focused and you don't need to re-run any of your fits.
+Indeed I imagine that if you check your MAP values for $A$, $l$, etc., they'd usually be consistent with any rigorously-motivated priors (since you get generally sensible results).
+Perhaps we could show in the follow-up paper how using the physically-motivated priors could lead to improved period inference (at least I hope that would be the case!).
+Any differential rotation tests could also be shifted to the other paper -- see my later remarks on this.
+\end{enumerate}}
+
 \subsubsection{Period}
@@
 but rather we identify several \emph{candidate} periods and define
-a weighting scheme in order to create a non-commital, though useful,
+a weighting scheme in order to create a noncommital, though useful,
 multimodal prior.  While this procedure does not avoid heuristic choices,
 the fact that its goal is simply to create a \emph{prior} for a
-probabilitistically justified inference scheme rather than
+\vrcomment{probabilistically [VR: spelling]} justified inference scheme rather than
 to identify a single correct period softens the potential impact
@@
 0.4, and 0.2 respectively.
-We run fifty steps of the sampler at a time, checking for convergence after
-each iteration, up to a maximum of fifty iterations.
+We run \vrcomment{50} steps of the sampler at a time, checking for convergence after
+each iteration, up to a maximum of \vrcomment{50} iterations.
 We declare convergence if the total effective chain length is at least
 8$\times$ the maximum autocorrelation time.
-When convergence is achieved, we drop the first two autocorrelation lengths in
+When convergence is achieved, we \vrcomment{discard} the first two autocorrelation lengths in
 the chain as a burn-in, and randomly choose 5000 samples as representative of
@@
 full set of 1000 light curves.
+\vrcomment{[VR: a few months ago I did a bunch of tests with kernels I constructed specifically to account for differential rotation.
+I need to dig up that work but I remember the clear conclusion at which I arrived was that there seems to be an unbreakable degeneracy between any extra hyperparameter(s) to account for differential evolution, and the standard QP hyperparameters.
+This might change if enforcing $l\gg P$; I can't remember.
+Anyway, I mention this now just as a little warning that we might not be able to count on the GP approach yielding many useful insights as far as differential rotation goes.]}
 
@@
 We thus caution that other data sets might have different noise properties,
-potentially requiring modifications to these priors 
+potentially requiring modifications to these priors \vrcomment{[VR: might want to reword or remove these remarks; see my earlier comments]}
 (e.g., see Section \ref{sec:kepler}).
@@
 \percentgpMAD \% and \percentgpMADnp \% respectively.
-The marginal posterior distributions of the QP kernel hyperparameters for the
+The marginal posterior distributions of the QP kernel hyperparameters\vrcomment{,} for the
 example simulated light curve in figure \ref{fig:demo_lc}, are shown in
@@
     light curve in figure \ref{fig:demo_lc}.
-    The green line in the period panel shows the injected period.
+    The green line in the period panel shows the injected period. \vrcomment{[VR: might it be possible to make the green line darker? It took me a little while to find it.]}
     This figure was made using \texttt{corner.py} \citep{Corner}.}
@@
 more representative uncertainties. \tdmcomment{Not sure saying ``a perfect
-model would give better results'' is worth saying...}
+model would give better results'' is worth saying...} \vrcomment{VR: I agree this is probably not worth saying.}
 
@@
 light curve.
-It adequately describes the data, captures that all-important periodic quality
-and is relatively simple, with only a few hyper-parameters.
+It adequately describes the data, captures that all-important periodic quality \vrcomment{[VR: I think this is a bit vague...a sinusoidal model would capture `periodic quality'.
+Maybe say something like: `it manages to describe a wide range of quasi-periodic behavior'.]}
+and is relatively simple, with only a few hyperparameters.
 It also satisfies the requirement to produce positive semi-definite covariance
-matrices.
+matrices. \vrcomment{[VR: I don't think it's worth mentioning the PSD thing. It's not as though this is some advantage the QP kernel has over any other kernel function.]}
 Whilst the QP kernel function evidently captures the periodic qualities of
-light curves adequately, it is still a somewhat arbitrary choice.
+light curves adequately, it is still a somewhat arbitrary choice \vrcomment{[VR: repetitive...maybe just say 'Nevertheless, it is still a somewhat arbitrary choice.']}.
 Another valid choice would be a squared cosine function multiplied by a
@@
 k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2}\right)
-\cos\left(\frac{2\pi}{P}\right)
+\cos\left(\frac{\pi\vrcomment{(x_i - x_j)}}{P}\right)
 \end{equation}
 \label{eq:cos_kernel}
-This function produces a positive semi-definite matrix and has the $P$
+\vrcomment{VR: missing $(x_i - x_j)$ term added above, and factor of $2$ removed.} This function produces a positive semi-definite matrix and has the $P$
 parameter of interest.
@@
 example, the relative precision of rotation periods recovered using
-alternative kernel functions could be tested.}
+alternative kernel functions could be tested.} \vrcomment{[VR: this is something I've started exploring using MultiNest. Computation times seem quite manageable.]}
 \item{To design and implement a physically motivated kernel function. We have
@@
 PDFs of stars' rotation periods.}
+\item \vrcomment{[VR: for the two points above, you might want to reword slightly depending on the extent to which you want to try to (i) incorporate my suggestions re prior constraints etc.\ etc.\ vs.\ (ii) deferring this to the future paper, but at least alluding to the possibilities and results.
+For example I think one needn't try to `design' a physically-motivated kernel -- rather it seems not too difficult just to come up with physically-motivated priors for the QP kernel parameters.
+I've tried to design a physically motivated kernel to capture differential rotation but as noted earlier, that came with its own slew of degeneracies and other problems...more on this another time.]}
 \item{To build in a noise model for \kepler\ data.

### 2017-02-13 documents/ms.tex
minor edit

--- 2017-02-06 documents/ms.tex
@@
 structure is present.
-Additionally, if $\Gamma$ is too large, the GP model becomes extremely
+Additionally, for large values of $\Gamma$ the GP model becomes extremely
 flexible and can fit the data without varying the period.
@@
 multimodal prior.  While this procedure does not avoid heuristic choices,
-the fact that its goal is simply to create a \emph{prior} for a
-probabilitistically justified inference scheme rather than
-to identify a single correct period softens the potential impact
-of these choices.
+we soften the potential impact of these choices because we are simply
+creating a \emph{prior} for probabilistically justified
+inference rather attempting to identify a single correct period.
 

### 2017-02-13 documents/ms.tex
beginning to incorporate Vinesh's comments

--- 2017-02-06 documents/ms.tex
@@
 and the finite lifetimes of these active regions and
-differential rotation on the stellar surface produce quasi-periodicity.
+differential rotation on the stellar surface produce quasi-periodicity
+\citep{Dumusque2011}.
 A strictly periodic sinusoid is therefore not necessarily a good model choice
@@
 suitability of the model choice: a sinusoid for the LS periodogram,
-and a choice of mother wavelet, assumed to desribe the data
+and a choice of mother wavelet, assumed to describe the data
 over a range of transpositions \citep[see, \eg][]{Carter2010},
@@
 A range of kernel functions can describe stellar variability.
-For example, the simplest and most commonly used kernel function, the `Squared
+For example, the most commonly used kernel function, the `Squared
 Exponential' (SE), defined as follows, could adequately fit the
@@
 \end{equation}
-Here $A$ is the amplitude of covariance, $l$ is the length scale of
+Here $A>0$ is the amplitude of covariance, $l$ is the length scale of
 exponential decay, and $x_i-x_j$ is the separation between data points.
@@
 k_{i,j} = A \exp \left[-\frac{(x_i - x_j)^2}{2l^2} -
-\Gamma^2 \sin^2\left(\frac{2\pi}{P}\right) \right].
+    \Gamma^2 \sin^2\left(\frac{\pi(x_i - x_j)}{P}\right) \right] + \sigma^2
+    \delta_{ij}.
 \end{equation}
@@
 controls the amplitude of the $\sin^2$ term.
-If $\Gamma$ is very small, only points almost exactly one period away are
+If $\Gamma$ is very large, only points almost exactly one period away are
 tightly correlated and points that are slightly more or less than one period
 away are very loosely correlated.
-If $\Gamma$ is large, points separated by one period are tightly
+If $\Gamma$ is small, points separated by one period are tightly
 correlated, and points separated by slightly more or less are still highly
 correlated, although less so.
+In other words, large values of $\Gamma$ lead to periodic variations with
+increasingly complex harmonic content.
 This kernel function allows two data points that are separated in time by one
@@
 and GP).
-The GP method is described in \textsection \ref{sec:method} and the ACF and LS
-periodogram methods are described in \textsection \ref{sec:comparison}.
+The GP method is described in \textsection \ref{sec:method}.
+Its performance is demonstrated and compared with literature methods in
+in \textsection \ref{sec:perf_and_comp}.
 In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data,
@@
 \begin{itemize}
-    \item The likelihood evaulation is computationally expensive;
+    \item The likelihood evaluation is computationally expensive;
     \item The GP model is very flexible, sometimes at the expense of
@@
 \begin{equation}
-\ln \mathcal{L} = -\frac{1}{2}\sum_{n=1}^N\frac{(y_n-\mu)^2}{\sigma_n^2}
-    - \frac{N}{2}\ln(2\pi\sigma_n^2),
+    \ln \mathcal{L} = -\frac{1}{2}\sum_{n=1}^N\left[\frac{(y_n-\mu)^2}{\sigma_n^2}
+    + \ln(2\pi\sigma_n^2)\right],
 \end{equation}
@@
 where $y_n$ are the data, $\mu$ is the mean model and $\sigma_n$ are the
-Gaussian uncertainites on the data.
+Gaussian uncertainties on the data.
 The equivalent equation in matrix notation is
@@
 \ln \mathcal{L} = -\frac{1}{2}\bf{r}^T\bf{C}^{-1}\bf{r}-\ln|\bf{C}|
-    + \mathrm{constant},
+    + \frac{N}{2}\log2\pi,
 \end{equation}
@@
 
-Evaluating this likeihood for a large number of points can be computationally
+Evaluating this likelihood for a large number of points can be computationally
 expensive.  For example, evaluating
@@
 flexible and can fit the data without varying the period.
-
-Managing this flexibility to focus on reliably
-retrieving the period parameter requires imposing priors 
-on the non-period GP parameters.
-While there are no strictly {\it a priori} reasons to believe
-that any of the hyperparameters should take values in any specific range,
-we have found in this work that reasonable priors for any particular
-dataset can be determined by experimentation.  
-We note this is a somewhat subjective art, with a goal of 
-not allowing the GP to become so flexible that it does not require
-periodicity to fit the data even in the presence of clear periodicity.
+Managing this flexibility to reliably retrieve the period parameter requires
+imposing informative priors on the non-period GP parameters.
+% While there are no strictly {\it a priori} reasons to believe
+% that any of the hyperparameters should take values in any specific range,
+% we have found in this work that reasonable priors for any particular
+% dataset can be determined by experimentation.
+% We note this is a somewhat subjective art, with a goal of
+% not allowing the GP to become so flexible that it does not require
+% periodicity to fit the data even in the presence of clear periodicity.
 In particular, we find it necessary to avoid large values of
@@
 the fact that its goal is simply to create a \emph{prior} for a
-probabilitistically justified inference scheme rather than
+probabilistically justified inference scheme rather than
 to identify a single correct period softens the potential impact
@@
 measured rotation periods.
-In future we intend to test to what extent we can recover 
-differential rotation using the GP method and will then use the 
-full set of 1000 light curves.
+% In future we intend to test to what extent we can recover
+% differential rotation using the GP method and will then use the
+% full set of 1000 light curves.
 
@@
 because long-term trends unrelated to rotation can cause excess power
-at long periods---a problem further exacerbated by instrumental noise. 
-\tdmcomment{I don't quite understand this---what do we mean here by 
-``instrumental noise'' as distinct from long-term trends?}
+at long periods.
+% a problem further exacerbated by instrumental noise.
 

### 2017-02-13 documents/ms.tex

--- 2017-02-12 documents/ms.tex
@@
 and the finite lifetimes of these active regions and
-differential rotation on the stellar surface produce quasi-periodicity \vrcomment{\citep{dumusque2012}}.
+differential rotation on the stellar surface produce quasi-periodicity
+\citep{Dumusque2011}.
 A strictly periodic sinusoid is therefore not necessarily a good model choice
@@
 suitability of the model choice: a sinusoid for the LS periodogram,
-and a choice of mother wavelet, assumed to \vrcomment{describe [VR: spelling]} the data
+and a choice of mother wavelet, assumed to describe the data
 over a range of transpositions \citep[see, \eg][]{Carter2010},
@@
 A range of kernel functions can describe stellar variability.
-For example, the simplest and most commonly used kernel function \vrcomment{[VR: it's true that it's the most commonly used kernel, but not true that it's the `simplest' -- a constant or linear kernel is simpler; an exponential is as `simple' in terms of number of hyper-parameters; etc.
-There are of course other motivations for using SE kernel; see e.g.
-Rasmussen \& Williams pg.\ 14, although these details are probably not relevant to this paper.
-Maybe it's the simplest stationary kernel function that admits infinitely-differentiable function draws, although I stand to be corrected on that.]}, the `Squared
+For example, the most commonly used kernel function, the `Squared
 Exponential' (SE), defined as follows, could adequately fit the
@@
 \end{equation}
-Here $A \vrcomment{>0}$ is the amplitude of covariance \vrcomment{[VR: either say $A>0$ or use $A^2$ in the definition of this kernel; need to ensure positive values so QP covariance matrix has nonnegative eigenvalues i.e.\ is positive semi-definite.]}, $l$ is the length scale of
+Here $A>0$ is the amplitude of covariance, $l$ is the length scale of
 exponential decay, and $x_i-x_j$ is the separation between data points.
@@
 k_{i,j} = A \exp \left[-\frac{(x_i - x_j)^2}{2l^2} -
-\Gamma^2 \sin^2\left(\frac{\vrcomment{\pi(x_i - x_j)}}{P}\right) \right].
+    \Gamma^2 \sin^2\left(\frac{\pi(x_i - x_j)}{P}\right) \right] + \sigma^2
+    \delta_{ij}.
 \end{equation}
-\vrcomment{[VR: there was a term missing from the above definition, viz.\ $(x_i - x_j)$ inside the $\sin()$.
-Without it, the $\sin()$ term reduces to a constant.
-Also, I changed $2\pi$ to $\pi$: if you define it with $2\pi$ then the period will differ from the physical period by a factor of $2$, given the squared term modifying the sine wave: ${\sin ^2}(\frac{\theta }{2}) = \frac{1}{2} - \frac{1}{2}\cos \theta $.
-This can be checked by plotting some covariance kernels with large $l$ and noting that peaks appear with expecting spacing $P$ only if you remove the factor of $2$ that originally appeared in the definition.]}   It is the product of the SE kernel function, which describes the overall
+It is the product of the SE kernel function, which describes the overall
 covariance decay, and an exponentiated, squared, sinusoidal kernel function
@@
 controls the amplitude of the $\sin^2$ term.
-If $\Gamma$ is very \vrcomment{large}, only points almost exactly one period away are
+If $\Gamma$ is very large, only points almost exactly one period away are
 tightly correlated and points that are slightly more or less than one period
 away are very loosely correlated.
-If $\Gamma$ is \vrcomment{small}, points separated by one period are tightly
+If $\Gamma$ is small, points separated by one period are tightly
 correlated, and points separated by slightly more or less are still highly
 correlated, although less so.
-\vrcomment{[VR: behaviour for large vs.\ small values of $\Gamma$ was described backwards; i.e.\ correct explanation is that large $\Gamma$ leads to only points $\sim$~one period apart being tightly correlated, with weak correlations for other separations.
-Another way of expressing this would be to say that large values of $\Gamma$ lead to periodic variations with increasingly complex harmonic content.
-Or in non-technical terms: more bumps and wiggles per function repetition.]}
+In other words, large values of $\Gamma$ lead to periodic variations with
+increasingly complex harmonic content.
 This kernel function allows two data points that are separated in time by one
@@
 \begin{itemize}
-    \item The likelihood \vrcomment{evaluation [VR: spelling]} is computationally expensive;
+    \item The likelihood evaluation is computationally expensive;
     \item The GP model is very flexible, sometimes at the expense of
@@
 where $y_n$ are the data, $\mu$ is the mean model and $\sigma_n$ are the
-Gaussian \vrcomment{uncertainties [VR: spelling]} on the data. \vrcomment{VR: the above expression is incorrect. It should read:
-\begin{equation}
-\ln \mathcal{L} =  - \frac{1}{2}\sum\limits_{n = 1}^N {\frac{{{{({y_n} - \mu )}^2}}}{{\sigma _n^2}} - } \frac{1}{2}\sum\limits_{n = 1}^N {\ln (2\pi \sigma _n^2)},
-\end{equation}
-or equivalently
-\begin{equation}
-\ln \mathcal{L} =  - \frac{1}{2}\sum\limits_{n = 1}^N {\left\{ {\frac{{{{({y_n} - \mu )}^2}}}{{\sigma _n^2}} + \ln (2\pi \sigma _n^2)} \right\}}.
-\end{equation}
-The problem with the expression as written above is that either the second term must be included in the summation, assuming $\sigma_n\neq\sigma$, in which case the factor of $N$ should not appear in the second term; or if $\sigma_n=\sigma$~$\forall n$ (which now no longer represents the general case), then the factor of $N$ is correct, but the subscript $n$ should be removed from $\sigma_n^2$.}
-
+Gaussian uncertainties on the data.
 The equivalent equation in matrix notation is
@@
 
-Evaluating this \vrcomment{likelihood [VR: spelling]} for a large number of points can be computationally
+Evaluating this likelihood for a large number of points can be computationally
 expensive.  For example, evaluating
@@
 the fact that its goal is simply to create a \emph{prior} for a
-\vrcomment{probabilistically [VR: spelling]} justified inference scheme rather than
+probabilistically justified inference scheme rather than
 to identify a single correct period softens the potential impact
@@
 measured rotation periods.
-In future we intend to test to what extent we can recover 
-differential rotation using the GP method and will then use the 
-full set of 1000 light curves.
-\vrcomment{[VR: a few months ago I did a bunch of tests with kernels I constructed specifically to account for differential rotation.
-I need to dig up that work but I remember the clear conclusion at which I arrived was that there seems to be an unbreakable degeneracy between any extra hyperparameter(s) to account for differential evolution, and the standard QP hyperparameters.
-This might change if enforcing $l\gg P$; I can't remember.
-Anyway, I mention this now just as a little warning that we might not be able to count on the GP approach yielding many useful insights as far as differential rotation goes.]}
+% In future we intend to test to what extent we can recover
+% differential rotation using the GP method and will then use the
+% full set of 1000 light curves.
 
@@
 where part of this work was completed.
+VR thanks Merton College and the National Research Foundation of South Africa
+for financial support.
 Some of the data presented in this paper were obtained from the Mikulski

### 2017-02-13 documents/ms.tex
incorporating Vinesh and Tim's comments

--- 2017-02-13 documents/ms.tex
@@
 that are not captured by the effective GP model.
-%In practice, this parameter should always be non-zero when performing
-%GP regression, because the covariance matrix must be positive definite and
-%slight deviations from positive definiteness can arise due to most
-%solvers using approximate, rather than exact, matrix inversion.
-%Including a small amount of extra variance in the model ensures enough
-%flexibility for matrix inversion algorithms to not run into such numerical
-%issues.
 We use this QP kernel function to produce the GP model
@@
 \ref{sec:} for two different examples of hyperparameter priors).
+More physically motivated priors are discussed in \textsection
+\ref{sec:discussion}.
+
+% \vrcomment{VR: in my opinion, this is the weakest part of the paper.
+% I don't agree that there are no \emph{a priori} reasons to believe that any of
+% the hyperparameters should take values in any specific range, and the phrase
+% `somewhat subjective art' seems likely to set alarm bells ringing for
+% referees/readers.
+% Going through the hyperparameters in turn, the following things come to mind.
+
+% \begin{itemize}
+% \item Amplitude, $A$: their is a closed form solution for the maximum-likelihood value of $A$ (solve ${\partial _A}\mathcal{L} = 0$).
+Steve Roberts alluded to this in his 2012 tutorial paper and I know I've worked through the mathematics myself -- offhand I can't remember the solution, but it's a fairly straightforward looking thing involving the precision matrix like $A\sim{{\mathbf{y}}^{\textrm{T}}}{{\mathbf{K}}^{ - 1}}{\mathbf{y}}$ with some scaling constant.
+I'll see if I can dig up my workings and/or find a reference.
+I think this means that if you're anyway going to use a flat prior on $A$, your MAP solution will end up corresponding to the ML solution.
+In which case you could speed things up a good deal by just setting $A$ to its optimal value \emph{ab initio}, and not bothering to explore the full posterior PDF.
+More importantly though, if for some reason the MAP value you use doesn't correspond more or less to the ML value (e.g.\ if something went wrong with the MCMC convergence) -- in other words, if you end up using a non-optimal value for $A$ -- I've found you can sometimes get rather bizarre behaviour.
+For example, when I fit data with an obvious period, sometimes I end up recovering a very wrong period if $A$ is also very wrong e.g.\ much too large.
+Such problems can be avoided quite straightforwardly by constraining $A$ to sensible values.
+% \item Harmonic complexity, $\Gamma$: for large values of $l$ ($l\gg P$), $\Gamma$ can be related to the typical number of stationary/non-stationary points of inflection, per period, found in functions drawn from the corresponding GP.
+This behaviour will be independent of the exact value of $l$.
+For `intermediate' values of $l$ ($l\gtrsim P$) this can still be done but the behaviour will be sensitive to the value of $l$.
+And small values of $l$ e.g.\ $l<P$ are probably not relevant -- see below.
+If useful I could make a plot or two illustrating this.
+Anyway, the point is that depending on the processes giving rise to observed stellar signals, you may not expect to require certain types of functions to model observed signals.
+In typical simulations e.g.\ of arbitrary numbers of spots on rotating stars, I think you're unlikely to end up with more than say three or four turning points per period in the disk-integrated photometric signal (as opposed to say $20$; of course $2$ turning points is the minimum for periodic functions).
+Such simulations, crude though they may be, can be used to come up with physically-motivated priors for $\Gamma$.
+% \item Evolution time-scale, $l$: it's sensible to enforce something like $l>P$ if we desire functions that have at least some fairly clear periodicity, as intuitively $l<P$ would mean the functions are allowed to evolve significantly over time scales shorter than one period.
+So a function $f(t+P)$ would look signficantly different to $f(t)$, in which case any claims of periodicity become dubious.
+Actually, you can come with a more rigorous criterion (which also takes into account dependencies on $\Gamma$) than this -- see below.
+% \item Suzanne and I have shown that provided
+% \begin{equation}
+% \frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} < \frac{{4\pi }}{3},
+% \end{equation}
+% the QP covariance function will have at least one inflection point other than at $|x_i-x_j|=0$ (the derivation is straightforward though a bit long).
+So $\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} = \frac{{4\pi }}{3}$ corresponds to a critical point at which secondary maxima disappear, and the behaviour of the kernel becomes dominated by the SE term rather than the periodic term.
+For $\Gamma\sim\mathcal{O}(1)$, this translates into $l\gtrsim P/2$.
+In practice satisfying this criterion doesn't absolutely guarantee that function draws will have clear evidence of periodicity, but it does give you a quantitative boundary between obviously periodic behaviour and not-so-obviously periodic behaviour (if you're trying to fit rotation periods then obviously you don't want to use an SE-like kernel to do so) and provides some justification for a more stringent criterion such as $l>P$, or $l>2P$ (ensure that functions evolve significantly over time-scales longer than 2 periods), etc.
+% \end{itemize}
+% The upshot of all of this is that I don't think constraining hyperparameters needs to be some sort of dark art or ad hockery, and it'd do the underlying method a disservice by suggesting as much.
+As I see it, there are at least two ways to proceed here.
+% \begin{enumerate}
+% \item Provide as much detail on constraining the hyperparameters in this paper as we possibly can, including quantitative justification, illustrations of qualitatively-different behaviours for different regions of the hyperparameter space, etc.
+Disadvantages: if we do this it'd be necessary to re-run your fits with the new priors; and the paper would be come a good deal longer, with the focus possibly shifted away from the essential idea of using a GP to infer rotation periods.
+% \item Alternatively, at least allude to some of these approaches to constraining hyperparameters (instead of saying that it's a subjective art) but defer a focused discussion on how to do that to a short, separate paper/research note.
+As I've already done much of the work for such a study, I'd be happy to take the lead on that at some point later this year.
+Advantage here would be your paper remains more focused and you don't need to re-run any of your fits.
+Indeed I imagine that if you check your MAP values for $A$, $l$, etc., they'd usually be consistent with any rigorously-motivated priors (since you get generally sensible results).
+Perhaps we could show in the follow-up paper how using the physically-motivated priors could lead to improved period inference (at least I hope that would be the case!).
+Any differential rotation tests could also be shifted to the other paper -- see my later remarks on this.
+% \end{enumerate}}
 
@@
 but rather we identify several \emph{candidate} periods and define
-a weighting scheme in order to create a non-commital, though useful,
+a weighting scheme in order to create a noncommital, though useful,
 multimodal prior.  While this procedure does not avoid heuristic choices,
@@
 0.4, and 0.2 respectively.
-We run fifty steps of the sampler at a time, checking for convergence after
-each iteration, up to a maximum of fifty iterations.
+We run \vrcomment{50} steps of the sampler at a time, checking for convergence after
+each iteration, up to a maximum of \vrcomment{50} iterations.
 We declare convergence if the total effective chain length is at least
 8$\times$ the maximum autocorrelation time.
-When convergence is achieved, we drop the first two autocorrelation lengths in
-the chain as a burn-in, and randomly choose 5000 samples as representative of
-the posterior.
+When convergence is achieved, we discard the first two autocorrelation lengths
+in the chain as a burn-in, and randomly choose 5000 samples as representative
+of the posterior.
 This fitting process takes several hours for a typical simulated light curve,
@@
 in Table \ref{tab:priors}.
-We note that we do not have any specific quantitative justification for the
-use of these priors, except that they produce good results for this particular
-test set of light curves.
-We thus caution that other data sets might have different noise properties,
-potentially requiring modifications to these priors
-(e.g., see Section \ref{sec:kepler}).
+These priors were determined experimentally --- see \textsection
+\ref{sec:discussion} for a discussion on more physically motivated priors.
+% We note that we do not have any specific quantitative justification for the
+% use of these priors, except that they produce good results for this particular
+% test set of light curves.
+% We thus caution that other data sets might have different noise properties,
+% potentially requiring modifications to these priors \vrcomment{[VR: might
+% want to reword or remove these remarks; see my earlier comments]}
+% (e.g., see Section \ref{sec:kepler}).
 For the period prior, we tried two different methods: an uninformed (log-flat)
@@
 \percentgpMAD \% and \percentgpMADnp \% respectively.
-The marginal posterior distributions of the QP kernel hyperparameters for the
+The marginal posterior distributions of the QP kernel hyperparameters, for the
 example simulated light curve in figure \ref{fig:demo_lc}, are shown in
@@
 50\% lie within 2 $\sigma$ and 66\% within 3 $\sigma$.
-\tdmcomment{What are these numbers for the uninformed-prior experiment?}
 The largest outlier is 114 $\sigma$ away from the true value.
-In the majority of these cases the uncertainties are underestimated due to the
-multi-modal nature of the period posterior PDFs, which makes them difficult to
-sample. \tdmcomment{Is there any evidence of this?  Doesn't sound convincing to me.
-I would think the multi-modality would only be relevant if the posterior is
-closer to the 1/2x or 2x period, but this isn't most of them.}
-The underestimated uncertainties can also be attributed to the model choice.
-Although a GP model is more appropriate than a sinusoid, it is still an
-imperfect, effective model.
-A perfect physical model of the star would still produce better results with
-more representative uncertainties. \tdmcomment{Not sure saying ``a perfect
-model would give better results'' is worth saying...}
+% In the majority of these cases the uncertainties are underestimated due to the
+% multi-modal nature of the period posterior PDFs, which makes them difficult to
+% sample. \tdmcomment{Is there any evidence of this?  Doesn't sound convincing to me.
+% I would think the multi-modality would only be relevant if the posterior is
+% closer to the 1/2x or 2x period, but this isn't most of them.}
+The underestimated uncertainties can be attributed to the model choice.
+Although a GP model is more appropriate than a sinusoid, it is still only an
+ effective model.
+A perfect physical model of the star would produce more representative
+uncertainties.
 
@@
 light curve.
-It adequately describes the data, captures that all-important periodic quality
-and is relatively simple, with only a few hyper-parameters.
-It also satisfies the requirement to produce positive semi-definite covariance
-matrices.
-Whilst the QP kernel function evidently captures the periodic qualities of
-light curves adequately, it is still a somewhat arbitrary choice.
+It manages to describe a wide range of quasi-periodic behavior and is
+relatively simple, with only a few hyperparameters.
+Nevertheless, it is still a somewhat arbitrary choice.
 Another valid choice would be a squared cosine function multiplied by a
@@
 k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2}\right)
-\cos\left(\frac{2\pi}{P}\right)
+\cos\left(\frac{\pi(x_i - x_j)}{P}\right)
 \end{equation}
@@
 parameter of interest.
-% It may in fact be even {\it more} suited to modelling stellar
-% light curves as it describes a Gaussian in frequency space.
-% It is easy to imagine a differentially rotating star with a period that is a
-% Gaussian in frequency space: the mean frequency would be the frequency at the
-% most active latitude, at or near which spots spend the majority of their time
-% and the tails would be occupied by spots that drift near the equator or poles.
 The main practical difference between this cosine and the QP function is that
@@
 take minutes.
-
-% Clearly, stellar rotation periods are well represented by the $P$ parameter in
-% our QP kernel function, as evidenced by the its impressive ability to recover
-% the true rotation periods from the simulated light curves in this work.
-% However, it may not be the {\it best} function to use.
-% It is, after all, an approximation to the shape of the covariance matrix of a
-% real stellar light curve.
-% There may be an alternative function which is better suited to capturing
-% stellar variability and is able to recover periods even more precisely than
-% the QP kernel.
-% There may also be an alternative function which is more physically motivated,
-% that captures not just the rotation period but also (for example) the spot
-% lifetime or differential surface rotation.
-% This is beyond the scope of this work but we hope that these questions will be
-% answered in the near future.
 
@@
 \item{To perform model selection with different kernel functions. Again, these
-tests have not been performed due to the cost of calculating the fully
-marginalised likelihood with GPs. This calculation may be prohibitively
-expensive, however simpler model selection tests can be performed. For
-example, the relative precision of rotation periods recovered using
-alternative kernel functions could be tested.}
-\item{To design and implement a physically motivated kernel function. We have
-only explored the physical interpretation of one parameter in our kernel
-function, $P$.
-However, the other parameters may also be related to some physical processes.
-For example, the overall timescale for covariance fall-off, $l$ may be
-related to spot lifetimes.
-A star with long spot lifetimes will show little variation in the overall
-shape and amplitude of its light curve between rotations and $l$ will be
-large.
-In contrast, the light curve of a star with short spot lifetimes may display
-non-repeating patterns and amplitudes that vary rapidly between rotations.
-In this case $l$ will be small.
-The $\Gamma$ parameter is related to the number of zero crossings within one
-rotation period: when $\Gamma$ is small there are many zero crossings and
-vice versa.
-Since the number of zero crossings per rotation period is related to the
-number of active regions on the surface of the star, this parameter may also
-be of physical interest.
-In addition, instead of interpreting the parameters of the QP kernel function
-used here, it may be possible to design an entirely new kernel function, based
-on the physical processes that drive the light curve variability.
-This idea is being explored by another member of my research group.}
-% \item{To develop a detection criterion to assess whether a rotation period was
-% measured.
-% Another general problem in rotation period inference is deciding whether a
-% real rotation period was measured at all.
-% Detection thresholds are usually set using the amplitudes of peaks in LS
-% periodograms or ACFs in order to eliminate contaminants.
-% Using a GP model it may be possible to do this via model selection: \ie\ are
-% the data better described by a periodic model, or a non-periodic model?
-% There is more than one way to implement such a model comparison, one way could
-% be to use a sum of two kernel functions: one periodic and one non-periodic,
-% each with its own amplitude parameter.}
-\item{To attempt to detect differential rotation.
-We did not test our code on the light curves simulated with differential
-rotation in \citet{Aigrain2015} since we were only interested in recovering
-the most precise measurements of rotation period possible.
-In future we intend to investigate the possibility of recovering
-differential rotation by searching for close double peaks in the posterior
-PDFs of stars' rotation periods.}
+    tests have not been performed thus far due to the cost of calculating the
+        fully marginalised likelihood with GPs.}
+\item{To design and implement more physically motivated priors.
+We have only explored the physical interpretation of one parameter in our
+        kernel function, $P$.
+For example, take the harmonic complexity parameter, $\Gamma$.
+        For large values of $l$ ($l\gg P$), $\Gamma$ can be related to the
+        typical number of stationary/non-stationary points of inflection, per
+        period, found in functions drawn from the corresponding GP.
+        This behaviour will be independent of the exact value of $l$.
+        For `intermediate' values of $l$ ($l\gtrsim P$) this can still be done
+        but the behaviour will be sensitive to the value of $l$.
+        Small values of $l$ e.g.\ $l<P$ are probably not relevant -- see
+        below.
+        % Depending on the processes giving rise to observed stellar signals,
+        % you may not expect to require certain types of functions to model
+        % observed signals.
+        In typical simulations \eg\ of arbitrary numbers of spots on rotating
+        stars, you're unlikely to end up with more than say three or four
+        turning points per period in the disk-integrated photometric signal
+        (as opposed to say $20$; of course $2$ turning points is the minimum
+        for periodic functions).
+        So there is likely to be some maximum $\Gamma$, related to the maximum
+        number of turning points typically observed in stellar light curves.
+        % Such simulations, crude though they may be, could be used to come up
+        % with physically-motivated priors for $\Gamma$.
+For the evolution time-scale, $l$, it is sensible to enforce $l>P$ if we
+        desire functions that have at least some fairly clear
+        periodicity.
+        Intuitively, asserting that $l<P$ will produce functions that can
+        evolve significantly over time scales shorter than one period.
+        So a function $f(t+P)$ would look signficantly different to $f(t)$,
+        in which case any claims of periodicity become dubious.
+        Actually, you can come with a more rigorous criterion (which also
+        takes into account dependencies on $\Gamma$) than this -- see below.
+        Provided
+        \begin{equation}
+        \frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} < \frac{{4\pi }}{3},
+        \end{equation}
+        the QP covariance function will have at least one inflection point
+        other than at $|x_i-x_j|=0$
+        So $\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} = \frac{{4\pi }}{3}$
+        corresponds to a critical point at which secondary maxima disappear,
+        and the behaviour of the kernel becomes dominated by the SE term
+        rather than the periodic term.
+        For $\Gamma\sim\mathcal{O}(1)$, this translates into $l\gtrsim P/2$.
+        In practice, satisfying this criterion doesn't absolutely guarantee
+        that function draws will have clear evidence of periodicity, but it
+        does give you a quantitative boundary between obviously periodic
+        behaviour and not-so-obviously periodic behaviour.
+        It also provides some justification for a more stringent criterion
+        such as $l>P$, or $l>2P$ (ensure that functions
+        evolve significantly over time-scales longer than 2 periods).}
+% However, the other parameters may also be related to some physical processes.
+% For example, the overall timescale for covariance fall-off, $l$ may be
+% related to spot lifetimes.
+% A star with long spot lifetimes will show little variation in the overall
+% shape and amplitude of its light curve between rotations and $l$ will be
+% large.
+% In contrast, the light curve of a star with short spot lifetimes may display
+% non-repeating patterns and amplitudes that vary rapidly between rotations.
+% In this case $l$ will be small.
+% The $\Gamma$ parameter is related to the number of zero crossings within one
+% rotation period: when $\Gamma$ is small there are many zero crossings and
+% vice versa.
+% Since the number of zero crossings per rotation period is related to the
+% number of active regions on the surface of the star, this parameter may also
+% be of physical interest.
+% In addition, instead of interpreting the parameters of the QP kernel function
+% used here, it may be possible to design an entirely new kernel function, based
+% on the physical processes that drive the light curve variability.
+% This idea is being explored by another member of my research group.}
+% \item \vrcomment{[VR: for the two points above, you might want to reword
+%     slightly depending on the extent to which you want to try to (i)
+%         incorporate my suggestions re prior constraints etc.\ etc.\ vs.\ (ii)
+%         deferring this to the future paper, but at least alluding to the
+%         possibilities and results. For example I think one needn't try to
+%         `design' a physically-motivated kernel -- rather it seems not too
+%         difficult just to come up with physically-motivated priors for the QP
+%         kernel parameters. I've tried to design a physically motivated kernel
+%         to capture differential rotation but as noted earlier, that came with
+%         its own slew of degeneracies and other problems...more on this another
+%         time.]}
 \item{To build in a noise model for \kepler\ data.
@@
 Although an improvement on competing methods, these uncertainties are still
-approximate because in many cases the posterior PDFs of the parameters are
-multi-modal.
-If it were possible to run an MCMC sampler for an infinite number of steps,
-the entire posterior PDF would eventually be sampled, and a more accurate
-uncertainty would result.
-Unfortunately, due to realistic computational resources and the limitations of
-MCMC sampling techniques for multi-modal posteriors, we are not always able to
-map out the entire posterior PDFs.
-\tdmcomment{I don't think this is the problem;
-I think the fits that we judge to have converged have actually converged.}
+underestimated in many cases because the GP model is an approximate, effective
+model.
+% in many cases the posterior PDFs of the parameters are
+% multi-modal.
+% If it were possible to run an MCMC sampler for an infinite number of steps,
+% the entire posterior PDF would eventually be sampled, and a more accurate
+% uncertainty would result.
+% Unfortunately, due to realistic computational resources and the limitations of
+% MCMC sampling techniques for multi-modal posteriors, we are not always able to
+% map out the entire posterior PDFs.
+% \tdmcomment{I don't think this is the problem;
+% I think the fits that we judge to have converged have actually converged.}
 

### 2017-02-13 documents/ms.tex
incorporating Suzanne's comments and remaking final figures

@@
 \newcommand{\lngpMAD}{0.02}
-\newcommand{\gpMAD}{0.38}
-\newcommand{\percentgpMAD}{1.85}
+\newcommand{\gpMAD}{0.39}
+\newcommand{\percentgpMAD}{1.96}
 
@@
 that fits the \Kepler\ light curve in in Figure \ref{fig:GP_example}.
-
-To infer a stellar rotation period $P$ from a light curve, we fit this QP-kernel
-GP model to the data.  As with any model-fitting exercise, the likelihood
+There are many ways to construct a periodic kernel and there are several
+non-periodic kernel choices, other than the SE kernel that could be combined
+with a periodic kernel in order to produce a QP kernel.
+The kernel presented above reproduces the behaviour of stellar light curves
+but it is not necessarily the only kernel choice that can do so.
+We do not attempt to test any other choices in this paper, noting only that
+this kernel function provides an adequeate model and leaving formal
+comparisons with other kernel function choices to a future publication.
+
+To infer a stellar rotation period $P$ from a light curve, we fit this
+QP-kernel GP model to the data.  As with any model-fitting exercise, the
+likelihood
 of the model could be maximized to find the maximum-likelihood value for $P$.
@@
 In order to recover a stellar rotation period from a light curve using a
-quasi-periodic Gaussian process (QP-GP), we sample the following posterior PDF:
+quasi-periodic Gaussian process (QP-GP), we sample the following posterior
+PDF:
 \begin{equation}
@@
 0.4, and 0.2 respectively.
-We run \vrcomment{50} steps of the sampler at a time, checking for convergence after
-each iteration, up to a maximum of \vrcomment{50} iterations.
+We run \vrcomment{50} steps of the sampler at a time, checking for convergence
+after each iteration, up to a maximum of \vrcomment{50} iterations.
 We declare convergence if the total effective chain length is at least
@@
 % \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_02_03.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_02_13.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_noprior_02_03.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_noprior_02_13.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/25.png}
+\includegraphics[width=6in, clip=true]{figures/corner_plot.pdf}
 \caption{Marginal posterior PDFs of the QP GP model parameters, fit to the
     light curve in figure \ref{fig:demo_lc}.
-    The green line in the period panel shows the injected period. \vrcomment{[VR: might it be possible to make the green line darker? It took me a little while to find it.]}
+    The blue line in the period panel shows the injected period.
     This figure was made using \texttt{corner.py} \citep{Corner}.}
@@
 Another valid choice would be a squared cosine function multiplied by a
-squared exponential,
+squared exponential, used by \citet{Brewer2009} to model asteroseismic
+pulsations,
 \begin{equation}
@@
 kernel function in future.
-If CPU time were not limited there may be some benefit to performing formal
-model comparison with different kernel functions.
-However, the evidence integral is an ambitious calculation for models with
-likelihood functions that take milliseconds to compute, let alone those
-involving GPs with light curves containing thousands of data points which can
-take minutes.
+% If CPU time were not limited there may be some benefit to performing formal
+% model comparison with different kernel functions.
+% However, the evidence integral is an ambitious calculation for models with
+% likelihood functions that take milliseconds to compute, let alone those
+% involving GPs with light curves containing thousands of data points which can
+% take minutes.
+The kernel we have chosen to use is only one of many possibilities, and it is
+already so flexible it does a fantastic job of reproducing more or less every
+stellar light curve we have thrown it at.
+This in turn implies we should be cautious in interpreting the
+hyperparameters as physical quantities.
+Even if the period is more meaningful than the others, there are certain
+regimes (depending on the value of the other hyperparameters relative to the
+period) where the period is not physically meaningful.
 

### 2017-02-13 documents/ms.tex

@@
 multimodal prior.  While this procedure does not avoid heuristic choices,
-the fact that its goal is simply to create a \emph{prior} for a
-probabilistically justified inference scheme rather than
-to identify a single correct period softens the potential impact
-of these choices.
+we soften the potential impact of these choices because we are simply
+creating a \emph{prior} for probabilistically justified
+inference rather attempting to identify a single correct period.
 

### 2017-02-13 documents/ms.tex
minor edits

@@
 \citet{Aigrain2015} simulated one thousand such light curves to test the
-ability of participating teams to recover both the stellar rotation periods
-and the rotational shear (the amplitude of surface differential rotation).
+ability of participating teams to recover both stellar rotation periods
+and rotational shear (the amplitude of surface differential rotation).
 % Since we are not interested in recovering differential rotation we only used
@@
 \citet{Aigrain2015} add real \kepler\ light curves with no obvious astrophysical
-variability to the theoretical rotational modulated light curves.
+variability to the theoretical rotationally modulated light curves.
 Figure \ref{fig:demo_lc} shows an example of a simulated light curve with a

### 2017-02-13 documents/ms.tex
small changes to the text

--- 2017-02-13 documents/ms.tex
@@
 \newcommand{\vrcomment}[1]{{\color{magenta}#1}}
+\newcommand{\racomment}[1]{{\color{green}#1}}
 
@@
 In this work, we introduce an appropriate {\it effective} model: a Gaussian
-    Process (GP) with a quasi-periodic covariance kernel function.
+    Process with a quasi-periodic covariance kernel function.
 Using this highly flexible model, we are able to sample from the posterior
@@
     are more accurate than both a sine-fitting periodogram and an
-    AutoCorrelation Function (ACF) method.
+    autocorrelation function method.
 We also demonstrate that it works effectively on real data, by
 inferring rotation periods for \nkois\ \Kepler\ stars.
-Because this method delivers posteriors PDFs it will enable hierarchical
+Because this method delivers posterior PDFs it will enable hierarchical
     studies involving stellar rotation, particularly those involving
@@
 peaks in a Lomb-Scargle \citep{Lomb1976, Scargle1982} (LS) periodogram
-\citep[e.g.][]{Reinhold2013}, Auto-Correlation Functions (ACFs),
-\citep{Mcquillan13b} and wavelet transforms \citep{Garcia2014}.
+\citep[\eg][]{Reinhold2013}, Auto-Correlation Functions (ACFs),
+\citep[\eg][]{Mcquillan13b} and wavelet transforms \citep[\eg][]{Garcia2014}.
 The precisions of the LS periodogram and wavelet methods are limited by the
@@
 estimate the posterior PDF of the rotation period, thereby producing a
-justified estimate of its uncertainty while simultaneously marginalizing over
-parameters which model correlated noise.
+justified estimate of its uncertainty.
+% while simultaneously marginalizing over
+% parameters which model correlated noise.
 
@@
 GP models parameterise the covariance between data points by means of a
-parametric kernel function that models the autocorrelation behaviour of the
-signal and defines the form of the covariance matrix.
+kernel function.
 As a qualitative demonstration, we present the time-series in Figure
@@
 widely separated are loosely correlated.
-A GP models this variation in correlation with the separation between data
-points; that is, it models the {\it covariance structure} rather than the data
-directly.
+A GP models this variation in correlation as a function of the separation
+between data points; that is, it models the {\it covariance structure} rather
+than the data directly.
 This lends GPs their flexibility---they can model any time
@@
 points separated by half a period to be weakly correlated.
-We also use an extra parameter $\sigma$ which is an additional white noise
+We also use an extra parameter, $\sigma$, which is an additional white noise
 term added to the diagonal elements of the covariance matrix.
@@
 that fits the \Kepler\ light curve in in Figure \ref{fig:GP_example}.
-There are many ways to construct a periodic kernel and there are several
-non-periodic kernel choices, other than the SE kernel that could be combined
-with a periodic kernel in order to produce a QP kernel.
-The kernel presented above reproduces the behaviour of stellar light curves
-but it is not necessarily the only kernel choice that can do so.
+There are many ways to construct a QP kernel function, involving a range of
+choices for both the periodic and a-periodic components of the model.
+The kernel function presented above reproduces the behaviour of stellar light
+curves but it is not necessarily the only model choice that can do so.
 We do not attempt to test any other choices in this paper, noting only that
-this kernel function provides an adequeate model and leaving formal
-comparisons with other kernel function choices to a future publication.
+this kernel function provides an adequate fit to the data.
+We leave formal comparisons with other kernel function choices to a future
+publication.
 
@@
 likelihood
-of the model could be maximized to find the maximum-likelihood value for $P$.
+of the model could be maximised to find the maximum-likelihood value for $P$.
 In this study, however, we explore the full posterior PDFs using a Markov
@@
 This paper is laid out as follows.
-In \textsection \ref{sec:simulations} we describe the simulated light curves
-used to test the three rotation period recovery methods (ACF, LS periodogram
-and GP).
 The GP method is described in \textsection \ref{sec:method}.
@@
 $A$ and $\Gamma$, and small values of $l$; though the exact details of what
-works well may differ among datasets (see \ref{sec:} and
-\ref{sec:} for two different examples of hyperparameter priors).
-More physically motivated priors are discussed in \textsection
-\ref{sec:discussion}.
+works well may differ among datasets.
+See \textsection \ref{sec:discussion} for a discussion of more physically
+motivated priors on the non-period parameters.
 
@@
 
-While using an uninformed prior on period (e.g., log-flat) often suffices,
-we also develop a method to construct an informed period prior, based on the
-autocorrelation function (ACF).
+% While using an uninformed prior on period (e.g., log-flat) often suffices,
+% we also develop a method to construct an informed period prior, based on the
+% autocorrelation function (ACF).
+We use an informed period prior, based on the autocorrelation function (ACF)
+of the light curve.
 The ACF has proven to be very useful for measuring stellar rotation periods
@@
 \end{equation}
-where and $t$ is the lag time,
+        \racomment{where and $t$ is the lag time, (some missing text here.)}
 and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
@@
 damped oscillator model and the actual ACF data.  The idea behind this
-quality metric to give a candidate period a higher score if
+quality metric is to give a candidate period a higher score if
 
@@
 
-We initialize 500 walkers with random samples from the prior and use a
+We initialise 500 walkers with random samples from the prior and use a
 weighted mixture of the KDE, DE, and snooker proposals with weights of 0.4,
@@
 Section \ref{sec:performance} demonstrates the performance of the
-QP-GP method; and Section \ref{sec:comparison} compares to the performance
+QP-GP method; and Section \ref{sec:comparison} compares it to the performance
 of the Lomb-Scargle periodogram and autocorrelation function methods.
@@
 Figures \ref{fig:compare_mcmc_acfprior} and \ref{fig:compare_mcmc_noprior}
-summarize our results compared to the injected
+summarise our results compared to the injected
 `true' stellar rotation periods, for the uninformed and ACF-informed priors on
@@
 technique.
+    Points are coloured by the peak-to-peak amplitude of the light curve, as
+    defined in \citet{Aigrain2015}.
 Since the posterior PDFs of rotation periods are often non-Gaussian,
-the points plotted here are the highest likelihood samples.
+    the points plotted here are maximum {\it a-posteriori} results.
 The uncertainties are the 16th and 84th percentiles.
@@
 The ACF-informed prior on rotation period used to generate these results is
-    described in the appendix.
-The MAD of these results is 0.38 in log-space, \ie\ 1.46 days.
+    described in \textsection \ref{sec:GP_prior}.
     }
@@
 simulated light curves vs the rotation periods measured using the GP
-technique with no .
+technique with an uninformative prior.
+    Points are coloured by the peak-to-peak amplitude of the light curve, as
+    defined in \citet{Aigrain2015}.
 Since the posterior PDFs of rotation periods are often non-Gaussian,
-the points plotted here are the highest likelihood samples.
+    the points plotted here are maximum {\it a-posteriori} results.
 The uncertainties are the 16th and 84th percentiles.
@@
     0.5 and 100 days was used to generate these results.
-The MAD of these results is 0.48 in log-space, \ie\ 1.61 days.
     }
@@
 \caption{Marginal posterior PDFs of the QP GP model parameters, fit to the
-    light curve in figure \ref{fig:demo_lc}.
+    simulated light curve in figure \ref{fig:demo_lc}.
     The blue line in the period panel shows the injected period.
@@
 periods, with the $2n$ and $\frac{1}{2}n$ harmonic lines as dashed lines.
-The MRD of the ACF-recovered periods is \percentacfMAD\ \% (see table
+The MRD of the ACF-recovered periods is \percentacfMAD\% (see table
 \ref{tab:MADs} for a side-by-side comparison with the other methods).
@@
 rotation periods are preferentially underestimated.
-This stems from how peak positions in the ACF are measured.
+This stems from peak position measurements in the ACF method.
 ACFs of stellar light curves have similar functional forms to the QP kernel
@@
 curves vs the rotation periods measured using the ACF technique.
+    Points are coloured by the peak-to-peak amplitude of the light curve, as
+    defined in \citet{Aigrain2015}.
     Several light curves have over-estimated rotation periods and some
@@
 curves vs the rotation periods measured using a LS periodogram technique.
+    Points are coloured by the peak-to-peak amplitude of the light curve, as
+    defined in \citet{Aigrain2015}.
 In many cases a large peak at a long period was present in the periodogram,
@@
     \caption{Median absolute (MAD) and median relative (MRD) deviations for
-    the ACF, LS periodogram and GP period recovery methods.}
+    the LS periodogram, ACF and GP period recovery methods.}
 \begin{tabular}{lccc}
@@
     \hline
+    LS periodogram & \pgramMAD\ days & \percentpgramMAD \% \\
     ACF & \acfMAD\ days & \percentacfMAD \% \\
-    LS periodogram & \pgramMAD\ days & \percentpgramMAD \% \\
-    GP (no prior) & \gpMADnp\ days & \percentgpMADnp \% \\
+    GP (uninformative prior) & \gpMADnp\ days & \percentgpMADnp \% \\
     GP (acf prior) & \gpMAD\ days & \percentgpMAD \% \\
@@
 host stars, \nkoimcq\ of which had rotation periods previously measured by
-the ACF method by \citet{Mcquillan2013}.  We use the pipeline-corrected flux
-(\texttt{pdcsap\_flux} column in the \Kepler\ light curve table), median-normalized
-and unit-subtracted, and mask out all known transiting planet candidate signals.
+the ACF method by \citet{Mcquillan2013}.
+We use the pipeline-corrected flux (\texttt{pdcsap\_flux} column in the
+\Kepler\ light curve table), median-normalised and unit-subtracted, and mask
+out all known transiting planet candidate signals.
 As with the simulated light curves, we randomly subsample each
@@
     \citet{Mcquillan2013}, defined as the interval between the 5th and 95th
-    percentiles of normalized flux per period bin in millimagnitudes.
-    The results show good agreement.}
+    percentiles of normalised flux per period bin in millimagnitudes.}
 \label{fig:mcquillan}
@@
 Of the rotation periods recovered from the simulated light curves, only 25\%
-of the measured periods lie within 1 $\sigma$ of the true period.
-50\% lie within 2 $\sigma$ and 66\% within 3 $\sigma$.
-The largest outlier is 114 $\sigma$ away from the true value.
+of the measured periods lie within 1$\sigma$ of the true period.
+50\% lie within 2$\sigma$ and 66\% within 3$\sigma$.
+The largest outlier is 114$\sigma$ away from the true value.
 % In the majority of these cases the uncertainties are underestimated due to the
@@
 k_{i,j} = A \exp \left(-\frac{(x_i - x_j)^2}{2l^2}\right)
-\cos\left(\frac{\pi(x_i - x_j)}{P}\right)
+\cos\left(\frac{\pi(x_i - x_j)}{P}\right).
 \end{equation}
@@
 % take minutes.
-The kernel we have chosen to use is only one of many possibilities, and it is
+The kernel we have chosen to use is only one of many possibilities, and is
 already so flexible it does a fantastic job of reproducing more or less every
-stellar light curve we have thrown it at.
-This in turn implies we should be cautious in interpreting the
-hyperparameters as physical quantities.
-Even if the period is more meaningful than the others, there are certain
-regimes (depending on the value of the other hyperparameters relative to the
-period) where the period is not physically meaningful.
+stellar light curve we have used it to model.
+% This in turn implies we should be cautious in interpreting the
+% hyperparameters as physical quantities.
+% Even if the period is more meaningful than the others, there are certain
+% regimes (depending on the value of the other hyperparameters relative to the
+% period) where the period is not physically meaningful.
 
@@
 We have only explored the physical interpretation of one parameter in our
-        kernel function, $P$.
+        kernel function, $P$ but some others may warrant physically motivated
+        priors.
 For example, take the harmonic complexity parameter, $\Gamma$.
@@
         typical number of stationary/non-stationary points of inflection, per
-        period, found in functions drawn from the corresponding GP.
+        period.
+        % found in functions drawn from the corresponding GP.
         This behaviour will be independent of the exact value of $l$.
-        For `intermediate' values of $l$ ($l\gtrsim P$) this can still be done
+        For `intermediate' values of $l$ ($l\gtrsim P$) this still holds
         but the behaviour will be sensitive to the value of $l$.
-        Small values of $l$ e.g.\ $l<P$ are probably not relevant -- see
+        Small values of $l$, \eg\ $l<P$, are probably not relevant -- see
         below.
@@
         % observed signals.
-        In typical simulations \eg\ of arbitrary numbers of spots on rotating
-        stars, you're unlikely to end up with more than say three or four
+        In typical light curves, observing more than say three or four
         turning points per period in the disk-integrated photometric signal
         (as opposed to say $20$; of course $2$ turning points is the minimum
-        for periodic functions).
+        for periodic functions) is highly unlikely.
         So there is likely to be some maximum $\Gamma$, related to the maximum
@@
         % with physically-motivated priors for $\Gamma$.
-For the evolution time-scale, $l$, it is sensible to enforce $l>P$ if we
-        desire functions that have at least some fairly clear
-        periodicity.
-        Intuitively, asserting that $l<P$ will produce functions that can
-        evolve significantly over time scales shorter than one period.
-        So a function $f(t+P)$ would look signficantly different to $f(t)$,
-        in which case any claims of periodicity become dubious.
-        Actually, you can come with a more rigorous criterion (which also
-        takes into account dependencies on $\Gamma$) than this -- see below.
+For the evolution time-scale, $l$, it is sensible to enforce $l>P$, such that
+        functions that have clear periodicity.
+        Intuitively, if $l<P$ the functions may evolve significantly over time
+        scales shorter than one period, so a function $f(t+P)$ would look
+        signficantly different to $f(t)$, in which case any claims of
+        periodicity become dubious.
+        A more rigorous criterion (which also
+        takes into account dependencies on $\Gamma$) is presented below.
         Provided
@@
 method works well on real \kepler\ data.
-
 Unlike the ACF and LS periodogram methods, the GP method provides posterior
@@
 % I think the fits that we judge to have converged have actually converged.}
-
 In addition, although the GP model used here is clearly a good one, it is
@@
 
-The main aim of this work is to develop a probabilistic rotation period
+The main aim of this work was to develop a probabilistic rotation period
 inference method.

### 2017-02-13 documents/ms.tex
minor wording

--- 2017-02-13 documents/ms.tex
@@
 We also demonstrate that it works effectively on real data, by
-inferring rotation periods for \nkois\ \Kepler\ stars.
-Because this method delivers posteriors PDFs it will enable hierarchical
+inferring rotation periods for \nkoimcq\ \Kepler\ stars with 
+previously measured periods.
+Because this method delivers posteriors PDFs, it will enable hierarchical
     studies involving stellar rotation, particularly those involving
@@
 more fall slightly below than above it, \ie\
-rotation periods are preferentially underestimated.
+rotation periods are systematically underestimated.
 This stems from how peak positions in the ACF are measured.
@@
 In order to test our rotation period inference method on real data,
-we apply it to a set of \nkois\ \Kepler\ Object of Interest (KOI)
-host stars, \nkoimcq\ of which had rotation periods previously measured by
+we apply it to a set of \nkoimcq\ \Kepler\ Object of Interest (KOI)
+host stars, that had rotation periods previously measured by
 the ACF method by \citet{Mcquillan2013}.  We use the pipeline-corrected flux
@@
 hyperparameters.
-Notably, below $\ln P$ of about 2.5, the GP method recovers periods
+Notably, the GP method recovers periods
 systematically slightly larger than \citet{Mcquillan2013}---a likely consequence

### 2017-02-13 documents/ms.tex

--- 2017-02-13 documents/ms.tex
@@
 We also demonstrate that it works effectively on real data, by
-inferring rotation periods for \nkois\ \Kepler\ stars.
-Because this method delivers posterior PDFs it will enable hierarchical
+inferring rotation periods for \nkoimcq\ \Kepler\ stars with 
+previously measured periods.
+Because this method delivers posteriors PDFs, it will enable hierarchical
     studies involving stellar rotation, particularly those involving
@@
 more fall slightly below than above it, \ie\
-rotation periods are preferentially underestimated.
+rotation periods are systematically underestimated.
 This stems from peak position measurements in the ACF method.
@@
 In order to test our rotation period inference method on real data,
-we apply it to a set of \nkois\ \Kepler\ Object of Interest (KOI)
-host stars, \nkoimcq\ of which had rotation periods previously measured by
-the ACF method by \citet{Mcquillan2013}.
-We use the pipeline-corrected flux (\texttt{pdcsap\_flux} column in the
-\Kepler\ light curve table), median-normalised and unit-subtracted, and mask
-out all known transiting planet candidate signals.
+we apply it to a set of \nkoimcq\ \Kepler\ Object of Interest (KOI)
+host stars, that had rotation periods previously measured by
+the ACF method by \citet{Mcquillan2013}.  We use the pipeline-corrected flux
+(\texttt{pdcsap\_flux} column in the \Kepler\ light curve table), median-normalized
+and unit-subtracted, and mask out all known transiting planet candidate signals.
 As with the simulated light curves, we randomly subsample each

### 2017-02-13 documents/ms.tex
changes to the prior discussion section

@@
 In addition to this, there are several other aspects of the GP method we are
-continuing to develop, listed as follows:.
+continuing to develop, listed as follows:
 \begin{itemize}
@@
         signficantly different to $f(t)$, in which case any claims of
-        periodicity become dubious.
-        A more rigorous criterion (which also
-        takes into account dependencies on $\Gamma$) is presented below.
-        Provided
-        \begin{equation}
-        \frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} < \frac{{4\pi }}{3},
-        \end{equation}
-        the QP covariance function will have at least one inflection point
-        other than at $|x_i-x_j|=0$
-        So $\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} = \frac{{4\pi }}{3}$
-        corresponds to a critical point at which secondary maxima disappear,
-        and the behaviour of the kernel becomes dominated by the SE term
-        rather than the periodic term.
-        For $\Gamma\sim\mathcal{O}(1)$, this translates into $l\gtrsim P/2$.
-        In practice, satisfying this criterion doesn't absolutely guarantee
-        that function draws will have clear evidence of periodicity, but it
-        does give you a quantitative boundary between obviously periodic
-        behaviour and not-so-obviously periodic behaviour.
-        It also provides some justification for a more stringent criterion
-        such as $l>P$, or $l>2P$ (ensure that functions
-        evolve significantly over time-scales longer than 2 periods).}
+        periodicity become dubious.}
+        % A more rigorous criterion (which also
+        % takes into account dependencies on $\Gamma$) is presented below.
+        % Provided
+        % \begin{equation}
+        % \frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} < \frac{{4\pi }}{3},
+        % \end{equation}
+        % the QP covariance function will have at least one inflection point
+        % other than at $|x_i-x_j|=0$, so $\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} =
+        % \frac{{4\pi }}{3}$
+        % corresponds to a critical point at which secondary maxima disappear,
+        % and the behaviour of the kernel becomes dominated by the SE term
+        % rather than the periodic term.
+        % For $\Gamma\sim\mathcal{O}(1)$, this translates into $l\gtrsim P/2$.
+        % In practice, satisfying this criterion doesn't absolutely guarantee
+        % that function draws will have clear evidence of periodicity, but it
+        % does provide a quantitative boundary between obviously periodic
+        % behaviour and not-so-obviously periodic behaviour.
+        % It also provides some justification for a more stringent criterion
+        % such as $l>P$, or $l>2P$ (ensure that functions
+        % evolve significantly over time-scales longer than 2 periods).}
 % However, the other parameters may also be related to some physical processes.

### 2017-02-13 documents/ms.tex
small change to conclusion

@@
 underestimated in many cases because the GP model is an approximate, effective
-model.
+model: we find that only a quarter of the reported uncertainties are accurate.
 % in many cases the posterior PDFs of the parameters are
@@
 covariance matrix, not the actual rotation period of a physical star.
-We find that only a quarter of the reported uncertainties are accurate.
 

### 2017-02-13 documents/ms.tex
minor edits

--- 2017-02-13 documents/ms.tex
@@
 
-The uncertainties produced by the GP method are more representative than those
-produced by the ACF and periodogram methods because the posterior PDF of the
-period is explored using MCMC.
-Whilst they are an improvement, unfortunately these uncertainties are still
-underestimated in the majority cases.
+The QP-GP inference method we present in this paper produces
+more accurate rotation periods than the ACF or periodogram methods.  
+Additionally, because it explores the posterior PDF with MCMC, 
+our method produces probabilistically justified uncertainties.  
+Unfortunately, these uncertainties still appear to be underestimated.
 Of the rotation periods recovered from the simulated light curves, only 25\%
@@
 The largest outlier is 114$\sigma$ away from the true value.
+\tdmcomment{I assume this is for the ACF-based prior run? (should probably be mentioned)  
+Are these numbers similar for the logflat prior?}
 % In the majority of these cases the uncertainties are underestimated due to the
@@
 % closer to the 1/2x or 2x period, but this isn't most of them.}
-The underestimated uncertainties can be attributed to the model choice.
-Although a GP model is more appropriate than a sinusoid, it is still only an
+We attribute these underestimated uncertainties to the model choice.
+Although more appropriate than a sinusoid, the GP is still only an
  effective model.

### 2017-02-13 documents/ms.tex
wording changes

@@
 Of the rotation periods recovered from the simulated light curves, only 25\%
-of the measured periods lie within 1$\sigma$ of the true period.
-50\% lie within 2$\sigma$ and 66\% within 3$\sigma$.
+of the measured periods lie within 1$\sigma$ of the true period,
+50\% lie within 2$\sigma$, and 66\% within 3$\sigma$.
 The largest outlier is 114$\sigma$ away from the true value.
@@
 
-The QP kernel function represents a simplistic effective model of a stellar
+The QP kernel function represents a simple effective model of a stellar
 light curve.
-It manages to describe a wide range of quasi-periodic behavior and is
+It can describe a wide range of quasi-periodic behavior and is
 relatively simple, with only a few hyperparameters.
@@
 This function produces a positive semi-definite matrix and has the $P$
-parameter of interest.
-The main practical difference between this cosine and the QP function is that
-the cosine function allows negative covariances and the QP function does not.
-Is is realistic to allow negative covariances?
+parameter of interest, but differs qualitatively from the QP function
+by allowing negative covariances.
+Is it realistic to allow negative covariances?
 In practice, the ACFs of \Kepler\ light curves often go negative.
 However, many stars have two active regions on opposite hemispheres that
-produce two brightness dips per rotation.
-If the covariance is forced to be negative for two data points that are
-separated by half a rotation period, those light curves with two peaks per
-rotation period may not be well modelled.
-In other words, we do not want to force anti-correlation of points that are
-$\frac{1}{2}$ a period apart.
+produce two brightness dips per rotation, and it
+may be difficult to model such stars with a kernel that forces
+anti-correlation of points $\frac{1}{2}$ a period apart.
 It would be very worthwhile to test this assumption and this alternative
-kernel function in future.
+kernel function (and others) in the future.
 % If CPU time were not limited there may be some benefit to performing formal
@@
 % take minutes.
-The kernel we have chosen to use is only one of many possibilities, and is
-already so flexible it does a fantastic job of reproducing more or less every
-stellar light curve we have used it to model.
+% The kernel we have chosen to use is only one of many 
+% possibilities, and is
+% already so flexible it does a fantastic job of reproducing more or less every
+% stellar light curve we have used it to model.
 % This in turn implies we should be cautious in interpreting the
@@
 Not all \kepler\ light curves show evidence of stellar rotation.
-In some cases perhaps the star has few or no active regions, it is rotating
-pole-on, or it rotates so slowly that the \kepler\ data detrending pipeline
+In some cases a star may have few or no active regions, be rotating
+pole-on, or be rotating so slowly that the \kepler\ data detrending pipeline
 removes any signal.

### 2017-02-13 documents/ms.tex
minor edits

@@
 asteroseismic oscillations in giants and even stellar activity cycles.
-For some of these there is little we can do save attempt to eliminate these
-astrophysical false positives via alternative methods, \eg\ apply colour cuts
-to avoid giant contamination.
-For some however, like the variable stars, they may have distinctive
-hyperparameters that identify them, for example long coherence timescales.
-Testing this is beyond the scope of this paper but may be an interesting
-follow-up study.
-As well as astrophysical contamination, there may also be instrumental sources
-of contaminating variability: \eg\ temperature variations or pointing shifts
+Identifying many of these astrophysical false positives 
+falls outside the scope of this GP method 
+(\eg\ applying colour cuts to avoid giant contamination);
+however, for some, such as variable stars, they may have distinctive
+hyperparameters (\eg\ long coherence timescales) that identify them.
+Testing this may be an interesting follow-up study.
+As well as astrophysical contamination, instrumental sources may contribute
+to contaminating variability, \eg\ temperature variations or pointing shifts
 of the \kepler\ spacecraft.
@@
 of hyperparameters.
-This is something that we hope to test in future.
-In addition to this, there are several other aspects of the GP method we are
-continuing to develop, listed as follows:
+We also hope to test this in the future.
+
+In addition, we are continuing to develop several other aspects of this GP method:
 \begin{itemize}
-\item{To perform model selection with different kernel functions. Again, these
-    tests have not been performed thus far due to the cost of calculating the
+\item{To perform model selection with different kernel functions. 
+    We have not performed these tests yet due to the cost of calculating the
         fully marginalised likelihood with GPs.}
@@
 We have only explored the physical interpretation of one parameter in our
-        kernel function, $P$ but some others may warrant physically motivated
+        kernel function ($P$), but others may warrant physically motivated
         priors.
 For example, take the harmonic complexity parameter, $\Gamma$.
-        For large values of $l$ ($l\gg P$), $\Gamma$ can be related to the
-        typical number of stationary/non-stationary points of inflection, per
+        For large values of $l$ ($l\gg P$), $\Gamma$ relates to the
+        typical number of stationary/non-stationary points of inflection per
         period.
@@
         % observed signals.
-        In typical light curves, observing more than say three or four
+        In typical light curves, observing more than perhaps three or four
         turning points per period in the disk-integrated photometric signal
-        (as opposed to say $20$; of course $2$ turning points is the minimum
+        (of course, $2$ turning points is the minimum
         for periodic functions) is highly unlikely.
@@
 \item{To build in a noise model for \kepler\ data.
-Another huge advantage of the GP method is, because it is a {\it generative}
-model of the data, the rotation period signal can be modelled at the same time
-as systematic noise features.
+Because it is a {\it generative} model of the data,
+the GP method models the rotation period at the same time as systematic
+noise features.
 One can then marginalise over the parameters of the noise model.
@@
 Marginalising over the noise model at the same time as inferring the
-parameters of interest will insure that the periodic signal is preserved.}
+parameters of interest will insure that the periodic signal is preserved. 
+\tdmcomment{Is this bullet point obselete, given that we have already successfully
+marginalized over \kepler\ noise to recover periods? Or is there
+something else we wanted to say here?}}
 \end{itemize}
@@
 
-We attempted to recover the rotation periods of \naigrain\ simulated
+We have attempted to recover the rotation periods of \naigrain\ simulated
 \kepler-like light curves for solid-body rotators \citep{Aigrain2015} using
-three different methods: a new Gaussian process method, a Lomb-Scargle
-periodogram method and an autocorrelation function method.
+three different methods: a new Gaussian process method, 
+an authocorrelation function method, and a Lomb-Scargle periodogram method.
 We demonstrate that the GP method produces the most accurate rotation periods
@@
 period due to a subtle feature of its peak detection algorithm.
-In addition, we compared the rotation periods of \kepler\ objects of
-interest measured using the GP method to those measured previously by
-\citet{Mcquillan2013}.
-The good agreement between the two sets of results demonstrates that the GP
+In addition, we measure the rotation periods of \nkoimcq\ \kepler\ objects of
+interest using the GP method, and find that these results compare well
+to those measured previously by \citet{Mcquillan2013}.
+The good agreement between these two sets of results demonstrates that the GP
 method works well on real \kepler\ data.
-Unlike the ACF and LS periodogram methods, the GP method provides posterior
+
+Unlike the ACF and L-S periodogram methods, the GP method provides posterior
 PDF samples which can be used to estimate rotation periods uncertainties.
@@
 underestimated in many cases because the GP model is an approximate, effective
-model: we find that only a quarter of the reported uncertainties are accurate.
+model. \tdm{Can we make a statement something like ``we believe the method underestimates
+uncertainties by about a factor of three'' or something like that?  This
+matches with 66\% of the fits being within ``3$\sigma''.
+Might be good to say something like this, both in the discussion and here.}
 % in many cases the posterior PDFs of the parameters are

### 2017-02-15 documents/ms.tex
adding Vinesh's new paragraph

--- 2017-02-13 documents/ms.tex
@@
 
-The QP-GP inference method we present in this paper produces
-more accurate rotation periods than the ACF or periodogram methods.  
-Additionally, because it explores the posterior PDF with MCMC, 
-our method produces probabilistically justified uncertainties.  
-Unfortunately, these uncertainties still appear to be underestimated.
+The uncertainties produced by the GP method are more representative than those
+produced by the ACF and periodogram methods because the posterior PDF of the
+period is explored using MCMC.
+Whilst they are an improvement, unfortunately these uncertainties are still
+underestimated in the majority cases.
 Of the rotation periods recovered from the simulated light curves, only 25\%
@@
 The largest outlier is 114$\sigma$ away from the true value.
-\tdmcomment{I assume this is for the ACF-based prior run? (should probably be mentioned)  
-Are these numbers similar for the logflat prior?}
 % In the majority of these cases the uncertainties are underestimated due to the
@@
 % closer to the 1/2x or 2x period, but this isn't most of them.}
-We attribute these underestimated uncertainties to the model choice.
-Although more appropriate than a sinusoid, the GP is still only an
+The underestimated uncertainties can be attributed to the model choice.
+Although a GP model is more appropriate than a sinusoid, it is still only an
  effective model.
@@
         fully marginalised likelihood with GPs.}
-\item{To design and implement more physically motivated priors.
-We have only explored the physical interpretation of one parameter in our
-        kernel function, $P$ but some others may warrant physically motivated
-        priors.
-For example, take the harmonic complexity parameter, $\Gamma$.
-        For large values of $l$ ($l\gg P$), $\Gamma$ can be related to the
-        typical number of stationary/non-stationary points of inflection, per
-        period.
-        % found in functions drawn from the corresponding GP.
-        This behaviour will be independent of the exact value of $l$.
-        For `intermediate' values of $l$ ($l\gtrsim P$) this still holds
-        but the behaviour will be sensitive to the value of $l$.
-        Small values of $l$, \eg\ $l<P$, are probably not relevant -- see
-        below.
-        % Depending on the processes giving rise to observed stellar signals,
-        % you may not expect to require certain types of functions to model
-        % observed signals.
-        In typical light curves, observing more than say three or four
-        turning points per period in the disk-integrated photometric signal
-        (as opposed to say $20$; of course $2$ turning points is the minimum
-        for periodic functions) is highly unlikely.
-        So there is likely to be some maximum $\Gamma$, related to the maximum
-        number of turning points typically observed in stellar light curves.
-        % Such simulations, crude though they may be, could be used to come up
-        % with physically-motivated priors for $\Gamma$.
-For the evolution time-scale, $l$, it is sensible to enforce $l>P$, such that
-        functions that have clear periodicity.
-        Intuitively, if $l<P$ the functions may evolve significantly over time
+\item{To discuss more physically motivated priors.
+        In this paper we have only explored the physical interpretation of one
+        parameter in our kernel function (P), which we related to stellar
+        rotation periods, but others also warrant physically motivated priors.
+        For example, the harmonic complexity parameter, $\Gamma$, defines the
+        typical number of inflection points per function period; suitable
+        priors may be informed by the typical properties of observed light
+        curves.
+        For the evolution time-scale, $l$,  it is sensible to enforce $l > P$,
+        such that functions that have clear periodicity.
+        Intuitively, if $l < P$, functions may evolve significantly over time
         scales shorter than one period, so a function $f(t+P)$ would look
-        signficantly different to $f(t)$, in which case any claims of
-        periodicity become dubious.}
-        % A more rigorous criterion (which also
-        % takes into account dependencies on $\Gamma$) is presented below.
-        % Provided
-        % \begin{equation}
-        % \frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} < \frac{{4\pi }}{3},
-        % \end{equation}
-        % the QP covariance function will have at least one inflection point
-        % other than at $|x_i-x_j|=0$, so $\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} =
-        % \frac{{4\pi }}{3}$
-        % corresponds to a critical point at which secondary maxima disappear,
-        % and the behaviour of the kernel becomes dominated by the SE term
-        % rather than the periodic term.
-        % For $\Gamma\sim\mathcal{O}(1)$, this translates into $l\gtrsim P/2$.
-        % In practice, satisfying this criterion doesn't absolutely guarantee
-        % that function draws will have clear evidence of periodicity, but it
-        % does provide a quantitative boundary between obviously periodic
-        % behaviour and not-so-obviously periodic behaviour.
-        % It also provides some justification for a more stringent criterion
-        % such as $l>P$, or $l>2P$ (ensure that functions
-        % evolve significantly over time-scales longer than 2 periods).}
-% However, the other parameters may also be related to some physical processes.
-% For example, the overall timescale for covariance fall-off, $l$ may be
-% related to spot lifetimes.
-% A star with long spot lifetimes will show little variation in the overall
-% shape and amplitude of its light curve between rotations and $l$ will be
-% large.
-% In contrast, the light curve of a star with short spot lifetimes may display
-% non-repeating patterns and amplitudes that vary rapidly between rotations.
-% In this case $l$ will be small.
-% The $\Gamma$ parameter is related to the number of zero crossings within one
-% rotation period: when $\Gamma$ is small there are many zero crossings and
-% vice versa.
-% Since the number of zero crossings per rotation period is related to the
-% number of active regions on the surface of the star, this parameter may also
-% be of physical interest.
-% In addition, instead of interpreting the parameters of the QP kernel function
-% used here, it may be possible to design an entirely new kernel function, based
-% on the physical processes that drive the light curve variability.
-% This idea is being explored by another member of my research group.}
-% \item \vrcomment{[VR: for the two points above, you might want to reword
-%     slightly depending on the extent to which you want to try to (i)
-%         incorporate my suggestions re prior constraints etc.\ etc.\ vs.\ (ii)
-%         deferring this to the future paper, but at least alluding to the
-%         possibilities and results. For example I think one needn't try to
-%         `design' a physically-motivated kernel -- rather it seems not too
-%         difficult just to come up with physically-motivated priors for the QP
-%         kernel parameters. I've tried to design a physically motivated kernel
-%         to capture differential rotation but as noted earlier, that came with
-%         its own slew of degeneracies and other problems...more on this another
-%         time.]}
+        significantly different to $f(t)$, in which case any claims of
+        periodicity become dubious.
+        More stringent criteria may be suggested by stellar astrophysical
+        considerations, e.g. knowledge of typical active region evolution
+        time-scales or persistence lifetimes (e.g. $l\sim2P$).
+        Finally, the amplitude hyper-parameter $A$ may be related to active
+        region coverage.
+        Unfortunately, because these hyper-parameters control the covariance
+        structure of functions, their interpretation is not always
+        straightforward; interpretation is further complicated by significant
+        degeneracies between the hyperparameters.
+        Therefore we defer such considerations, along with strategies for
+        efficient optimization or marginalization of hyperparameters, to a
+        separate paper (currently in prep.).
+    }
 \item{To build in a noise model for \kepler\ data.

### 2017-02-15 documents/ms.tex

--- 2017-02-13 documents/ms.tex
@@
         fully marginalised likelihood with GPs.}
-\item{To design and implement more physically motivated priors.
-We have only explored the physical interpretation of one parameter in our
-        kernel function ($P$), but others may warrant physically motivated
-        priors.
-For example, take the harmonic complexity parameter, $\Gamma$.
-        For large values of $l$ ($l\gg P$), $\Gamma$ relates to the
-        typical number of stationary/non-stationary points of inflection per
-        period.
-        % found in functions drawn from the corresponding GP.
-        This behaviour will be independent of the exact value of $l$.
-        For `intermediate' values of $l$ ($l\gtrsim P$) this still holds
-        but the behaviour will be sensitive to the value of $l$.
-        Small values of $l$, \eg\ $l<P$, are probably not relevant -- see
-        below.
-        % Depending on the processes giving rise to observed stellar signals,
-        % you may not expect to require certain types of functions to model
-        % observed signals.
-        In typical light curves, observing more than perhaps three or four
-        turning points per period in the disk-integrated photometric signal
-        (of course, $2$ turning points is the minimum
-        for periodic functions) is highly unlikely.
-        So there is likely to be some maximum $\Gamma$, related to the maximum
-        number of turning points typically observed in stellar light curves.
-        % Such simulations, crude though they may be, could be used to come up
-        % with physically-motivated priors for $\Gamma$.
-For the evolution time-scale, $l$, it is sensible to enforce $l>P$, such that
-        functions that have clear periodicity.
-        Intuitively, if $l<P$ the functions may evolve significantly over time
+\item{To discuss more physically motivated priors.
+        In this paper we have only explored the physical interpretation of one
+        parameter in our kernel function (P), which we related to stellar
+        rotation periods, but others also warrant physically motivated priors.
+        For example, the harmonic complexity parameter, $\Gamma$, defines the
+        typical number of inflection points per function period; suitable
+        priors may be informed by the typical properties of observed light
+        curves.
+        For the evolution time-scale, $l$,  it is sensible to enforce $l > P$,
+        such that functions that have clear periodicity.
+        Intuitively, if $l < P$, functions may evolve significantly over time
         scales shorter than one period, so a function $f(t+P)$ would look
-        signficantly different to $f(t)$, in which case any claims of
-        periodicity become dubious.}
-        % A more rigorous criterion (which also
-        % takes into account dependencies on $\Gamma$) is presented below.
-        % Provided
-        % \begin{equation}
-        % \frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} < \frac{{4\pi }}{3},
-        % \end{equation}
-        % the QP covariance function will have at least one inflection point
-        % other than at $|x_i-x_j|=0$, so $\frac{{{P^2}}}{{{l^2}{\Gamma ^2}}} =
-        % \frac{{4\pi }}{3}$
-        % corresponds to a critical point at which secondary maxima disappear,
-        % and the behaviour of the kernel becomes dominated by the SE term
-        % rather than the periodic term.
-        % For $\Gamma\sim\mathcal{O}(1)$, this translates into $l\gtrsim P/2$.
-        % In practice, satisfying this criterion doesn't absolutely guarantee
-        % that function draws will have clear evidence of periodicity, but it
-        % does provide a quantitative boundary between obviously periodic
-        % behaviour and not-so-obviously periodic behaviour.
-        % It also provides some justification for a more stringent criterion
-        % such as $l>P$, or $l>2P$ (ensure that functions
-        % evolve significantly over time-scales longer than 2 periods).}
-% However, the other parameters may also be related to some physical processes.
-% For example, the overall timescale for covariance fall-off, $l$ may be
-% related to spot lifetimes.
-% A star with long spot lifetimes will show little variation in the overall
-% shape and amplitude of its light curve between rotations and $l$ will be
-% large.
-% In contrast, the light curve of a star with short spot lifetimes may display
-% non-repeating patterns and amplitudes that vary rapidly between rotations.
-% In this case $l$ will be small.
-% The $\Gamma$ parameter is related to the number of zero crossings within one
-% rotation period: when $\Gamma$ is small there are many zero crossings and
-% vice versa.
-% Since the number of zero crossings per rotation period is related to the
-% number of active regions on the surface of the star, this parameter may also
-% be of physical interest.
-% In addition, instead of interpreting the parameters of the QP kernel function
-% used here, it may be possible to design an entirely new kernel function, based
-% on the physical processes that drive the light curve variability.
-% This idea is being explored by another member of my research group.}
-% \item \vrcomment{[VR: for the two points above, you might want to reword
-%     slightly depending on the extent to which you want to try to (i)
-%         incorporate my suggestions re prior constraints etc.\ etc.\ vs.\ (ii)
-%         deferring this to the future paper, but at least alluding to the
-%         possibilities and results. For example I think one needn't try to
-%         `design' a physically-motivated kernel -- rather it seems not too
-%         difficult just to come up with physically-motivated priors for the QP
-%         kernel parameters. I've tried to design a physically motivated kernel
-%         to capture differential rotation but as noted earlier, that came with
-%         its own slew of degeneracies and other problems...more on this another
-%         time.]}
+        significantly different to $f(t)$, in which case any claims of
+        periodicity become dubious.
+        More stringent criteria may be suggested by stellar astrophysical
+        considerations, e.g. knowledge of typical active region evolution
+        time-scales or persistence lifetimes (e.g. $l\sim2P$).
+        Finally, the amplitude hyper-parameter $A$ may be related to active
+        region coverage.
+        Unfortunately, because these hyper-parameters control the covariance
+        structure of functions, their interpretation is not always
+        straightforward; interpretation is further complicated by significant
+        degeneracies between the hyperparameters.
+        Therefore we defer such considerations, along with strategies for
+        efficient optimization or marginalization of hyperparameters, to a
+        separate paper (currently in prep.).
+    }
 \item{To build in a noise model for \kepler\ data.

### 2017-02-15 documents/ms.tex
fix typo \tdm -> \tdmcomment

@@
 underestimated in many cases because the GP model is an approximate, effective
-model. \tdm{Can we make a statement something like ``we believe the method underestimates
+model. \tdmcomment{Can we make a statement something like ``we believe the method underestimates
 uncertainties by about a factor of three'' or something like that?  This

### 2017-02-15 documents/ms.tex
fix another typo
fix typo

@@
 \item{We assign a quality metric $Q_i$ to each of these candidate
-periods, as follows.  First, we model the ACF as a
-damped oscillator with fixed period $T_i$:
+periods, as follows.  First, we model the ACF as a function of lag
+time $t$ as a damped oscillator with fixed period $T_i$:
 \begin{equation}
@@
 \end{equation}
-        \racomment{where and $t$ is the lag time, (some missing text here.)}
 and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
@@
 uncertainties by about a factor of three'' or something like that?  This
-matches with 66\% of the fits being within ``3$\sigma''.
+matches with 66\% of the fits being within ``3$\sigma$''.
 Might be good to say something like this, both in the discussion and here.}

### 2017-02-16 documents/ms.tex
updating urls and including a comment about the noprior uncertainties

@@
 \newcommand{\oxford}{3}
-\newcommand{\sagan}{4}
-\newcommand{\uw}{5}
+\newcommand{\uw}{4}
+\newcommand{\sagan}{5}
 \newcommand{\princeton}{2}
@@
     exoplanet systems, or characterising star-planet interactions.
+The code is available at \url{https://github.com/RuthAngus/GProtation/} .
 % Thehe ACF method is not applicable to unevenly spaced data,
@@
 figure \ref{fig:gp_posteriors}.
+The posterior samples of the model parameters for the simulated light curves
+are available online:
+\url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
 
@@
 The final priors and parameter ranges that we used are in Table \ref{tab:koipriors}.
+The posterior samples of the model parameters for the \Kepler\ objects of
+interest are available online:
+\url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
 
@@
 Unfortunately, these uncertainties still appear to be underestimated.
-Of the rotation periods recovered from the simulated light curves, only 25\%
-of the measured periods lie within 1$\sigma$ of the true period,
-50\% lie within 2$\sigma$, and 66\% within 3$\sigma$.
+Of the rotation periods recovered from the simulated light curves, using the
+ACF-informed prior, only 25\% of the measured periods lie within 1$\sigma$ of
+the true period, 50\% lie within 2$\sigma$, and 66\% within 3$\sigma$.
 The largest outlier is 114$\sigma$ away from the true value.
-\tdmcomment{I assume this is for the ACF-based prior run? (should probably be mentioned)
-Are these numbers similar for the logflat prior?}
+The median uncertainty is underestimated by around a factor of two.
+These numbers are similar for the uninformative prior results which also
+provides underestimated uncertainties.
 % In the majority of these cases the uncertainties are underestimated due to the
@@
 inference currently available.
+The code used in this project is available at
+\url{https://github.com/RuthAngus/GProtation/}.
 

### 2017-02-16 code/plot_tim.py

@@ -90,6 +90,8 @@ def plot(prior, nbins):
 
     plt.clf()
     resids = np.log(x[l]) - np.log(recovered[l])
+    linresids = x[l] - recovered[l]
+    linerrs = .5*(lnerrp[l]*recovered[l] + lnerrm[l]*recovered[l])
     plt.hist(resids, 80, histtype="stepfilled", color="w")
     plt.xlabel("$\ln(\mathrm{True~Period}) - \ln(\mathrm{GP~Period})$")
     plt.axvline(np.percentile(resids, 16), color=".5", ls="--")
@@ -104,20 +106,11 @@ def plot(prior, nbins):
         plt.savefig("gp_hist_noprior.pdf")
 
     print(len(x[l]), "stars")
-    # print("percentiles:")
-    # print(np.percentile(resids, 16), np.percentile(resids, 84))
-    # print("diff = ",
-    #       np.abs(np.percentile(resids, 16) - np.percentile(resids, 84)))
-    # print("median_err = ", median_err)
-
     print("MAD = ", np.median(np.abs(x[l] - recovered[l])))
     print("MAD (log) = ", np.median(np.abs(np.log(x[l]) -
                                            np.log(recovered[l]))))
     print("MAD relative % = ", np.median((np.abs(x[l] -
                                                  recovered[l]))/x[l])*100)
-    # print("MAD relative (log) = ", np.median(np.abs(np.log(x[l]) -
-    #                                            np.log(recovered[l]))
-    #                                          /np.log(x[l]))*100j)
 
     errs = .5*(lnerrp[l] + lnerrm[l])
     plt.clf()
@@ -129,14 +122,25 @@ def plot(prior, nbins):
 
     plt.clf()
     nsigma_diff = np.abs(resids - errs)/errs
-    plt.hist(nsigma_diff, 100, histtype="stepfilled", color="w")
-    plt.axvline(np.percentile(nsigma_diff, 66), color="r", ls="--")
+    plt.hist(nsigma_diff, 100)
+    # plt.axvline(np.percentile(nsigma_diff, 66), color="r", ls="--")
+    print(np.percentile(nsigma_diff, 50), "= median")
+    "50% are 1.88 sigma off. But you'd expect "
     print(np.percentile(nsigma_diff, 66))
     print(max(nsigma_diff))
     if prior:
         plt.savefig("err_resid_ratio_hist")
     else:
         plt.savefig("err_resid_ratio_hist_noprior")
+
+    if prior:
[…]

### 2017-02-16 documents/ms.tex
adding paragraph about tel aviv results and RMS values

@@
 \newcommand{\percentgpMAD}{1.96}
+
+\newcommand{\lntelavivMAD}{0.02}
+\newcommand{\telavivMAD}{0.34}
+\newcommand{\percenttelavivMAD}{1.75}
+
+\newcommand{\acfRMS}{1.75}
+\newcommand{\telavivRMS}{25.09}
+\newcommand{\pgramRMS}{19.50}
+\newcommand{\gpRMSnp}{7.48}
+\newcommand{\gpRMS}{7.12}
 
@@
 
+We also compare our results to those of the Tel Aviv team in
+\citet{Aigrain2015} who use the ACF method described in \citet{Mcquillan2013}.
+Although we follow the same overall ACF method in this paper, the exact
+implementation of \citet{Mcquillan2013} code is slightly different and
+therefore produces different results.
+The MRD of their results relative to the true rotation periods is
+\percenttelavivMAD\% which is smaller than the results produced by the GP
+method: their results are more precise.
+However, they produce a larger number of large outliers --- they predict
+periods that are significantly different from the true period in more cases
+than the GP method.
+This is revealed by comparing the root-mean-squared (RMS) of the two sets of
+results, which is more sensitive to outliers: the RMS of the Tel Aviv results
+is 25 days and the RMS of GP results is 7 days.
+The Tel Aviv ACF results are therefore more precise but are catastrophically
+wrong in a larger number of cases.
+% This can be understood as a consequence of the ACF method's sensitivity to
+% noise.
+% The ACF method will predict a period according to the positions of the highest
+% peaks in the ACF which, on occasion, are produced by noise in the data rather
+% than the periodic signal.
+% On the other hand, the priors on the GP parameters ensure that a `reasonable'
+% rotation period is always recovered --- it is therefore less sensitive to
+% noise and rarely infers periods that are .
+
 \begin{figure*}
@@
 \begin{center}
-    \caption{Median absolute (MAD) and median relative (MRD) deviations for
-    the LS periodogram, ACF and GP period recovery methods.}
-\begin{tabular}{lccc}
-Method & MAD & MRD \\
+    \caption{Median absolute (MAD), median relative (MRD) deviations and
+    root-mean-squared (RMS) for
+    the LS periodogram, ACF, \citet{Mcquillan2013} ACF and GP period recovery
+    methods.}
+\begin{tabular}{lcccc}
+    Method & MAD & MRD & RMS \\
     \hline
-    LS periodogram & \pgramMAD\ days & \percentpgramMAD \% \\
-    ACF & \acfMAD\ days & \percentacfMAD \% \\
-    GP (uninformative prior) & \gpMADnp\ days & \percentgpMADnp \% \\
-    GP (acf prior) & \gpMAD\ days & \percentgpMAD \% \\
+    LS periodogram & \pgramMAD\ days & \percentpgramMAD \% & \pgramRMS\ days\\
+    ACF & \acfMAD\ days & \percentacfMAD \% & \acfRMS\ days \\
+    ACF \citep{Mcquillan2013} & \telavivMAD\ days & \percenttelavivMAD \%
+    & \telavivRMS\ days \\
+    GP (uninformative prior) & \gpMADnp\ days & \percentgpMADnp \% &
+    \gpRMSnp\ days \\
+    GP (acf prior) & \gpMAD\ days & \percentgpMAD \% & \gpRMS\ days \\
 \end{tabular}

### 2017-02-16 code/plot_tim.py

@@ -111,6 +111,7 @@ def plot(prior, nbins):
                                            np.log(recovered[l]))))
     print("MAD relative % = ", np.median((np.abs(x[l] -
                                                  recovered[l]))/x[l])*100)
+    print("RMS = ", (np.mean((x[l] - recovered[l])**2))**.5)
 
     errs = .5*(lnerrp[l] + lnerrm[l])
     plt.clf()
@@ -155,6 +156,7 @@ def plot(prior, nbins):
 if __name__ == "__main__":
     FIG_DIR = "/Users/ruthangus/projects/GProtation/documents/figures"
     print("ACF prior")
-    plot(True, 100)  # with ACF prior
+    nbins = 95
+    plot(True, nbins)  # with ACF prior
     print("\n", "No prior")
-    plot(False, 100)  # without ACF prior
+    plot(False, nbins)  # without ACF prior

### 2017-02-16 documents/ms.tex
adding some words to the conclusion

@@
 \citet{Aigrain2015} who use the ACF method described in \citet{Mcquillan2013}.
-Although we follow the same overall ACF method in this paper, the exact
-implementation of \citet{Mcquillan2013} code is slightly different and
-therefore produces different results.
-The MRD of their results relative to the true rotation periods is
-\percenttelavivMAD\% which is smaller than the results produced by the GP
-method: their results are more precise.
-However, they produce a larger number of large outliers --- they predict
+Although we follow the same general ACF method in this paper, the exact
+implementation of the \citet{Mcquillan2013} code is slightly different and
+produces different results.
+The MRD of their periods, relative to the true rotation periods, is
+\percenttelavivMAD\% --- smaller than the results produced by the GP
+method.
+However, their results contain a larger number of extreme outliers --- they
+predict
 periods that are significantly different from the true period in more cases
@@
 This is revealed by comparing the root-mean-squared (RMS) of the two sets of
-results, which is more sensitive to outliers: the RMS of the Tel Aviv results
-is 25 days and the RMS of GP results is 7 days.
-The Tel Aviv ACF results are therefore more precise but are catastrophically
-wrong in a larger number of cases.
+results since the RMS is more sensitive to outliers.
+The RMS of the Tel Aviv method results is \telavivRMS\ days and the RMS of GP
+method results is \gpRMS\ days.
 % This can be understood as a consequence of the ACF method's sensitivity to
@@
 of the three techniques, providing a large improvement over the LS periodogram
-method and a moderate one over the ACF method.
-We also find that the standard version of the ACF method, most commonly
-implemented in the literature, often slightly under-predicts the rotation
-period due to a subtle feature of its peak detection algorithm.
-In addition, we measure the rotation periods of \nkoimcq\ \kepler\ objects of
-interest using the GP method, and find that these results compare well
-to those measured previously by \citet{Mcquillan2013}.
+method and a moderate one over our basic implementation of the ACF method.
+We also find that our basic ACF method which is similar to methods implemented
+in the literature, often slightly under-predicts the rotation period due to a
+subtle feature of its peak detection algorithm.
+Comparing our results with the results of the \citet{Aigrain2015} Tel Aviv
+team who implement the method of \citet{Mcquillan2013}, a slightly different
+version of the ACF method to the one we use, we find that their results are
+slightly more precise than the GP method results overall.
+They do however have a larger number of significant outliers.
+We measured the rotation periods of \nkoimcq\ \kepler\ objects of interest
+using the GP method, and find that these results compare well to those
+measured previously by \citet{Mcquillan2013}.
 The good agreement between these two sets of results demonstrates that the GP

### 2017-02-17 documents/ms.tex
removing comments
adding words to the conclusion

@@
 
-\newcommand{\lnacfMAD}{0.04}
-\newcommand{\acfMAD}{0.60}
-\newcommand{\percentacfMAD}{3.65}
-\newcommand{\lnpgramMAD}{0.04}
-\newcommand{\pgramMAD}{0.98}
-\newcommand{\percentpgramMAD}{4.32}
+\newcommand{\lnacfMAD}{0.14}
+\newcommand{\acfMAD}{2.36}
+\newcommand{\percentacfMAD}{13.02}
+\newcommand{\acfRMS}{26.09}
+
+\newcommand{\lnpgramMAD}{0.05}
+\newcommand{\pgramMAD}{1.13}
+\newcommand{\percentpgramMAD}{4.70}
+\newcommand{\pgramRMS}{18.98}
+
 \newcommand{\lngpMADnp}{0.02}
@@
 \newcommand{\percentgpMADnp}{2.45}
+\newcommand{\gpRMSnp}{7.48}
+
 \newcommand{\lngpMAD}{0.02}
@@
 \newcommand{\percentgpMAD}{1.96}
+\newcommand{\gpRMS}{7.12}
 
@@
 \newcommand{\percenttelavivMAD}{1.75}
-
-\newcommand{\acfRMS}{1.75}
 \newcommand{\telavivRMS}{25.09}
-\newcommand{\pgramRMS}{19.50}
-\newcommand{\gpRMSnp}{7.48}
-\newcommand{\gpRMS}{7.12}
 
@@
 of the three techniques, providing a large improvement over the LS periodogram
-method and a moderate one over our basic implementation of the ACF method.
-We also find that our basic ACF method which is similar to methods implemented
-in the literature, often slightly under-predicts the rotation period due to a
-subtle feature of its peak detection algorithm.
+method and a moderate one over the ACF method.
+We also find that the ACF method often slightly under-predicts the rotation
+period due to a feature of its peak detection algorithm.
 Comparing our results with the results of the \citet{Aigrain2015} Tel Aviv
@@
 
-Unlike the ACF and L-S periodogram methods, the GP method provides posterior
+Unlike the ACF and LS periodogram methods, the GP method provides posterior
 PDF samples which can be used to estimate rotation periods uncertainties.
-Although an improvement on competing methods, these uncertainties are still
-underestimated in many cases because the GP model is an approximate, effective
-model. \tdmcomment{Can we make a statement something like ``we believe the method underestimates
-uncertainties by about a factor of three'' or something like that?  This
-matches with 66\% of the fits being within ``3$\sigma$''.
-Might be good to say something like this, both in the discussion and here.}
-% in many cases the posterior PDFs of the parameters are
-% multi-modal.
-% If it were possible to run an MCMC sampler for an infinite number of steps,
-% the entire posterior PDF would eventually be sampled, and a more accurate
-% uncertainty would result.
-% Unfortunately, due to realistic computational resources and the limitations of
-% MCMC sampling techniques for multi-modal posteriors, we are not always able to
-% map out the entire posterior PDFs.
-% \tdmcomment{I don't think this is the problem;
-% I think the fits that we judge to have converged have actually converged.}
-In addition, although the GP model used here is clearly a good one, it is
-still only an effective model, not an accurate physical model.
-It can only capture the posterior PDF of the periodic component of the
-covariance matrix, not the actual rotation period of a physical star.
-
+In general it provides more accurate rotation periods than either method and
+produces significantly fewer outliers.
+The GP method can also be applied to non-uniformly sampled light curves,
+unlike the ACF method.
+
+Although an improvement on competing methods, the uncertainties are still
+underestimated by the GP method in many cases because it is an approximate
+model.
+Although the GP model used here is clearly a good one, as it has been
+demonstrated to reliably recover rotation periods, it is still only an
+effective model, not an accurate physical model.
+It captures the posterior PDF of the periodic component of the covariance
+matrix of a time series, not the actual rotation period of a physical star.
+The marginalised posterior PDF does not therefore exactly reflect the
+probability of a rotation period, given the data.
+Similarly, the uncertainty, calculated from the 16th and 84th percentiles of
+the posterior PDF, is not exactly the standard deviation of a physical
+rotation period.
+
+Although underestimated uncertainties are a drawback of this method, they are
+at least well motivated and well defined which is not the case for
+uncertainties calculated from the ACF and LS periodogram methods.
 The main aim of this work was to develop a probabilistic rotation period
@@
 This paper includes data collected by the Kepler mission. Funding for the
-Kepler mission is provided by the NASA Science Mission directorate.
-
-% \appendix
-
-% \section{An ACF-informed Prior for Rotation Period}
-
-% The autocorrelation function (ACF) has proven to be very useful
-% for measuring the rotation periods of stars from time-series
-% photometry.
-% However, as discussed in the body of this paper, the
-% ACF-based method has several shortcomings,
-% most notably the inability to deliver uncertainties, but also
-% the necessity of several heuristic choices,
-% such as a timescale on which to smooth the ACF,
-% how to define a peak, whether the first or second peak
-% gets selected, and what constitutes a secure detection.
-% While this paper presents a new MCMC-based method
-% that avoids these shortcomings,
-% it seems prudent to still use information available from the ACF.
-% This Appendix describes how we use the ACF to define a prior on period.
-
-% Our strategy here is not to attempt to decide which single
-% peak in the ACF best represents the true rotation period,
-% but rather to identify several \emph{candidate} periods and define
-% a weighting scheme in order to create a non-commital, though useful,
-% multimodal prior.  While this procedure does not avoid heuristic choices,
-% the fact that its goal is simply to create a \emph{prior} for a
-% probabilitistically justified inference scheme rather than
-% to identify a single correct period softens the potential impact
-% of these choices.
-
-% As another innovation beyond what ACF methods in the literature have
-% presented, we also bandpass filter the light curves (using a 5th order
-% Butterworth filter, as implemented in \texttt{SciPy}) before calculating the
-% autocorrelation.  This suppresses power on timescales shorter than a chosen
-% minimum period $P_{\rm min}$ and longer than a chosen maximum $P_{\rm max}$,
-% producing a cleaner autocorrelation signal than an unfiltered light curve.
-
-% We use the following procedure to construct a prior for rotation period
-% given a light curve:
-% \begin{enumerate}
-% \item{For each value of $P_i$, where $i = \{1, 3, 5, 10, 30, 50, 100\}$\,d,
-% we apply a bandpass filter to the light curve using $P_{\rm min}=0.1$\,d
-% and $P_{\rm max} = P_i$.  We then calculate the ACF of the filtered
-% light curve out to a maximum lag of $2P_i$ and smooth it with a boxcar
-% filter of width $P_i/10$.}
-
-% \item{We identify the time lag corresponding to the
-% first peak of each of these ACFs, as well as the first peak's
-% trough-to-peak height, creating a set of candidate periods
-% $T_i$ and heights $h_i$.}
-
-% \item{We assign a quality metric $Q_i$ to each of these candidate
-% periods, as follows.  First, we model the ACF as a
-% damped oscillator with fixed period $T_i$:
-% \begin{equation}
-% y = A e^{-t/\tau} \cos{\frac{2\pi t}{T_i} },
-% \end{equation}
-% where and $t$ is the lag time,
-% and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
-% least squares minimization procedure.  We then define the
-% following heuristic quality metric:
-% \begin{equation}
-% \label{eq:quality}
-% Q_i = \left(\frac{\tau_i}{T_i}\right) \left(\frac{N_i h_i}{R_i}\right),
-% \end{equation}
-% where $h_i$ is the height of the ACF peak at $T_i$,
-% $N_i$ is the length of the lag vector in the ACF (directly proportional
-% to the maximum allowed period $P_i$),
-% and $R_i$ is the sum of squared residuals between the
-% damped oscillator model and the actual ACF data.  The idea behind this
-% quality metric to give a candidate period a higher score if
-
-%     \begin{enumerate}[(a)]
-
-%     \item{it has many regular sinusoidal peaks, such that the decay
-%         time $\tau_i$ is long compared to the oscillation period $T_i$,}
-%     \item{the ACF peak height is high, and}
-%     \item{the damped oscillator model is a good fit (in a $\chi^2$
-%       sense) to the ACF, with extra bonus for being
-%       a good fit over more points (larger $N_i$, or longer $P_i$).}
-
-%     \end{enumerate}
-% }
-
-% \item{Given this set of candidate periods $T_i$ and quality metrics $Q_i$,
-% we finally construct a multimodal prior on the $P$ parameter of the GP
-% model as a weighted mixture of Gaussians:
-% \begin{equation}
-% \label{eq:mixture}
-% p(\ln P) = \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) +
-%                                           0.05\mathcal N(\ln (T_i/2), 0.2) +
-%                                           0.05\mathcal N(\ln (2 T_i), 0.2) \right)}
-%                 {\sum_i Q_i}.
-% \end{equation}
-% That is, in addition to taking the candidate periods themselves as mixture
-% components, we also mix in twice and half each candidate period at a lower level,
-% which compensates for the possibility that the first peak in the ACF may actually
-% represent half or twice the actual rotation period.  The period width of 0.2 in
-% log space (corresponding to roughly 20\% uncertainty) is again a heuristic choice,
-% balancing a healthy specificity with the desire to not have the results of
-% the inference overly determined by the ACF prior.
-% }
-
-% Incidentally, while we use the procedure described here to create a prior on
-% $P$ which we use while inferring the parameters of the quasi-periodic GP model,
-% this same procedure may also be used in the service of a
-% rotation-period estimating procedure all on its own, perhaps being even more
-% robust and accurate than the traditional ACF method.  We leave exploration of
-% this possibility to future work.
+\Kepler\ mission is provided by the NASA Science Mission directorate.
 

### 2017-02-17 documents/ms.tex
wording changes to abstract and intro

@@
 \begin{abstract}
-The light curves of spotted, rotating stars often vary in a non-sinusoidal and
-quasi-periodic fashion---spots move on the stellar surface and have finite
-lifetimes, causing stellar flux variations to slowly shift in phase.
-A strictly periodic sinusoid therefore cannot accurately model a
-rotationally modulated stellar light curve.
-Physical models of the stellar surface also have many drawbacks preventing
-    effective inference, such as highly degenerate or high-dimensional
-    parameter spaces.
-% \citep[\eg][]{Russell1906, Jeffers2009, Kipping2012}.
+Variability in the light curves of spotted, rotating stars is often
+    non-sinusoidal and quasi-periodic --- spots move on the stellar surface
+    and have finite lifetimes, causing stellar flux variations to slowly shift
+    in phase.
+A strictly periodic sinusoid therefore cannot accurately model a rotationally
+    modulated stellar light curve.
+Physical models of stellar surfaces have many drawbacks preventing effective
+    inference, such as highly degenerate or high-dimensional parameter spaces.
 In this work, we introduce an appropriate {\it effective} model: a Gaussian
@@
     autocorrelation function method.
-We also demonstrate that it works effectively on real data, by
-inferring rotation periods for \nkoimcq\ \Kepler\ stars with
-previously measured periods.
-Because this method delivers posteriors PDFs, it will enable hierarchical
+We also demonstrate that it works effectively on real data, by inferring
+    rotation periods for \nkoimcq\ \Kepler\ stars with previously measured
+    periods.
+Because this method delivers posterior PDFs, it will enable hierarchical
     studies involving stellar rotation, particularly those involving
@@
     exoplanet systems, or characterising star-planet interactions.
-The code is available at \url{https://github.com/RuthAngus/GProtation/} .
-% Thehe ACF method is not applicable to unevenly spaced data,
+The code used to implement this new method is available at
+    \url{https://github.com/RuthAngus/GProtation/} .
+% The ACF method is not applicable to unevenly spaced data,
 % therefore it is inappropriate for the ~$10^7$ stellar light curves expected
@@
 
-The brightness of a spotted, rotating star often varies in a non-sinusoidal and
-Quasi-Periodic (QP) manner, due to active regions on its surface which
-rotate in and out of view.
-Complicated surface spot patterns produce non-sinusoidal variations,
-and the finite lifetimes of these active regions and
-differential rotation on the stellar surface produce quasi-periodicity
-\citep{Dumusque2011}.
+The disk-integrated flux of a spotted, rotating star often varies in a
+non-sinusoidal and Quasi-Periodic (QP) manner, due to active regions on its
+surface which rotate in and out of view.
+Complicated surface spot patterns produce non-sinusoidal variations, and the
+finite lifetimes of these active regions and differential rotation on the
+stellar surface produce quasi-periodicity \citep{Dumusque2011}.
 A strictly periodic sinusoid is therefore not necessarily a good model choice
@@
 
-Standard methods to measure rotation periods include detecting
-peaks in a Lomb-Scargle \citep{Lomb1976, Scargle1982} (LS) periodogram
-\citep[\eg][]{Reinhold2013}, Auto-Correlation Functions (ACFs),
+Standard non-inference based methods to measure rotation periods include
+detecting peaks in a Lomb-Scargle \citep{Lomb1976, Scargle1982} (LS)
+periodogram \citep[\eg][]{Reinhold2013}, Auto-Correlation Functions (ACFs),
 \citep[\eg][]{Mcquillan13b} and wavelet transforms \citep[\eg][]{Garcia2014}.
 The precisions of the LS periodogram and wavelet methods are limited by the
-suitability of the model choice: a sinusoid for the LS periodogram,
-and a choice of mother wavelet, assumed to describe the data
-over a range of transpositions \citep[see, \eg][]{Carter2010},
-for the wavelet method.
-In contrast, the ACF method is much better suited to signals that are
-non-sinusoidal.
+suitability of the model choice: a sinusoid for the LS periodogram, and a
+choice of mother wavelet, assumed to describe the data over a range of
+transpositions \citep[see, \eg][]{Carter2010}, for the wavelet method.
+In contrast, since it does not rely on a generative model, the ACF method is
+much better suited to signals that are non-sinusoidal.
 In fact, as long as the signal is approximately periodic the ACF will
 display a peak at the rotation period, no matter its shape.
-A drawback of the ACF method, however, is that it requires data to be
+A drawback of the ACF method however, is that it requires data to be
 evenly-spaced\footnote{\citet{Edelson1988} describe a method for computing
@@
 light curves (although in many cases it can be approximated as uniformly
-sampled) and not nearly the case for most ground-based measurements, such
-as the future Large Synoptic Survey Telescope (\LSST).
+sampled) and not nearly the case for most ground-based measurements, such as
+the future Large Synoptic Survey Telescope (\LSST).
 An ACF is also an operation performed on the data rather than a generative
@@
 formally propagated to constraints on the rotation period.
-% Many rotation periods in the literature have been inferred by measuring the
-% position of the first peak in an ACF, however this approach can be dangerous.
-% The exponential decay in correlation can shift the peak position short-wards
-% of its true value, leading to an underestimate of the period.
-% We return to this point in section \textsection \ref{sec:discussion}.
-
-In this work, we introduce an {\it effective} model for rotationally
-modulated stellar light curves
-which captures the salient behaviour but is not
-physically motivated---although some parameters may indeed be
-{\it interpreted} as physical ones.
-An ideal effective model should have a small number of non-degenerate parameters
-and be flexible enough to perfectly capture non-sinusoidal and QP behaviour.
+
+In this work, we introduce an {\it effective} model for rotationally modulated
+stellar light curves which captures the salient behaviour but is not
+physically motivated --- although some parameters may indeed be {\it
+interpreted} as physical ones.
+An ideal effective model should have a small number of non-degenerate
+parameters and be flexible enough to perfectly capture non-sinusoidal and QP
+behaviour.
 A Gaussian process (GP) model fulfills these requirements. We thus use a GP
 as the generative model at the core of a method to probabilistically
-infer accurate and precise stellar rotation periods.  This enables us to
-estimate the posterior PDF of the rotation period, thereby producing a
-justified estimate of its uncertainty.
-% while simultaneously marginalizing over
-% parameters which model correlated noise.
+infer accurate and precise stellar rotation periods.
+This enables us to estimate the posterior PDF of the rotation period, thereby
+producing a justified estimate of its uncertainty.
 

### 2017-03-13 documents/ms.tex
converting to MNRAS format

@@
-\documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
+% \documentclass[useAMS, usenatbib, preprint, 12pt]{aastex}
+\documentclass[a4paper,fleqn,usenatbib,useAMS]{mnras}
 \usepackage{cite, natbib}
-\usepackage{float}
-\usepackage{epsfig}
-\usepackage{cases}
-\usepackage[section]{placeins}
+% \usepackage{float}
+% \usepackage{epsfig}
+% \usepackage{cases}
+% \usepackage[section]{placeins}
 \usepackage{graphicx, subfigure}
@@
 \usepackage{bm}
-\usepackage{enumerate}
+\usepackage{amsmath}
+% \usepackage{enumerate}
+
+% From mnras_guide.tex
+\usepackage[T1]{fontenc}
+\usepackage{ae,aecompl}
+\usepackage{newtxtext,newtxmath}
 
@@
 
-\newcommand{\lnacfMAD}{0.14}
-\newcommand{\acfMAD}{2.36}
-\newcommand{\percentacfMAD}{13.02}
-\newcommand{\acfRMS}{26.09}
-
+\newcommand{\acfRMS}{4.58}
+\newcommand{\acfMAD}{0.37}
+\newcommand{\lnacfMAD}{0.02}
+\newcommand{\percentacfMAD}{1.75}
+
+\newcommand{\pgramRMS}{18.98}
+\newcommand{\pgramMAD}{1.13}
 \newcommand{\lnpgramMAD}{0.05}
-\newcommand{\pgramMAD}{1.13}
 \newcommand{\percentpgramMAD}{4.70}
-\newcommand{\pgramRMS}{18.98}
-
-\newcommand{\lngpMADnp}{0.02}
+
+\newcommand{\gpRMS}{0.25}
+\newcommand{\gpMAD}{0.40}
+\newcommand{\lngpMAD}{0.02}
+\newcommand{\percentgpMAD}{1.97}
+
+\newcommand{\gpRMSnp}{0.44}
 \newcommand{\gpMADnp}{0.48}
-\newcommand{\percentgpMADnp}{2.45}
-\newcommand{\gpRMSnp}{7.48}
-
-\newcommand{\lngpMAD}{0.02}
-\newcommand{\gpMAD}{0.39}
-\newcommand{\percentgpMAD}{1.96}
-\newcommand{\gpRMS}{7.12}
-
+\newcommand{\lngpMADnp}{0.03}
+\newcommand{\percentgpMADnp}{2.62}
+
+\newcommand{\telavivRMS}{4.58}
+\newcommand{\telavivMAD}{0.37}
 \newcommand{\lntelavivMAD}{0.02}
-\newcommand{\telavivMAD}{0.34}
 \newcommand{\percenttelavivMAD}{1.75}
-\newcommand{\telavivRMS}{25.09}
 
@@
 
+% \begin{document}
+
+% \title{Inferring probabilistic stellar rotation periods using Gaussian
+% processes}
+
+% \author{%
+%    Ruth Angus\altaffilmark{\columbia},
+%    Timothy Morton\altaffilmark{\princeton},
+%    Suzanne Aigrain\altaffilmark{\oxford},
+%    Daniel Foreman-Mackey\altaffilmark{\uw,\sagan}
+%    \& Vinesh Rajpaul\altaffilmark{\oxford}
+% }
+% \altaffiltext{\columbia}{Simons Fellow, Department of Astronomy, Columbia
+% University, NY, NY, RuthAngus@gmail.com}
+% \altaffiltext{\princeton}{Department of Astrophysical Sciences, Princeton
+% University,  Princeton, NJ}
+% \altaffiltext{\oxford}{Subdepartment of Astrophysics, University of Oxford,
+% UK}
+% \altaffiltext{\uw}{Department of Astronomy, University of
+% Washington, Seattle, WA}
+% \altaffiltext{\sagan}{Sagan Fellow}
+
+\title[GP rotation periods]{Inferring probabilistic stellar rotation periods
+using Gaussian processes}
+
+\author[R.~Angus \emph{et al.}]{
+    Ruth~Angus,$^1$\thanks{Contact e-mail:
+\href{mailto:ruthangus@astro.columbia.edu}{ruthangus@astro.columbia.edu}}
+    Timothy Morton$^2$,
+    Suzanne Aigrain$^3$,
+    Daniel Foreman-Mackey$^4$,
+    \newauthor
+    \& Vinesh Rajpaul$^3$
+    \\
+    $^1$Simons Fellow, Department of Astronomy, Columbia University, NY, NY \\
+    $^2$Department of Astrophysical Sciences, Princeton University,
+    Princeton, NJ \\
+    $^3$Subdepartment of Astrophysics, University of Oxford, UK \\
+    $^4$Sagan Fellow, Department of Astronomy, University of Washington,
+    Seattle, WA}
+
+\date{Last updated 2017 March 13; in original form 2017 March 13}
+\pubyear{2017}
 \begin{document}
-
-\title{Inferring probabilistic stellar rotation periods using Gaussian
-processes}
-
-\author{%
-   Ruth Angus\altaffilmark{\columbia},
-   Timothy Morton\altaffilmark{\princeton},
-   Suzanne Aigrain\altaffilmark{\oxford},
-   Daniel Foreman-Mackey\altaffilmark{\uw,\sagan}
-   \& Vinesh Rajpaul\altaffilmark{\oxford}
-}
-\altaffiltext{\columbia}{Simons Fellow, Department of Astronomy, Columbia
-University, NY, NY, RuthAngus@gmail.com}
-\altaffiltext{\princeton}{Department of Astrophysical Sciences, Princeton
-University,  Princeton, NJ}
-\altaffiltext{\oxford}{Subdepartment of Astrophysics, University of Oxford,
-UK}
-\altaffiltext{\uw}{Department of Astronomy, University of
-Washington, Seattle, WA}
-\altaffiltext{\sagan}{Sagan Fellow}
+\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
 
@@
     exoplanet systems, or characterising star-planet interactions.
-The code used to implement this new method is available at
-    \url{https://github.com/RuthAngus/GProtation/} .
+The code used to implement this new method is available online.
 % The ACF method is not applicable to unevenly spaced data,
@@
 \end{abstract}
+
+\begin{keywords}
+    stars --- rotation,
+    stars --- starspots,
+    stars --- solar-type,
+    methods --- data-analysis,
+    methods --- statistical,
+    techniques --- photometric.
+\end{keywords}
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/koi_lc_demo.pdf}
+\includegraphics[width=\columnwidth, clip=true]{figures/koi_lc_demo.pdf}
 \caption[A light curve with a GP model.]
@@
 model as a weighted mixture of Gaussians:
-\begin{equation}
+\begin{multline}
 \label{eq:mixture}
-p(\ln P) = \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) +
+p(\ln P) = \\ \frac {\displaystyle \sum_i Q_i \left(0.9\mathcal N(\ln T_i, 0.2) +
                                           0.05\mathcal N(\ln (T_i/2), 0.2) +
@@
                 {\sum_i Q_i}.
-\end{equation}
+\end{multline}
 That is, in addition to taking the candidate periods themselves as mixture
@@
 0.4, and 0.2 respectively.
-We run \vrcomment{50} steps of the sampler at a time, checking for convergence
-after each iteration, up to a maximum of \vrcomment{50} iterations.
+We run 50 steps of the sampler at a time, checking for convergence
+after each iteration, up to a maximum of 50 iterations.
 We declare convergence if the total effective chain length is at least
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/period_hist.pdf}
+\includegraphics[width=\columnwidth, clip=true]{figures/period_hist.pdf}
 \caption{A histogram of the rotation periods used to generate the \naigrain\
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/demo_lc.pdf}
+\includegraphics[width=\columnwidth, clip=true]{figures/demo_lc.pdf}
 \caption[A simulated light curve.]
@@
 figure \ref{fig:gp_posteriors}.
-The posterior samples of the model parameters for the simulated light curves
-are available online:
-\url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
+% The posterior samples of the model parameters for the simulated light curves
+% are available online:
+% \url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
 
@@
 % \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_02_13.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_03_10.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 
-We measure an ACF-based period for each light curve,
-following the method of
-\citet{Mcquillan2013}.
-After calculating the ACF for a light curve, we smooth it by convolving with a
-Gaussian filter ($\sigma=9$\,d), and select the rotation period as the
-lag-time of the highest peak in the ACF less than 100\,d.
-This is not always the first peak---the second can be larger than the first if
-two active regions are at or near opposite longitudes on the surface of
-the star, producing a light curve with two dips per rotation period.
-We restrict our search to periods less than 100\,d because instrumental
-noise and light-curve systematics significantly distort \kepler\ light
-curves on longer timescales, rendering longer periods realistically unrecoverable.
-Figure \ref{fig:demo_lc} contains an example ACF of the light curve in
-Figure \ref{fig:demo_acf}.
+% We measure an ACF-based period for each light curve,
+% following the method of
+% \citet{Mcquillan2013}.
+% After calculating the ACF for a light curve, we smooth it by convolving with a
+% Gaussian filter ($\sigma=9$\,d), and select the rotation period as the
+% lag-time of the highest peak in the ACF less than 100\,d.
+% This is not always the first peak---the second can be larger than the first if
+% two active regions are at or near opposite longitudes on the surface of
+% the star, producing a light curve with two dips per rotation period.
+% We restrict our search to periods less than 100\,d because instrumental
+% noise and light-curve systematics significantly distort \kepler\ light
+% curves on longer timescales, rendering longer periods realistically unrecoverable.
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/demo_ACF.pdf}
+\includegraphics[width=\columnwidth, clip=true]{figures/demo_ACF.pdf}
 \caption[ACF of a simulated light curve.]
@@
 disadvantage.
-
-We apply the ACF method to the sample of \naigrain\ simulated light
-curves.
+Figure \ref{fig:demo_lc} contains an example ACF of the light curve in
+Figure \ref{fig:demo_acf}.
+
+We compare our results to those of the Tel Aviv team in \citet{Aigrain2015}
+who use the ACF method described in \citet{Mcquillan2013}.
 Figure \ref{fig:compare_acf} shows ACF-measured versus true rotation
 periods, with the $2n$ and $\frac{1}{2}n$ harmonic lines as dashed lines.
+The MRD of their periods, relative to the true rotation periods, is
+\percenttelavivMAD\% --- smaller than the results produced by the GP
+method.
+However, their results contain a larger number of extreme outliers --- they
+predict
+periods that are significantly different from the true period in more cases
+than the GP method.
+This is revealed by comparing the root-mean-squared (RMS) of the two sets of
+results since the RMS is more sensitive to outliers.
+The RMS of the Tel Aviv method results is \telavivRMS\ days and the RMS of GP
+method results is \gpRMS\ days.
 The MRD of the ACF-recovered periods is \percentacfMAD\% (see table
@@
 exponentials.
-In such functions the peak positions can be shifted towards the left
-(towards shorter periods), because the decaying exponential raises the left side of
+In such functions the peak positions can be shifted towards the left (towards
+shorter periods), because the decaying exponential raises the left side of
 each peak more than the right.
@@
 simply measures the peak position without taking it into account.
-We adopt this standard practice here to faithfully compare
-our new method to that used in the literature.
-
-We also compare our results to those of the Tel Aviv team in
-\citet{Aigrain2015} who use the ACF method described in \citet{Mcquillan2013}.
-Although we follow the same general ACF method in this paper, the exact
-implementation of the \citet{Mcquillan2013} code is slightly different and
-produces different results.
-The MRD of their periods, relative to the true rotation periods, is
-\percenttelavivMAD\% --- smaller than the results produced by the GP
-method.
-However, their results contain a larger number of extreme outliers --- they
-predict
-periods that are significantly different from the true period in more cases
-than the GP method.
-This is revealed by comparing the root-mean-squared (RMS) of the two sets of
-results since the RMS is more sensitive to outliers.
-The RMS of the Tel Aviv method results is \telavivRMS\ days and the RMS of GP
-method results is \gpRMS\ days.
+
 % This can be understood as a consequence of the ACF method's sensitivity to
@@
     defined in \citet{Aigrain2015}.
-    Several light curves have over-estimated rotation periods and some
-    are drastically underestimated.}
+    Several light curves have drastically over- and under-estimated rotation
+    periods.
+    For the points that lie close to the $x=y$ line, there is excess scatter
+    beneath the line due to a tendency of the ACF method to slightly
+    underpredict rotation periods.  }
 \label{fig:compare_acf}
@@
 
-For each simulated light curve, we compute a LS periodogram
-\footnote{LS periodograms were calculated using the gatspy Python module:
-\url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.}
-over a grid of 10,000 periods, evenly spaced in frequency,
-between 1 and 100 days.
+The \citet{Aigrain2015} Tel Aviv team were the only group to report periods
+for all \naigrain\ stars, and are therefore the only team with which we
+compare results from this paper.
+The other participating groups chose to report only the period measurements
+they were confident in and are likely to have omitted large outliers.
+So, in order to test our results against an additional commonly used
+technique, we implemented a simple LS periodogram method.
+
+For each simulated light curve, we compute a LS periodogram\footnote{LS
+periodograms were calculated using the gatspy Python module:
+\url{https://github.com/astroML/gatspy/tree/master/gatspy/periodic}.} over a
+grid of 10,000 periods, evenly spaced in frequency, between 1 and 100 days.
 We adopt the period of the highest peak in the periodogram as the rotation
@@
     LS periodogram & \pgramMAD\ days & \percentpgramMAD \% & \pgramRMS\ days\\
-    ACF & \acfMAD\ days & \percentacfMAD \% & \acfRMS\ days \\
-    ACF \citep{Mcquillan2013} & \telavivMAD\ days & \percenttelavivMAD \%
+    ACF & \telavivMAD\ days & \percenttelavivMAD \%
     & \telavivRMS\ days \\
@@
 The final priors and parameter ranges that we used are in Table \ref{tab:koipriors}.
-The posterior samples of the model parameters for the \Kepler\ objects of
-interest are available online:
-\url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
+% The posterior samples of the model parameters for the \Kepler\ objects of
+% interest are available online:
+% \url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
 
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
@@
 \end{center}
-\end{figure}
+\end{figure*}
 

### 2017-03-14 documents/ms.tex
last few edits

@@
 choice of mother wavelet, assumed to describe the data over a range of
-transpositions \citep[see, \eg][]{Carter2010}, for the wavelet method.
+transpositions and scales \citep[see, \eg][]{Carter2010}, for the wavelet
+method.
 In contrast, since it does not rely on a generative model, the ACF method is
@@
 \citep[see \eg][]{Gibson2012, Haywood2014, Dawson2014, Barclay2015,
-Haywood2015, Evans2015, Rajpaul2015, Rajpaul2016, Aigrain2016}.
+Haywood2015, Evans2015, Rajpaul2015, Czekala2015, Rajpaul2016, Aigrain2016}.
 They are useful in regression problems involving any stochastic process,
@@
 
-A range of kernel functions can describe stellar variability.
+A range of covariance models, or kernel functions, can describe stellar
+variability.
 For example, the most commonly used kernel function, the `Squared
@@
 that are not captured by the effective GP model.
-We use this QP kernel function to produce the GP model
-that fits the \Kepler\ light curve in in Figure \ref{fig:GP_example}.
+
+We use this QP kernel function, defined in equation \ref{eqn:QP} to produce
+the GP model that fits the \Kepler\ light curve in in Figure
+\ref{fig:GP_example}.
 There are many ways to construct a QP kernel function, involving a range of
@@
 ${\bf K}$ replaces ${\bf C}$ in the above equation (for our purposes, the QP
-kernel of Equation \ref{eq:QP}).
+kernel of Equation \ref{eq:QP} generates ${\bf K}$).
 
 Evaluating this likelihood for a large number of points can be computationally
-expensive.  For example, evaluating
-$\mathcal L$ for an entire \Kepler\ lightcurve
-($\sim$40,000 points) takes about $\sim$5\,s--- too slow
-to perform inference on large numbers of light curves\footnote{All
-computational times cited in this section are based on evaluations on a
-single core of a 2015 Macbook Pro, 3.1 GHz Intel Core i7.}.
-The matrix operations necessary to evaluate the GP likelihood\footnote{These
-operations use the fast matrix solver HODLR \citep{Ambikasaran2014},
-implemented in the {\tt george} \citep{George} python package.} scale as
+expensive.
+For example, evaluating $\mathcal L$ for an entire \Kepler\ lightcurve
+($\sim$40,000 points) takes about $\sim$5\,s--- too slow to perform inference
+on large numbers of light curves\footnote{All computational times cited in
+this section are based on evaluations on a single core of a 2015 Macbook Pro,
+3.1 GHz Intel Core i7.}.
+The matrix operations necessary to evaluate the GP likelihood scale as
 $N\ln(N)^2$, where $N$ is the number of data points in the light curve.
+These operations use the fast matrix solver HODLR \citep{Ambikasaran2014},
+implemented in the {\tt george} \citep{George} python package.
 
@@
 works well may differ among datasets.
-See \textsection \ref{sec:discussion} for a discussion of more physically
-motivated priors on the non-period parameters.
+% See \textsection \ref{sec:discussion} for a discussion of more physically
+% motivated priors on the non-period parameters.
 
@@
 
-In order to benchmark this new rotation period recovery method, we apply
-it to a set of simulated light curves
-and compare to the performance of established literature methods.
+In order to test this new rotation period recovery method, we apply it to a
+set of simulated light curves and compare to the performance of established
+literature methods.
 Section \ref{sec:simulations} describes the simulated data we use;
@@
 in Table \ref{tab:priors}.
-These priors were determined experimentally --- see \textsection
-\ref{sec:discussion} for a discussion on more physically motivated priors.
+% These priors were determined experimentally --- see \textsection
+% \ref{sec:discussion} for a discussion on more physically motivated priors.
 % We note that we do not have any specific quantitative justification for the
@@
 Figures \ref{fig:compare_mcmc_acfprior} and \ref{fig:compare_mcmc_noprior}
-summarise our results compared to the injected
-`true' stellar rotation periods, for the uninformed and ACF-informed priors on
-period, respectively.
-To assess the performance of the QP-GP and other period recovery
-methods, we compute Median Absolute Deviations (MADs) of the results, relative
-to the input periods.
-We also compute the Median Relative Deviation (MRD), as a percentage.
+summarise our results compared to the injected `true' stellar rotation
+periods, for the uninformed and ACF-informed priors on period, respectively.
+The periods inferred using the GP method show good agreement with the true
+periods.
+To assess the performance of the QP-GP and other period recovery methods, we
+compute Median Absolute Deviations (MADs) of the results, relative to the
+input periods.
+We also compute the Median Relative Deviation (MRD), as a percentage and the
+root-mean-squared (RMS).
 These metrics are presented for the three different methods tested in this
@@
 simulated light curves vs the rotation periods measured using the GP
-technique.
+technique with an informed, ACF-based prior.
     Points are coloured by the peak-to-peak amplitude of the light curve, as
@@
 Another advantage is its fast implementation speed.
-However, because the ACF method is non-probabilistic, ACF-estimated
-rotation period uncertainties are poorly defined---a clear
-disadvantage.
-Figure \ref{fig:demo_lc} contains an example ACF of the light curve in
-Figure \ref{fig:demo_acf}.
+However, because the ACF method is non-probabilistic, ACF-estimated rotation
+period uncertainties are poorly defined~---~a clear disadvantage.
+It also requires evenly-spaced data, therefore is not applicable to
+ground-based light curves.
+Figure \ref{fig:demo_acf} contains an example ACF of the light curve in
+Figure \ref{fig:demo_lc}.
 
@@
 periods, with the $2n$ and $\frac{1}{2}n$ harmonic lines as dashed lines.
-The MRD of their periods, relative to the true rotation periods, is
-\percenttelavivMAD\% --- smaller than the results produced by the GP
-method.
+The injected and recovered rotation periods generally agree well: the MRD of
+their periods, relative to the true rotation periods, is \percenttelavivMAD\%
+--- smaller than the results produced by the GP method.
 However, their results contain a larger number of extreme outliers --- they
-predict
-periods that are significantly different from the true period in more cases
-than the GP method.
-This is revealed by comparing the root-mean-squared (RMS) of the two sets of
-results since the RMS is more sensitive to outliers.
+predict periods that are significantly different from the true period in more
+cases than the GP method.
+This is revealed by comparing the RMS of the two sets of results since the RMS
+is more sensitive to outliers.
 The RMS of the Tel Aviv method results is \telavivRMS\ days and the RMS of GP
 method results is \gpRMS\ days.
-The MRD of the ACF-recovered periods is \percentacfMAD\% (see table
-\ref{tab:MADs} for a side-by-side comparison with the other methods).
-The injected and recovered rotation periods generally agree well,
-though with a few drastic over- or underestimates.
-Additionally, of the points clustered close to the 1:1 line,
-more fall slightly below than above it, \ie\
+Another negative feature of the ACF method is revealed by examining the points
+clustered close to the 1:1 line: more fall slightly below than above it, \ie\
 rotation periods are systematically underestimated.
@@
 for all \naigrain\ stars, and are therefore the only team with which we
-compare results from this paper.
-The other participating groups chose to report only the period measurements
-they were confident in and are likely to have omitted large outliers.
-So, in order to test our results against an additional commonly used
+compare results.
+The other participating groups chose to report only the period measurements in
+which they were confident and are likely to have omitted large outliers.
+In order therefore to test our results against another commonly used
 technique, we implemented a simple LS periodogram method.
@@
 grid of 10,000 periods, evenly spaced in frequency, between 1 and 100 days.
-We adopt the period of the highest peak in the periodogram as the rotation
-period.
+We adopt the period of the highest peak in the periodogram as the measured
+rotation period.
 Figure \ref{fig:pgram_compare} shows the resulting recovered rotation periods
 as a function of true period.
-The MRD of the periodogram-recovered periods is \percentpgramMAD\% (see table
-\ref{tab:MADs} for a side-by-side comparison with the other methods).
-This method drastically overestimates many rotation periods,
-because long-term trends unrelated to rotation can cause excess power
-at long periods.
-% a problem further exacerbated by instrumental noise.
+The MRD of the periodogram-recovered periods is \percentpgramMAD\%,
+significantly worse than the ACF and GP methods (see table \ref{tab:MADs} for
+a side-by-side comparison with the other methods).
+The LS periodogram method drastically overestimates many rotation periods
+because long-term trends unrelated to rotation can cause excess power at long
+periods.
 
@@
     root-mean-squared (RMS) for
-    the LS periodogram, ACF, \citet{Mcquillan2013} ACF and GP period recovery
+    the LS periodogram, \citet{Mcquillan2013} ACF and GP period recovery
     methods.}
@@
     \hline
-    LS periodogram & \pgramMAD\ days & \percentpgramMAD \% & \pgramRMS\ days\\
+    GP (acf prior) & \gpMAD\ days & \percentgpMAD \% & \gpRMS\ days \\
+    GP (uninformative prior) & \gpMADnp\ days & \percentgpMADnp \% &
+    \gpRMSnp\ days \\
     ACF & \telavivMAD\ days & \percenttelavivMAD \%
     & \telavivRMS\ days \\
-    GP (uninformative prior) & \gpMADnp\ days & \percentgpMADnp \% &
-    \gpRMSnp\ days \\
-    GP (acf prior) & \gpMAD\ days & \percentgpMAD \% & \gpRMS\ days \\
+    LS periodogram & \pgramMAD\ days & \percentpgramMAD \% & \pgramRMS\ days\\
 \end{tabular}
@@
 KOIs.
-This comparison shows generally very good agreement, with only a few
-exceptions, demonstrating that this method works not only on simulated data,
-but also on real data---with the caveat that for any particular data set, some
-care is needed regarding the setting the priors and ranges for the GP
-hyperparameters.
+This comparison shows generally very good agreement, demonstrating that this
+method works not only on simulated data, but also on real data---with the
+caveat that for any particular data set, some care is needed regarding the
+setting the priors and ranges for the GP hyperparameters.
 Notably, the GP method recovers periods
@@
 Unfortunately, these uncertainties still appear to be underestimated.
-Of the rotation periods recovered from the simulated light curves, using the
+Of the rotation periods recovered from the simulated light curves using the
 ACF-informed prior, only 25\% of the measured periods lie within 1$\sigma$ of
@@
 The median uncertainty is underestimated by around a factor of two.
-These numbers are similar for the uninformative prior results which also
-provides underestimated uncertainties.
+These numbers are similar for the uninformative prior results.
 % In the majority of these cases the uncertainties are underestimated due to the
@@
 We attribute these underestimated uncertainties to the model choice.
-Although more appropriate than a sinusoid, the GP is still only an
+Although more appropriate than a sinusoid, the QP-GP is still only an
  effective model.
@@
 \begin{itemize}
-\item{To perform model selection with different kernel functions.
-    We have not performed these tests yet due to the cost of calculating the
-        fully marginalised likelihood with GPs.}
 \item{To discuss more physically motivated priors.
@@
         efficient optimization or marginalization of hyperparameters, to a
-        separate paper (currently in prep.).
-    }
+        separate paper (currently in prep.).}
 \item{To build in a noise model for \kepler\ data.
-Because it is a {\it generative} model of the data,
-the GP method models the rotation period at the same time as systematic
-noise features.
-One can then marginalise over the parameters of the noise model.
+        Because it is a {\it generative} model of the data, the GP method
+        models the rotation period at the same time as systematic noise
+        features.  One can then marginalise over the parameters of the noise
+        model.
         This approach would be extremely advantageous for \kepler\ data since
-long-term trends are often removed by the \kepler\ detrending pipeline.
-Marginalising over the noise model at the same time as inferring the
-parameters of interest will insure that the periodic signal is preserved.
-\tdmcomment{Is this bullet point obselete, given that we have already successfully
-marginalized over \kepler\ noise to recover periods? Or is there
-something else we wanted to say here?}}
+        long-term trends are often removed by the \kepler\ detrending
+        pipeline.  Marginalising over the noise model at the same time as
+        inferring the parameters of interest will insure that the periodic
+        signal is preserved.}
 \end{itemize}
@@
 three different methods: a new Gaussian process method,
-an authocorrelation function method, and a Lomb-Scargle periodogram method.
+an autocorrelation function method, and a Lomb-Scargle periodogram method.
 We demonstrate that the GP method produces the most accurate rotation periods
@@
 probability of a rotation period, given the data.
-Similarly, the uncertainty, calculated from the 16th and 84th percentiles of
-the posterior PDF, is not exactly the standard deviation of a physical
+Similarly, the 16th and 84th percentile ranges calculated from the posterior
+PDF of the QP-GP model, are not exactly the uncertainty of a physical
 rotation period.
-
 Although underestimated uncertainties are a drawback of this method, they are
@@
 uncertainties calculated from the ACF and LS periodogram methods.
+
 The main aim of this work was to develop a probabilistic rotation period

### 2017-03-14 code/plot_koi.py
updating figures

@@ -41,16 +41,16 @@ def load_and_plot(data):
 
     plt.clf()
     xs = np.log(np.linspace(0, 100, 100))
-    plt.plot(xs, xs, "-", color=".7", zorder=0)
-    plt.plot(xs, xs + 2./3, "--", color=".7", zorder=0)
-    plt.plot(xs, xs - 2./3, "--", color=".7", zorder=0)
+    plt.plot(xs, xs, "-", color=".7", lw=.8, zorder=0)
+    plt.plot(xs, xs + 2./3, "--", color=".7", lw=.8, zorder=0)
+    plt.plot(xs, xs - 2./3, "--", color=".7", lw=.8, zorder=0)
 
     plt.errorbar(np.log(x[l]), np.log(recovered[l]), yerr=[lnerrp[l],
                  lnerrm[l]], xerr=lnxerr[l], fmt="k.", zorder=1, capsize=0,
-                 ecolor=".7", alpha=.5, ms=.1)
+                 ecolor=".7", alpha=.5, ms=.1, elinewidth=.8)
     plt.scatter(np.log(x[l]), np.log(recovered[l]), c=np.log(amp[l]),
                 edgecolor=".5", cmap="GnBu_r", vmin=min(np.log(amp[l])),
-                vmax=max(np.log(amp[l])), s=20, zorder=2, lw=.2)
+                vmax=max(np.log(amp[l])), s=10, zorder=2, lw=.2)
 
     plt.ylim(0, 4)
     plt.xlim(0, 4)
@@ -58,6 +58,7 @@ def load_and_plot(data):
     cbar.ax.set_ylabel("$\ln(\mathrm{R}_{\mathrm{var}})$")
     plt.xlabel("$\ln(\mathrm{McQuillan~{\it et~al.,}~2013~Period})$")
     plt.ylabel("$\ln(\mathrm{Recovered~Period})$")
+    plt.subplots_adjust(bottom=.15)
     plt.savefig("comparison_koi")
     plt.savefig(os.path.join(FIG_DIR, "comparison_koi_02_03.pdf"))
 

### 2017-03-14 documents/ms.tex

@@
 evenly-spaced\footnote{\citet{Edelson1988} describe a method for computing
-ACFs for unevenly-spaced data.}, which is not exactly the case with \Kepler\
-light curves (although in many cases it can be approximated as uniformly
-sampled) and not nearly the case for most ground-based measurements, such as
-the future Large Synoptic Survey Telescope (\LSST).
+ACFs for unevenly-spaced data.}.
+Most ground-based measurements, such as the future Large Synoptic Survey 
+Telescope (\LSST), do not nearly satisfy this criterion, and even 
+light curves from the \Kepler\ space telescope, which has provided the richest
+stellar rotation data to date, can only be approximated as uniformly sampled.
 An ACF is also an operation performed on the data rather than a generative
@@
 points separated by half a period to be weakly correlated.
-We also use an extra parameter, $\sigma$, which is an additional white noise
-term added to the diagonal elements of the covariance matrix.
+The additional parameter $\sigma$ captures white noise by adding
+a term to the diagonal of the covariance matrix.
 This can be interpreted to represent underestimation of observational
 uncertainties~---~if the uncertainties reported on the data are too small, it
-will be non-zero~---~or it can capture any remaining ``jitter'' or residuals
-that are not captured by the effective GP model.
-
-We use this QP kernel function, defined in equation \ref{eqn:QP} to produce
-the GP model that fits the \Kepler\ light curve in in Figure
+will be non-zero~---~or it can capture any remaining ``jitter,'' or residuals
+not captured by the effective GP model.
+We use this QP kernel function (Equation \ref{eq:QP}) to produce
+the GP model that fits the \Kepler\ light curve in Figure
 \ref{fig:GP_example}.
+
 There are many ways to construct a QP kernel function, involving a range of
-choices for both the periodic and a-periodic components of the model.
+choices for both the periodic and aperiodic components of the model.
 The kernel function presented above reproduces the behaviour of stellar light
-curves but it is not necessarily the only model choice that can do so.
-We do not attempt to test any other choices in this paper, noting only that
+curves, but other model choices can do so as well. 
+We do not attempt to test any other models in this paper, noting only that
 this kernel function provides an adequate fit to the data.
@@
 The GP likelihood is similar to the simple Gaussian likelihood
-function used for optimisation problems where the uncertainties are
+function where the uncertainties are
 Gaussian and uncorrelated. The latter can be written
@@
 In the case where the uncertainties are uncorrelated, the noise is `white',
-(which is a frequent assumption made by astronomers and is sometimes
-justified) and the off-diagonal elements of the covariance matrix are zero.
+(an assumption frequently made by astronomers, sometimes justified) 
+and the off-diagonal elements of the covariance matrix are zero.
 However, in the case where there is evidence for correlated
@@
 \begin{enumerate}
-\item{For each value of $P_i$, where $i = \{1, 3, 5, 10, 30, 50, 100\}$\,d,
+\item{For each value of $P_i$, where $i = \{1, 2, 4, 8, 16, 32, 64, 128\}$\,d,
 we apply a bandpass filter to the light curve using $P_{\rm min}=0.1$\,d
@@
 and find the best-fitting parameters $A_i$ and $\tau_i$ by a non-linear
-least squares minimization procedure.  We then define the
+least squares minimization procedure--though enforcing a maximum
+value of $\tau = 20 * P_i/T_i$ to avoid unreasonably
+large solutions of $\tau.  We then define the
 following heuristic quality metric:
@@
 their periods, relative to the true rotation periods, is \percenttelavivMAD\%
---- smaller than the results produced by the GP method.
+--- slightly smaller than the results produced by the GP method.
 However, their results contain a larger number of extreme outliers --- they
@@
 The \citet{Aigrain2015} Tel Aviv team were the only group to report periods
-for all \naigrain\ stars, and are therefore the only team with which we
+for all \naigrain\ simulated stars, and are therefore the only team with which we
 compare results.
@@
  effective model.
-A perfect physical model of the star would produce more representative
+A perfect physical model of the star would produce more accurate
 uncertainties.
@@
 anti-correlation of points $\frac{1}{2}$ a period apart.
-It would be very worthwhile to test this assumption and this alternative
+It would be worthwhile to test this assumption and this alternative
 kernel function (and others) in the future.
@@
 \begin{itemize}
-\item{To discuss more physically motivated priors.
+\item{To investigate more physically motivated priors.
         In this paper we have only explored the physical interpretation of one
@@
 of the three techniques, providing a large improvement over the LS periodogram
-method and a moderate one over the ACF method.
+method and a slight one over the ACF method.
+Comparing our results with the results of the \citet{Aigrain2015} Tel Aviv
+team who implement the ACF method of \citet{Mcquillan2013},  
+we find their results to be
+slightly more precise in a median sense than the GP method results, but with
+a larger number of significant outliers.
 We also find that the ACF method often slightly under-predicts the rotation
 period due to a feature of its peak detection algorithm.
-Comparing our results with the results of the \citet{Aigrain2015} Tel Aviv
-team who implement the method of \citet{Mcquillan2013}, a slightly different
-version of the ACF method to the one we use, we find that their results are
-slightly more precise than the GP method results overall.
-They do however have a larger number of significant outliers.
-We measured the rotation periods of \nkoimcq\ \kepler\ objects of interest
+Additionally, we measure the rotation periods of \nkoimcq\ 
+\kepler\ objects of interest
 using the GP method, and find that these results compare well to those

### 2017-03-14 documents/ms.tex
cleaning up figure directory and getting ready to submit

@@
 value of $\tau = 20 * P_i/T_i$ to avoid unreasonably
-large solutions of $\tau.  We then define the
-following heuristic quality metric:
+large solutions of $\tau$.
+We then define the following heuristic quality metric:
 \begin{equation}

### 2017-05-08 documents/ms.tex
adding Littlefair, vanderburg and stressing the not novel

@@
 
-\newcommand{\pgramRMS}{18.98}
-\newcommand{\pgramMAD}{1.13}
-\newcommand{\lnpgramMAD}{0.05}
-\newcommand{\percentpgramMAD}{4.70}
+\newcommand{\oldpgramRMS}{18.98}
+\newcommand{\oldpgramMAD}{1.13}
+\newcommand{\oldlnpgramMAD}{0.05}
+\newcommand{\oldpercentpgramMAD}{4.70}
+
+\newcommand{\pgramRMS}{1.04}
+\newcommand{\pgramMAD}{0.52}
+\newcommand{\lnpgramMAD}{0.03}
+\newcommand{\percentpgramMAD}{2.82}
 
@@
 
-In this work, we introduce an {\it effective} model for rotationally modulated
+In this work, we test an {\it effective} model for rotationally modulated
 stellar light curves which captures the salient behaviour but is not
@@
 producing a justified estimate of its uncertainty.
+{\bf Using a quasi-periodic GP model to infer a rotation period is not a new idea.
+We used this method in \citet{Vanderburg2015} to measure the rotation period
+of an exoplanet host.
+Previously to that, \citet{Haywood2014} used the quasi-periodic GP model
+to disentangle the correlated noise produced by stellar activity and rotation
+from the radial velocity signature of an exoplanet.
+In addition, \citet{Littlefair2017} use a quasi-periodic GP to establish the
+quasi-periodicity of variability in the light curves of brown dwarfs.
+What we present here is the first test of the GP method, placed alongside
+alternatives.
+We also provide code allowing others to easily apply this technique
+themselves.}
 
@@
 \citep[see \eg][]{Gibson2012, Haywood2014, Dawson2014, Barclay2015,
-Haywood2015, Evans2015, Rajpaul2015, Czekala2015, Rajpaul2016, Aigrain2016}.
+Haywood2015, Evans2015, Rajpaul2015, Czekala2015, Vanderburg2015, Rajpaul2016,
+Aigrain2016, Littlefair2017}.
 They are useful in regression problems involving any stochastic process,

### 2017-05-09 documents/ms.tex
edits to the abstract

@@
     inference, such as highly degenerate or high-dimensional parameter spaces.
-In this work, we introduce an appropriate {\it effective} model: a Gaussian
+In this work, we test an appropriate {\it effective} model: a Gaussian
     Process with a quasi-periodic covariance kernel function.
-Using this highly flexible model, we are able to sample from the posterior
-    Probability Density Function (PDF) of the periodic parameter,
-    marginalising over the other kernel hyperparameters using a Markov Chain
-    Monte Carlo approach.
+This highly flexible model allows sampling of the posterior Probability
+    Density Function (PDF) of the periodic parameter, marginalising over the
+    other kernel hyperparameters using a Markov Chain Monte Carlo approach.
 To test the effectiveness of this method, we infer rotation periods from
-\naigrain\ simulated stellar light curves, demonstrating that our results
-    are more accurate than both a sine-fitting periodogram and an
-    autocorrelation function method.
+    \naigrain\ simulated stellar light curves, demonstrating that the Gaussian
+    process method produces periods that are slightly more accurate than both
+    a sine-fitting periodogram and an autocorrelation function method.
 We also demonstrate that it works effectively on real data, by inferring

### 2017-05-09 documents/ms.tex
adding description of the ACF method and removing underprediction comments

@@
 We compare our results to those of the Tel Aviv team in \citet{Aigrain2015}
-who use the ACF method described in \citet{Mcquillan2013}.
+who use the ACF method described in \citet{Mcquillan2014} and
+\citet{Aigrain2015}.
+{\bf This method involves the calculation of an autocorrelation function, a
+measure of the self-similarity of the light curve over a range of lags, which
+is then smoothed with a Gaussian kernel.
+Periodic variability in the light curve produces a peak at the period of the
+signal in the ACF and a decaying sequence of peaks at integer multiples of the
+period.
+\citet{McQuillan2014} first pre-processed the light curves by removing very
+long-term variations.
+They then calculated an ACF for each star and measured the positions of up to
+the first four peaks in the ACF.
+They fit a straight line to the periods of these peaks as a function of their
+integer multiple and adopted the gradient of that line as the rotation
+period.}
 Figure \ref{fig:compare_acf} shows ACF-measured versus true rotation
@@
 method results is \gpRMS\ days.
-Another negative feature of the ACF method is revealed by examining the points
-clustered close to the 1:1 line: more fall slightly below than above it, \ie\
-rotation periods are systematically underestimated.
-This stems from peak position measurements in the ACF method.
-ACFs of stellar light curves have similar functional forms to the QP kernel
-function in Equation \ref{eq:QP}: periodic functions added to decaying
-exponentials.
-In such functions the peak positions can be shifted towards the left (towards
-shorter periods), because the decaying exponential raises the left side of
-each peak more than the right.
-It is possible to model this effect; however, standard practice
-simply measures the peak position without taking it into account.
+% Another negative feature of the ACF method is revealed by examining the points
+% clustered close to the 1:1 line: more fall slightly below than above it, \ie\
+% rotation periods are systematically underestimated.
+% This stems from peak position measurements in the ACF method.
+% ACFs of stellar light curves have similar functional forms to the QP kernel
+% function in Equation \ref{eq:QP}: periodic functions added to decaying
+% exponentials.
+% In such functions the peak positions can be shifted towards the left (towards
+% shorter periods), because the decaying exponential raises the left side of
+% each peak more than the right.
+% It is possible to model this effect; however, standard practice
+% simply measures the peak position without taking it into account.
 
@@
     periods.
-    For the points that lie close to the $x=y$ line, there is excess scatter
-    beneath the line due to a tendency of the ACF method to slightly
-    underpredict rotation periods.  }
+    % For the points that lie close to the $x=y$ line, there is excess scatter
+    % beneath the line due to a tendency of the ACF method to slightly
+    % underpredict rotation periods.  }
 \label{fig:compare_acf}
@@
 a larger number of significant outliers.
-We also find that the ACF method often slightly under-predicts the rotation
-period due to a feature of its peak detection algorithm.
+% We also find that the ACF method often slightly under-predicts the rotation
+% period due to a feature of its peak detection algorithm.
 Additionally, we measure the rotation periods of \nkoimcq\

### 2017-05-11 documents/ms.tex
adding discussion of outliers

@@
 integer multiple and adopted the gradient of that line as the rotation
-period.}
+period.
+The uncertainty on the ACF period is the uncertainty on the best-fit slope.
+}
 Figure \ref{fig:compare_acf} shows ACF-measured versus true rotation
@@
     % beneath the line due to a tendency of the ACF method to slightly
-    % underpredict rotation periods.  }
+    % underpredict rotation periods.
+    }
 \label{fig:compare_acf}
@@
 The \citet{Aigrain2015} Tel Aviv team were the only group to report periods
-for all \naigrain\ simulated stars, and are therefore the only team with which we
-compare results.
+for all \naigrain\ simulated stars, and are therefore the only team with which
+we compare results.
 The other participating groups chose to report only the period measurements in
@@
 
+{\bf We first applied a high-pass filter to the light curves, removing
+long-term trends that could produce large peaks at long periods in the
+periodogram.
+We used the 3rd order Butterworth filter in {\tt scipy} python with a cut-off
+of 35 days, attenuating signals with periods greater than this threshold.
+This filter was applied in frequency space, rather than period space, and full
+attenuation was attained at zero frequency.
+In other words, the level of attenuation increased smoothly with period, for
+periods above 35 days but no periods were fully suppressed.
+Experimentally, we found the 35 day cut off removed the largest number of long
+period outliers while preserving shorter period signals.
+We also tried 50, 45, 40 and 30 day cut-offs but the 35 day cut-off value
+maximised the RMS and MAD of the LS periodogram results.}
 For each simulated light curve, we compute a LS periodogram\footnote{LS
@@
 rotation period.
+{\bf The uncertainties on the rotation periods were calculated using the
+following equation for the standard deviation of the frequency
+\citep{Horne1986, Kovacs1981}:
+\begin{equation}
+    \sigma_{\nu} = \frac{3\pi\sigma_N}{2N_O^{1/2}TA},
+\end{equation}
+where $A$ is the amplitude of the signal of highest power, $\sigma_N$ is the
+variance of the time series, with the signal of highest power removed, $N_O$
+is the number of variations and $T$ is the timespan of the data.
+These formal uncertainties are only valid in the case that the noise is white,
+the data are evenly sampled and there is only one signal present.
+Since there are multiple signals present in these light curves, this formal
+uncertainty is an underestimate of the true uncertainty.
+}
 Figure \ref{fig:pgram_compare} shows the resulting recovered rotation periods
 as a function of true period.
-The MRD of the periodogram-recovered periods is \percentpgramMAD\%,
-significantly worse than the ACF and GP methods (see table \ref{tab:MADs} for
+{\bf The MRD of the periodogram-recovered periods is \percentpgramMAD\%,
+slightly worse than the ACF and GP methods (see table \ref{tab:MADs} for
 a side-by-side comparison with the other methods).
-The LS periodogram method drastically overestimates many rotation periods
-because long-term trends unrelated to rotation can cause excess power at long
-periods.
+The filter suppresses signals with periods greater than 35 days but doesn't
+eliminate them altogether, thus allowing high amplitude frequencies to be
+detectable.
+For this reason, some erroneous long periods are recovered by the periodogram
+method.
+These results still show a marked improvement in comparison to periodograms
+calculated from unfiltered data.
+Simply adopting the highest peak of a periodogram calculated from the
+unfiltered simulated data resulted in an RMS of \oldpgramRMS\ and relative MAD
+of \oldpercentpgramMAD.}
+% The LS periodogram method drastically overestimates some rotation periods
+% because long-term trends unrelated to rotation still produced excess power at
+% long periods, despite our attempts to remove these trends via filtering.
+% {\bf Filtering a light curve before attempting to measure a rotation period is
+% not always advisable because it may remove real signals from the data.
+% The simulated light curves used in this experiment had rotation periods of up
+% to 50 days, and thus applying a high pass filter with a 35 day cut off left
+% most of these periods unaffected.
+% Of course, this would not be the case if there were many slower rotators in
+% this sample.
+% In some cases a signal greater than 35 days was recovered, sometimes correctly
+% and sometimes erroneously.
+% }
+% Filtering is not analogous to using priors in Bayesian inference as filtering
+% completely removes signals; if a periodic signal is filtered out there is no
+% hope of recovering it.
+% Using priors that favour shorter periods over longer ones still allows for the
+% detection of long periods if they are strongly supported by the data.}
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
+% \includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
+\includegraphics[width=6in, clip=true]{figures/compare_pgram_filtered_35.pdf}
 \caption[LS periodogram results.]
@@
 host stars, that had rotation periods previously measured by
-the ACF method by \citet{Mcquillan2013}.  We use the pipeline-corrected flux
-(\texttt{pdcsap\_flux} column in the \Kepler\ light curve table), median-normalized
-and unit-subtracted, and mask out all known transiting planet candidate signals.
+the ACF method by \citet{Mcquillan2013}.
+We use the pipeline-corrected flux (\texttt{pdcsap\_flux} column in the
+\Kepler\ light curve table), median-normalized and unit-subtracted, and mask
+out all known transiting planet candidate signals.
 As with the simulated light curves, we randomly subsample each
 light curve by a factor of 30 and split it into segments of about 300 points
-for the purposes of evaluating the likelihood.  We also follow the same MCMC
-fitting procedure as with the simulated data, using the ACF-based prior as before.
+for the purposes of evaluating the likelihood.
+We also follow the same MCMC fitting procedure as with the simulated data,
+using the ACF-based prior as before.
 
@@
 as for the simulated light curves (Table \ref{tab:priors}).
-However, we found that $\ln l$ and $\ln \Gamma$ tended toward slightly different
-values than the simulations.  We also found that the allowed hyperparameter
-range that we used for the simulations was too large for the KOI population,
-as maybe $\sim$15\% of the fits tended toward corners in the hyperparameter
-space, resulting in poor period measurements.  As a result, after this initial
-test, we subsequently adjusted the priors and re-fit all the KOIs.
-The final priors and parameter ranges that we used are in Table \ref{tab:koipriors}.
+However, we found that $\ln l$ and $\ln \Gamma$ tended toward slightly
+different values than the simulations.
+We also found that the allowed hyperparameter range that we used for the
+simulations was too large for the KOI population, as maybe $\sim$15\% of the
+fits tended toward corners in the hyperparameter space, resulting in poor
+period measurements.
+As a result, after this initial test, we subsequently adjusted the priors and
+re-fit all the KOIs.
+The final priors and parameter ranges that we used are in Table
+\ref{tab:koipriors}.
 % The posterior samples of the model parameters for the \Kepler\ objects of
@@
 setting the priors and ranges for the GP hyperparameters.
-Notably, the GP method recovers periods
-systematically slightly larger than \citet{Mcquillan2013}---a likely consequence
-of correcting for the ACF peak measurement bias discussed in Section \ref{sec:acf}.
+% Notably, the GP method recovers periods systematically slightly larger than
+% \citet{Mcquillan2013}---a likely consequence of correcting for the ACF peak
+% measurement bias discussed in Section \ref{sec:acf}.
 
@@
 
-The QP-GP inference method we present in this paper produces
-more accurate rotation periods than the ACF or periodogram methods.
-Additionally, because it explores the posterior PDF with MCMC,
-our method produces probabilistically justified uncertainties.
+{\bf The majority of long period outliers seen in the periodogram comparison
+figure (figure \ref{fig:compare_pgram} results are produced when the rotation
+period signal is highly non-sinusoidal.
+This can be understood intuitively as the LS periodogram fits a series of sine
+waves to the data; if the data are not sinusoidal, a sine wave will not
+produce a good fit.
+It is difficult to assess the reason for the ACF incorrect recoveries because
+we do not have the individual ACFs computed by the \citet{Aigrain2015} Tel
+Aviv team, however it seems that the ACF method is more sensitive than the LS
+periodogram method to noise: this particular implementation of the LS
+periodogram method is capable of detecting periodic signals at lower
+amplitudes that the ACF.
+We do not suggest that the LS periodogram method is better than the ACF method
+in general.
+It was made clear in \citet{Aigrain2015} that the performance of each of these
+rotation period detection algorithms depends sensitively on the exact
+implementations.
+The exact steps performed in the pre-processing of the light curves and the
+post-processing of the periodogram or ACF seems to hugely affect the results.
+This is also true for the GP method.
+Using a different MCMC sampler, different priors, different subsampling, or a
+different kernel would all alter the results.
+Clearly, none of these rotation period measurement methods is perfect and they
+all require a lot of background engineering and tuning.
+This is simply due to inherent difficulty of inferring periods, or any
+parameter that produces a highly multimodal probability distribution, from
+noisy data.}
+
+The QP-GP inference method we test in this paper produces slightly more
+accurate rotation periods than the ACF or periodogram methods.
+Additionally, because it explores the posterior PDF with MCMC, our method
+produces probabilistically justified uncertainties.
 Unfortunately, these uncertainties still appear to be underestimated.

### 2017-05-11 documents/ms.tex
adding celerite

@@
 
-In addition, we are continuing to develop several other aspects of this GP method:
+In addition, we are continuing to develop several other aspects of this GP
+method:
 \begin{itemize}
+    \item{{\bf To use an updated method for calculating the GP likelihood and
+    drastically improve computational efficiency.
+        {\tt celerite}, a new method for rapid computation of the 'solve' and
+        determinant calculations of covariance matrices necessary for GP
+        inference has recently been developed \citep{Foremanmackey2017}.
+        {\tt celerite} scales with the number of data points, $\mathcal{O}(N)$
+        rather than the $N\log(N)^2$ scaling of the {\tt HODLR} algorithm used
+        in this analysis.
+        It requires a slightly different functional form for the kernel as the
+        increased speed is obtained by using mathematical properties of
+        exponential functions.
+        Kernels must be written as the sums of exponentials in order to be
+        implemented in {\tt celerite}, however a close approximation to the
+        kernel function used here is demonstrated to recover rotation periods
+        in \citet{Foremanmackey2017}.}}
 \item{To investigate more physically motivated priors.

### 2017-05-11 documents/ms.tex
adding more words about celerite to the conclusion

@@
 alternative methods.
-We therefore argue that the GP method is the best method for rotation period
-inference currently available.
+{\bf Although GP inference is slow as it requires a `solve' and a matrix
+determinant calculation at every step of an MCMC, a new GP solver, {\tt
+celerite} \citep{Foremanmackey2017} shows promise for increasing the
+computional speed of GP rotation period inference by several orders of
+magnitude.
+We intend to integrate {\tt celerite} into our code in the near future.}
 The code used in this project is available at

### 2017-05-11 documents/ms.tex
adding discussion of computation time and hierarchical inference

@@
 magnitude.
-We intend to integrate {\tt celerite} into our code in the near future.}
+We intend to integrate {\tt celerite} into our code in the near future.
+The issue of computational time is a personal one.
+GP period inference is trivially parallelisable, so if one has access to
+thousands of computer cores, one can infer rotation periods of thousands of
+stars in a few hours (perhaps just a few minutes with {\tt celerite}).
+For single stars GP rotation inference takes a reasonable amount of time and
+has already been used on exoplanet hosts \citep[\eg][]{Haywood2014,
+Vanderburg2015}.
+
+Is it worth spending the extra computation power to get slightly more
+accurate, probabilistic rotation periods for a large number of stars?
+A growing number of astronomers are interested in hierarchical probabilistic
+modelling, particularly in the stellar and exoplanet communities.
+These individuals are mindful of carefully inferring probabilistic parameters
+wherever possible.
+Probabilistic inference also allows you to marginalise over, or ‘bypass’,
+certain parameters.
+For example, imagine trying to infer the distribution of rotation periods of a
+sample of stars.
+One could simply calculate rotation periods from the light curves and produce
+a histogram of the rotation values.
+This is `adding posteriors' which is not technically legal in probability
+theory, you are only allowed to add (or sum) likelihoods.
+In order to add likelihoods you have to jointly infer the rotation period of
+each individual star and the parameters of the distribution, which you might
+model as a Gaussian.
+Here, the individual rotation periods of the stars are nuisance parameters:
+you don’t care about their values, you only compute them in order to infer the
+mean and variance of the period distribution.
+Using hierarchical inference allows you to marginalise over individual
+rotation periods and infer distribution parameters directly from the light
+curves.
+Many people are interested in the distribution of stellar ages across the
+galaxy and rotation periods can be used as an age proxy via gyrochronology.
+Probabilistic rotation period inference is therefore highly relevant to a
+large number of galactic archaeology and exoplanet population studies which
+require rotation period (or age) distributions.}
+
 The code used in this project is available at

### 2017-06-08 documents/ms.tex
finishing off - still need to read over once more

--- 2017-05-11 documents/ms.tex
@@
 
+\newcommand{\nkoi}{1132}
+
 \newcommand{\naigrain}{333}
@@
 This highly flexible model allows sampling of the posterior Probability
-    Density Function (PDF) of the periodic parameter, marginalising over the
+    Density Function of the periodic parameter, marginalising over the
     other kernel hyperparameters using a Markov Chain Monte Carlo approach.
@@
     periods.
-Because this method delivers posterior PDFs, it will enable hierarchical
-    studies involving stellar rotation, particularly those involving
-    population modelling, such as inferring stellar ages, obliquities in
-    exoplanet systems, or characterising star-planet interactions.
+Because this method delivers posterior probability density functions, it will
+    enable hierarchical studies involving stellar rotation, particularly those
+    involving population modelling, such as inferring stellar ages,
+    obliquities in exoplanet systems, or characterising star-planet
+    interactions.
+    {\bf We provide a table of rotation periods for \nkoi\ \Kepler\ objects of
+    interest and full posterior probability density function samples.}
     The code used to implement this new method is available online.
@@
 quasi-periodicity of variability in the light curves of brown dwarfs.
-What we present here is the first test of the GP method, placed alongside
-alternatives.
+What we present here is a test of the GP method and a comparison with
+alternative rotation period measurement methods.
 We also provide code allowing others to easily apply this technique
@@
 is then smoothed with a Gaussian kernel.
-Periodic variability in the light curve produces a peak at the period of the
-signal in the ACF and a decaying sequence of peaks at integer multiples of the
-period.
-\citet{McQuillan2014} first pre-processed the light curves by removing very
+Periodic variability in the light curve produces a peak in the ACF at the
+period of the signal and a decaying sequence of peaks at integer multiples of
+the period.
+\citet{Mcquillan2014} first pre-processed the light curves by removing very
 long-term variations.
@@
 periodogram.
-We used the 3rd order Butterworth filter in {\tt scipy} python with a cut-off
+We used the 3rd order Butterworth filter in {\tt scipy} with a cut-off
 of 35 days, attenuating signals with periods greater than this threshold.
@@
 maximised the RMS and MAD of the LS periodogram results.}
-For each simulated light curve, we compute a LS periodogram\footnote{LS
+For each simulated light curve, we computed a LS periodogram\footnote{LS
 periodograms were calculated using the gatspy Python module:
@@
 grid of 10,000 periods, evenly spaced in frequency, between 1 and 100 days.
-We adopt the period of the highest peak in the periodogram as the measured
+We adopted the period of the highest peak in the periodogram as the measured
 rotation period.
@@
 \begin{equation}
-    \sigma_{\nu} = \frac{3\pi\sigma_N}{2N_O^{1/2}TA},
+    \sigma_{\nu} = \frac{3\pi\sigma_N}{2N^{1/2}TA},
 \end{equation}
 where $A$ is the amplitude of the signal of highest power, $\sigma_N$ is the
-variance of the time series, with the signal of highest power removed, $N_O$
-is the number of variations and $T$ is the timespan of the data.
+variance of the time series, with the signal of highest power removed, $N$
+is the number of observations and $T$ is the timespan of the data.
 These formal uncertainties are only valid in the case that the noise is white,
@@
 calculated from unfiltered data.
-Simply adopting the highest peak of a periodogram calculated from the
-unfiltered simulated data resulted in an RMS of \oldpgramRMS\ and relative MAD
-of \oldpercentpgramMAD.}
+Simply adopting the highest peak of a periodogram calculated from the {\it
+unfiltered} simulated data resulted in an RMS of \oldpgramRMS\ and relative
+MAD of \oldpercentpgramMAD.}
 % The LS periodogram method drastically overestimates some rotation periods
-% because long-term trends unrelated to rotation still produced excess power at
-% long periods, despite our attempts to remove these trends via filtering.
+% because long-term trends unrelated to rotation still produced excess power
+% at long periods, despite our attempts to remove these trends via filtering.
 % {\bf Filtering a light curve before attempting to measure a rotation period is
@@
 % measurement bias discussed in Section \ref{sec:acf}.
+{\bf The new GP rotation periods calculated for \nkoi\ KOIs are provided in an
+online table accompanying this paper.
+The first few lines of this table is shown below
+(table \ref{tab:koi_periods}).
+Samples of the rotation period posterior PDFs are available online:
+\url{https://doi.org/10.5281/zenodo.804440}}
+
+\begin{table*}
+\begin{center}
+    \caption{Rotation periods and \Kepler\ input catalogue (KIC) properties
+    for the \Kepler\ objects of interest. The full table is available online.}
+\begin{tabular}{lcccc}
+    KOI number & Period (days) & KIC T$_{\mathrm{eff}}$ & KIC $\log(g)$ & KIC
+    $[$Fe$/$H$]$ \\
+    \hline
+    1 & 25$\pm1$ & $5126^{+160}_{-138}$ & $4.61^{+0.03}_{-0.1}$ &
+    $-0.1\pm0.3$ \\
+    2 & 20$\pm1$ & $5850\pm50$ & $4.46\pm0.03$ & $-0.2\pm0.1$ \\
+    3 & 28$\pm1$ & $6350\pm80$ & $4.02\pm0.01$ & $0.26\pm0.08$\\
+    6 & 23$\pm1$ & $6225^{+114}_{-158}$ & $4.17^{+0.06}_{-0.05}$ &
+    $-0.0^{+0.1}_{-0.2}$\\
+    7 & 25$\pm1$ & $5543\pm79 $ & $4.08\pm0.01$ & $0.4\pm0.1$ \\
+    ... & ... & ... & ... & ... \\
+\end{tabular}
+\end{center}
+\end{table*}
+\label{tab:koi_periods}
 
@@
 {\bf The majority of long period outliers seen in the periodogram comparison
-figure (figure \ref{fig:compare_pgram} results are produced when the rotation
-period signal is highly non-sinusoidal.
+figure (figure \ref{fig:pgram_compare}) are produced when the rotation period
+signal is highly non-sinusoidal.
 This can be understood intuitively as the LS periodogram fits a series of sine
@@
 Aviv team, however it seems that the ACF method is more sensitive than the LS
-periodogram method to noise: this particular implementation of the LS
-periodogram method is capable of detecting periodic signals at lower
-amplitudes that the ACF.
+periodogram method to noise.
+This particular implementation of the LS periodogram method is capable of
+detecting periodic signals at lower amplitudes that the ACF.
 We do not suggest that the LS periodogram method is better than the ACF method
@@
 It was made clear in \citet{Aigrain2015} that the performance of each of these
-rotation period detection algorithms depends sensitively on the exact
-implementations.
+rotation period detection algorithms depends sensitively on the
+implementation technique.
 The exact steps performed in the pre-processing of the light curves and the
@@
     drastically improve computational efficiency.
-        {\tt celerite}, a new method for rapid computation of the 'solve' and
+        {\tt celerite}, a new method for rapid computation of the `solve' and
         determinant calculations of covariance matrices necessary for GP
@@
 magnitude.
-We intend to integrate {\tt celerite} into our code in the near future.}
+We intend to integrate {\tt celerite} into our code in the near future.
+How much computational time is too much computational time is a question that
+depends on the available resources of each individual and their scientific
+priorities.
+GP period inference is trivially parallelisable, so if one has access to
+thousands of computer cores, one can infer rotation periods of thousands of
+stars in a few hours (perhaps just a few minutes with {\tt celerite}).
+For single stars GP rotation inference takes a reasonable amount of time and
+has already been used on exoplanet hosts \citep[\eg][]{Haywood2014,
+Vanderburg2015}.
+
+Is it worth spending the extra computation power to get slightly more
+accurate, probabilistic rotation periods for a large number of stars?
+A growing number of astronomers are interested in hierarchical probabilistic
+modelling, particularly in the stellar and exoplanet communities.
+These individuals are mindful of carefully inferring probabilistic parameters
+wherever possible.
+Probabilistic inference allows one to marginalise over nuisance
+parameters while formally taking any uncertainty into account when
+measuring the quantity of interest -- the star's rotation period, in
+this case.
+This becomes especially important as the signal-to-noise ratio
+decreases and the rotation period is no longer tightly constrained,
+and as the sampling becomes uneven.
+One of our goals is to perform population inferences with a catalog of
+rotation periods.
+Throughout the astronomical literature, hierarchical Bayesian
+inference has been demonstrated to be a robust method of population
+inference \citep[\eg][]{Hogg2010, Foremanmackey2015, Wolfgang2015, Rogers2015}.
+Probabilistic measurements, with full rigorous propagation of
+uncertainties, are a crucial ingredient for any hierarchical inference
+where the target is the population level distribution.
+Inferring the ages of stars in the galaxy via gyrochronology (age-rotation
+relations) \citep[\eg][]{Skumanich1972, Kawaler1989, Barnes2003, Barnes2007}
+is an example of a hierarchical inference problem using stellar rotation.
+Hierarchical inference allows you to marginalise over individual rotation
+periods and infer age distribution parameters directly from the light curves.
+Probabilistic rotation period inference is highly relevant to a large number
+of galactic archaeology and exoplanet population studies which require
+rotation period (or age) distributions.}
+
 The code used in this project is available at
-\url{https://github.com/RuthAngus/GProtation/}.
+\url{https://github.com/RuthAngus/GProtation/} and samples of the KOI rotation
+period posterior PDFs can be found at
+\url{https://doi.org/10.5281/zenodo.804440}.
 

### 2017-06-09 documents/ms.tex
sent to coauthors

@@
     Process with a quasi-periodic covariance kernel function.
-This highly flexible model allows sampling of the posterior Probability
-    Density Function of the periodic parameter, marginalising over the
+This highly flexible model allows sampling of the posterior probability
+    density function of the periodic parameter, marginalising over the
     other kernel hyperparameters using a Markov Chain Monte Carlo approach.
@@
     {\bf We provide a table of rotation periods for \nkoi\ \Kepler\ objects of
-    interest and full posterior probability density function samples.}
+    interest and their posterior probability density function samples.}
     The code used to implement this new method is available online.
@@
 in \textsection \ref{sec:perf_and_comp}.
-In \textsection \ref{sec:kepler} we apply our method to real \kepler\ data,
+In \textsection \ref{sec:kepler} we apply the GP method to real \kepler\ data,
 and the results are discussed in \textsection \ref{sec:discussion}.
@@
 Here, the parameters describing the correlated noise are what we are
-interested in and our mean model is simply a straight line at $y=0$.}, as in the
+interested in and the mean model is simply a straight line at $y=0$.}, as in the
 case of \Kepler\ light curves, those off-diagonal elements are non-zero.
@@
 accurate rotation periods than the ACF or periodogram methods.
-Additionally, because it explores the posterior PDF with MCMC, our method
+Additionally, because it explores the posterior PDF with MCMC, the GP method
 produces probabilistically justified uncertainties.
@@
         In this paper we have only explored the physical interpretation of one
-        parameter in our kernel function (P), which we related to stellar
+        parameter in the kernel function ($P$), which we related to stellar
         rotation periods, but others also warrant physically motivated priors.
@@
 
-The main aim of this work was to develop a probabilistic rotation period
-inference method.
-Probabilistic periods are necessary for hierarchical Bayesian inference,
-particularly when performing population analysis.
-This new GP method is capable of generating a catalogue of probabilistic
-rotation periods which could, for example, reveal the period distribution
-(and therefore potentially the age distribution) of stars in the Milky Way.
-There is still room for improvement since the uncertainty estimates are not
-yet truly representative and it is still only an `effective' model, not a
-physical one.
-However, it is probabilistic and provides more accurate periods than
-alternative methods.
+% The main aim of this work was to develop a probabilistic rotation period
+% inference method.
+% Probabilistic periods are necessary for hierarchical Bayesian inference,
+% particularly when performing population analysis.
+% This new GP method is capable of generating a catalogue of probabilistic
+% rotation periods which could, for example, reveal the period distribution
+% (and therefore potentially the age distribution) of stars in the Milky Way.
+% There is still room for improvement since the uncertainty estimates are not
+% yet truly representative and it is still only an `effective' model, not a
+% physical one.
+% However, it is probabilistic and provides more accurate periods than
+% alternative methods.
 {\bf Although GP inference is slow as it requires a `solve' and a matrix

### 2017-06-09 code/make_acf_comparison_plot.py

@@ -43,11 +43,16 @@ def acf_plot(truths_e,
     plt.plot(np.log(xs), np.log(xs) + 2./3, "k--", alpha=.3, lw=.8, zorder=0)
 
     plt.errorbar(np.log(true), np.log(acfs), yerr=(acf_errs/acfs), fmt="k.",
-                 capsize=0, ecolor=".7", alpha=.4, ms=1, zorder=1,
+                 capsize=0, ecolor=".2", alpha=.4, ms=1, zorder=1,
                  elinewidth=.8)
     plt.scatter(np.log(true), np.log(acfs), c=np.log(amp), edgecolor=".5",
                 cmap="GnBu_r", vmin=min(np.log(amp)), vmax=max(np.log(amp)),
                 s=10, lw=.2, zorder=2)
+
+    # badinds, varinds = get_bad_inds(N)
+    # plt.plot(np.log(true)[badinds], np.log(acfs)[badinds], "ro")
+    # plt.plot(np.log(true)[varinds], np.log(acfs)[varinds], "bo")
+
     cbar = plt.colorbar()
     cbar.ax.set_ylabel("$\ln\mathrm{(Amplitude)}$")
     plt.xlim(0, 4)
@@ -63,8 +68,33 @@ def acf_plot(truths_e,
     plt.xlabel("$\ln(\mathrm{ACF~Period}) - \ln(\mathrm{True~Period})$")
     plt.savefig("acf_hist.pdf")
 
+    dff = np.log(acfs) - np.log(true)
+    plt.clf()
+    m = dff > 1
+    plt.plot(np.log(true), dff, "k.")
+    print(N[m])
+    plt.savefig("diff_acf")
+
     return MAD(np.log(true), np.log(acfs)), MAD(true, acfs), \
         MAD_rel(true, acfs), RMS(true, acfs)
+
+
+def get_bad_inds(N):
+    bad = np.array([121, 164, 180, 188, 189, 227, 239, 246, 253, 269, 273,
+                    328, 336, 338, 361, 377, 397, 470, 479, 556, 573, 608,
+                    615, 671])
+    variable = np.array([345, 350, 356, 370, 406, 477, 450, 663, 682, 692])
+    badinds, varinds = [], []
+
+    for i, n in enumerate(N):
+        mb = n == bad
+        mv = n == variable
+        if len(bad[mb]):
+            badinds.append(i)
+        if len(variable[mv]):
+            varinds.append(i)
[…]

### 2017-06-09 code/make_gp_comparison_plot.py

@@ -4,6 +4,8 @@ import os
 import numpy as np
 import pandas as pd
 import matplotlib.pyplot as plt
+from make_acf_comparison_plot import get_bad_inds
+
 
 plotpar = {'axes.labelsize': 18,
            'font.size': 10,
@@ -36,6 +38,7 @@ def plot(prior, nbins):
     # remove differential rotators and take just the first 100
     m = truths.DELTA_OMEGA.values == 0
     truths = truths.iloc[m]
+    N = truths.N.values
 
     recovered = np.zeros(len(truths.N.values))
     errp, errm = [np.zeros(len(truths.N.values)) for i in range(2)]
@@ -67,10 +70,15 @@ def plot(prior, nbins):
 
     plt.errorbar(np.log(x[l]), np.log(recovered[l]),
                  yerr=[lnerrp[l], lnerrm[l]], fmt="k.", zorder=1, capsize=0,
-                 ecolor=".7", alpha=.5, ms=.1, elinewidth=.8)
+                 ecolor=".2", alpha=.4, ms=.1, elinewidth=.8)
     plt.scatter(np.log(x[l]), np.log(recovered[l]), c=np.log(amp[l]),
                 edgecolor=".5", cmap="GnBu_r", vmin=min(np.log(amp[l])),
                 vmax=max(np.log(amp[l])), s=10, zorder=2, lw=.2)
+
+    # badinds, varinds = get_bad_inds(N[l])
+    # plt.plot(np.log(x[l])[badinds], np.log(recovered[l])[badinds], "ro")
+    # plt.plot(np.log(x[l])[varinds], np.log(recovered[l])[varinds], "bo")
+
     plt.xlim(0, 4)
     plt.ylim(-2, 6)
 
@@ -147,6 +155,14 @@ def plot(prior, nbins):
     66% are within 3 sigma.
     Largest outlier is 114 sigma off.
     """
+
+    dff = np.log(recovered[l]) - np.log(x[l])
+    plt.clf()
+    m = dff > 1
+    m = dff < -.5
+    plt.plot(np.log(x[l]), dff, "k.")
+    print(N[m])
+    plt.savefig("diff_gp")
 
     print((np.median(np.abs(recovered[l] - x[l]))))
 

### 2017-06-09 code/make_pgram_comparison_plot.py

@@ -7,6 +7,8 @@ import matplotlib.pyplot as plt
 from multiprocessing import Pool
 import h5py
 
+from calc_mad import MAD, MAD_rel, RMS
+
 plotpar = {'axes.labelsize': 18,
            'font.size': 10,
            'legend.fontsize': 15,
@@ -157,7 +159,7 @@ def mcmc_plots(truths, DIR):
 
     cbar = plt.colorbar()
     cbar.ax.set_ylabel("$\ln\mathrm{(Amplitude)}$")
-    plt.savefig(os.path.join(DIR, "compare_mcmc.pdf"))
+    # plt.savefig(os.path.join(DIR, "compare_mcmc.pdf"))
 
     # make convergence plot.
     plt.clf()
@@ -175,7 +177,7 @@ def mcmc_plots(truths, DIR):
                 s=20, lw=.2, zorder=2)
     cbar = plt.colorbar()
     cbar.ax.set_ylabel("$<\mathrm{(Autocorrelation~Time)}>$")
-    plt.savefig(os.path.join(DIR, "compare_mcmc_convergence"))
+    # plt.savefig(os.path.join(DIR, "compare_mcmc_convergence"))
 
 
     # mcmc plot with samples
@@ -193,7 +195,7 @@ def mcmc_plots(truths, DIR):
         if samples != None:
             plt.plot(np.log(np.ones(100) * true[i]), np.random.choice(samples,
                      100), "k.", ms=1)
-    plt.savefig(os.path.join(DIR, "compare_mcmc_samples.pdf"))
+    # plt.savefig(os.path.join(DIR, "compare_mcmc_samples.pdf"))
     return (np.median((maxlike - true)**2))**.5, \
             (np.median((np.exp(lnp) - true)**2))**.5
 
@@ -220,8 +222,8 @@ def acf_plot(truths, DIR):
     plt.plot(np.log(xs), np.log(xs) - 2./3, "k--", alpha=.3, zorder=0)
     plt.plot(np.log(xs), np.log(xs) + 2./3, "k--", alpha=.3, zorder=0)
 
-#     plt.errorbar(np.log(true), np.log(acfs), yerr=(acf_errs/acfs), fmt="k.",
-#                  capsize=0, ecolor=".7", alpha=.4, ms=1, zorder=1)
+    # plt.errorbar(np.log(true), np.log(acfs), yerr=(acf_errs/acfs), fmt="k.",
[…]

### 2017-06-09 code/plot_fits.py

@@ -52,13 +52,18 @@ def make_plots(id, RESULTS_DIR="/Users/ruthangus/projects/GProtation/code/" \
     plt.plot(x, y, "k.")
     plt.plot(xs, mu)
     plt.xlim(0, 100)
-    v = np.std(y)
-    plt.ylim(-10*v, 10*v)
+    # v = np.std(y)
+    # plt.ylim(-10*v, 10*v)
     plt.savefig("{}_fit".format(id))
 
-    """ plot corner """
-    fig = corner.corner(df)
-    fig.savefig("{}_corner".format(id))
+    # """ plot corner """
+    # fig = corner.corner(df)
+    # fig.savefig("{}_corner".format(id))
 
 if __name__ == "__main__":
     make_plots(337)
+    make_plots(361)
+    make_plots(892)
+    make_plots(977)
+    make_plots(993)
+    make_plots(996)

### 2017-06-09 gprotation/filtering.py

@@ -11,20 +11,20 @@ def butter_bandpass(lowcut, fs, order=5):
     return b, a
 
 
-def butter_bandpass_filter(data, lowcut, fs, order=5):
+def butter_bandpass_filter(data, lowcut, fs, order=5, plot=False):
     b, a = butter_bandpass(lowcut, fs, order=order)
     y = lfilter(b, a, data)
-    plot_bandpass(lowcut, fs, order)
+    if plot:
+        plot_bandpass(lowcut, fs, order)
     return y
 
 
 def plot_bandpass(lowcut, fs, order):
     # Plot the frequency response for a few different orders.
     plt.clf()
-    for order in [3, 6, 9]:
-        b, a = butter_bandpass(lowcut, fs, order=order)
-        w, h = freqz(b, a, worN=2000)
-        plt.plot((fs * 0.5 / np.pi) * w, abs(h), label="order = %d" % order)
+    b, a = butter_bandpass(lowcut, fs, order=3)
+    w, h = freqz(b, a, worN=5000)
+    plt.plot((fs * 0.5 / np.pi) * w, abs(h), label="order = %d" % order)
 
     plt.plot([0, 0.5 * fs], [np.sqrt(0.5), np.sqrt(0.5)],
              '--', label='sqrt(0.5)')
@@ -32,6 +32,7 @@ def plot_bandpass(lowcut, fs, order):
     plt.ylabel('Gain')
     plt.grid(True)
     plt.legend(loc='best')
+    plt.xlim(0, 1)
     plt.savefig("butter_bandpass")
 
 

### 2017-06-09 gprotation/pgram_filtered.py

@@ -153,6 +153,7 @@ def calc_p_init(x, y, yerr, id, RESULTS_DIR="pgram_filtered_results_35",
                            "pgram_period": [pgram_period],
                            "pgram_period_err": [pgram_period_err]})
         df.to_csv(fname)
+
     return acf_period, err, pgram_period, pgram_period_err
 
 
@@ -173,4 +174,4 @@ if __name__ == "__main__":
     for i, id in enumerate(truths.N.values):
         x, y = load_suzanne_lcs(id, os.path.join(DIR, "final"))
         yerr = np.ones(len(y)) * 1e-5
-        calc_p_init(x, y, yerr, id, clobber=True)
+        _, _, pgram, _ = calc_p_init(x, y, yerr, id, clobber=True)

### 2017-06-16 documents/ms.tex
adding response macro

@@
 
+\newcommand{\response}[1]{{\bf #1}}
 \newcommand{\Kepler}{{\it Kepler}}
@@
     \naigrain\ simulated stellar light curves, demonstrating that the Gaussian
-    process method produces periods that are slightly more accurate than both
+    process method produces periods that are more accurate than both
     a sine-fitting periodogram and an autocorrelation function method.
-We also demonstrate that it works effectively on real data, by inferring
+We also demonstrate that it works well on real data, by inferring
     rotation periods for \nkoimcq\ \Kepler\ stars with previously measured
     periods.
+    \response{We provide a table of rotation periods for these \nkoi\ \Kepler\ objects of
+    interest and their posterior probability density function samples.}
 Because this method delivers posterior probability density functions, it will
@@
     interactions.
-    {\bf We provide a table of rotation periods for \nkoi\ \Kepler\ objects of
-    interest and their posterior probability density function samples.}
-    The code used to implement this new method is available online.
+    The code used to implement this method is available online.
 % The ACF method is not applicable to unevenly spaced data,
@@
 A physically realistic model of the stellar surface
-would, ideally,  perfectly capture the complexity of shapes
+would perfectly capture the complexity of shapes
 within stellar light curves as well as the quasi-periodic nature, allowing for
@@
 A drawback of the ACF method however, is that it requires data to be
-evenly-spaced\footnote{\citet{Edelson1988} describe a method for computing
-ACFs for unevenly-spaced data.}.
+evenly-spaced.
 Most ground-based measurements, such as the future Large Synoptic Survey
@@
 producing a justified estimate of its uncertainty.
-{\bf Using a quasi-periodic GP model to infer a rotation period is not a new idea.
-We used this method in \citet{Vanderburg2015} to measure the rotation period
-of an exoplanet host.
+\response{Using a quasi-periodic GP model to infer a rotation period is not a new
+idea.
+For example, we used this method in \citet{Vanderburg2015} to measure the
+rotation period of an exoplanet host.
 Previously to that, \citet{Haywood2014} used the quasi-periodic GP model
@@
 kernel function.
-As a qualitative demonstration, we present the time-series in Figure
+As a qualitative demonstration, we present the time-series in figure
 \ref{fig:GP_example}: the \kepler\ light curve of KIC \kepexample.
@@
 We use this QP kernel function (Equation \ref{eq:QP}) to produce
-the GP model that fits the \Kepler\ light curve in Figure
+the GP model that fits the \Kepler\ light curve in figure
 \ref{fig:GP_example}.
@@
 
+\response{Performing inference with GPs is computationally expensive.
+Is it worth spending the extra computation power to get slightly more
+accurate, probabilistic rotation periods for a large number of stars?
+A growing number of astronomers are interested in hierarchical probabilistic
+modelling, particularly in the stellar and exoplanet communities.
+These individuals are mindful of carefully inferring probabilistic parameters
+wherever possible.
+Probabilistic inference allows one to marginalise over nuisance
+parameters while formally taking any uncertainty into account when
+measuring the quantity of interest -- the star's rotation period, in
+this case.
+This becomes especially important as the signal-to-noise ratio
+decreases and the rotation period is no longer tightly constrained,
+and as the sampling becomes uneven.
+% One of our goals is to perform population inferences with a catalog of
+% rotation periods.
+Throughout the astronomical literature, hierarchical Bayesian
+inference has been demonstrated to be a robust method of population
+inference \citep[\eg][]{Hogg2010, Foremanmackey2014, Wolfgang2015, Rogers2015}.
+Probabilistic measurements with full, rigorous propagation of
+uncertainties are a crucial ingredient for any hierarchical inference
+where the target is the population level distribution.
+Inferring the ages of stars in the galaxy via gyrochronology (age-rotation
+relations) \citep[\eg][]{Skumanich1972, Kawaler1989, Barnes2003, Barnes2007}
+is an example of a hierarchical inference problem using stellar rotation.
+Hierarchical inference allows you to marginalise over individual rotation
+periods and infer age distribution parameters directly from the light curves.
+Probabilistic rotation period inference is highly relevant to a large number
+of galactic archaeology and exoplanet population studies which require
+rotation period (or age) distributions.}
+
 This paper is laid out as follows.
@@
 \label{eq:posterior}
-p({\bm \theta}\,|\,y) \propto \mathcal L(y\,|\,{\bm \theta}) p({\bm \theta}),
+p({\bm \theta}\,|\,y) \propto \mathcal L({\bm \theta}) p({\bm \theta}),
 \end{equation}
@@
 \begin{equation}
-\ln \mathcal{L} = -\frac{1}{2}\bf{r}^T\bf{C}^{-1}\bf{r}-\ln|\bf{C}|
+    \ln \mathcal{L} = -\frac{1}{2}\bf{r}^T\bf{C}^{-1}\bf{r}-\frac{1}{2}\ln|\bf{C}|
     + \frac{N}{2}\log2\pi,
@@
 The matrix operations necessary to evaluate the GP likelihood scale as
-$N\ln(N)^2$, where $N$ is the number of data points in the light curve.
-These operations use the fast matrix solver HODLR \citep{Ambikasaran2014},
+$N\ln(N)^2$, where $N$ is the number of data points in the light curve, using
+the fast matrix solver HODLR \citep{Ambikasaran2014},
 implemented in the {\tt george} \citep{George} python package.
@@
 producing a large number of accurate rotation period measurements
-(see, e.g., their Figure 8).
+(see, e.g., their figure 8).
 Another advantage is its fast implementation speed.
@@
 \citet{Aigrain2015}.
-{\bf This method involves the calculation of an autocorrelation function, a
+\response{This method involves the calculation of an autocorrelation function, a
 measure of the self-similarity of the light curve over a range of lags, which
@@
 
-{\bf We first applied a high-pass filter to the light curves, removing
+\response{We first applied a high-pass filter to the light curves, removing
 long-term trends that could produce large peaks at long periods in the
@@
 We also tried 50, 45, 40 and 30 day cut-offs but the 35 day cut-off value
-maximised the RMS and MAD of the LS periodogram results.}
+minimised the RMS and MAD of the LS periodogram results.}
 For each simulated light curve, we computed a LS periodogram\footnote{LS
@@
 rotation period.
-{\bf The uncertainties on the rotation periods were calculated using the
+\response{The uncertainties on the rotation periods were calculated using the
 following equation for the standard deviation of the frequency
@@
 as a function of true period.
-{\bf The MRD of the periodogram-recovered periods is \percentpgramMAD\%,
+\response{The MRD of the periodogram-recovered periods is \percentpgramMAD\%,
 slightly worse than the ACF and GP methods (see table \ref{tab:MADs} for
@@
 % measurement bias discussed in Section \ref{sec:acf}.
-{\bf The new GP rotation periods calculated for \nkoi\ KOIs are provided in an
+\response{The new GP rotation periods calculated for \nkoi\ KOIs are provided in an
 online table accompanying this paper.
@@
 
-{\bf The majority of long period outliers seen in the periodogram comparison
+\response{The majority of long period outliers seen in the periodogram comparison
 figure (figure \ref{fig:pgram_compare}) are produced when the rotation period
@@
 The exact steps performed in the pre-processing of the light curves and the
-post-processing of the periodogram or ACF seems to hugely affect the results.
+post-processing of the periodogram or ACF seem to affect the results strongly.
 This is also true for the GP method.
@@
 \begin{itemize}
-    \item{{\bf To use an updated method for calculating the GP likelihood and
+    \item{\response{To use an updated method for calculating the GP likelihood and
     drastically improve computational efficiency.
@@
 \kepler-like light curves for solid-body rotators \citep{Aigrain2015} using
-three different methods: a new Gaussian process method,
+three different methods: a Gaussian process method,
 an autocorrelation function method, and a Lomb-Scargle periodogram method.
 We demonstrate that the GP method produces the most accurate rotation periods
-of the three techniques, providing a large improvement over the LS periodogram
-method and a slight one over the ACF method.
+of the three techniques.
 Comparing our results with the results of the \citet{Aigrain2015} Tel Aviv
@@
 method works well on real \kepler\ data.
+Samples of the full set of \nkoi\ KOI rotation period posterior PDFs can be
+found at \url{https://doi.org/10.5281/zenodo.804440}.
 
@@
 model.
-Although the GP model used here is clearly a good one, as it has been
-demonstrated to reliably recover rotation periods, it is still only an
-effective model, not an accurate physical model.
+A quasi-periodic Gaussian process is clearly a good model, as demonstrated
+here, however it is still only an effective model, not an accurate physical
+one.
 It captures the posterior PDF of the periodic component of the covariance
@@
 % alternative methods.
-{\bf Although GP inference is slow as it requires a `solve' and a matrix
+\response{Although GP inference is slow as it requires a `solve' and a matrix
 determinant calculation at every step of an MCMC, a new GP solver, {\tt
@@
 We intend to integrate {\tt celerite} into our code in the near future.
-How much computational time is too much computational time is a question that
-depends on the available resources of each individual and their scientific
-priorities.
+How much computational time is too much is a question that depends on the
+available resources of each individual and their scientific priorities.
 GP period inference is trivially parallelisable, so if one has access to
@@
 has already been used on exoplanet hosts \citep[\eg][]{Haywood2014,
-Vanderburg2015}.
-
-Is it worth spending the extra computation power to get slightly more
-accurate, probabilistic rotation periods for a large number of stars?
-A growing number of astronomers are interested in hierarchical probabilistic
-modelling, particularly in the stellar and exoplanet communities.
-These individuals are mindful of carefully inferring probabilistic parameters
-wherever possible.
-Probabilistic inference allows one to marginalise over nuisance
-parameters while formally taking any uncertainty into account when
-measuring the quantity of interest -- the star's rotation period, in
-this case.
-This becomes especially important as the signal-to-noise ratio
-decreases and the rotation period is no longer tightly constrained,
-and as the sampling becomes uneven.
-One of our goals is to perform population inferences with a catalog of
-rotation periods.
-Throughout the astronomical literature, hierarchical Bayesian
-inference has been demonstrated to be a robust method of population
-inference \citep[\eg][]{Hogg2010, Foremanmackey2015, Wolfgang2015, Rogers2015}.
-Probabilistic measurements, with full rigorous propagation of
-uncertainties, are a crucial ingredient for any hierarchical inference
-where the target is the population level distribution.
-Inferring the ages of stars in the galaxy via gyrochronology (age-rotation
-relations) \citep[\eg][]{Skumanich1972, Kawaler1989, Barnes2003, Barnes2007}
-is an example of a hierarchical inference problem using stellar rotation.
-Hierarchical inference allows you to marginalise over individual rotation
-periods and infer age distribution parameters directly from the light curves.
-Probabilistic rotation period inference is highly relevant to a large number
-of galactic archaeology and exoplanet population studies which require
-rotation period (or age) distributions.}
+Vanderburg2015}.}
 
 The code used in this project is available at
-\url{https://github.com/RuthAngus/GProtation/} and samples of the KOI rotation
-period posterior PDFs can be found at
-\url{https://doi.org/10.5281/zenodo.804440}.
+\url{https://github.com/RuthAngus/GProtation/}.
 
@@
 for financial support.
+We would like to thank the anonymous referee who's comments significantly
+improved this paper.
 Some of the data presented in this paper were obtained from the Mikulski

### 2017-06-21 documents/ms.tex
removing data file

@@
 
-\newcommand{\response}[1]{{\bf #1}}
+\newcommand{\response}[1]{{#1}}
 \newcommand{\Kepler}{{\it Kepler}}
@@
     Suzanne Aigrain$^3$,
-    Daniel Foreman-Mackey$^4$,
+    Daniel Foreman-Mackey$^4$
     \newauthor
@@
     periods.
-    \response{We provide a table of rotation periods for these \nkoi\ \Kepler\ objects of
-    interest and their posterior probability density function samples.}
+    % \response{We provide a table of rotation periods for these \nkoi\ \Kepler\ objects of
+    % interest and their posterior probability density function samples.}
 Because this method delivers posterior probability density functions, it will
@@
 % measurement bias discussed in Section \ref{sec:acf}.
-\response{The new GP rotation periods calculated for \nkoi\ KOIs are provided in an
-online table accompanying this paper.
-The first few lines of this table is shown below
-(table \ref{tab:koi_periods}).
-Samples of the rotation period posterior PDFs are available online:
-\url{https://doi.org/10.5281/zenodo.804440}}
-
-\begin{table*}
-\begin{center}
-    \caption{Rotation periods and \Kepler\ input catalogue (KIC) properties
-    for the \Kepler\ objects of interest. The full table is available online.}
-\begin{tabular}{lcccc}
-    KOI number & Period (days) & KIC T$_{\mathrm{eff}}$ & KIC $\log(g)$ & KIC
-    $[$Fe$/$H$]$ \\
-    \hline
-    1 & 25$\pm1$ & $5126^{+160}_{-138}$ & $4.61^{+0.03}_{-0.1}$ &
-    $-0.1\pm0.3$ \\
-    2 & 20$\pm1$ & $5850\pm50$ & $4.46\pm0.03$ & $-0.2\pm0.1$ \\
-    3 & 28$\pm1$ & $6350\pm80$ & $4.02\pm0.01$ & $0.26\pm0.08$\\
-    6 & 23$\pm1$ & $6225^{+114}_{-158}$ & $4.17^{+0.06}_{-0.05}$ &
-    $-0.0^{+0.1}_{-0.2}$\\
-    7 & 25$\pm1$ & $5543\pm79 $ & $4.08\pm0.01$ & $0.4\pm0.1$ \\
-    ... & ... & ... & ... & ... \\
-\end{tabular}
-\end{center}
-\end{table*}
-\label{tab:koi_periods}
+
+% \response{The new GP rotation periods calculated for \nkoi\ KOIs are provided in an
+% online table accompanying this paper.
+% The first few lines of this table is shown below
+% (table \ref{tab:koi_periods}).
+% Samples of the rotation period posterior PDFs are available online:
+% \url{https://doi.org/10.5281/zenodo.804440}}
+
+% \begin{table*}
+% \begin{center}
+%     \caption{Rotation periods and \Kepler\ input catalogue (KIC) properties
+%     for the \Kepler\ objects of interest. The full table is available online.}
+% \begin{tabular}{lcccc}
+%     KOI number & Period (days) & KIC T$_{\mathrm{eff}}$ & KIC $\log(g)$ & KIC
+%     $[$Fe$/$H$]$ \\
+%     \hline
+%     1 & 25$\pm1$ & $5126^{+160}_{-138}$ & $4.61^{+0.03}_{-0.1}$ &
+%     $-0.1\pm0.3$ \\
+%     2 & 20$\pm1$ & $5850\pm50$ & $4.46\pm0.03$ & $-0.2\pm0.1$ \\
+%     3 & 28$\pm1$ & $6350\pm80$ & $4.02\pm0.01$ & $0.26\pm0.08$\\
+%     6 & 23$\pm1$ & $6225^{+114}_{-158}$ & $4.17^{+0.06}_{-0.05}$ &
+%     $-0.0^{+0.1}_{-0.2}$\\
+%     7 & 25$\pm1$ & $5543\pm79 $ & $4.08\pm0.01$ & $0.4\pm0.1$ \\
+%     ... & ... & ... & ... & ... \\
+% \end{tabular}
+% \end{center}
+% \end{table*}
+% \label{tab:koi_periods}
 
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_koi_02_03.pdf}
+\includegraphics[width=6in, clip=true]{figures/comparison_koi_01_23.pdf}
 \caption[Comparison with McQuillan results.]

### 2017-06-22 code/plot_koi.py
changing errorbar bit

@@ -18,18 +18,20 @@ plt.rcParams.update(plotpar)
 
 def load_and_plot(data):
 
-    recovered, errp, errm, lnerrp, lnerrm = [np.zeros(len(data.koi_id.values))
-                                             for i in range(5)]
+    recovered, errp, errm, lnerrp, lnerrm = [], [], [], [], []
     for i, kid in enumerate(data.koi_id.values):
         fn = os.path.join(RESULTS_DIR, "KOI-{}.h5".format(int(kid)))
         if os.path.exists(fn):
             df = pd.read_hdf(fn, key="samples")
-            ln_p = np.median(df.ln_period.values)
-            recovered[i] = np.exp(ln_p)
-            lnerrp[i] = np.percentile(df.ln_period.values, 84) - ln_p
-            lnerrm[i] = ln_p - np.percentile(df.ln_period.values, 16)
-            errp[i] = np.exp(lnerrp[i]/ln_p)
-            errm[i] = np.exp(lnerrm[i]/ln_p)
+
+            samps = np.exp(df.ln_period.values)
+            recovered[i] = np.median(samps)
+            upper = np.percentile(samps, 84)
+            lower = np.percentile(samps, 16)
+            errp.append(upper - recovered[i])
+            errm.append(recovered[i] - lower)
+
+    errp, errm = np.array(errp), np.array(errm)
 
     x = data.P_rot.values
     xerr = data.P_rot_err.values
@@ -46,23 +48,23 @@ def load_and_plot(data):
     plt.plot(xs, xs - 2./3, "--", color=".7", lw=.8, zorder=0)
 
     print(lnerrp[l])
-    assert 0
     plt.errorbar(np.log(x[l]), np.log(recovered[l]), yerr=[lnerrp[l],
                  lnerrm[l]], xerr=lnxerr[l], fmt="k.", zorder=1, capsize=0,
                  ecolor=".7", alpha=.5, ms=.1, elinewidth=.8)
-    plt.scatter(np.log(x[l]), np.log(recovered[l]), c=np.log(amp[l]),
-                edgecolor=".5", cmap="GnBu_r", vmin=min(np.log(amp[l])),
-                vmax=max(np.log(amp[l])), s=10, zorder=2, lw=.2)
[…]

### 2017-06-22 documents/ms.tex
removing the table stubb and updating the table

@@
 
-\newcommand{\nkoi}{1132}
+\newcommand{\nkoi}{1102}
 
@@
     periods.
-    % \response{We provide a table of rotation periods for these \nkoi\ \Kepler\ objects of
-    % interest and their posterior probability density function samples.}
+    \response{We provide a table of rotation periods for these \nkoi\ \Kepler\ objects of
+    interest and their posterior probability density function samples.}
 Because this method delivers posterior probability density functions, it will
@@
 \ref{tab:koipriors}.
-% The posterior samples of the model parameters for the \Kepler\ objects of
-% interest are available online:
-% \url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
+The posterior samples of the model parameters for the \Kepler\ objects of
+interest are available online:
+\url{https://zenodo.org/record/292340\#.WKWpiBIrJE4}.
 
@@
 % measurement bias discussed in Section \ref{sec:acf}.
-
-% \response{The new GP rotation periods calculated for \nkoi\ KOIs are provided in an
-% online table accompanying this paper.
+\response{The new GP rotation periods calculated for \nkoi\ KOIs are provided in an
+online table accompanying this paper.
 % The first few lines of this table is shown below
 % (table \ref{tab:koi_periods}).
-% Samples of the rotation period posterior PDFs are available online:
-% \url{https://doi.org/10.5281/zenodo.804440}}
+Samples of the rotation period posterior PDFs are available online:
+\url{https://doi.org/10.5281/zenodo.804440}}
 

### 2017-06-22 documents/ms.tex
removing figures in path for arxiv

@@
 \begin{center}
-\includegraphics[width=\columnwidth, clip=true]{figures/koi_lc_demo.pdf}
+\includegraphics[width=\columnwidth, clip=true]{koi_lc_demo.pdf}
 \caption[A light curve with a GP model.]
@@
 \begin{center}
-\includegraphics[width=\columnwidth, clip=true]{figures/period_hist.pdf}
+\includegraphics[width=\columnwidth, clip=true]{period_hist.pdf}
 \caption{A histogram of the rotation periods used to generate the \naigrain\
@@
 \begin{center}
-\includegraphics[width=\columnwidth, clip=true]{figures/demo_lc.pdf}
+\includegraphics[width=\columnwidth, clip=true]{demo_lc.pdf}
 \caption[A simulated light curve.]
@@
 \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_03_10.pdf}
+% \includegraphics[width=6in, clip=true]{compare_mcmc.pdf}
+\includegraphics[width=6in, clip=true]{comparison_acfprior_03_10.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_noprior_02_13.pdf}
+\includegraphics[width=6in, clip=true]{comparison_noprior_02_13.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/corner_plot.pdf}
+\includegraphics[width=6in, clip=true]{corner_plot.pdf}
 \caption{Marginal posterior PDFs of the QP GP model parameters, fit to the
@@
 \begin{center}
-\includegraphics[width=\columnwidth, clip=true]{figures/demo_ACF.pdf}
+\includegraphics[width=\columnwidth, clip=true]{demo_ACF.pdf}
 \caption[ACF of a simulated light curve.]
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/compare_acf.pdf}
+\includegraphics[width=6in, clip=true]{compare_acf.pdf}
 \caption[ACF results.]
@@
 \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
-\includegraphics[width=6in, clip=true]{figures/compare_pgram_filtered_35.pdf}
+% \includegraphics[width=6in, clip=true]{compare_pgram.pdf}
+\includegraphics[width=6in, clip=true]{compare_pgram_filtered_35.pdf}
 \caption[LS periodogram results.]
@@
 \begin{center}
-\includegraphics[width=6in, clip=true]{figures/comparison_koi_01_23.pdf}
+\includegraphics[width=6in, clip=true]{comparison_koi_01_23.pdf}
 \caption[Comparison with McQuillan results.]

### 2017-07-06 documents/ms.tex
second referee response

--- 2017-06-22 documents/ms.tex
@@
     periods.
-    \response{We provide a table of rotation periods for these \nkoi\ \Kepler\ objects of
-    interest and their posterior probability density function samples.}
+    \response{We provide a table of rotation periods for these and many
+    more, altogether \nkoi \Kepler\ objects of interest and their posterior
+    probability density function samples.}
 Because this method delivers posterior probability density functions, it will
@@
 of galactic archaeology and exoplanet population studies which require
-rotation period (or age) distributions.}
+rotation period (or age) distributions.
+It is important to note however that gyrochronology may have limited
+applicability for galactic field stars: growing evidence points to weakened
+magnetic braking at late ages \citep{Angus2015, vanSaders2016}.}
 
@@
 \begin{center}
-\caption{Ranges and distributions of parameters used to simulate light curves
-in \citet{Aigrain2015}}
+\caption{\label{tab:simulation_parameters} Ranges and distributions of
+    parameters used to simulate light curves in \citet{Aigrain2015}}
 \begin{tabular}{lcc}
@@
 \end{table*}
-\label{tab:simulation_parameters}
 
@@
 to each of these 333 simulated light curves.  As discussed in Section
-\ref{sec:nonperiod_prior}, reliable inference
+\ref{sec:GP_prior}, reliable inference
 requires defining a useful set of priors on the non-period
@@
 \begin{center}
-\caption{Priors and bounds on the natural logarithms of the GP model
-    parameters.}
+\caption{\label{tab:priors} Priors and bounds on the natural logarithms of
+    the GP model parameters.}
 \begin{tabular}{lcc}
@@
 \end{table*}
-\label{tab:priors}
 
@@
 \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/compare_mcmc.pdf}
+% \includegraphics[width=6in, clip=true]{compare_mcmc.pdf}
 \includegraphics[width=6in, clip=true]{figures/comparison_acfprior_03_10.pdf}
@@
 \begin{center}
-% \includegraphics[width=6in, clip=true]{figures/compare_pgram.pdf}
+% \includegraphics[width=6in, clip=true]{compare_pgram.pdf}
 \includegraphics[width=6in, clip=true]{figures/compare_pgram_filtered_35.pdf}
@@
 \begin{center}
-    \caption{Median absolute (MAD), median relative (MRD) deviations and
-    root-mean-squared (RMS) for
-    the LS periodogram, \citet{Mcquillan2013} ACF and GP period recovery
-    methods.}
+    \caption{\label{tab:MADs} Median absolute (MAD), median relative (MRD)
+    deviations and root-mean-squared (RMS) for the LS periodogram,
+    \citet{Mcquillan2013} ACF and GP period recovery methods.}
 \begin{tabular}{lcccc}
@@
 \end{table*}
-\label{tab:MADs}
 
@@
 \begin{center}
-\caption{Priors and bounds on the natural logarithms of the GP model parameters,
-        for \Kepler\ light curves}
+\caption{\label{tab:koipriors} Priors and bounds on the natural logarithms of
+    the GP model parameters, for \Kepler\ light curves}
 \begin{tabular}{lcc}
@@
 \end{table*}
-\label{tab:koipriors}
 
@@
 
-\response{The majority of long period outliers seen in the periodogram comparison
-figure (figure \ref{fig:pgram_compare}) are produced when the rotation period
-signal is highly non-sinusoidal.
+\response{Apon examining the light curves that produce incorrect, long period
+measurements seen in the periodogram comparison figure (figure
+\ref{fig:pgram_compare}), it appears that the majority of these
+overpredictions are produced when the rotation period signal is highly
+non-sinusoidal.
+In these cases the peak corresponding to the true period has a reduced
+amplitude and is not the highest peak in the periodogram.
 This can be understood intuitively as the LS periodogram fits a series of sine
@@
 produce a good fit.
+There may be ways to further improve the precision of the periodogram method,
+for example modelling the detailed shapes of peaks in the periodogram instead
+of adopting the point closest to the derivative zero-point.
+We consider it likely that such alterations would produce only slight changes
+to the results, for example because peaks in the LS periodograms of Kepler
+light curves are usually narrow.
 It is difficult to assess the reason for the ACF incorrect recoveries because
@@
 Nevertheless, it is still a somewhat arbitrary choice.
-Another valid choice would be a squared cosine function multiplied by a
+Another valid choice would be a cosine function multiplied by a
 squared exponential, used by \citet{Brewer2009} to model asteroseismic

### 2017-07-07 documents/ms.tex
fixing typos. About to submit

@@
     \response{We provide a table of rotation periods for these and many
-    more, altogether \nkoi \Kepler\ objects of interest and their posterior
+    more, altogether \nkoi\ \Kepler\ objects of interest, and their posterior
     probability density function samples.}
@@
 of adopting the point closest to the derivative zero-point.
-We consider it likely that such alterations would produce only slight changes
-to the results, for example because peaks in the LS periodograms of Kepler
-light curves are usually narrow.
+However, we consider it likely that such alterations would produce only slight
+changes to the results, for example because peaks in the LS periodograms of
+Kepler light curves are usually narrow.
 It is difficult to assess the reason for the ACF incorrect recoveries because

### 2017-07-07 gprot/fit.py

@@ -24,6 +24,7 @@ class Emcee3Model(emcee3.Model):
 def write_samples(mod, df, resultsdir='results', true_period=None):
     """df is dataframe of samples, mod is model
     """
+    print("writing samples")
 
     if not os.path.exists(resultsdir):
         os.makedirs(resultsdir)
@@ -41,7 +42,7 @@ def write_samples(mod, df, resultsdir='results', true_period=None):
     figfile = os.path.join(resultsdir, '{}.png'.format(mod.name))
     try:
         if true_period is None:
-            true_period = (np.log(mod.lc.sim_params.P_MIN), 
+            true_period = (np.log(mod.lc.sim_params.P_MIN),
                            np.log(mod.lc.sim_params.P_MAX))
     except AttributeError:
         pass
@@ -53,7 +54,7 @@ def fit_emcee3(mod, nwalkers=500, verbose=False, nsamples=5000, targetn=6,
                 iter_chunksize=10, pool=None, overwrite=False,
                 maxiter=100, sample_directory='mcmc_chains',
                 nburn=3, mixedmoves=True, resultsdir='results', **kwargs):
-    """fit model using Emcee3 
+    """fit model using Emcee3
 
     modeled after https://github.com/dfm/gaia-kepler/blob/master/fit.py
 
@@ -104,7 +105,7 @@ def fit_emcee3(mod, nwalkers=500, verbose=False, nsamples=5000, targetn=6,
         neff = s.backend.niter / tau_max - nburn
         if verbose:
             print("Maximum autocorrelation time: {0}".format(tau_max))
-            print("N_eff: {0}\n".format(neff * nwalkers))            
+            print("N_eff: {0}\n".format(neff * nwalkers))
         return tau_max, neff
 
     done = False
@@ -147,11 +148,11 @@ def fit_emcee3(mod, nwalkers=500, verbose=False, nsamples=5000, targetn=6,
 
     df = pd.DataFrame(samples, columns=mod.param_names)
     write_samples(mod, df, resultsdir=resultsdir)
-    
+
     return df
     # return sampler
 
-def fit_mnest(mod, basename=None, test=False, 
+def fit_mnest(mod, basename=None, test=False,
                 verbose=False, resultsdir='results', overwrite=False, **kwargs):
     import pymultinest
 
@@ -163,14 +164,14 @@ def fit_mnest(mod, basename=None, test=False,
     if test:
         print('Will run multinest on star {}..., basename={}'.format(i, basename))
     else:
[…]

### 2017-07-07 documents/ms.tex
removing figures path

@@
 % \includegraphics[width=6in, clip=true]{compare_mcmc.pdf}
-\includegraphics[width=6in, clip=true]{figures/comparison_acfprior_03_10.pdf}
+\includegraphics[width=6in, clip=true]{comparison_acfprior_03_10.pdf}
 \caption{The `true' rotation periods used to generate \naigrain\
@@
 % \includegraphics[width=6in, clip=true]{compare_pgram.pdf}
-\includegraphics[width=6in, clip=true]{figures/compare_pgram_filtered_35.pdf}
+\includegraphics[width=6in, clip=true]{compare_pgram_filtered_35.pdf}
 \caption[LS periodogram results.]

### 2017-08-03 documents/ms.tex
adding the table stubb back in

--- 2017-07-07 documents/ms.tex
@@
 \usepackage{newtxtext,newtxmath}
+
+% Figure list:
+% koi_lc_demo.pdf
+% period_hist.pdf
+% demo_lc.pdf
+% comparison_acfprior_03_10.pdf
+% comparison_noprior_02_13.pdf
+% corner_plot.pdf
+% demo_ACF.pdf
+% compare_acf.pdf
+% compare_pgram_filtered_35.pdf
+% comparison_koi_01_23.pdf
 
@@
 $\sim$ \kepexampleperiod\ days.
-The blue line shows a fit to the data using a Gaussian process model with a QP
+The solid line shows a fit to the data using a Gaussian process model with a QP
 covariance kernel function.}
@@
     simulated light curve in figure \ref{fig:demo_lc}.
-    The blue line in the period panel shows the injected period.
+    The solid line in the period panel shows the injected period.
     This figure was made using \texttt{corner.py} \citep{Corner}.}
@@
 \ref{fig:demo_lc}.
-The vertical blue line shows the period measured using the ACF method (20.4
+The vertical solid blue line shows the period measured using the ACF method (20.4
 days) and the pink dashed line shows the period that was used to simulate the
@@
 online table accompanying this paper.
-% The first few lines of this table is shown below
-% (table \ref{tab:koi_periods}).
+The first few lines of this table is shown below
+(table \ref{tab:koi_periods}).
 Samples of the rotation period posterior PDFs are available online:
@@
 
-% \begin{table*}
-% \begin{center}
-%     \caption{Rotation periods and \Kepler\ input catalogue (KIC) properties
-%     for the \Kepler\ objects of interest. The full table is available online.}
-% \begin{tabular}{lcccc}
-%     KOI number & Period (days) & KIC T$_{\mathrm{eff}}$ & KIC $\log(g)$ & KIC
-%     $[$Fe$/$H$]$ \\
-%     \hline
-%     1 & 25$\pm1$ & $5126^{+160}_{-138}$ & $4.61^{+0.03}_{-0.1}$ &
-%     $-0.1\pm0.3$ \\
-%     2 & 20$\pm1$ & $5850\pm50$ & $4.46\pm0.03$ & $-0.2\pm0.1$ \\
-%     3 & 28$\pm1$ & $6350\pm80$ & $4.02\pm0.01$ & $0.26\pm0.08$\\
-%     6 & 23$\pm1$ & $6225^{+114}_{-158}$ & $4.17^{+0.06}_{-0.05}$ &
-%     $-0.0^{+0.1}_{-0.2}$\\
-%     7 & 25$\pm1$ & $5543\pm79 $ & $4.08\pm0.01$ & $0.4\pm0.1$ \\
-%     ... & ... & ... & ... & ... \\
-% \end{tabular}
-% \end{center}
-% \end{table*}
-% \label{tab:koi_periods}
+\begin{table*}
+\begin{center}
+    \caption{Rotation periods and \Kepler\ input catalogue (KIC) properties
+    for the \Kepler\ objects of interest. The full table is available online.}
+\begin{tabular}{lcccc}
+    KOI number & Period (days) & KIC T$_{\mathrm{eff}} (K)$ & KIC $\log(g)$ &
+    KIC $[$Fe$/$H$]$ \\
+    \hline
+1 & 25$^{+23}_{-6}$ & $5126^{+160}_{-138}$ & $4.61^{+0.03}_{-0.11}$ & $-0.12\pm0.3$\\
+2 & 20$\pm4$ & $5850\pm50$ & $4.46\pm0.03$ & $-0.15\pm0.1$\\
+3 & 28$^{+3}_{-0.2}$ & $6350\pm80$ & $4.02\pm0.01$ & $0.26\pm0.08$\\
+6 & 23$^{+3}_{-15}$ & $6225^{+114}_{-158}$ & $4.17^{+0.06}_{-0.05}$ & $-0.04^{+0.1}_{-0.2}$\\
+7 & 25$\pm2$ & $5543\pm79$ & $4.08\pm0.01$ & $0.44\pm0.1$\\
+    ... & ... & ... & ... & ... \\
+\end{tabular}
+\end{center}
+\end{table*}
+\label{tab:koi_periods}
 
@@
     \citet{Mcquillan2013}, defined as the interval between the 5th and 95th
-    percentiles of normalised flux per period bin in millimagnitudes.}
+    percentiles of normalized flux per period bin in millimagnitudes.}
 \label{fig:mcquillan}
@@
 
-\bibliographystyle{plainnat}
+% \bibliographystyle{plainnat}
+\bibliographystyle{hapj}
 \bibliography{GProtation}

### 2017-08-03 code/make_period_table.py

@@ -63,11 +63,9 @@ def get_kic_properties(kois):
 def make_tex_table(kois, p, errp, errm, teff, teff_errp, teff_errm, logg,
                    logg_errp, logg_errm, feh, feh_errp, feh_errm):
     with open("period_table.tex", "w") as f:
-        f.write("KOI & period (days) \\\ \n")
-        f.write("\\hline \n")
         for i, koi in enumerate(kois[:10]):
-            f.write("{0} & {1:.2f}$^{{+{2:.1g}}}_{{-{3:.1g}}}$ & "
-                    "${4}^{{+{5}}}_{{{6}}}$ & ${7:.2f}^{{+{8:.2g}}}_{{{9:.2f}}}$ & "
+            f.write("{0} & {1:.0f}$^{{+{2:.1f}}}_{{-{3:.1f}}}$ & "
+                    "${4}^{{+{5}}}_{{{6}}}$ & ${7:.2f}^{{+{8:.2f}}}_{{{9:.2f}}}$ & "
                     "${10}^{{+{11:.1g}}}_{{{12:.1g}}}$"
                     "\\\ \n"
                     .format(koi, p[i], errp[i], errm[i], int(teff[i]),

### 2017-08-09 documents/ms.tex
getting ready to resubmit

@@
 Since there are multiple signals present in these light curves, this formal
-uncertainty is an underestimate of the true uncertainty.
+uncertainty is a noisy estimate of the true uncertainty.
 }
@@
     \caption{Rotation periods and \Kepler\ input catalogue (KIC) properties
-    for the \Kepler\ objects of interest. The full table is available online.}
+    for \nkoi\ \Kepler\ objects of interest. The full table is available online.}
 \begin{tabular}{lcccc}
-    KOI number & Period (days) & KIC T$_{\mathrm{eff}} (K)$ & KIC $\log(g)$ &
+    KOI number & Period & KIC T$_{\mathrm{eff}} $ & KIC $\log(g)$ &
     KIC $[$Fe$/$H$]$ \\
+    & (days) & (K) & dex & dex \\
     \hline
